### 2020-10-21 paper/paper.tex
paper outline

%
% File acl2020.tex
%
%% Based on the style files for ACL 2020, which were
%% Based on the style files for ACL 2018, NAACL 2018/19, which were
%% Based on the style files for ACL-2015, with some improvements
%%  taken from the NAACL-2016 style
%% Based on the style files for ACL-2014, which were, in turn,
%% based on ACL-2013, ACL-2012, ACL-2011, ACL-2010, ACL-IJCNLP-2009,
%% EACL-2009, IJCNLP-2008...
%% Based on the style files for EACL 2006 by 
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[11pt,a4paper]{article}
\usepackage[hyperref]{acl2020}
\usepackage{times}
\usepackage{latexsym}
\renewcommand{\UrlFont}{\ttfamily\small}

% This is not strictly necessary, and may be commented out,
% but it will improve the layout of the manuscript,
% and will typically save some space.
\usepackage{microtype}

%\aclfinalcopy % Uncomment this line for the final submission
%\def\aclpaperid{***} %  Enter the acl Paper ID here

%\setlength\titlebox{5cm}
% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

\newcommand\BibTeX{B\textsc{ib}\TeX}

\title{Community-Conditioned Language Models}

\author{Anon.}

\date{}

\begin{document}
\maketitle
\begin{abstract}
\end{abstract}

\section{Introduction}

We propose to test the following hypotheses:

 - Different communities have different levels of linguistic complexity
 - different language models are able to recognize such differences
 - level at which the community embedding is taken into account in the language model influences this.
 - community vectors are correlated with extra-linguistic information.
 - stability is correlated with language complexity

 \section{Experimental setup}
 \subsection{Data Sets}
 \subsection{Models}
 \subsection{Training scheme}

 \section{Experimental setup}
 \section{Results and analysis}

- Computation of loss/community/model (show all)
- Computation of confusion matrix for community inference per model (eg. show only best and worst model) (on test set)
- Computation of H(c,m) entropy of community identification (on test set)
- Correlation between H(c,m) for various models. Identification of specific communities by hand
- Correlation between H(c,m) and Stability(c).

\section{Discussion and Conclusion}

- Comparison with topic modelling

\end{document}

### 2020-10-21 paper/paper.tex
analysis notes

@@
 
- - Different communities have different levels of linguistic complexity
- - different language models are able to recognize such differences
- - level at which the community embedding is taken into account in the language model influences this.
- - community vectors are correlated with extra-linguistic information.
- - stability is correlated with language complexity
+\begin{enumerate}
+\item Different communities have different levels of linguistic
+  complexity
+\item different language models are able to recognize such differences
+\item level at which the community embedding is taken into account in the language model influences this.
+\item community vectors are correlated with extra-linguistic information.
+\item stability is correlated with language complexity
+\end{enumerate}
 
@@
 
- \section{Experimental setup}
 \section{Results and analysis}
 
-- Computation of loss/community/model (show all)
-- Computation of confusion matrix for community inference per model (eg. show only best and worst model) (on test set)
-- Computation of H(c,m) entropy of community identification (on test set)
+% Computation of loss/community/model (show all)
+\begin{table}
+  \centering
+  \begin{tabular}{ccc}
+    todo &todo &todo 
+  \end{tabular}
+  \caption{Model losses per community}
+  \label{tab:losses}
+\end{table}
+TODO commentary
+
+TODO: say why we can't use the losses directly!
+Turn a CCLM into a community classifier using Bayes theorem: The
+probability that a given post $m$ belongs to a community $c_i$ can be
+calculated as:
+\[P(c=c_i | m) = P(m | c_i)\frac {P(m)} {P(c=c_i)}\]
+
+$P(c=c_i)$ is the frequency of community $c_i$ in the dataset, and
+$P(m | c_i)$ is given by the CCLM. $P(m)$ is our normalisation factor, which can be set so that
+\[\sum_i P(c=c_i | m) = 1\].
+
+Knowing that $m$ \emph{actually} comes from community $c_j$, we test
+the recall and precision abilities of our classifier.  if we let
+$C_{ij} = average_{m ∈ Posts(c_j)}(P(c=c_i | m))$ then $C$ is a
+confusion matrix for it.
+
+(TODO: on what set?)
+
+TODO: show a few relevant matrices. (eg. show only best and worst model) and commentary.
+
+TODO: why don't we use F scores or such bullshit?
+
+TODO:Define \(D_j\) as the discrete probability distribution given by \(C_{ij}\) for fixed \(j\).
+
+We then compute the perplexity of the classifier, per given community
+$j$, as the exponential of the entropy of the distribution \(D_j\):
+
+\[Ppl_j = e^{-\sum_i C_{ij} log(C_{ij})}\]
+
+\(Ppl_j\) is a measure of the capability (or inability) of the model
+to correctly identify a message as coming from a certain
+community. (ie. it aggregates confusion on a column of the matrix).
+
+We can then analyse the correlation between \(Ppl^M_j\) for various models $M$.
+
 - Correlation between H(c,m) for various models. Identification of specific communities by hand

### 2020-10-23 paper/paper.tex
finish drafting analysis section

@@
 
+\subsection{Losses of CCLM}
+
 % Computation of loss/community/model (show all)
@@
 \end{table}
-TODO commentary
+TODO commentary. Why are some communities low/high.
 
-TODO: say why we can't use the losses directly!
-Turn a CCLM into a community classifier using Bayes theorem: The
-probability that a given post $m$ belongs to a community $c_i$ can be
-calculated as:
+\subsection{Analysis of community embeddings}
+
+- PCA plots: not conclusive:
+
+- Correlation with co-occurence matrix (not conclusive)
+
+- Analysis: probably because the model is not using dimensions
+with \emph{a priori} equal variance in all dimensions. (Both the LSTM and transformer are highly non-linear complex models)
+
+\subsection{Implicit language-model-based community classifiers}
+
+(TODO: Insert a the correlation plot between two models, on the exponential scale so that it's the same as the other later plots)
+TODO: So we can't use the losses directly. Why is this so?
+
+However, we are not so much interested in the ability of the CCLM to
+precisely model the language used in a community.
+
+We now turn our interest to language-model-based community classifiers
+(LMCC). Given a post $m$, and LMCC gives the probability that it
+belongs in a community $c_i$ Using the Bayes theorem, we can turn a
+CCLM into a LMCC. Indeed, the probability that a given post $m$
+belongs to a community $c_i$ can be calculated as follows:
 \[P(c=c_i | m) = P(m | c_i)\frac {P(m)} {P(c=c_i)}\]
-
+In the above,
 $P(c=c_i)$ is the frequency of community $c_i$ in the dataset, and
@@
 
-Knowing that $m$ \emph{actually} comes from community $c_j$, we test
-the recall and precision abilities of our classifier.  if we let
-$C_{ij} = average_{m ∈ Posts(c_j)}(P(c=c_i | m))$ then $C$ is a
-confusion matrix for it.
-
-(TODO: on what set?)
+Knowing that $m$ \emph{actually} comes from community $c_j$, we can
+measure the the precision our classifier, as a confusion matrix $C$,
+which can be calculated as follows for a given set of posts:(TODO: what set exactly?)
+\[C_{ij} = average_{m ∈ Posts(c_j)}(P(c=c_i | m))\].
 
@@
 
-TODO: why don't we use F scores or such bullshit?
+TODO: why don't we use F scores?
 
-TODO:Define \(D_j\) as the discrete probability distribution given by \(C_{ij}\) for fixed \(j\).
-
-We then compute the perplexity of the classifier, per given community
-$j$, as the exponential of the entropy of the distribution \(D_j\):
-
+While the confusion matrices provide a lot of insight, they are
+somewhat \emph{too} fine-grained. For this reason, we will be
+interested in confusing the posts from community $c_j$ are to the
+LMCC. For this purpose, we consider the perplexity of the LMCC for a
+post coming from community $j$, defined as the exponential of the
+entropy of the distribution \(D_j\), which is the (discrete)
+probability distribution of predicted community for a random post from
+$c_j$. $D_j$ is simply the $j$th row in \(C\), and so we obtain the formula:
 \[Ppl_j = e^{-\sum_i C_{ij} log(C_{ij})}\]
 
-\(Ppl_j\) is a measure of the capability (or inability) of the model
-to correctly identify a message as coming from a certain
-community. (ie. it aggregates confusion on a column of the matrix).
+To recap, \(Ppl_j\) is a measure of the capability (or rather
+inability) of the model to correctly identify a message as coming from
+a certain community $c_j$.
 
@@
 
-- Correlation between H(c,m) for various models. Identification of specific communities by hand
-- Correlation between H(c,m) and Stability(c).
+TODO: show a scatter plot, compute r2, analyse out-of-balance points; discuss why they stand where they do.
+
+Finally, we investigate the correlation between the stability of a
+community and the perplexity of the LMCC for it. Stability is defined as TODO.
+
+- computer r2.
+
+We see a positive correlation: this suggests that the more stable a
+community is, the more difficult it is to identify posts as coming
+from it. What this suggests is that stable communities tend to use a
+varied and standard, subset of English, while unstable communities
+resort to more formulaic posts. This may suggests that to 'fit' in an
+unstable community, one must use more obvious linguistic cues than in
+a stable community. In fact, it is reasonable to think that a new
+member will ostensibly use community-specifc language, but once
+established in a community, a member would tend not to do so.
 

### 2020-10-27 paper/paper.tex
add Data sets section

@@
 \usepackage[hyperref]{acl2020}
+\usepackage[T1]{fontenc}
 \usepackage{times}
 \usepackage{latexsym}
+\usepackage{siunitx}
 \renewcommand{\UrlFont}{\ttfamily\small}
@@
 \subsection{Data Sets}
+
+To investigate variation across communities, we use comments collected 
+from the social media website Reddit.\footnote{Comments were obtained 
+from the archive at \href{https://pushshift.io/}{pushshift.io} \cite{Baumgartner2020}.}
+
+Reddit is divided into forums called \textit{subreddits}, 
+which are typically organized around a topic of interest. 
+Our dataset includes comments from \num{46} subreddits, 
+randomly selected from the set of all forums 
+with at least \num{15000} comments per month for each month 
+in the three years between 2015 and 2017.\footnote{
+We initially selected \num{50} subreddits, 
+but excluded four from further analysis: 
+two which were primarily non-English and two with particularly short average comment lengths.}
+Each community corpus consist of \num{50000} randomly selected comments from the year 2015.
+
+Comments were preprocessed as follows: 
+We removed markdown formatting, extracting only rendered text. 
+We exclude the content of block quotes, code blocks, and tables. 
+We tokenized the comments using the default English model for the SpaCy tokenizer,\footnote{Version 2.2.3}
+and lower-cased all tokens.
+
 \subsection{Models}
@@
 
+\bibliography{paper}
+\bibliographystyle{acl_natbib}
+
 \end{document}

### 2020-10-27 paper/paper.tex
Start of model section

@@
 \section{Experimental setup}
-\subsection{Data Sets}
+
+\subsection{Data sets}
 
@@
 \subsection{Models}
+
+At the the core of the community-conditioned language model (CCLM) 
+is a standard $n$-layer neural sequence encoder 
+with token embedding (we experiment with LSTM and transformer encoders).
+To condition the model, we additionally train a \emph{community embedding},
+which is concatenated with the hidden layer of the sequence encoder
+for some layer $l_c \leq n$ and passed through a linear layer, 
+which projects the resulting vector back down to the origial hidden size.
+For $l_c=0$, the community embedding is concatenated directly to the token embedding.
+For this reason, the hidden size of the sequence encoder 
+and the size of the token embedding are always the same.
+
 \subsection{Training scheme}

### 2020-10-29 paper/paper.tex
training scheme

@@
 
+We combined the 46 community corproa and split the combined corpus,
+reserving 10\% for testing, 10\% for validation and the rest for training.
+Splits were stratified by community so that the same number of training examples
+were available for each community.
+
+We used a vocabulary size of \num{40000} tokens, with a special out-of-vocabulary token.
+The vocabulary consisted of the most frequent tokens at the comment level,
+meaning that we only counted each token once per comment.
+The main reason for this was to avoid over-counting otherwise rare tokens that were spammed
+many times in a single comment.
+
+We trained the models on a simple left-to-right language modeling task 
+with cross entropy loss over the vocabulary 
+and AdamW \citep{Loshchilov2019} optimization, with an inital learning rate of \num{0.001}.
+We used a batch size of \num{256} and a maximum sequence length of \num{64} tokens.
+During training, a dropout rate of $0.1$ was applied between encoder layers 
+and after each linear layer.
+
+All experiments used models with \num{3} sequence encoder layers, 
+each with hidden (and token embedding) size of \num{256}. 
+The transformer models had \num{8} attention heads per layer.\footnote{
+  This number of attention heads was chosen to give the LSTM and transformer
+  models a comparable number of paramters 
+  (\num{22171203} and \num{21779523}, respectively).}
+The conditional models were given a community embedding with \num{16} dimensions.
+
+We trained the models until the validation loss was smaller than the smallest
+validation loss for two epochs in a row. 
+The weights from the epoch with the smallest validation loss were used for testing.
+
 \section{Results and analysis}

### 2020-10-30 paper/paper.tex
some tables
add really big table
Losses of CCLM

@@
 \usepackage{siunitx}
+\usepackage{booktabs}
 \renewcommand{\UrlFont}{\ttfamily\small}
@@
 which is concatenated with the hidden layer of the sequence encoder
-for some layer $l_c \leq n$ and passed through a linear layer, 
+for some layer $c \leq n$ and passed through a linear layer, 
 which projects the resulting vector back down to the origial hidden size.
-For $l_c=0$, the community embedding is concatenated directly to the token embedding.
-For this reason, the hidden size of the sequence encoder 
-and the size of the token embedding are always the same.
+For $c = n$, the output of this linear layer is passed directly to the softmax function,
+just as the final hidden layer of the sequence encoder is in other models.
+For $c=0$, the community embedding is concatenated with the token embedding.
+For this reason, we set the hidden size of the sequence encoder 
+and the size of the token to be embedding equal for all models.
 
@@
 The transformer models had \num{8} attention heads per layer.\footnote{
-  This number of attention heads was chosen to give the LSTM and transformer
+  This number of attention heads was chosen in part to give the LSTM and transformer
   models a comparable number of paramters 
@@
 
+\begin{table}
+  \centering
+  \input{floats/best_epoch.tex}
+  \label{tab:best-epoch}
+  \caption{Epoch with the lowest validation loss.}
+\end{table}
+
 \section{Results and analysis}
@@
 
-% Computation of loss/community/model (show all)
-\begin{table}
+As one would expect, 
+the conditioned models have lower perplexity 
+than their respective unconditioned baseline models,
+but this effect is different in magnitude
+for models with different community embedding depths .
+For the LSTM encoder, 
+the best model ($c=1$) concatenates the community embedding after the first encoder layer,
+with two more LSTM layers after that.
+The best transformer model ($c=3$), on the other hand,
+incorporates the community information last,
+after all the sequence encoder layers
+(Table \ref{tab:model-ppls}).
+
+TODO: For space reasons, it might be better to put the community-level table in the appendix
+and have the table in body just show the means (and perhaps medians/stddevs?) for each architecture.
+
+We also note that the conditioned models perform differently across different communities---\
+for some communities, the test perplexity of the conditioned models is a lot lowe, 
+whereas for others it is the same or even slightly higher.
+(TODO: talk about this in terms of mutual information instead?)
+As an anecdotal observation, we note 
+that communities where conditioning has the least effect
+tend to be more organized around more ``general interest'' topics.
+This makes sense intuitively, since communities with more niche topics
+would tend to have more specialized vocabulary and speech patterns.
+
+The pattern of how much the perplexity falls across different communities
+is similar across architectures, but with some differences.
+It could be interesting to further investigate these differences in future work,
+since it could reveal differences in the kind of linguistic variation the different
+model architectures capture.
+
+\begin{table*}
+  \footnotesize
   \centering
-  \begin{tabular}{ccc}
-    todo &todo &todo 
-  \end{tabular}
-  \caption{Model losses per community}
-  \label{tab:losses}
-\end{table}
-TODO commentary. Why are some communities low/high.
+  \input{floats/model_ppls.tex}
+  \caption{Mean model test perplexity per community.}
+  \label{tab:model-ppls}
+\end{table*}
 

### 2020-10-30 paper/paper.tex
couple of questions

@@
 \usepackage{booktabs}
+\usepackage{todonotes}
 \renewcommand{\UrlFont}{\ttfamily\small}
+
+\newcommand\jp[1]{\todo[backgroundcolor=blue!10]{JP: #1}}
 
@@
 We also note that the conditioned models perform differently across different communities---\
-for some communities, the test perplexity of the conditioned models is a lot lowe, 
+for some communities, the test perplexity of the conditioned models is a lot lower,\jp{Where can I see this?}
 whereas for others it is the same or even slightly higher.
@@
 that communities where conditioning has the least effect
-tend to be more organized around more ``general interest'' topics.
+tend to be more organized around more ``general interest'' topics.\jp{Is there a way to quantify general interest topics? Community size in terms of number of users?}
 This makes sense intuitively, since communities with more niche topics

### 2020-10-30 paper/paper.tex
add pca plots

--- 2020-10-30 paper/paper.tex
@@
 \usepackage{booktabs}
+\usepackage{pgfplots}
 \renewcommand{\UrlFont}{\ttfamily\small}
+\usepgfplotslibrary{groupplots}
+\pgfplotsset{compat=1.17}
 
@@
 \newcommand\BibTeX{B\textsc{ib}\TeX}
+
+\input{plot_macros}
 
@@
 
-- PCA plots: not conclusive:
+As a qualitative assesment of the community embeddings, 
+we identified several topic categories into which at least three
+of the selected communities fall. 
+We plotted the two-component PCA projection of the embeddings,
+annotated with these topic categories.
+We see that there is some evidence of clustering according to
+the topic categories we identified,
+suggesting that the embedding space learned by the model is meaningful.
+
+\begin{figure*}
+\begin{tikzpicture}
+  \commpcagroupplot{floats/lstm-3-1_pca.csv}{floats/transformer-3-3_pca.csv}
+\end{tikzpicture}
+\caption{Community embedding PCA of the best LSTM (left, $c=1$) and transformer (right, $c=3$) models.}
+\label{fig:lstm-pca}
+\end{figure*}
 

### 2020-10-30 paper/plot_macros.tex
add pca plots

\newcommand{\commpcaplot}[1]{%
  \addplot[
    scatter, only marks,
    font=\tiny,
    scatter src=explicit symbolic,
    scatter/classes={
      general-interest={mark=*, blue},
      videogames={mark=*, red},
      technology={mark=*, green},
      sports={mark=*, cyan},
      female-focused={mark=*, magenta},
      other={mark=o, black}%
    },
    nodes near coords*={\community},
    visualization depends on={value \thisrow{community} \as \community}
  ] table [x=pca1, y=pca2, meta=category] {#1};
}

\newcommand{\commpcagroupplot}[2]{%
  \begin{groupplot}[
    group style={group size=2 by 1}, 
    yticklabels={,,}, xticklabels={,,},
    legend style={
      scatter legend, anchor=north, legend columns=1, font=\tiny
    }
    ]
    \nextgroupplot
    \commpcaplot{#1}
    \nextgroupplot
    \commpcaplot{#2}
    \legend{General interest,Videogames,Technology,Sports,Female-focused,Other}
  \end{groupplot}
}

### 2020-10-30 paper/paper.tex

--- 2020-10-30 paper/paper.tex
@@
 \usepackage{booktabs}
+\usepackage{pgfplots}
 \usepackage{todonotes}
 \renewcommand{\UrlFont}{\ttfamily\small}
+\usepgfplotslibrary{groupplots}
+\pgfplotsset{compat=1.17}
 

### 2020-10-31 eval_lm.py
fix comm inference bug
we were computing the P(c_j|m) by normalizing over (-log P(m|c)) vector
but that is *not* equivallent to normalizing P(m|c). Need to eponentiate
first. Problem: underflow since those probs are v. small. Needed
to use the exponentiation trick.

@@ -1,141 +1,0 @@
-from model import CommunityConditionedLM
-from data import load_data_and_fields
-import torch
-import torch.nn as nn
-import torch.nn.functional as F
-import torchtext as tt
-import pandas as pd
-import numpy as np
-import random
-import json
-import os
-import click
-
-@click.group()
-def cli():
-    pass
-
-def test_perplexity():
-
-    def perplexity_per_word(y, y_hat, lengths, text_pad_idx):
-        vocab_size = y_hat.shape[2]
-        nll = F.nll_loss(y_hat.view(-1,vocab_size), y.view(-1), ignore_index=text_pad_idx, reduction='none').view(y.shape)
-        nll_word = nll.sum(axis=0) / lengths.to(torch.float) # negative log-likelihood per word
-        return nll_word.exp() # perplexity per word
-
-    def test_ppl(lm, test_batches, text_pad_idx):
-        ppls = []
-        for i, batch in enumerate(test_batches):
-            print(f'{i+1}/{len(test_batches)}', end='\r')
-            x_comm = batch.community
-            text, lengths = batch.text
-            x_text = text[:-1]
-            y = text[1:]
-            y_hat = lm(x_text, x_comm)
-            ppl_word = perplexity_per_word(y, y_hat, lengths, text_pad_idx)
-            ppls += [ppl.item() for ppl in ppl_word]
-        return ppls
-
-    df = pd.DataFrame([(example.community, ' '.join(example.text)) 
-        for example in test_data], columns=['community', 'comment'])
-    models = {}, []
-    for name in architectures:
-        ppls = test_ppl(lm, test_iterator, text_pad_idx)
-        model_names.append(name)
-        models[name] = lm
-        print(f"Done {name}")
-        df[name] = ppls
-    df.to_pickle(results_file)
-
-@cli.command()
-@click.argument('model_family_dir', type=click.Path(exists=False))
-@click.argument('data_dir', type=click.Path(exists=True))
-@click.option('--max-seq-len', default=64)
-@click.option('--file-limit', type=int, default=50000,
-        help="Number of examples per file (community).")
-def write_test_set_df(model_family_dir, data_dir, max_seq_len, file_limit):
-
-    random.seed(42)
-    random_state = random.getstate()
-    dataset, fields = load_data_and_fields(data_dir, model_family_dir, max_seq_len, file_limit)
[…]

### 2020-10-31 results.py

@@ -1,0 +1,215 @@
+""" Use the trained models on test data, testing both CCLM and LMCC.
+    Produce results tabels that will be read by analysis.py to create plots, tables, etc.
+"""
+
+from model import CommunityConditionedLM
+from data import load_data_and_fields
+import torch
+import torch.nn as nn
+import torch.nn.functional as F
+import torchtext as tt
+import pandas as pd
+import numpy as np
+import random
+import json
+import os
+import click
+
+architectures = [
+    'lstm-3', 'lstm-3-0', 'lstm-3-1', 'lstm-3-2', 'lstm-3-3',
+    'transformer-3', 'transformer-3-0', 'transformer-3-1', 'transformer-3-2', 'transformer-3-3'
+    ]
+conditioned_architectures = [m for m in architectures if not m in ('lstm-3', 'transformer-3')]
+
+def load_test_data(data_dir, model_family_dir, max_seq_len, file_limit,
+        device, batch_size=512):
+    dataset, fields = load_data_and_fields(data_dir, model_family_dir,
+            max_seq_len, file_limit)
+    fields['text'].include_lengths = True
+    random.seed(42)
+    random_state = random.getstate()
+    _, _, test_data = dataset.split(split_ratio=[0.8,0.1,0.1],
+            stratified=True, strata_field='community', random_state=random_state)
+    test_iterator = tt.data.BucketIterator(
+        test_data,
+        device=device,
+        batch_size=batch_size,
+        sort=False,
+        shuffle=False,
+        train=False)
+    return test_data, test_iterator, fields
+
+def exp_normalize(x, axis):
+    """ https://timvieira.github.io/blog/post/2014/02/11/exp-normalize-trick/ """
+    b = x.max(axis=axis)
+    y = np.exp(x - b)
+    return y / y.sum(axis)
+
+def mask_seqs(max_seq_len, batch_size, lengths):
+    idxs = torch.arange(max_seq_len).expand(batch_size, max_seq_len).to(lengths.device)
+    mask = idxs < lengths.unsqueeze(1)
+    return mask.T.float()
+
+def batch_nll(model, batch, comm=None):
+    """
+    Compute the negative log likelihood of the batch for the given community
+    If comm is None, then the actual communities for the batch will be used.
+    """
+    if comm is None:
+        comm = batch.community
+    text, lengths = batch.text
+    max_seq_len = text.shape[0] - 1
+    x_text = text[:-1]
+    y = text[1:]
[…]

### 2020-10-31 paper/plot_macros.tex

@@
     yticklabels={,,}, xticklabels={,,},
-    legend style={
-      scatter legend, anchor=north, legend columns=1, font=\tiny
-    }
+    legend style={anchor=north, legend columns=1, font=\tiny}%
     ]

### 2020-11-02 paper/paper.tex
tweaks

@@
 \usepgfplotslibrary{groupplots}
-\pgfplotsset{compat=1.17}
+\pgfplotsset{compat=1.16}
 
@@
 which are typically organized around a topic of interest. 
-Our dataset includes comments from \num{46} subreddits, 
+Our dataset includes \emph{comments}\jp{Comments/Post terminology; make sure its consistens} from \num{46} subreddits, 
 randomly selected from the set of all forums 
 with at least \num{15000} comments per month for each month
-in the three years between 2015 and 2017.\footnote{
+in the three years between 2015 and 2017. 
 We initially selected \num{50} subreddits, 
-but excluded four from further analysis: 
-two which were primarily non-English and two with particularly short average comment lengths.}
+but excluded four of them from further analysis: 
+two which were primarily non-English and two with particularly\jp{quantify?} short average comment lengths.
 Each community corpus consist of \num{50000} randomly selected comments from the year 2015.
@@
 Comments were preprocessed as follows: 
-We removed markdown formatting, extracting only rendered text. 
-We exclude the content of block quotes, code blocks, and tables. 
-We tokenized the comments using the default English model for the SpaCy tokenizer,\footnote{Version 2.2.3}
+We excluded the content of block quotes, code blocks, and tables.
+We removed any markup (formatting) commands from the remaining text, extracting only rendered text.
+We tokenized the comments using the default English model for the SpaCy tokenizer Version 2.2.3\jp{citation?}
 and lower-cased all tokens.
@@
 
-We combined the 46 community corproa and split the combined corpus,
+We combined the 46 community corpora and split the combined corpus\jp{unclear if there is balancing between },
 reserving 10\% for testing, 10\% for validation and the rest for training.
@@
 We plotted the two-component PCA projection of the embeddings,
-annotated with these topic categories.
+annotated with these topic categories (Figure~\ref{comm-pca}).
 We see that there is some evidence of clustering according to
@@
 \caption{Community embedding PCA of the best LSTM (left, $c=1$) and transformer (right, $c=3$) models.}
-\label{fig:lstm-pca}
+\label{fig:comm-pca}
 \end{figure*}

### 2020-11-02 paper/paper.tex
tweaks and notes
notes

@@
 At the the core of the community-conditioned language model (CCLM) 
-is a standard $n$-layer neural sequence encoder 
+is a standard $n$\jp{with n=3}-layer neural sequence encoder
 with token embedding (we experiment with LSTM and transformer encoders).
-To condition the model, we additionally train a \emph{community embedding},
-which is concatenated with the hidden layer of the sequence encoder
+To condition the model, we additionally train a \emph{community embedding} parameter.
+This parameter is concatenated with the hidden layer of the sequence encoder
 for some layer $c \leq n$ and passed through a linear layer, 
@@
 
-We combined the 46 community corpora and split the combined corpus\jp{unclear if there is balancing between },
+We combined the 46 community corpora and split the combined corpus,
 reserving 10\% for testing, 10\% for validation and the rest for training.
 Splits were stratified by community so that the same number of training examples
-were available for each community.
+were available for each community.\jp{is it the same for testing/validation? How can you maintain this split?}
 
-We used a vocabulary size of \num{40000} tokens, with a special out-of-vocabulary token.
+We used\jp{past tense is a bit awkward?} a vocabulary size of \num{40000} tokens, including a special out-of-vocabulary token.
 The vocabulary consisted of the most frequent tokens at the comment level,
 meaning that we only counted each token once per comment.
-The main reason for this was to avoid over-counting otherwise rare tokens that were spammed
+The main reason for this choice was to avoid over-counting otherwise rare tokens that were spammed
 many times in a single comment.
 
-We trained the models on a simple left-to-right language modeling task 
+We trained the models on a simple left-to-right\jp{I don't think that this applies for the transformer model} language modeling task 
 with cross entropy loss over the vocabulary 
 and AdamW \citep{Loshchilov2019} optimization, with an inital learning rate of \num{0.001}.
-We used a batch size of \num{256} and a maximum sequence length of \num{64} tokens.
+We used a batch size of \num{256} and a maximum sequence length of \num{64} tokens.\jp{What happens with longer sequences? Are sequences split at sentence boundaries?}
 During training, a dropout rate of $0.1$ was applied between encoder layers 
@@
 
-We trained the models until the validation loss was smaller than the smallest
-validation loss for two epochs in a row. 
+We trained the models until the validation loss stopped decreasing for two epochs in a row.
+\jp{How many epochs does that turn out to be? Is there a meaningful difference between lstm/}
 The weights from the epoch with the smallest validation loss were used for testing.
@@
 which can be calculated as follows for a given set of posts:(TODO: what set exactly?)
-\[C_{ij} = average_{m ∈ Posts(c_j)}(P(c=c_i | m))\].
+\[C_{ij} = average_{\{m ∈ Posts(c_j)\}}(P(c=c_i | m))\].
 

### 2020-11-02 comm_author_embed.py
remove qualified path (assume 20 iter)

@@ -46,5 +46,5 @@ np.save('model/reddit2015/comm_author_counts', author_comm_mat)
 
 svd = TruncatedSVD(n_components=16, n_iter=20) #, algorithm='arpack')
 comm_author_embed = svd.fit_transform(author_comm_mat)
-np.save('model/reddit2015/comm_author_embed_svd16dim.20iter', comm_author_embed)
+np.save('model/reddit2015/comm_author_embed_svd16dim', comm_author_embed)
 

### 2020-11-02 paper/paper.tex
possible intro structure

@@
 \usepackage{todonotes}
+\usepackage{enumitem}
 \renewcommand{\UrlFont}{\ttfamily\small}
@@
 
+\newlist{hypotheses}{enumerate}{3}
+\setlist[hypotheses,1]{parsep=0pt,itemsep=1pt,font=\bfseries,label=(H\arabic*)}
+
 \input{plot_macros}
@@
 
+- What are linguistic communities
+  - Linguistic complexity
+- Conditioned language models
+
 We propose to test the following hypotheses:
 
-\begin{enumerate}
+\begin{hypotheses}
 \item Different communities have different levels of linguistic
   complexity
-\item different language models are able to recognize such differences
-\item level at which the community embedding is taken into account in the language model influences this.
-\item community vectors are correlated with extra-linguistic information.
-\item stability is correlated with language complexity
-\end{enumerate}
+\item Different language models are able to recognize such differences
+\item The level at which the community embedding is taken into account
+  in the language model influences this.
+\item Community embedding are correlated with extra-linguistic
+  properties of the community.
+\item Community stability is correlated with language complexity
+\end{hypotheses}
 
@@
 
-- computer r2.
+- compute pearson correlation coefficient
 

### 2020-11-02 paper/plot_macros.tex
add community infreence confusion plots

@@
 }
+
+\newcommand{\confusionplot}[1]{%
+    \addplot[
+        matrix plot*,
+        shader=faceted,
+        mesh/cols=46,
+        point meta=explicit,
+] table [x=actual_comm, y=confered_to, meta=avg_prob] {#1};
+}
+
+\newcommand{\confusiongroupplot}[2]{%
+  \begin{groupplot}[
+  group style={group size=2 by 1}, 
+  enlargelimits=false,
+  axis line style={draw=none},
+  %axis equal,
+  xmin=0, xmax=46,
+  ymin=0, ymax=46,
+  width=8cm,
+  xticklabels={Advice,AskWomen,BabyBumps,CFB,Drugs,EDH,EarthPorn,Fantasy,GameDeals,GlobalOffensive,Jokes,Kappa,KerbalSpaceProgram,KotakuInAction,LifeProTips,MLS,MMA,MaddenUltimateTeam,TwoXChromosomes,Warframe,airsoft,bodybuilding,breakingmom,cars,cringe,eu4,exjw,explainlikeimfive,femalefashionadvice,food,heroesofthestorm,jailbreak,justneckbeardthings,oculus,pcmasterrace,photography,reddevils,relationships,rupaulsdragrace,stopdrinking,streetwear,techsupport,todayilearned,toronto,videos,xxfitness},
+  yticklabels=none,
+  xticklabel style={font=\tiny, rotate=90},
+  xtick=data
+    ]
+    \nextgroupplot
+    \confusionplot{#1}
+    \nextgroupplot[colorbar right]
+    \confusionplot{#2}
+  \end{groupplot}
+}

### 2020-11-02 paper/paper.tex
add community infreence confusion plots
address some comments

@@
 \newcommand\jp[1]{\todo[backgroundcolor=blue!10]{JP: #1}}
+\newcommand\bn[1]{\todo[backgroundcolor=green!10]{BN: #1}}
 
@@
 which are typically organized around a topic of interest. 
-Our dataset includes \emph{comments}\jp{Comments/Post terminology; make sure its consistens} from \num{46} subreddits, 
+Users create \textit{posts}, which consist of a link, image, 
+or text, along with a \emph{comment} section. 
+Comments are threaded: a comment can be made directly on a post,
+or appear as a reply to another comment.
+
+Our dataset includes comments from \num{46} subreddits, 
 randomly selected from the set of all forums 
@@
 We removed any markup (formatting) commands from the remaining text, extracting only rendered text.
-We tokenized the comments using the default English model for the SpaCy tokenizer Version 2.2.3\jp{citation?}
+We tokenized the comments using the default English model for the SpaCy tokenizer Version 2.2.3 \citep{Honnibal2017}
 and lower-cased all tokens.
@@
 At the the core of the community-conditioned language model (CCLM) 
-is a standard $n$\jp{with n=3}-layer neural sequence encoder
-with token embedding (we experiment with LSTM and transformer encoders).
+is a standard $n$-layer neural sequence encoder
+with token embedding (we experiment with LSTM and transformer encoders and set $n=3$ for all models).
 To condition the model, we additionally train a \emph{community embedding} parameter.
@@
 
-We used\jp{past tense is a bit awkward?} a vocabulary size of \num{40000} tokens, including a special out-of-vocabulary token.
+We use a vocabulary size of \num{40000} tokens, including a special out-of-vocabulary token.
 The vocabulary consisted of the most frequent tokens at the comment level,
@@
 and AdamW \citep{Loshchilov2019} optimization, with an inital learning rate of \num{0.001}.
-We used a batch size of \num{256} and a maximum sequence length of \num{64} tokens.\jp{What happens with longer sequences? Are sequences split at sentence boundaries?}
+We used a batch size of \num{256} and a maximum sequence length of \num{64} tokens, truncating longer comments (16.8\% of comments were longer than \num{64} tokens).
 During training, a dropout rate of $0.1$ was applied between encoder layers 
@@
 
-We trained the models until the validation loss stopped decreasing for two epochs in a row.
-\jp{How many epochs does that turn out to be? Is there a meaningful difference between lstm/}
-The weights from the epoch with the smallest validation loss were used for testing.
+We trained the models until the validation loss stopped decreasing 
+for two epochs in a row,
+and used the weights from the epoch with the smallest validation loss for testing.
+In general, transformer models trained for about half as long as the LSTM models (Table~\ref{tab:best-epoch}).
 
@@
   \input{floats/best_epoch.tex}
+  \caption{Epoch with the lowest validation loss.}
   \label{tab:best-epoch}
-  \caption{Epoch with the lowest validation loss.}
 \end{table}
@@
 
-TODO: For space reasons, it might be better to put the community-level table in the appendix
-and have the table in body just show the means (and perhaps medians/stddevs?) for each architecture.
+%TODO: For space reasons, it might be better to put the community-level table in the appendix and have the table in body just show the means (and perhaps medians/stddevs?) for each architecture.}
 
 We also note that the conditioned models perform differently across different communities---\
-for some communities, the test perplexity of the conditioned models is a lot lower,\jp{Where can I see this?}
+for some communities, the test perplexity of the conditioned models is a lot lower (Table~\ref{tab:model-ppls},
 whereas for others it is the same or even slightly higher.
-(TODO: talk about this in terms of mutual information instead?)
+\bn{TODO: talk about this in terms of mutual information instead?}
 As an anecdotal observation, we note 
@@
 We plotted the two-component PCA projection of the embeddings,
-annotated with these topic categories (Figure~\ref{comm-pca}).
+annotated with these topic categories (Figure~\ref{fig:comm-pca}).
 We see that there is some evidence of clustering according to
@@
 We now turn our interest to language-model-based community classifiers
-(LMCC). Given a post $m$, and LMCC gives the probability that it
+(LMCC). Given a comment $m$, and LMCC gives the probability that it
 belongs in a community $c_i$ Using the Bayes theorem, we can turn a
-CCLM into a LMCC. Indeed, the probability that a given post $m$
+CCLM into a LMCC. Indeed, the probability that a given comment $m$
 belongs to a community $c_i$ can be calculated as follows:
@@
 Knowing that $m$ \emph{actually} comes from community $c_j$, we can
-measure the the precision our classifier, as a confusion matrix $C$,
-which can be calculated as follows for a given set of posts:(TODO: what set exactly?)
+measure the precision our classifier, as a confusion matrix $C$,
+which can be calculated as follows for a given set of comment (TODO: what set exactly?)
 \[C_{ij} = average_{\{m ∈ Posts(c_j)\}}(P(c=c_i | m))\].
@@
 TODO: show a few relevant matrices. (eg. show only best and worst model) and commentary.
+
+\begin{figure*}
+\begin{tikzpicture}
+  \confusiongroupplot{floats/lstm-3-1_comm_infer_confusion.csv}{floats/transformer-3-3_comm_infer_confusion.csv}
+\end{tikzpicture}
+\caption{LMCC confusion matrices for the best LSTM (left, $c=1$) and transformer (right, $c=3$) models.}
+\label{fig:confusion}
+\end{figure*}
 
@@
 somewhat \emph{too} fine-grained. For this reason, we will be
-interested in confusing the posts from community $c_j$ are to the
+interested in confusing the comments from community $c_j$ are to the
 LMCC. For this purpose, we consider the perplexity of the LMCC for a
-post coming from community $j$, defined as the exponential of the
+comment coming from community $j$, defined as the exponential of the
 entropy of the distribution \(D_j\), which is the (discrete)
-probability distribution of predicted community for a random post from
+probability distribution of predicted community for a random comment from
 $c_j$. $D_j$ is simply the $j$th row in \(C\), and so we obtain the formula:
@@
 We see a positive correlation: this suggests that the more stable a
-community is, the more difficult it is to identify posts as coming
+community is, the more difficult it is to identify comment as coming
 from it. What this suggests is that stable communities tend to use a
 varied and standard, subset of English, while unstable communities
-resort to more formulaic posts. This may suggests that to 'fit' in an
+resort to more formulaic language. This may suggests that to 'fit' in an
 unstable community, one must use more obvious linguistic cues than in

### 2020-11-03 comm_author_embed.py
download author/comm data from pushshift & SVD counts matrix

@@ -8,43 +8,30 @@ from collections import Counter
 from scipy.sparse import coo_matrix
 from sklearn.decomposition import TruncatedSVD
 import numpy as np
+import glob
 
-dataset, fields = data.load_data_and_fields('data/reddit2015', 'model/reddit2015', 64, 5)
-comms = fields['community'].vocab.itos
-comms = comms[1:]
+count_files = glob.glob('data/reddit2015_author_comm_counts/2015_author_sub_counts-*.csv')
 
-def iter_comments(comm):
-    for month in range(1,13):
-        filename = f'data/subreddit_comments/{comm}/2015-{month:02d}.csv'
-        with open(filename, 'r') as f:
+def iter_counts():
+    for file in count_files:
+        print(f"reading {file}")
+        with open(file, 'r') as f:
             reader = csv.DictReader(f)
-            for comment in reader:
-                yield comment
-
-authors = set()
-for c_idx,c in enumerate(comms):
-    print(f'{c_idx+1}/{len(comms)}',end='\r')
-    for comment in iter_comments(c):
-        authors.add(comment['author'])
-authors = list(authors)
-author_dict = dict(zip(authors, range(len(authors))))
-
-author_counts = Counter()
-for c_idx, c in enumerate(comms):
-    print(f'{c_idx+1}/{len(comms)}',end='\r')
-    for comment in iter_comments(c):
-        author_idx = author_dict[comment['author']]
-        author_counts[(c_idx, author_idx)] += 1
+            for item in reader:
+                yield item['author'], item['subreddit'], item['comment_count']
 
 data, row, col = [], [], []
-for (comm, author), count in author_counts.items():
+for author, comm, count in iter_counts():
     data.append(count)
     row.append(comm)
     col.append(author)
-author_comm_mat = coo_matrix((data, (row, col))).tocsr()
-np.save('model/reddit2015/comm_author_counts', author_comm_mat)
 
+print("Creating counts matrix.")
+author_comm_mat = coo_matrix((data, (row, col))).tocsr()
+np.save('model/reddit2015/comm_author_counts_full', author_comm_mat)
+
+print("Computing SVD.")
 svd = TruncatedSVD(n_components=16, n_iter=20) #, algorithm='arpack')
 comm_author_embed = svd.fit_transform(author_comm_mat)
-np.save('model/reddit2015/comm_author_embed_svd16dim', comm_author_embed)
+np.save('model/reddit2015/comm_author_embed_full_svd16dim', comm_author_embed)
 

### 2020-11-03 query_comm_author_data.py

@@ -1,0 +1,60 @@
+import click
+import os
+import google.auth
+from google.cloud import bigquery
+from google.cloud import bigquery_storage_v1beta1
+from google.cloud.exceptions import NotFound
+
+
+# Processes ~15GB of data (1TB free as of 11/2020 )
+query = """
+SELECT author, subreddit, count(*) as comment_count
+FROM `fh-bigquery.reddit_comments.2015_*`
+GROUP BY author, subreddit
+"""
+
+@click.command()
+@click.argument('google_credentials_file', type=click.Path(exists=True))
+@click.option('--google-project', type=str, default="bill-gu-research")
+@click.option('--google-dataset-id', type=str, default="reddit_sample")
+def cli(google_credentials_file, google_project, google_dataset_id):
+    os.environ['GOOGLE_APPLICATION_CREDENTIALS'] = google_credentials_file
+    credentials, project_id = google.auth.default(scopes=["https://www.googleapis.com/auth/cloud-platform"])
+    bq = bigquery.Client(credentials=credentials, project=project_id)
+    bq_storage = bigquery_storage_v1beta1.BigQueryStorageClient(credentials=credentials)
+
+    table_id = f"2015_author_sub_counts"
+    dataset_ref = bq.dataset(google_dataset_id, project=google_project)
+    table_ref = dataset_ref.table(table_id)
+
+    try:
+        bq.get_table(table_ref)  # Make an API request.
+        print(f"{table_ref} already exists")
+    except NotFound:
+        print(f"Creating f{table_ref}")
+        table = bigquery.Table(table_ref)
+        bq.create_table(table)
+
+    job_config = bigquery.QueryJobConfig(
+        allow_large_results=True,
+        destination=table_ref,
+        use_legacy_sql=False,
+        write_disposition='WRITE_TRUNCATE'
+    )
+
+    query_job = bq.query(query, job_config=job_config)  # Make an API request.
+
+    destination_uri = "gs://{}/{}/{}".format(
+        google_project, 'CondLM', f"{table_id}-*.csv"
+    )
+    extract_job = bq.extract_table(
+        table_ref,
+        destination_uri,
+        location="US",  # Location must match that of the source table.
+    )  # API request
+    extract_job.result()  # Waits for job to complete.
+
+    print(f"Exported {google_project}:{google_dataset_id}.{table_id} to {destination_uri}")
+
+if __name__ == '__main__':
+    cli()

### 2020-11-03 query_comm_author_data.py
add google cloud deps

@@ -1,3 +1,11 @@
+"""
+Queries the pushshift.io data on BigQuery to create author/community cooccurance counts for 2015.
+The results are saved to a Google Storange bucket and can be downloaded as follows:
+    gsutil -m cp gs://bill-gu-research/CondLM/2015_author_sub_counts-*.csv ./data/
+
+See: https://cloud.google.com/docs/authentication/getting-started#auth-cloud-implicit-python
+"""
+
 import click
 import os
 import google.auth

### 2020-11-03 comm_author_embed.py
make comm/author dicts

@@ -18,16 +18,41 @@ def iter_counts():
         with open(file, 'r') as f:
             reader = csv.DictReader(f)
             for item in reader:
-                yield item['author'], item['subreddit'], item['comment_count']
+                yield item['author'], item['subreddit'], int(item['comment_count'])
 
+author_stoi, author_itos = {}, []
+comm_stoi, comm_itos = {}, []
+
+print("Compiling counts.")
 data, row, col = [], [], []
 for author, comm, count in iter_counts():
+    if author in author_stoi:
+        author_idx = author_stoi[author]
+    else:
+        author_itos.append(author)
+        author_idx = len(author_itos)
+        author_stoi[author] = author_idx
+    if comm in comm_stoi:
+        comm_idx = comm_stoi[comm]
+    else:
+        comm_itos.append(comm)
+        comm_idx = len(comm_itos)
+        comm_stoi[comm] = comm_idx
     data.append(count)
-    row.append(comm)
-    col.append(author)
+    row.append(comm_idx)
+    col.append(author_idx)
+
+def save_list(l, file):
+    with open(file, 'w') as f:
+        f.write('\n'.join(l))
+
+save_list(author_itos, 'model/reddit2015/authors_full.txt')
+save_list(comm_itos, 'model/reddit2015/comms_full.txt')
+
 
 print("Creating counts matrix.")
-author_comm_mat = coo_matrix((data, (row, col))).tocsr()
+author_comm_mat = coo_matrix((data, (row, col)))
+author_comm_mat = author_comm_mat.tocsr()
 np.save('model/reddit2015/comm_author_counts_full', author_comm_mat)
 
 print("Computing SVD.")

### 2020-11-05 paper/paper.tex
related work stub

@@
 
+\section{Related work}
+
+\paragraph{Vector representations of online communities encode semantic relationships}
+https://www.aclweb.org/anthology/W17-2904.pdf
+
+\paragraph{Language models for sociolinguistic variation}
+- Del Tredici, M., & Fernández, R. (2017). Semantic Variation in Online Communities of Practice. https://www.aclweb.org/anthology/W17-6804
+https://www.aclweb.org/anthology/W17-2904.pdf
+
+\paragraph{Topic modelling}
+- Discovering Discrete Latent Topics with Neural Variational Inference https://arxiv.org/pdf/1706.00359.pdf
+- JeyHan Lau's work
+
+\paragraph{Pre-trained language models}
+- Unsupervised Domain Clusters in Pretrained Language Models https://www.aclweb.org/anthology/2020.acl-main.692/
+
 \section{Discussion and Conclusion}

### 2020-11-08 paper/paper.tex
other potentially relevant paper

@@
 
+\paragraph{Reddit Mining to Understand Gendered Movements}
+http://ceur-ws.org/Vol-2578/DARLIAP3.pdf
+
 \paragraph{Language models for sociolinguistic variation}
-- Del Tredici, M., & Fernández, R. (2017). Semantic Variation in Online Communities of Practice. https://www.aclweb.org/anthology/W17-6804
-https://www.aclweb.org/anthology/W17-2904.pdf
+\cite{DelTredici2017}
 

### 2020-11-10 comm_author_embed.py
add community2vec embeddings
maybe there is a bug.

@@ -1,16 +1,22 @@
 """
 Creates a community embedding for the set of subreddits from the author/community cooccurance matrix
+as in community2vec: https://www.aclweb.org/anthology/W17-2904/
 """
 
 import data
 import csv
 from collections import Counter
-from scipy.sparse import coo_matrix
-from sklearn.decomposition import TruncatedSVD
+from scipy.sparse import coo_matrix, save_npz, load_npz
+from sklearn.decomposition import TruncatedSVD, PCA
 import numpy as np
 import glob
+import os
 
 count_files = glob.glob('data/reddit2015_author_comm_counts/2015_author_sub_counts-*.csv')
+counts_matrix_file = 'model/reddit2015/comm_author_counts_full.npz'
+authors_file = 'model/reddit2015/authors_full.txt'
+comms_file = 'model/reddit2015/comms_full.txt'
+comms_sorted_file =  'model/reddit2015/comms_full_sorted.txt'
 
 def iter_counts():
     for file in count_files:
@@ -20,43 +26,128 @@ def iter_counts():
             for item in reader:
                 yield item['author'], item['subreddit'], int(item['comment_count'])
 
-author_stoi, author_itos = {}, []
-comm_stoi, comm_itos = {}, []
-
-print("Compiling counts.")
-data, row, col = [], [], []
-for author, comm, count in iter_counts():
-    if author in author_stoi:
-        author_idx = author_stoi[author]
-    else:
-        author_itos.append(author)
-        author_idx = len(author_itos)
-        author_stoi[author] = author_idx
-    if comm in comm_stoi:
-        comm_idx = comm_stoi[comm]
-    else:
-        comm_itos.append(comm)
-        comm_idx = len(comm_itos)
-        comm_stoi[comm] = comm_idx
-    data.append(count)
-    row.append(comm_idx)
-    col.append(author_idx)
-
 def save_list(l, file):
     with open(file, 'w') as f:
         f.write('\n'.join(l))
 
-save_list(author_itos, 'model/reddit2015/authors_full.txt')
-save_list(comm_itos, 'model/reddit2015/comms_full.txt')
+def load_list(file):
+    with open(file, 'r') as f:
+        return f.read().split('\n') 
+
+def ppmi(counts, cds=False, neg=1, smooth=0):
+    """ Based on (copied from) histwords/representations/ppmigen.py """
+
+    # compute marginal probs
+    smooth = counts.sum() * smooth
[…]

### 2020-11-11 comm_author_embed.py
add SNAP embedding

@@ -13,10 +13,10 @@ import glob
 import os
 
 count_files = glob.glob('data/reddit2015_author_comm_counts/2015_author_sub_counts-*.csv')
-counts_matrix_file = 'model/reddit2015/comm_author_counts_full.npz'
-authors_file = 'model/reddit2015/authors_full.txt'
-comms_file = 'model/reddit2015/comms_full.txt'
-comms_sorted_file =  'model/reddit2015/comms_full_sorted.txt'
+counts_matrix_file = 'embedding/comm_author_counts_full.npz'
+authors_file = 'embedding/authors_full.txt'
+comms_file = 'embedding/comms_full.txt'
+comms_sorted_file =  'embedding/comms_full_sorted.txt'
 
 def iter_counts():
     for file in count_files:
@@ -107,13 +107,25 @@ def pca(w, n_components):
     pca = PCA(n_components=n_components)
     return pca.fit_transform(w)
 
-def load_comm_embed(embedding_path, comms=None):
+def load_comm_embed(embedding_path, selected_comms=None):
     comms_sorted = load_list(comms_sorted_file)
-    if not comms:
-        comms = comms_sorted
+    comms = selected_comms if selected_comms else comms_sorted
     comm_idxs = [comms_sorted.index(comm) for comm in comms]
     embed = np.load(embedding_path)
     return embed[comm_idxs]
+
+def load_snap_comm_embed(selected_comms):
+    embedding_path = 'embedding/web-redditEmbeddings-subreddits.csv'
+    embedding = np.zeros((len(selected_comms), 300), float)
+    selected_comms = [c.lower() for c in selected_comms]
+    with open(embedding_path, 'r') as f:
+        reader = csv.reader(f)
+        for line in reader:
+            comm = line[0]
+            vector = np.array([float(x) for x in line[1:]])
+            if comm in selected_comms:
+                embedding[selected_comms.index(comm)] = vector
+    return embedding 
 
 if __name__ == '__main__':
 
@@ -147,7 +159,7 @@ if __name__ == '__main__':
 
 
     save_list(comms_sorted, comms_sorted_file)
-    np.save('model/reddit2015/comm_embed_explicit.npy', comm_embed_explicit)
-    np.save('model/reddit2015/comm_embed_pca.npy', comm_embed_pca)
+    np.save('embedding/comm_embed_explicit.npy', comm_embed_explicit)
+    np.save('embedding/comm_embed_pca.npy', comm_embed_pca)
 
 

### 2020-11-11 paper/paper.tex
revise hypotheses

@@
   - Linguistic complexity
-- Conditioned language models
+- Community Conditioned language models
+- Latent Community Classifier
 
@@
   complexity
-\item Different language models are able to recognize such differences
-\item The level at which the community embedding is taken into account
-  in the language model influences this.
-\item Community embedding are correlated with extra-linguistic
-  properties of the community.
-\item Community stability is correlated with language complexity
+
+\item LCC works: the communty-specific linguistic features are recognized by CCLM.
+\item The layer at which the community embedding is taken into account
+  in CCLM influences its perplexity.
+\item The layer at which the community embedding is taken into account
+  influences the LCC.
+\item Correlation between community language complexity and recognisability. (False)
+\item Community embedding are correlated with coocurrence of users
+  between communities.
+% \item Community stability is correlated with language complexity
 \end{hypotheses}

### 2020-11-11 paper/paper.tex
update in light of new results

@@
 
-- Correlation with co-occurence matrix (not conclusive)
+As a quantitative ....
 
-- Analysis: probably because the model is not using dimensions
-with \emph{a priori} equal variance in all dimensions. (Both the LSTM and transformer are highly non-linear complex models)
+- Correlation with community embeddings based on co-occurence;
 
-\subsection{Implicit language-model-based community classifiers}
+- Description of the snap data (JP)
 
-(TODO: Insert a the correlation plot between two models, on the exponential scale so that it's the same as the other later plots)
-TODO: So we can't use the losses directly. Why is this so?
+- Insert scatterplot (select 1 model)
+  + pearson r for ``all'' models.
+
+\subsection{Latent language-model-based community classifiers}
+
+- Two different use of the model
+1. ``absolute'' perplexity
+2. latent ability to recognize a community
 
@@
 \[C_{ij} = average_{\{m ∈ Posts(c_j)\}}(P(c=c_i | m))\].
-
-TODO: show a few relevant matrices. (eg. show only best and worst model) and commentary.
 
@@
 
+TODO: scatterplot/r2 between LMCC and CCLM perplexity  ---> no correlation.
+
 We can then analyse the correlation between \(Ppl^M_j\) for various models $M$.
 
-TODO: show a scatter plot, compute r2, analyse out-of-balance points; discuss why they stand where they do.
+TODO: show a scatter plot, compute r2 --> almost perfect correlation
 
-Finally, we investigate the correlation between the stability of a
-community and the perplexity of the LMCC for it. Stability is defined as TODO.
+% Finally, we investigate the correlation between the stability of a
+% community and the perplexity of the LMCC for it. Stability is defined as TODO.
 
-- compute pearson correlation coefficient
+% - compute pearson correlation coefficient
 
-We see a positive correlation: this suggests that the more stable a
-community is, the more difficult it is to identify comment as coming
-from it. What this suggests is that stable communities tend to use a
-varied and standard, subset of English, while unstable communities
-resort to more formulaic language. This may suggests that to 'fit' in an
-unstable community, one must use more obvious linguistic cues than in
-a stable community. In fact, it is reasonable to think that a new
-member will ostensibly use community-specifc language, but once
-established in a community, a member would tend not to do so.
+% We see a positive correlation: this suggests that the more stable a
+% community is, the more difficult it is to identify comment as coming
+% from it. What this suggests is that stable communities tend to use a
+% varied and standard, subset of English, while unstable communities
+% resort to more formulaic language. This may suggests that to 'fit' in an
+% unstable community, one must use more obvious linguistic cues than in
+% a stable community. In fact, it is reasonable to think that a new
+% member will ostensibly use community-specifc language, but once
+% established in a community, a member would tend not to do so.
 

### 2020-11-11 paper/paper.tex
add pairwise similarity plot/table

@@
 
-\subsection{Analysis of community embeddings}
+\subsection{Analysis of CCLM community embeddings}
 
 As a qualitative assesment of the community embeddings, 
-we identified several topic categories into which at least three
+we manually identified topic categories into which at least three
 of the selected communities fall. 
 We plotted the two-component PCA projection of the embeddings,
-annotated with these topic categories (Figure~\ref{fig:comm-pca}).
+annotated with these categories (Figure~\ref{fig:comm-pca}).
 We see that there is some evidence of clustering according to
@@
 
-As a quantitative ....
-
-- Correlation with community embeddings based on co-occurence;
+As a quantitative assessment,
+we compare the CCLM-learned community embedding with the community embedding from \citet{Kumar2018},
+which is based on community membership.
 
@@
 
-- Insert scatterplot (select 1 model)
-  + pearson r for ``all'' models.
+We correlate the pairwise cosine similarity of the
+CCLM embeddings with the same metric on the membership embedding,
+restricted to the 46 communities chosen for our study
+(Figure~\ref{fig:pairwise-comm-sim}).
+
+\begin{figure}
+  \begin{tikzpicture}
+  \begin{axis}[]
+  \addplot[
+    scatter, only marks,
+    font=\tiny,
+    scatter src=explicit symbolic,
+    scatter/classes={
+      general-interest={mark=*, blue},
+      videogames={mark=*, red},
+      technology={mark=*, green},
+      sports={mark=*, cyan},
+      female-focused={mark=*, magenta},
+      other={mark=o, black},
+      different={mark=x}
+    },
+    %mark size=1pt,
+    %nodes near coords*={\community},
+    %visualization depends on={value \thisrow{community} \as \community}
+  ] table [x=snap, y=lstm-3-1, meta=meta] {floats/comm_sim.csv};
+  \end{axis}
+  \end{tikzpicture}
+  \caption{Pairwise cosine similarity between community vectors for 
+    the membership-based embedding (x-axis) and
+  the LSTM CCLM embedding with $c=1$ (y-axis). 
+  Pairs from the same category are colored according to Figure~\ref{fig:comm-pca}.
+  We also mark pairs from different (\texttimes) and pairs where one or both
+  communities belong to the ``Other'' category (\textbigcircle).
+  }
+  \label{fig:pairwise-comm-sim}
+\end{figure}
+
+\begin{table}
+  \centering
+  \input{floats/comm_sim.tex}
+  \caption{Pearson's r between the pairwise similarity of community vectors
+  in the CCLM models and the membership-based embedding of \citet{Kumar2018}.
+  ($p<0.001$ for all models).
+  }
+  \label{tab:pairwise-comm-sim}
+\end{table}
 

### 2020-11-11 paper/plot_macros.tex
add pairwise similarity plot/table

--- 2020-10-31 paper/plot_macros.tex
@@
     yticklabels={,,}, xticklabels={,,},
-    legend style={anchor=north, legend columns=1, font=\tiny}%
+    legend style={anchor=north, legend columns=1, font=\tiny},
+    every node near coord/.append style={font=\sffamily\tiny}%
     ]
@@
 }
+
+\newcommand{\confusionplot}[1]{%
+    \addplot[
+        matrix plot*,
+        shader=faceted,
+        mesh/cols=46,
+        point meta=explicit,
+] table [x=actual_comm, y=confered_to, meta=avg_prob] {#1};
+}
+
+\newcommand{\confusiongroupplot}[2]{%
+  \begin{groupplot}[
+  group style={group size=2 by 1}, 
+  enlargelimits=false,
+  axis line style={draw=none},
+  %axis equal,
+  xmin=-1, xmax=46,
+  ymin=-1, ymax=46,
+  width=8cm,
+  xticklabels={Advice,AskWomen,BabyBumps,CFB,Drugs,EDH,EarthPorn,Fantasy,GameDeals,GlobalOffensive,Jokes,Kappa,KerbalSpaceProgram,KotakuInAction,LifeProTips,MLS,MMA,MaddenUltimateTeam,TwoXChromosomes,Warframe,airsoft,bodybuilding,breakingmom,cars,cringe,eu4,exjw,explainlikeimfive,femalefashionadvice,food,heroesofthestorm,jailbreak,justneckbeardthings,oculus,pcmasterrace,photography,reddevils,relationships,rupaulsdragrace,stopdrinking,streetwear,techsupport,todayilearned,toronto,videos,xxfitness},
+  yticklabels=none,
+  xticklabel style={font=\sffamily\tiny, rotate=90},
+  xtick=data,
+  legend style={font=\sffamily\small},
+    ]
+    \nextgroupplot
+    \confusionplot{#1}
+    \nextgroupplot[colorbar right]
+    \confusionplot{#2}
+  \end{groupplot}
+}

### 2020-11-11 analysis.py

@@ -2,6 +2,8 @@ import pandas as pd
 import os
 import numpy as np
 import torch
+from scipy.stats import pearsonr
+
 
 model_dir = 'model/reddit2015'
 floats_dir = 'paper/floats'
@@ -32,8 +34,15 @@ def model_params_from_name(model_name):
         lc = layers_str[-1]
     return encoder_arch, lc 
 
-models = [model_params_to_name(arch, lc) for arch in architectures for lc in lcs]
+models = [model_params_to_name(arch, lc) for arch in architectures for lc in lcs] 
 conditioned_models = [model_params_to_name(arch, lc) for arch in architectures for lc in lcs if lc is not None]
+
+comms = torch.load(open(os.path.join(model_dir, 'community.field'), 'rb')).vocab.itos[1:]
+comms = pd.Series(comms, name='community')
+
+
+cclm_comm_embed = {model: np.load(os.path.join(model_dir, f'{model}/comm_embed.npy'))[1:]
+        for model in conditioned_models}
 
 ##### BEST EPOCH
 
@@ -43,6 +52,7 @@ def get_best_epoch(model_name):
 best_epoch = pd.DataFrame([{str(lc): get_best_epoch(model_params_to_name(arch, lc))
     for lc in lcs} for arch in architectures], index=architectures)
 best_epoch.to_latex(os.path.join(floats_dir, 'best_epoch.tex'))
+
 
 ##### Perplexity by model
 
@@ -59,13 +69,10 @@ ppl_mean_comm = ppl_mean_comm.reindex(rows)
 ppl_mean_comm.to_latex(os.path.join(floats_dir, 'model_ppls.tex'), 
         float_format="%.2f", multicolumn_format='c')
 
+
 ##### Community embedding PCA
 
 from sklearn.decomposition import PCA
-
-comms = torch.load(open(os.path.join(model_dir, 'community.field'), 'rb')).vocab.itos[1:]
-comms = pd.Series(comms, name='community')
-comms_alpha = sorted(comms)
 
 # Manually assign communities to different types/subjects
 # NOTE this shouldn't really be a partition. Some communities clearly belong to multiple types,
@@ -91,7 +98,7 @@ def pca(w):
     return pca.fit_transform(w)
 
 def write_pca_table(model_name):
-    comm_embed = np.load(os.path.join(model_dir, f'{model_name}/comm_embed.npy'))[1:]
[…]

### 2020-11-12 paper/paper.tex
lmcc and cclm ppl plots

@@
 \begin{figure}
-  \begin{tikzpicture}
-  \begin{axis}[]
-  \addplot[
-    scatter, only marks,
-    font=\tiny,
-    scatter src=explicit symbolic,
-    scatter/classes={
-      general-interest={mark=*, blue},
-      videogames={mark=*, red},
-      technology={mark=*, green},
-      sports={mark=*, cyan},
-      female-focused={mark=*, magenta},
-      other={mark=o, black},
-      different={mark=x}
-    },
-    %mark size=1pt,
-    %nodes near coords*={\community},
-    %visualization depends on={value \thisrow{community} \as \community}
-  ] table [x=snap, y=lstm-3-1, meta=meta] {floats/comm_sim.csv};
-  \end{axis}
-  \end{tikzpicture}
+  \pairwisecommsim{lstm-3-1} 
   \caption{Pairwise cosine similarity between community vectors for 
@@
 
-\begin{figure*}
-\begin{tikzpicture}
-  \confusiongroupplot{floats/lstm-3-1_comm_infer_confusion.csv}{floats/transformer-3-3_comm_infer_confusion.csv}
-\end{tikzpicture}
-\caption{LMCC confusion matrices for the best LSTM (left, $c=1$) and transformer (right, $c=3$) models.}
-\label{fig:confusion}
-\end{figure*}
+%\begin{figure*}
+%\begin{tikzpicture}
+  %\confusiongroupplot{floats/lstm-3-1_comm_infer_confusion.csv}{floats/transformer-3-3_comm_infer_confusion.csv}
+%\end{tikzpicture}
+%\caption{LMCC confusion matrices for the best LSTM (left, $c=1$) and transformer (right, $c=3$) models.}
+%\label{fig:confusion}
+%\end{figure*}
 
@@
 
-TODO: scatterplot/r2 between LMCC and CCLM perplexity  ---> no correlation.
+  %\begin{tikzpicture}
+  %\pplscatter{lstm-3-1_lmcc}{lstm-3-1_cclm}
+  %\end{tikzpicture}
+
+\begin{figure}
+  \pplscatter{lstm-3-1_cclm}{lstm-3-1_lmcc}
+  \caption{%
+    Mean CCLM (x-axis) and LMCC (y-axis) perplexity per community (Pearson's r = \num{0.30}, p = \num{0.0444}).
+    Shown here for the LSTM model with $c = 1$, but other models exhibit similar behavior.
+  }
+  \label{fig:cclm_lmcc_ppl}
+\end{figure}
 
@@
 
-TODO: show a scatter plot, compute r2 --> almost perfect correlation
+\begin{figure}
+  \pplscatter{lstm-3-1_lmcc}{transformer-3-3_lmcc}
+  \caption{%
+    Mean CCLM perplexity per community for the best LSTM ($c = 1$) and transformer ($c = 3$) models
+    (Pearson's r = \num{0.99} p < \num{0.0001}).
+  }
+  \label{fig:lmcc_ppl}
+\end{figure}
 

### 2020-11-12 paper/plot_macros.tex
lmcc and cclm ppl plots

@@
 }
+
+\newcommand{\pairwisecommsim}[1]{%
+  \begin{tikzpicture}
+  \begin{axis}[]
+  \addplot[
+    scatter, only marks,
+    font=\tiny,
+    scatter src=explicit symbolic,
+    scatter/classes={
+      general-interest={mark=*, blue},
+      videogames={mark=*, red},
+      technology={mark=*, green},
+      sports={mark=*, cyan},
+      female-focused={mark=*, magenta},
+      other={mark=o, black},
+      different={mark=x}
+    },
+  ] table [x=snap, y=#1, meta=meta] {floats/comm_sim.csv};
+  \end{axis}
+  \end{tikzpicture}
+}
+
+\newcommand{\pplscatter}[2]{%
+  \begin{tikzpicture}
+  \begin{axis}[]
+  \addplot[
+    only marks, black,
+  ] table [x=#1, y=#2] {floats/cclm_lmcc_ppl.csv};
+  \end{axis}
+  \end{tikzpicture}
+}

### 2020-11-12 analysis.py

@@ -44,6 +44,8 @@ comms = pd.Series(comms, name='community')
 cclm_comm_embed = {model: np.load(os.path.join(model_dir, f'{model}/comm_embed.npy'))[1:]
         for model in conditioned_models}
 
+cclm_ppl = pd.read_pickle(os.path.join(model_dir, 'test_ppl.pickle'))
+
 ##### BEST EPOCH
 
 def get_best_epoch(model_name):
@@ -56,12 +58,11 @@ best_epoch.to_latex(os.path.join(floats_dir, 'best_epoch.tex'))
 
 ##### Perplexity by model
 
-ppl = pd.read_pickle(os.path.join(model_dir, 'test_ppl.pickle'))
 
-ppl_mean = ppl[models].mean()
+ppl_mean = cclm_ppl[models].mean()
 ppl_mean.index = ppl_mean.index.map(model_params_from_name)
 
-ppl_mean_comm = ppl.groupby('community').mean().sort_values('lstm-3')
+ppl_mean_comm = cclm_ppl.groupby('community').mean().sort_values('lstm-3')
 ppl_mean_comm.columns = ppl_mean_comm.columns.map(model_params_from_name)
 ppl_mean_comm.loc['Mean'] = ppl_mean_comm.mean()
 rows = ['Mean'] + [c for c in ppl_mean_comm.index if not c == 'Mean']
@@ -133,6 +134,7 @@ for model_name in conditioned_models:
     C = model_comm_confusion_matrix(model_name)
     C.to_csv(os.path.join(floats_dir, f'{model_name}_comm_infer_confusion.csv'), sep='\t', float_format="% 5.4f", index=False)
 
+
 #### Pairwise community similarity scatter
 
 from itertools import combinations
@@ -180,7 +182,46 @@ def pivot_model_table(df):
     df['arch'] = df['arch'].apply({'lstm': 'LSTM', 'transformer': 'Transformer'}.get)
     df['c'] = df['index'].apply(lambda x: x.split('-')[-1])
     df = df.drop('index', axis=1)
-    return df.pivot(index='arch', columns='c')
+    df = df.pivot(index='arch', columns='c')
+    df.index.name = None
+    return df
 
-pivot_model_table(df_r)['r'].to_latex(os.path.join(floats_dir, 'comm_sim.tex')
[…]

### 2020-11-12 analysis.py
drip

@@ -190,38 +190,69 @@ pivot_model_table(df_r)['r'].to_latex(os.path.join(floats_dir, 'comm_sim.tex'),
         float_format="%.2f", multicolumn_format='c')
 
 
-##### Compare LMCC and CCLM entropy by community
+##### Compare LMCC and CCLM perplexity by community
 
 from scipy.special import entr
 
-cclm_entropy = cclm_ppl[models].apply(np.log)
-cclm_entropy['community'] = ppl['community']
+def ppl(x):
+    return np.exp(entr(x).sum())
 
-lmcc_entropy = []
+# This is what we were looking at in the notebook.
+# It gives the "perplexity" of the confusion matrix row.
+# The problem is, since this is an average over the (comment-level) CC distributions,
+# it is not itself a probability distribution.
+lmcc_confusion_ppl = []
 for model in conditioned_models:
-    model_entropy = pd.read_pickle(os.path.join(os.path.join(model_dir, model), 'comm_probs.pickle'))[comms].apply(lambda x: entr(x).sum(), axis=1)
-    model_entropy.name = model
-    lmcc_entropy.append(model_entropy) 
+    model_ppl = pd.read_pickle(os.path.join(os.path.join(model_dir, model), 'comm_probs.pickle')).groupby('actual_comm').mean().apply(ppl, axis=1)[comms]
+    model_ppl.name = model
+    lmcc_confusion_ppl.append(model_ppl)
 
-lmcc_entropy = pd.concat(lmcc_entropy, axis=1)
-lmcc_entropy['community'] = cclm_entropy['community']
+lmcc_confusion_ppl = pd.concat(lmcc_confusion_ppl, axis=1)
+lmcc_confusion_ppl.index.name = 'community'
 
-lmcc_ppl = lmcc_entropy[conditioned_models].apply(np.exp)
-lmcc_ppl['community'] = lmcc_entropy['community']
+# I think this is what we should be looking at for lmcc_ppl instead (and what is defined in the paper, if I understand correctly).
+# It computes the CC perplexity for each comment and averages it over the (actual) community.
+lmcc_ppl = []
[…]

### 2020-11-12 analysis.py
better explanation of the latent model in comment

@@ -197,12 +197,38 @@ from scipy.special import entr
 def ppl(x):
     return np.exp(entr(x).sum())
 
+
 # This is what we were looking at in the notebook.
 # It gives the "perplexity" of the confusion matrix row.
 # The problem is, since this is an average over the (comment-level) CC distributions,
 # it is not itself a probability distribution.
 lmcc_confusion_ppl = []
 for model in conditioned_models:
+    # Why is there a single comm_probs.pickle file? Isn't it model-dependent? What is in this file?
+    
+    # We should: 
+    # 1. Compute P(c_i | m)
+    #   Given: P(m | c_i); ie. a probability assigned assigned to m, when setting the CCLM to c_i
+    #   (this is just the exponential of the negative loss.)
+    #   The actual community of m is given by the "actual_comm" column in the 'dataframe'.
+    # Then we can to compute P(c_i | m) by normalising the list
+    #  of P(m | c_i) (varying i but keeping j fixed) so that ∑_i P(c_i|m) = 1
+    # 2. We have too many numbers, so we take the accuracy of the LMCC prediction per *actual* community
+    #    P(c_i | c_j)
+    #      = P(c_i|m) for a random message m in community c_j.
+    #                  (but we take every message to be equally "probable" in c_j)
+    #      = ∑_m P(c_i|m) / |c_j|
+    #    Is this a probability distribution (when varying c_j?):
+    #    ∑_i P(c_i | c_j)
+    #                                                                by def.
+    #      = ∑_i (∑_m P(c_i|m) / |c_j|)
+    #                                                                by linearity of sum.
+    #      = ∑_m ∑_i P(c_i|m) / |c_j|
+    #                                                                total probability
+    #      = ∑_m 1 / |c_j|
+    #                                                                size of the domain of m
+    #      = |c_j| / |c_j|
+    #      = 1
     model_ppl = pd.read_pickle(os.path.join(os.path.join(model_dir, model), 'comm_probs.pickle')).groupby('actual_comm').mean().apply(ppl, axis=1)[comms]
     model_ppl.name = model
     lmcc_confusion_ppl.append(model_ppl)

### 2020-11-12 analysis.py
oopses

@@ -216,9 +216,11 @@ for model in conditioned_models:
     # 2. We have too many numbers, so we take the accuracy of the LMCC prediction per *actual* community
     #    P(c_i | c_j)
     #      = P(c_i|m) for a random message m in community c_j.
-    #                  (but we take every message to be equally "probable" in c_j)
+    #      = ∑_m P(c_i|m) P(m|c_j)
+    #                   but here we take every message to be equally "probable" in c_j, so P(m|c_j)=1/|c_j|
+    #                   if we take |c_j| to be the number of messages in community c_j 
     #      = ∑_m P(c_i|m) / |c_j|
-    #    Is this a probability distribution (when varying c_j?):
+    #    Is this a probability distribution (when varying c_i?):
     #    ∑_i P(c_i | c_j)
     #                                                                by def.
     #      = ∑_i (∑_m P(c_i|m) / |c_j|)

### 2020-11-12 paper/paper.tex
oopses
comment/message

@@
 from the social media website Reddit.\footnote{Comments were obtained
-from the archive at \href{https://pushshift.io/}{pushshift.io} \cite{Baumgartner2020}.}
+  from the archive at \href{https://pushshift.io/}{pushshift.io}
+  \cite{Baumgartner2020}.}
 
@@
 or appear as a reply to another comment.
+%
+Hereafter we refer to such comments as ``messages'', to match our
+convention in mathematical formulas: the letter $c$ stands for a
+community, and $m$ stands for a message.
 
-Our dataset includes comments from \num{46} subreddits, 
+Our dataset includes messages from \num{46} subreddits, 
 randomly selected from the set of all forums 
-with at least \num{15000} comments per month for each month
+with at least \num{15000} messages per month for each month
 in the three years between 2015 and 2017. 
@@
 but excluded four of them from further analysis: 
-two which were primarily non-English and two with particularly\jp{quantify?} short average comment lengths.
-Each community corpus consist of \num{50000} randomly selected comments from the year 2015.
+two which were primarily non-English and two with particularly\jp{quantify?} short average message lengths.
+Each community corpus consist of \num{50000} randomly selected messages from the year 2015.
 
-Comments were preprocessed as follows: 
+Messages were preprocessed as follows: 
 We excluded the content of block quotes, code blocks, and tables.
 We removed any markup (formatting) commands from the remaining text, extracting only rendered text.
-We tokenized the comments using the default English model for the SpaCy tokenizer Version 2.2.3 \citep{Honnibal2017}
+We tokenized the messages using the default English model for the SpaCy tokenizer Version 2.2.3 \citep{Honnibal2017}
 and lower-cased all tokens.
@@
 We use a vocabulary size of \num{40000} tokens, including a special out-of-vocabulary token.
-The vocabulary consisted of the most frequent tokens at the comment level,
-meaning that we only counted each token once per comment.
+The vocabulary consisted of the most frequent tokens at the message level,
+meaning that we only counted each token once per message.
 The main reason for this choice was to avoid over-counting otherwise rare tokens that were spammed
-many times in a single comment.
+many times in a single message.
 
@@
 and AdamW \citep{Loshchilov2019} optimization, with an inital learning rate of \num{0.001}.
-We used a batch size of \num{256} and a maximum sequence length of \num{64} tokens, truncating longer comments (16.8\% of comments were longer than \num{64} tokens).
+We used a batch size of \num{256} and a maximum sequence length of \num{64} tokens, truncating longer messages (16.8\% of messages were longer than \num{64} tokens).
 During training, a dropout rate of $0.1$ was applied between encoder layers 
@@
 We now turn our interest to language-model-based community classifiers
-(LMCC). Given a comment $m$, and LMCC gives the probability that it
+(LMCC). Given a message $m$, and LMCC gives the probability that it
 belongs in a community $c_i$ Using the Bayes theorem, we can turn a
-CCLM into a LMCC. Indeed, the probability that a given comment $m$
+CCLM into a LMCC. Indeed, the probability that a given message $m$
 belongs to a community $c_i$ can be calculated as follows:
@@
 measure the precision our classifier, as a confusion matrix $C$,
-which can be calculated as follows for a given set of comment (TODO: what set exactly?)
-\[C_{ij} = average_{\{m ∈ Posts(c_j)\}}(P(c=c_i | m))\].
+which can be calculated as follows for a given set of messages (TODO: what set exactly?)
+\[C_{ij} = average_{m \in Messages(c_j)}(P(c=c_i | m))\].
 
@@
 somewhat \emph{too} fine-grained. For this reason, we will be
-interested in confusing the comments from community $c_j$ are to the
+interested in confusing the messages from community $c_j$ are to the
 LMCC. For this purpose, we consider the perplexity of the LMCC for a
-comment coming from community $j$, defined as the exponential of the
+message coming from community $j$, defined as the exponential of the
 entropy of the distribution \(D_j\), which is the (discrete)
-probability distribution of predicted community for a random comment from
+probability distribution of predicted community for a random message from
 $c_j$. $D_j$ is simply the $j$th row in \(C\), and so we obtain the formula:
@@
 % We see a positive correlation: this suggests that the more stable a
-% community is, the more difficult it is to identify comment as coming
+% community is, the more difficult it is to identify message as coming
 % from it. What this suggests is that stable communities tend to use a

### 2020-11-12 paper/paper.tex
take a stab at intro

@@
 
-- What are linguistic communities
-  - Linguistic complexity
-- Community Conditioned language models
-- Latent Community Classifier
+TODO: What is a linguistic community.
 
-We propose to test the following hypotheses:
+Broadly construed, we want to investigate two questions.  First, we
+are interested in the variation of to what extent are various
+communities exhibiting various level of linguistic complexity.
 
+TODO: insert some anecdotes (perhaps some particularly stupid/clever
+comments from reddit here.)
+
+It is folklore that certain communities are more complex
+linguistically: here we want to quantify this phenomenon.
+
+More precisely, we investigate if this degree of complexity be
+captured in a computational model.  For this purpose define various
+Community Conditioned Language Models (CCLM for short), which can be
+modeled as an estimator $M$ what a given message $m$ is acceptable in
+a given community $c$: \(P_M(m \mid c)\).
+
+Second, we want to find out how, and how much, CCLMs can be attuned to
+the specificity of communities. We test whether 1. the linguistic
+models are correlated with a model of community based only on
+co-occurence of users between communities and 2. if latent community
+classifiers based on the CCLM can accurately identify that a given message
+comes from a given community.
+
+To sum up, we propose to test the following hypotheses:
 \begin{hypotheses}
@@
   complexity
-
-\item LCC works: the communty-specific linguistic features are recognized by CCLM.
 \item The layer at which the community embedding is taken into account
   in CCLM influences its perplexity.
+\item CCLM can be used to recognize communities based on
+  latent bayesian classification.
 \item The layer at which the community embedding is taken into account
-  influences the LCC.
-\item Correlation between community language complexity and recognisability. (False)
-\item Community embedding are correlated with coocurrence of users
-  between communities.
+  influences CCLM and its associated latent classifier.
+  % \item There is a correlation between community language complexity and recognizability.
+  % This is more of a justification for studying the latent classifier than a real "user-facing" hypothesis.
+\item The model representation of communities are correlated with
+  co-ocurrence of users between communities.
 % \item Community stability is correlated with language complexity
@@
 but excluded four of them from further analysis: 
-two which were primarily non-English and two with particularly\jp{quantify?} short average message lengths.
+two because they were primarily non-English and two with particularly\jp{quantify?} short average message lengths.
 Each community corpus consist of \num{50000} randomly selected messages from the year 2015.

### 2020-11-12 paper/paper.tex
tweaks

@@
 
-At the the core of the community-conditioned language model (CCLM) 
-is a standard $n$-layer neural sequence encoder
-with token embedding (we experiment with LSTM and transformer encoders and set $n=3$ for all models).
-To condition the model, we additionally train a \emph{community embedding} parameter.
-This parameter is concatenated with the hidden layer of the sequence encoder
-for some layer $c \leq n$ and passed through a linear layer, 
-which projects the resulting vector back down to the origial hidden size.
+We experiment with two kinds of architectures for our models: an LSTM
+and a Transformer architecture\jp{citations}.  In either case, they
+are organised as a standard $n$-layer neural sequence encoder with
+token embedding (we set $n=3$ for all models). As standard, words are predicted
+using a softmax projection layer.
+Such encoders form the
+core of the community-conditioned language models (CCLM).
+
+To condition the model, we add a \emph{community embedding} parameter to the models.
+This parameter is concatenated with the hidden layer of the sequence encoder,
+for some layer $c \leq n$, and passed through a linear layer 
+which projects the resulting vector back to the origial hidden layer size.
 For $c = n$, the output of this linear layer is passed directly to the softmax function,
@@
 Splits were stratified by community so that the same number of training examples
-were available for each community.\jp{is it the same for testing/validation? How can you maintain this split?}
+were available for each community.\jp{is it the same for testing/validation? How do you maintain this split?}
 
 We use a vocabulary size of \num{40000} tokens, including a special out-of-vocabulary token.
-The vocabulary consisted of the most frequent tokens at the message level,
-meaning that we only counted each token once per message.
+The vocabulary consists of the most frequent tokens at the message level,
+meaning that we only count each token once per message.
 The main reason for this choice was to avoid over-counting otherwise rare tokens that were spammed
@@
 
-All experiments used models with \num{3} sequence encoder layers,
+All experiments use models with \num{3} sequence encoder layers,
 each with hidden (and token embedding) size of \num{256}. 

### 2020-11-12 paper/paper.tex
what is 'snap' data

@@
 co-occurence of users between communities and 2. if latent community
-classifiers based on the CCLM can accurately identify that a given message
-comes from a given community.
+classifiers based on the CCLM can accurately identify that a given
+message comes from a given community. Additionally, one may
+intuitively assume that linguistically poor communities are easier to
+recognize: the less linguistic diversity, the more specific the
+language become and the easier it is to identify.
 
@@
   influences CCLM and its associated latent classifier.
-  % \item There is a correlation between community language complexity and recognizability.
   % This is more of a justification for studying the latent classifier than a real "user-facing" hypothesis.
@@
 % \item Community stability is correlated with language complexity
+\item Linguistically poor communities are easier to identify than linguistically rich communities.
 \end{hypotheses}
+
+\jp{TODO: Outline of the paper.}
 
@@
 
-As a quantitative assessment,
-we compare the CCLM-learned community embedding with the community embedding from \citet{Kumar2018},
-which is based on community membership.
-
-- Description of the snap data (JP)
-
+As a quantitative assessment, we compare the CCLM-learned community
+embedding with the community embedding created by \citet{Kumar2018},
+who generated them from the comment-community co-occurence matrix,
+using data which spans January 2014 to April 2017. We refer the reader
+to \citet{Kumar2018}\jp{I actually cannot find in their paper what
+  they do.} for details, but the important point is that no linguistic
+information is used to create these embeddings: they only depend on
+whether and how much a given user participates in a given
+community. In sharp contrasts, CCLM community embeddings depend in no
+way on which user is the author of any given message: we only use the
+contents of messages, not authorship data. 
+%
 We correlate the pairwise cosine similarity of the
@@
 
+\jp{Go through hypotheses again}
+
 - Comparison with topic modelling

### 2020-11-12 paper/paper.tex
clarificatino

@@
 
-However, we are not so much interested in the ability of the CCLM to
-precisely model the language used in a community.
+% However, we are not so much interested in the ability of the CCLM to precisely model the language used in a community.
 
-We now turn our interest to language-model-based community classifiers
-(LMCC). Given a message $m$, and LMCC gives the probability that it
-belongs in a community $c_i$ Using the Bayes theorem, we can turn a
-CCLM into a LMCC. Indeed, the probability that a given message $m$
+We now turn our interest to the Bayesian classifier induced by a CCLM, which we call a language-model-based community classifier
+(LMCC). Indeed, the purpose of an LMCC is to estimate the probability $P(m|c_i)$ that a message $m$
+belong in a community $c_i$. Using the Bayes theorem, we can turn a
+CCLM into a LMCC: the probability that a given message $m$
 belongs to a community $c_i$ can be calculated as follows:
@@
 $P(c=c_i)$ is the frequency of community $c_i$ in the dataset, and
-$P(m | c_i)$ is given by the CCLM. $P(m)$ is our normalisation factor, which can be set so that
-\[\sum_i P(c=c_i | m) = 1\].
+$P(m | c_i)$ is given by the CCLM. $P(m)$ is the absolute
+probability of a message, which can be computed as
+\[P(m) = \frac 1 {\sum_i P(c=c_i) P(m|c=c_i ) }\].
 
-Knowing that $m$ \emph{actually} comes from community $c_j$, we can
-measure the precision our classifier, as a confusion matrix $C$,
-which can be calculated as follows for a given set of messages (TODO: what set exactly?)
+We can characterise the quality the LMCC by a confusion matrix
+$C$, defined by $C_{ij} = P(c=c_i | m)$ for a random message $m$ in
+community $c_j$.  That is:
 \[C_{ij} = average_{m \in Messages(c_j)}(P(c=c_i | m))\].
+\jp{Is this test set or training set?}
 

### 2020-11-12 paper/paper.tex
revisions

@@
 
+\newcommand\Ppl{\mathsf{Ppl}}
+\newcommand\Ind{I}
 \newcommand\BibTeX{B\textsc{ib}\TeX}
@@
 just as the final hidden layer of the sequence encoder is in other models.
-For $c=0$, the community embedding is concatenated with the token embedding.
+For $c=0$\jp{it's unfortunate that the letter $c$ is clashing. To Fix!}, the community embedding is concatenated with the token embedding.
 For this reason, we set the hidden size of the sequence encoder 
@@
 We also note that the conditioned models perform differently across different communities---\
-for some communities, the test perplexity of the conditioned models is a lot lower (Table~\ref{tab:model-ppls},
+for some communities, the test \jp{and training?}perplexity of the conditioned models is a lot lower (Table~\ref{tab:model-ppls},
 whereas for others it is the same or even slightly higher.
@@
 This makes sense intuitively, since communities with more niche topics
-would tend to have more specialized vocabulary and speech patterns.
+would tend to have more specialized vocabulary and speech\jp{can you say speech for written language?} patterns.
 
 The pattern of how much the perplexity falls across different communities
-is similar across architectures, but with some differences.
+is similar across architectures, but with some differences.\jp{Either say something specific here or move the whole paragraph to future work.}
 It could be interesting to further investigate these differences in future work,
@@
 
-- Two different use of the model
-1. ``absolute'' perplexity
-2. latent ability to recognize a community
-
-% However, we are not so much interested in the ability of the CCLM to precisely model the language used in a community.
-
-We now turn our interest to the Bayesian classifier induced by a CCLM, which we call a language-model-based community classifier
-(LMCC). Indeed, the purpose of an LMCC is to estimate the probability $P(m|c_i)$ that a message $m$
-belong in a community $c_i$. Using the Bayes theorem, we can turn a
+The perplexity of a CCLM tells is a measure of linguistic complexity,
+but it is not a measure of how linguistically \emph{specific} the
+community is. However, the CCLM induces a latent 
+a latent classifier induced by a CCLM, which we call a language-model-based community classifier
+(LMCC). Indeed, using the Bayes theorem, we can turn a
 CCLM into a LMCC: the probability that a given message $m$
@@
 $P(c=c_i)$ is the frequency of community $c_i$ in the dataset, and
-$P(m | c_i)$ is given by the CCLM. $P(m)$ is the absolute
+$P(m | c_i)$ (the probability that a message $m$
+belongs in a community $c_i$. ) is estimated by the CCLM. $P(m)$ is the absolute
 probability of a message, which can be computed as
@@
 \[C_{ij} = average_{m \in Messages(c_j)}(P(c=c_i | m))\].
+As a sanity check, we observe that the diagonal dominates the matrix: on average the LMCC picks the right community for a message.
+
 \jp{Is this test set or training set?}
@@
 
-TODO: why don't we use F scores?
-
-While the confusion matrices provide a lot of insight, they are
-somewhat \emph{too} fine-grained. For this reason, we will be
-interested in confusing the messages from community $c_j$ are to the
-LMCC. For this purpose, we consider the perplexity of the LMCC for a
+However, we want to measure the ability of the LMCC to correctly
+recognize messages coming from a given community $c_j$. We could use
+something like an F-score, but we prefer to use the perplexity metric
+(again), which is has stronger information-theoretic grounds.
+However, this time, the perplexity corresponds to the difficulty of
+identifying a community, not predicting a word.
+Technically, we consider the average perplexity of the LMCC for a
 message coming from community $j$, defined as the exponential of the
@@
 probability distribution of predicted community for a random message from
-$c_j$. $D_j$ is simply the $j$th row in \(C\), and so we obtain the formula:
-\[Ppl_j = e^{-\sum_i C_{ij} log(C_{ij})}\]
+$c_j$. $D_j$ is simply the $j$th row in \(C\). In sum, we have the formula:
+\[\Ind_j = e^{-\sum_i C_{ij} log(C_{ij})}\]
 
-To recap, \(Ppl_j\) is a measure of the capability (or rather
-inability) of the model to correctly identify a message as coming from
-a certain community $c_j$.
+We call \(\Ind_j\) the (linguistic) indiscernibility of the community
+$c_j$, and it is a measure of how unsure a model is about which
+community a message belongs to --- when the message in question is
+actually coming from a community $c_j$.
 
@@
 
-\begin{figure}
-  \pplscatter{lstm-3-1_cclm}{lstm-3-1_lmcc}
-  \caption{%
-    Mean CCLM (x-axis) and LMCC (y-axis) perplexity per community (Pearson's r = \num{0.30}, p = \num{0.0444}).
-    Shown here for the LSTM model with $c = 1$, but other models exhibit similar behavior.
-  }
-  \label{fig:cclm_lmcc_ppl}
-\end{figure}
-
-We can then analyse the correlation between \(Ppl^M_j\) for various models $M$.
-
+Indiscernibility is a robust measure: it is almost perfectly
+correlated for various CCLM (transformer, lstm and various $c$) $M$.
+\jp{other r's}
 \begin{figure}
@@
   \label{fig:lmcc_ppl}
+\end{figure}
+
+However we observe that the CCLM perplexity is only weakly correlated with the
+indiscernibility to the corresponding latent LMCC
+(\cref{fig:cclm_lmcc_ppl}).
+
+\begin{figure}
+  \pplscatter{lstm-3-1_cclm}{lstm-3-1_lmcc}
+  \caption{%
+    Mean linguistic perplexity (x-axis) and community indiscernibility
+    (y-axis) per community (Pearson's r = \num{0.30}, p =
+    \num{0.0444}).  Shown here for the LSTM model with $c = 1$, but
+    other models exhibit similar behavior.  }
+  \label{fig:cclm_lmcc_ppl}
 \end{figure}

### 2020-11-12 analysis.py
questions about LMCC perplexity metric

@@ -205,6 +205,9 @@ def ppl(x):
 lmcc_confusion_ppl = []
 for model in conditioned_models:
     # Why is there a single comm_probs.pickle file? Isn't it model-dependent? What is in this file?
+    #  > There is one for each model in the model directory.
+    #  > This file has a row for each comment with the columns "actual_comm" and the remaining columns
+    #  > are a probability distribution over communities (the normalisation of P(c_i | m), as below).
     
     # We should: 
     # 1. Compute P(c_i | m)
@@ -231,6 +234,45 @@ for model in conditioned_models:
     #                                                                size of the domain of m
     #      = |c_j| / |c_j|
     #      = 1
+    # 
+    # > OK, I see that you're right. The mean of discrete probability distributions 
+    # > is itself a probability distribution. I'm not sure it's right to call this
+    # > distribution accuracy, however, since I would a prediction (via argmax, for
+    # > example) to be involved.
+    # > 
+    # > It's possible I've convinced myself there's a problem where there isn't one,
+    # > but I see two problems with using the entropy of the average distribution:
+    # >
+    # > 1. This metric can be the same for communities for which the classifier has 
+    # > different accuracy.
+    # >
+    # >   comm_prob =
+    # >   m | p(c1 | m) | p(c2 | m ) | actual_comm 
+    # >   ----------------------------------------
+    # >   0 |       0.7 |       0.3  | c1
+    # >   1 |       0.8 |       0.2  | c1
+    # >   2 |       0.7 |       0.3  | c2
+    # >   3 |       0.8 |       0.2  | c2
+    # >
+    # >   comm_prob.groupby('actual_comm').mean() =
+    # >   actual_comm | p(c1 | m) | p(c2 | m ) 
+    # >   ----------------------------------------
+    # >            c1 |      0.75 |       0.25  
+    # >            c2 |      0.75 |       0.25
+    # >
+    # >   comm_prob.groupby('actual_comm').mean().apply(ppl) =
+    # >   actual_comm | ppl 
+    # >   ------------------
+    # >            c1 | 1.754    
+    # >            c2 | 1.754   
+    # >
+    # > Even though the classifier is more accurace for c1 than c2, the perplexity is
+    # > the same for both.
+    # >
+    # > 2. Why should we care about the perplexity of the mean distriubtion rather
[…]

### 2020-11-12 analysis.py
consistent notation

@@ -247,7 +247,7 @@ for model in conditioned_models:
     # > different accuracy.
     # >
     # >   comm_prob =
-    # >   m | p(c1 | m) | p(c2 | m ) | actual_comm 
+    # >   m | p(c1 | m) | p(c2 | m ) | cj (actual_comm) 
     # >   ----------------------------------------
     # >   0 |       0.7 |       0.3  | c1
     # >   1 |       0.8 |       0.2  | c1
@@ -255,16 +255,16 @@ for model in conditioned_models:
     # >   3 |       0.8 |       0.2  | c2
     # >
     # >   comm_prob.groupby('actual_comm').mean() =
-    # >   actual_comm | p(c1 | m) | p(c2 | m ) 
+    # >            cj | p(c1 | cj) | p(c2 | cj ) 
     # >   ----------------------------------------
     # >            c1 |      0.75 |       0.25  
     # >            c2 |      0.75 |       0.25
     # >
     # >   comm_prob.groupby('actual_comm').mean().apply(ppl) =
-    # >   actual_comm | ppl 
-    # >   ------------------
-    # >            c1 | 1.754    
-    # >            c2 | 1.754   
+    # >    cj  | ppl 
+    # >   ----------
+    # >     c1 | 1.754    
+    # >     c2 | 1.754   
     # >
     # > Even though the classifier is more accurace for c1 than c2, the perplexity is
     # > the same for both.

### 2020-11-12 paper/paper.tex
idea
use better vocabulary

@@
 \usepackage{enumitem}
+\usepackage{cleveref}
 \renewcommand{\UrlFont}{\ttfamily\small}
@@
 recognize messages coming from a given community $c_j$. We could use
-something like an F-score, but we prefer to use the perplexity metric
-(again), which is has stronger information-theoretic grounds.
-However, this time, the perplexity corresponds to the difficulty of
-identifying a community, not predicting a word.
-Technically, we consider the average perplexity of the LMCC for a
-message coming from community $j$, defined as the exponential of the
+something like an F-score, but we prefer to use the entropy metric
+which is has stronger information-theoretic grounds, and is a measure
+of the difficulty of identifying a community, as estimated by the
+model. We call this measure the (linguistic) indiscernibility of the
+community $c_j$, and use \(\Ind_j\) to stand for it in formulas.
+
+Technically, it is defined as the exponential of the
 entropy of the distribution \(D_j\), which is the (discrete)
@@
 \[\Ind_j = e^{-\sum_i C_{ij} log(C_{ij})}\]
-
-We call \(\Ind_j\) the (linguistic) indiscernibility of the community
-$c_j$, and it is a measure of how unsure a model is about which
-community a message belongs to --- when the message in question is
-actually coming from a community $c_j$.
 
@@
 \end{figure}
+
+\jp{We could also compute the entropy of rows of Cij = snap cosine similarity w/ softmax.}
 

### 2020-11-12 paper/paper.tex
coherent term

@@
   \caption{%
-    Mean CCLM perplexity per community for the best LSTM ($c = 1$) and transformer ($c = 3$) models
+    Correlation between indeterminacy for the best LSTM ($c = 1$) and transformer ($c = 3$) models
     (Pearson's r = \num{0.99} p < \num{0.0001}).

### 2020-11-12 paper/paper.tex
for real this time

@@
   \caption{%
-    Correlation between indeterminacy for the best LSTM ($c = 1$) and transformer ($c = 3$) models
+    Correlation between indiscernibility for the best LSTM ($c = 1$) and transformer ($c = 3$) models
     (Pearson's r = \num{0.99} p < \num{0.0001}).

### 2020-11-12 analysis.py
resolve LMCC entropy question

@@ -4,7 +4,6 @@ import numpy as np
 import torch
 from scipy.stats import pearsonr
 
-
 model_dir = 'model/reddit2015'
 floats_dir = 'paper/floats'
 
@@ -39,7 +38,6 @@ conditioned_models = [model_params_to_name(arch, lc) for arch in architectures f
 
 comms = torch.load(open(os.path.join(model_dir, 'community.field'), 'rb')).vocab.itos[1:]
 comms = pd.Series(comms, name='community')
-
 
 cclm_comm_embed = {model: np.load(os.path.join(model_dir, f'{model}/comm_embed.npy'))[1:]
         for model in conditioned_models}
@@ -194,111 +192,29 @@ pivot_model_table(df_r)['r'].to_latex(os.path.join(floats_dir, 'comm_sim.tex'),
 
 from scipy.special import entr
 
+def entropy(x):
+    return entr(x).sum()
+
 def ppl(x):
-    return np.exp(entr(x).sum())
+    return np.exp(etropy(x))
 
-
-# This is what we were looking at in the notebook.
-# It gives the "perplexity" of the confusion matrix row.
-# The problem is, since this is an average over the (comment-level) CC distributions,
-# it is not itself a probability distribution.
-lmcc_confusion_ppl = []
+lmcc_mean_entr = []
 for model in conditioned_models:
-    # Why is there a single comm_probs.pickle file? Isn't it model-dependent? What is in this file?
-    #  > There is one for each model in the model directory.
-    #  > This file has a row for each comment with the columns "actual_comm" and the remaining columns
-    #  > are a probability distribution over communities (the normalisation of P(c_i | m), as below).
-    
-    # We should: 
-    # 1. Compute P(c_i | m)
-    #   Given: P(m | c_i); ie. a probability assigned assigned to m, when setting the CCLM to c_i
-    #   (this is just the exponential of the negative loss.)
-    #   The actual community of m is given by the "actual_comm" column in the 'dataframe'.
-    # Then we can to compute P(c_i | m) by normalising the list
-    #  of P(m | c_i) (varying i but keeping j fixed) so that ∑_i P(c_i|m) = 1
-    # 2. We have too many numbers, so we take the accuracy of the LMCC prediction per *actual* community
-    #    P(c_i | c_j)
[…]

### 2020-11-12 paper/paper.tex
note
first stab at speech community paragraph

@@
 
-TODO: What is a linguistic community.
+The \emph{speech community} is an important concept 
+in the study of linguistic variation.
+Any social group exhibiting some degree of linguistic idiosyncrasy
+may be considered a speech community \citep{Gumperz1972}.
+Linguistic communication requires that participants share 
+\emph{common ground} \citep{Stalnaker2002} knowledge of certain norms, 
+including the meaning of words, syntactic structure, and patterns of interaction.
+Speakers often assume that these norms are common ground based on joint membership
+in some community \cite{Clark1996}.
+
+This basis allows for variation in linguistic convention accross communities.
 
@@
 embedding with the community embedding created by \citet{Kumar2018},
-who generated them from the comment-community co-occurence matrix,
+who generated them from the message-community co-occurence matrix,
 using data which spans January 2014 to April 2017. We refer the reader
@@
 (\cref{fig:cclm_lmcc_ppl}).
-
+\jp{actually, if we had many more communities, then the lines could be ``more blurred'' between various communities, and we may have a stronger correlation here.}
 \begin{figure}

### 2020-11-12 paper/paper.tex
second attempt

@@
 
+Linguistic communication requires that speakers share 
+knowledge of certain linguistic conventions, such as syntactic structure, 
+word meaings, and accepted patterns of interaction.
+Speakers assume that these conventions are common ground based on joint membership
+in some community \cite{Clark1996}.
+Such a community may range in size and specificity from the very large and general
+such as \emph{speakers of a language} to very small and specific, such as 
+of a dialectical region
+
+\emph{common ground} \citep{Stalnaker2002}
+
+This basis allows for variation in linguistic convention accross communities.
+
 The \emph{speech community} is an important concept 
@@
 may be considered a speech community \citep{Gumperz1972}.
-Linguistic communication requires that participants share 
-\emph{common ground} \citep{Stalnaker2002} knowledge of certain norms, 
-including the meaning of words, syntactic structure, and patterns of interaction.
-Speakers often assume that these norms are common ground based on joint membership
-in some community \cite{Clark1996}.
-
-This basis allows for variation in linguistic convention accross communities.
 
 Broadly construed, we want to investigate two questions.  First, we
-are interested in the variation of to what extent are various
-communities exhibiting various level of linguistic complexity.
+are interested in variation in the level of linguistic complexity
+exhibited by different communities.
 

### 2020-11-12 paper/paper.tex
addressing some comments

@@
 knowledge of certain linguistic conventions, such as syntactic structure, 
-word meaings, and accepted patterns of interaction.
-Speakers assume that these conventions are common ground based on joint membership
-in some community \cite{Clark1996}.
-Such a community may range in size and specificity from the very large and general
-such as \emph{speakers of a language} to very small and specific, such as 
-of a dialectical region
+word meanings, and accepted patterns of interaction.
+Speakers assume that these conventions are \emph{common ground} among their interlocutors,
+based on joint membership in a community \cite{Stalnaker2002, Clark1996}.
+Such a community may be of any size,
+from the very small, like members of a particular friend group,
+to the very large, like speakers of English.
+Because it is the source of linguistic convention, the \emph{speech community} 
+is an important concept in the study of linguistic variation \citep{Gumperz1972}.
 
-\emph{common ground} \citep{Stalnaker2002}
-
-This basis allows for variation in linguistic convention accross communities.
-
-The \emph{speech community} is an important concept 
-in the study of linguistic variation.
-Any social group exhibiting some degree of linguistic idiosyncrasy
-may be considered a speech community \citep{Gumperz1972}.
-
+In this work, we investigate linguistic variation across online communities.
 Broadly construed, we want to investigate two questions.  First, we
@@
 
-It is folklore that certain communities are more complex
-linguistically: here we want to quantify this phenomenon.
-
-More precisely, we investigate if this degree of complexity be
+It is folklore that certain communities are more linguistically complex than others.
+Here we want to quantify this phenomenon.
+More precisely, we investigate if the degree of complexity can be
 captured in a computational model.  For this purpose define various
 Community Conditioned Language Models (CCLM for short), which can be
-modeled as an estimator $M$ what a given message $m$ is acceptable in
-a given community $c$: \(P_M(m \mid c)\).
+modeled as an estimator $M$ of the degree to which a given message $m$ 
+is acceptable in a given community $c$: \(P_M(m \mid c)\).
 
@@
 from the social media website Reddit.\footnote{Comments were obtained
-  from the archive at \href{https://pushshift.io/}{pushshift.io}
+  from the archive at \url{https://pushshift.io/}.
   \cite{Baumgartner2020}.}
@@
 
-We experiment with two kinds of architectures for our models: an LSTM
-and a Transformer architecture\jp{citations}.  In either case, they
-are organised as a standard $n$-layer neural sequence encoder with
-token embedding (we set $n=3$ for all models). As standard, words are predicted
-using a softmax projection layer.
-Such encoders form the
-core of the community-conditioned language models (CCLM).
+We experiment with two kinds of model architecture: LSTM \citep{Hochreiter1997}
+and Transformer \citep{Vaswani2017}.  In either case, the model
+is organised as a standard $n$-layer neural sequence encoder with
+token embedding (we set $n=3$ for all models). As is standard, words are predicted
+using a softmax projection layer over a pre-determined vocabulary.
+These encoders form the core of the community-conditioned language models (CCLM).
 
@@
 This parameter is concatenated with the hidden layer of the sequence encoder,
-for some layer $c \leq n$, and passed through a linear layer 
-which projects the resulting vector back to the origial hidden layer size.
-For $c = n$, the output of this linear layer is passed directly to the softmax function,
+for some layer $l_c \leq n$, and passed through a linear layer 
+which projects the resulting vector back to the original hidden layer size.
+For $l_c = n$, the output of this linear layer is passed directly to the softmax function,
 just as the final hidden layer of the sequence encoder is in other models.
-For $c=0$\jp{it's unfortunate that the letter $c$ is clashing. To Fix!}, the community embedding is concatenated with the token embedding.
+For $l_c=0$, the community embedding is concatenated with the token embedding.
 For this reason, we set the hidden size of the sequence encoder 
@@
 reserving 10\% for testing, 10\% for validation and the rest for training.
-Splits were stratified by community so that the same number of training examples
-were available for each community.\jp{is it the same for testing/validation? How do you maintain this split?}
+Splits were stratified by community so that each community had \num{5000} examples
+for testing, \num{5000} for evaluation, and \num{40000} for training.
 
 We use a vocabulary size of \num{40000} tokens, including a special out-of-vocabulary token.
-The vocabulary consists of the most frequent tokens at the message level,
+The vocabulary consists of the most frequent tokens at the message level (across all communities),
 meaning that we only count each token once per message.
-The main reason for this choice was to avoid over-counting otherwise rare tokens that were spammed
+The main reason for this is to avoid over-counting otherwise rare tokens that were spammed
 many times in a single message.
 
-We trained the models on a simple left-to-right\jp{I don't think that this applies for the transformer model} language modeling task 
-with cross entropy loss over the vocabulary 
-and AdamW \citep{Loshchilov2019} optimization, with an inital learning rate of \num{0.001}.
-We used a batch size of \num{256} and a maximum sequence length of \num{64} tokens, truncating longer messages (16.8\% of messages were longer than \num{64} tokens).
+We trained the models on a simple left-to-right language modeling task
+with cross entropy loss over the vocabulary. 
+Since the transformer operates on all tokens in the sequence at once,
+this was achieved by masking and incrementally un-masking input tokens
+in a left-to-right fashion. \bn{does this address it?} %\jp{I don't think that this applies for the transformer model} 
+We used the AdamW \citep{Loshchilov2019} optimization algorithm, 
+with an initial learning rate of \num{0.001}.
+We used a batch size of \num{256} and a maximum sequence length of \num{64} tokens, 
+truncating longer messages (16.8\% of messages were longer than \num{64} tokens).
 During training, a dropout rate of $0.1$ was applied between encoder layers 
@@
 The conditional models were given a community embedding with \num{16} dimensions. 
+We experimented with every possible value for $l_c$ in a three-layer model ($l_c\in\{0,1,2,3\}$).
 
@@
 For the LSTM encoder, 
-the best model ($c=1$) concatenates the community embedding after the first encoder layer,
+the best model ($l_c=1$) concatenates the community embedding after the first encoder layer,
 with two more LSTM layers after that.
-The best transformer model ($c=3$), on the other hand,
+The best transformer model ($l_c=3$), on the other hand,
 incorporates the community information last,
@@
 We also note that the conditioned models perform differently across different communities---\
-for some communities, the test \jp{and training?}perplexity of the conditioned models is a lot lower (Table~\ref{tab:model-ppls},
+for some communities, perplexity of the conditioned models is a lot lower (Table~\ref{tab:model-ppls},\bn{is this ok took ``test'' out but made clear in the caption that it's the test set}%\jp{and training?}
 whereas for others it is the same or even slightly higher.
@@
 that communities where conditioning has the least effect
-tend to be more organized around more ``general interest'' topics.\jp{Is there a way to quantify general interest topics? Community size in terms of number of users?}
+tend to be more organized around more ``general interest'' topics,
+such as /r/relationships, where the subject matter is relevant to a broad
+range of people.
+\bn{I really mean this as a quality of the topic, rather than the community itself.}% \jp{Is there a way to quantify general interest topics?
+Community size in terms of number of users?}
+Conditioning the model on community has the most benefit
+for narrower special-interest subreddits such as those organized
+around a certain videogame, sports team, or subculture.
+
 This makes sense intuitively, since communities with more niche topics
-would tend to have more specialized vocabulary and speech\jp{can you say speech for written language?} patterns.
+would tend to have more specialized vocabulary and speech patterns.\bn{i think it's ok---i've seen this broader sense used in academic papers}%\jp{can you say speech for written language?}
 
@@
   \input{floats/model_ppls.tex}
-  \caption{Mean model test perplexity per community.}
+  \caption{Mean model perplexity for messages in the test set, broken down by community.}
   \label{tab:model-ppls}

### 2020-11-13 paper/paper.tex
small changes

@@
 \usepackage{cleveref}
-\renewcommand{\UrlFont}{\ttfamily\small}
+
 \usepgfplotslibrary{groupplots}
 \pgfplotsset{compat=1.16}
+
+\renewcommand{\UrlFont}{\ttfamily\small}
 
@@
 
-We also note that the conditioned models perform differently across different communities---\
-for some communities, perplexity of the conditioned models is a lot lower (Table~\ref{tab:model-ppls},\bn{is this ok took ``test'' out but made clear in the caption that it's the test set}%\jp{and training?}
-whereas for others it is the same or even slightly higher.
-\bn{TODO: talk about this in terms of mutual information instead?}
-As an anecdotal observation, we note 
-that communities where conditioning has the least effect
+We also note that the conditioned models perform differently across different communities---%
+for some communities, perplexity of the conditioned models is a lot lower,\bn{is this ok? took ``test'' out but made clear in the caption that it's the test set}%\jp{and training?}
+whereas for others it is the same or even slightly higher (Table~\ref{tab:model-ppls}).\bn{talk about this in terms of mutual information instead?}
+As an informal observation, we note 
+that across all the models we tested, 
+communities where conditioning has the least effect
 tend to be more organized around more ``general interest'' topics,
@@
 range of people.
-\bn{I really mean this as a quality of the topic, rather than the community itself.}% \jp{Is there a way to quantify general interest topics?
+\bn{I really mean this as a quality of the topic, rather than the community itself, although the two are probably correlated}% \jp{Is there a way to quantify general interest topics?
 Community size in terms of number of users?}
@@
 around a certain videogame, sports team, or subculture.
-
 This makes sense intuitively, since communities with more niche topics
 would tend to have more specialized vocabulary and speech patterns.\bn{i think it's ok---i've seen this broader sense used in academic papers}%\jp{can you say speech for written language?}
-
-The pattern of how much the perplexity falls across different communities
-is similar across architectures, but with some differences.\jp{Either say something specific here or move the whole paragraph to future work.}
-It could be interesting to further investigate these differences in future work,
-since it could reveal differences in the kind of linguistic variation the different
-model architectures capture.
 
@@
 
-The perplexity of a CCLM tells is a measure of linguistic complexity,
+The perplexity of a CCLM gives us a measure of linguistic complexity,
 but it is not a measure of how linguistically \emph{specific} the
-community is. However, the CCLM induces a latent 
-a latent classifier induced by a CCLM, which we call a language-model-based community classifier
-(LMCC). Indeed, using the Bayes theorem, we can turn a
+community is. However, the CCLM induces a latent classifier,
+which we call a language model-based community classifier
+(LMCC). Using the Bayes theorem, we can turn a
 CCLM into a LMCC: the probability that a given message $m$
-belongs to a community $c_i$ can be calculated as follows:
+belongs to a community $c_i$ is calculated as follows:
 \[P(c=c_i | m) = P(m | c_i)\frac {P(m)} {P(c=c_i)}\]
@@
 We can characterise the quality the LMCC by a confusion matrix
-$C$, defined by $C_{ij} = P(c=c_i | m)$ for a random message $m$ in
-community $c_j$.  That is:
+$C$, defined by $C_{ij} = P(c=c_i | m)$ for a given message $m$ 
+from the community $c_j$ portion of the test set.  That is:
 \[C_{ij} = average_{m \in Messages(c_j)}(P(c=c_i | m))\].
@@
 
-\jp{Is this test set or training set?}
-
-%\begin{figure*}
-%\begin{tikzpicture}
-  %\confusiongroupplot{floats/lstm-3-1_comm_infer_confusion.csv}{floats/transformer-3-3_comm_infer_confusion.csv}
-%\end{tikzpicture}
-%\caption{LMCC confusion matrices for the best LSTM (left, $c=1$) and transformer (right, $c=3$) models.}
-%\label{fig:confusion}
-%\end{figure*}
+\begin{figure*}
+\begin{tikzpicture}
+  \confusiongroupplot{floats/lstm-3-1_comm_infer_confusion.csv}{floats/transformer-3-3_comm_infer_confusion.csv}
+\end{tikzpicture}
+\caption{LMCC confusion matrices for the best LSTM (left, $c=1$) and transformer (right, $c=3$) models.}
+\label{fig:confusion}
+\end{figure*}
 
@@
 \[\Ind_j = e^{-\sum_i C_{ij} log(C_{ij})}\]
-
-  %\begin{tikzpicture}
-  %\pplscatter{lstm-3-1_lmcc}{lstm-3-1_cclm}
-  %\end{tikzpicture}
 
@@
 
+% future work
+The pattern of how much the perplexity falls across different communities
+is similar across architectures, but with some differences.\jp{Either say something specific here or move the whole paragraph to future work.}
+It could be interesting to further investigate these differences in future work,
+since it could reveal differences in the kind of linguistic variation the different
+model architectures capture.
+
 \bibliography{paper}

### 2020-11-17 paper/paper.tex
quantify left out sub comment lengths

@@
 but excluded four of them from further analysis: 
-two because they were primarily non-English and two with particularly\jp{quantify?} short average message lengths.
+two because they were primarily non-English and two with particularly short average message lengths.\footnote{
+The macro-average message length was \num{39.3} ($\sigma=15.7$) and the two removed communities had average comment lengths of \num{10.7} and \num{18.1}.}
 Each community corpus consist of \num{50000} randomly selected messages from the year 2015.

### 2020-11-17 paper/paper.tex
use cleveref everywhere

@@
 and used the weights from the epoch with the smallest validation loss for testing.
-In general, transformer models trained for about half as long as the LSTM models (Table~\ref{tab:best-epoch}).
+In general, transformer models trained for about half as long as the LSTM models (\cref{tab:best-epoch}).
 
@@
 after all the sequence encoder layers
-(Table \ref{tab:model-ppls}).
+(\cref{tab:model-ppls}).
 
@@
 for some communities, perplexity of the conditioned models is a lot lower,\bn{is this ok? took ``test'' out but made clear in the caption that it's the test set}%\jp{and training?}
-whereas for others it is the same or even slightly higher (Table~\ref{tab:model-ppls}).\bn{talk about this in terms of mutual information instead?}
+whereas for others it is the same or even slightly higher (\cref{tab:model-ppls}).\bn{talk about this in terms of mutual information instead?}
 As an informal observation, we note 
@@
 We plotted the two-component PCA projection of the embeddings,
-annotated with these categories (Figure~\ref{fig:comm-pca}).
+annotated with these categories (\cref{fig:comm-pca}).
 We see that there is some evidence of clustering according to
@@
 restricted to the 46 communities chosen for our study
-(Figure~\ref{fig:pairwise-comm-sim}).
+(\cref{fig:pairwise-comm-sim}).
 
@@
   the LSTM CCLM embedding with $c=1$ (y-axis). 
-  Pairs from the same category are colored according to Figure~\ref{fig:comm-pca}.
+  Pairs from the same category are colored according to \cref{fig:comm-pca}.
   We also mark pairs from different (\texttimes) and pairs where one or both
@@
 However we observe that the CCLM perplexity is only weakly correlated with the
-indiscernibility to the corresponding latent LMCC
-(\cref{fig:cclm_lmcc_ppl}).
+indiscernibility of the corresponding latent LMCC(\cref{fig:cclm_lmcc_ppl}).
+
 \jp{actually, if we had many more communities, then the lines could be ``more blurred'' between various communities, and we may have a stronger correlation here.}

### 2020-11-17 paper/paper.tex
add correlation/p-value table

@@
 
-We can characterise the quality the LMCC by a confusion matrix
-$C$, defined by $C_{ij} = P(c=c_i | m)$ for a given message $m$ 
-from the community $c_j$ portion of the test set.  That is:
-\[C_{ij} = average_{m \in Messages(c_j)}(P(c=c_i | m))\].
-As a sanity check, we observe that the diagonal dominates the matrix: on average the LMCC picks the right community for a message.
+We can characterise the quality the LMCC with a confusion matrix
+$C$, defined by $C_{ij} = P(c=c_i | m)$ for a given (test set) message $m$ 
+from community $c_j$. 
+That is: \[C_{ij} = average_{m \in Messages(c_j)}(P(c=c_i | m))\].
+As a sanity check, we observe that the diagonal dominates the matrix: on average the LMCC picks the right community for a message (\cref{fig:confusion}).
 
@@
 recognize messages coming from a given community $c_j$. We could use
-something like an F-score, but we prefer to use the entropy metric
+something like an F-score, but we prefer to use the entropy metric,
 which is has stronger information-theoretic grounds, and is a measure
-of the difficulty of identifying a community, as estimated by the
-model. We call this measure the (linguistic) indiscernibility of the
-community $c_j$, and use \(\Ind_j\) to stand for it in formulas.
-
-Technically, it is defined as the exponential of the
-entropy of the distribution \(D_j\), which is the (discrete)
-probability distribution of predicted community for a random message from
-$c_j$. $D_j$ is simply the $j$th row in \(C\). In sum, we have the formula:
+of the uncertainty of the LMCC, given messages from a particular 
+community. 
+Technically, the linguistic indiscernibility of community $c_j$ is defined 
+as the exponential of the entropy of the $j$ row in \(C\),
+which is a probability distribution 
+of predicted community for a random message from $c_j$.
+In sum, we have the formula: \bn{Took out $D$. is it needed elsewhere?}
 \[\Ind_j = e^{-\sum_i C_{ij} log(C_{ij})}\]
@@
 Indiscernibility is a robust measure: it is almost perfectly
-correlated for various CCLM (transformer, lstm and various $c$) $M$.
-\jp{other r's}
+correlated for various CCLM (transformer, LSTM and various $c$) $M$.
+In fact, we observe a correlation between the indiscernibility 
+measures by every pair of models with Pearson's r $\geq 0.98$ 
+and $p < 0.0001$ (see, for example, \cref{fig:lmcc_ppl}).
+
 \begin{figure}
@@
     Correlation between indiscernibility for the best LSTM ($c = 1$) and transformer ($c = 3$) models
-    (Pearson's r = \num{0.99} p < \num{0.0001}).
+    (Pearson's $r = 0.99$, $p < 0.0001$).
   }
@@
 
-However we observe that the CCLM perplexity is only weakly correlated with the
-indiscernibility of the corresponding latent LMCC(\cref{fig:cclm_lmcc_ppl}).
+However, we observe that community-wise CCLM perplexity is only weakly (if at all) \bn{Changed the indiscernability (y-axis) from entropy to Ppl and the p-values all shot up above 0.1.
+However the r is still very weakly positive for every model.}
+correlated with community indiscernibility under the corresponding latent 
+LMCC (see \cref{fig:cclm_lmcc_ppl} and \cref{tab:cclm_lmcc_ppl}).
+In other words, the linguistic complexity of a given community
+is a poor predictor of how distinct its language is.
 
@@
   \caption{%
-    Mean linguistic perplexity (x-axis) and community indiscernibility
-    (y-axis) per community (Pearson's r = \num{0.30}, p =
-    \num{0.0444}).  Shown here for the LSTM model with $c = 1$, but
-    other models exhibit similar behavior.  }
+    Mean community-wise linguistic perplexity (x-axis) 
+    and indiscernibility (y-axis) 
+    (Pearson's r = \num{0.16}; p = \num{0.2991}).  
+    Shown here for the LSTM with $c = 1$.}
   \label{fig:cclm_lmcc_ppl}
 \end{figure}
+
+\begin{table}
+  \centering
+  \input{floats/cclm_lmcc_ppl.tex}
+  \caption{
+  Pearson's r correlation between community-wise linguistic perplexity and indescernibility for each model.
+  }
+  \label{tab:cclm_lmcc_ppl}
+\end{table}
 
@@
 
-% future work
 The pattern of how much the perplexity falls across different communities

### 2020-11-17 analysis.py

@@ -108,34 +108,73 @@ def write_pca_table(model_name):
 for model in conditioned_models:
     write_pca_table(model)
 
-
-#### Community inference confusion matrix
+#### LMCC confusion matrix
 
 from comm_author_embed import load_snap_comm_embed
+from itertools import combinations
 
 comms_alpha = sorted(list(comms))
 
 def model_comm_confusion_matrix(model_name):
     """ C[i,j] = average_{Posts(cj)}(P(c=ci|m))"""
-    P = pd.read_pickle(os.path.join(os.path.join(model_dir, model_name), 'comm_probs.pickle'))
+    P = pd.read_pickle(os.path.join(model_dir, model_name), 'comm_probs.pickle')
     C = P.groupby('actual_comm').mean()
     C = C.T # transpose to (prob assigned, actual comm), as in the paper
     C = C.sort_index() # sort the rows alphabetically
     C = C[C.index] # sort the columns alphabetically too
-    C = C.unstack().reset_index()
-    C = C.rename(columns={'level_1': 'confered_to', 0: 'avg_prob'})
-    C['actual_comm'] = C['actual_comm'].apply(comms_alpha.index)
-    C['confered_to'] = C['confered_to'].apply(comms_alpha.index)
     return C
 
-for model_name in conditioned_models:
-    C = model_comm_confusion_matrix(model_name)
-    C.to_csv(os.path.join(floats_dir, f'{model_name}_comm_infer_confusion.csv'), sep='\t', float_format="% 5.4f", index=False)
+C = {model_name: model_comm_confusion_matrix(model_name) for model_name in conditioned_models}
 
+for model_name in conditioned_models:
+    Cm = C[model_name]
+    Cm = Cm.unstack().reset_index()
+    Cm = Cm.rename(columns={'level_1': 'confered_to', 0: 'avg_prob'})
+    Cm['actual_comm'] = Cm['actual_comm'].apply(comms_alpha.index)
+    Cm['confered_to'] = Cm['confered_to'].apply(comms_alpha.index)
[…]

### 2020-11-18 paper/paper.tex
draft conclusion

--- 2020-11-17 paper/paper.tex
@@
 
-In this work, we investigate linguistic variation across online communities.
+In the present work, we investigate linguistic variation across online communities.
 Broadly construed, we want to investigate two questions.  First, we
@@
 
-TODO: insert some anecdotes (perhaps some particularly stupid/clever
-comments from reddit here.)
+\jp{
+  Some examples (perhaps include if there is space):
 
-It is folklore that certain communities are more linguistically complex than others.
+  (CFB)
+  It was against SMU in the 47-48 Cotton bowl. They had cancelled a
+  game the year before against Miami for the same reason. There's
+  debate on if that's really the origin for the phrase but the story
+  itself is true.
+
+  (Videos)
+  Reminds me of Y2K computer bug. A lot of work went into
+  pre-emptively fixing it, and a lot of bugs were found and fixed,
+  some of which would have been catastrophic. Unfortunately, because
+  the work was successful, a lot of people now describe it as a waste
+  of time/money and a big deal made out of nothing.  }
+
+It is folklore that certain communities exhibit more linguistic complexity than others.
 Here we want to quantify this phenomenon.
@@
 the specificity of communities. We test whether 1. the linguistic
-models are correlated with a model of community based only on
+models are correlated with a model of communities based only on
 co-occurence of users between communities and 2. if latent community
@@
 \begin{hypotheses}
-\item Different communities have different levels of linguistic
+\item \label{hyp:varying-complexity} Different communities have different levels of linguistic
   complexity
-\item The layer at which the community embedding is taken into account
-  in CCLM influences its perplexity.
-\item CCLM can be used to recognize communities based on
+\item \label{hyp:layer-effect} The layer at which the community embedding is taken into account
+  in a CCLM influences its perplexity.
+\item \label{hyp:LMCC-works} CCLM can be used to recognize communities based on
   latent bayesian classification.
-\item The layer at which the community embedding is taken into account
-  influences CCLM and its associated latent classifier.
+% \item \label{hyp:LMCC-works} The layer at which the community embedding is taken into account
+%   influences CCLM and its associated latent classifier.
   % This is more of a justification for studying the latent classifier than a real "user-facing" hypothesis.
-\item The model representation of communities are correlated with
+\item \label{hyp:extra-linguistic-correlation} The CCLM representation of communities are correlated with
   co-ocurrence of users between communities.
 % \item Community stability is correlated with language complexity
-\item Linguistically poor communities are easier to identify than linguistically rich communities.
+\item \label{hyp:rich-harder-to-identify} Linguistically poor communities are easier to identify than
+  linguistically rich communities.
 \end{hypotheses}
 
-\jp{TODO: Outline of the paper.}
+\jp{Perhaps: Outline of the paper.}
 
@@
 
-To investigate variation across communities, we use comments collected
-from the social media website Reddit.\footnote{Comments were obtained
+We investigate linguistic variation across various communities present
+on the social media website Reddit.\footnote{Comments were obtained
   from the archive at \url{https://pushshift.io/}.
   \cite{Baumgartner2020}.}
-
+%
 Reddit is divided into forums called \textit{subreddits}, 
@@
 
-We experiment with two kinds of model architecture: LSTM \citep{Hochreiter1997}
+We experiment with two kinds of model architecture: LSTM \citep{Hochreiter1997}\jp{Is it really a simple or a bi-lstm?
+Even If it's simple it may be worth specifying due to the prevalence of bilstm everywhere.}
 and Transformer \citep{Vaswani2017}.  In either case, the model
@@
 meaning that we only count each token once per message.
-The main reason for this is to avoid over-counting otherwise rare tokens that were spammed
+The main reason for this choice is to avoid over-counting otherwise rare tokens that were spammed
 many times in a single message.
 
-We trained the models on a simple left-to-right language modeling task
+For the LSTM, we trained the models on a simple left-to-right language modeling task
 with cross entropy loss over the vocabulary. 
-Since the transformer operates on all tokens in the sequence at once,
-this was achieved by masking and incrementally un-masking input tokens
-in a left-to-right fashion. \bn{does this address it?} %\jp{I don't think that this applies for the transformer model} 
+Because the transformer operates on all tokens in the sequence at once,
+the language modeling was achieved by masking and incrementally un-masking input tokens
+in a left-to-right fashion, following \citet{BERT}\jp{Or some other standard paper}.  
 We used the AdamW \citep{Loshchilov2019} optimization algorithm, 
-with an initial learning rate of \num{0.001}.
+with an initial learning rate of \num{0.001}.\jp{Also for the transformer? No simulated annealing?}
 We used a batch size of \num{256} and a maximum sequence length of \num{64} tokens, 
@@
 but this effect is different in magnitude
-for models with different community embedding depths .
+for models with different community embedding depths.
 For the LSTM encoder, 
@@
 
-\jp{Go through hypotheses again}
+We find that various communities exhibit widely varying levels of
+linguistic complexity (\cref{hyp:varying-complexity}), with perplexity
+per-word ranging from less than 53 to 104 for the best model
+architecture.
+
+With indiscernibility values of 1.5 to 3.5 (out of a maximum of 46),
+we find that CCLMs can be used to recognize communities based on
+latent bayesian classification (\cref{hyp:LMCC-works}). Even the most
+generic communities appear to exhibit fairly specific linguistic
+patterns. Furthermore, these patterns are just as recognizable by LSTM
+or Transformer models, when properly configured and trained.
+
+We find a weak correlation (TODO; r , p) between the perplexity of the
+CCLM and the indiscernibility of the corresponding community
+(\cref{hyp:rich-harder-to-identify}). Because we have relatively few,
+and only large communities in our test set, it is difficult to
+provide a conclusive answer, but this kind of correlation would suggest that
+community-specific complexity can be partially attributed to using
+more varied patterns. However, even the most complex communities
+(perplexity per word above 100) have an indiscernibility of about 3
+(out of a maximum of 46), so they retain fairly specific patterns:
+they are not simply an even mixture of the kind of comments found in
+other communities.
+
+We find that the layer at which the community embedding is taken into
+account in a CCLM influences its perplexity (\cref{hyp:layer-effect}),
+albeit weakly. For LSTM models, the perplexity per word, averaged over
+all communities, varies between 78.64 and 80.70. It is steadily
+increasing when $l_c$ decreases.  For transformer models, it varies
+between 89.38 and 122.56. Here perplexity per-word is not a monotonous
+function of $l_c$: the maximum is reached at $l_c=2$. We did not
+expect non-mononicity, and leave studying this effect to future work.
+
+The representation of communities built by CCLM is exhibiting a clear
+correlation with user-coocurrence patterns
+(\cref{hyp:extra-linguistic-correlation}, TODO; r , p). A natural
+question that may arise is how much this correlation can be explain by
+user-specifc linguistic patterns. We plan to control for this variable
+in future work, for example by building author-conditioned language
+models and testing if community vector embeddings can be determined by
+the set of the vector embeddings of its members.
 

### 2020-11-18 paper/paper.tex
info gain table/plot

--- 2020-11-17 paper/paper.tex
@@
 
+On the other hand, when we consider the \emph{information gain}\bn{define earlier}
+of the CCLM over the corresponding (same encoder architecture)
+un-conditioned
+
+\begin{figure}
+  \pplscatter{lstm-3-1_gain}{lstm-3-1_lmcc}
+  \caption{%
+    Mean community-wise information gain (x-axis) 
+    and indiscernibility (y-axis) 
+    %(Pearson's r = \num{0.16}; p = \num{0.2991}).  
+    Shown here for the LSTM with $c = 1$.}
+  \label{fig:cclm_lmcc_ppl}
+\end{figure}
+
+\begin{table}
+  \centering
+  \input{floats/lmcc_gain.tex}
+  \caption{
+    Pearson's r correlation between community-wise information gain (of the CCLM over the un-conditioned model) and indescernibility.
+  }
+  \label{tab:lmcc_gain}
+\end{table}
+
 % Finally, we investigate the correlation between the stability of a

### 2020-11-18 analysis.py

@@ -3,6 +3,7 @@ import os
 import numpy as np
 import torch
 from scipy.stats import pearsonr
+from scipy.special import entr
 
 model_dir = 'model/reddit2015'
 floats_dir = 'paper/floats'
@@ -10,6 +11,12 @@ floats_dir = 'paper/floats'
 lcs = [None, 0, 1, 2, 3]
 architectures = ['LSTM', 'Transformer']
 lcs_str = [str(i) for i in lcs]
+
+def entropy(x):
+    return entr(x).sum()
+
+def ppl(x):
+    return np.exp(entropy(x))
 
 def model_params_to_name(encoder_arch, lc):
     if encoder_arch == "Transformer":
@@ -117,7 +124,7 @@ comms_alpha = sorted(list(comms))
 
 def model_comm_confusion_matrix(model_name):
     """ C[i,j] = average_{Posts(cj)}(P(c=ci|m))"""
-    P = pd.read_pickle(os.path.join(model_dir, model_name), 'comm_probs.pickle')
+    P = pd.read_pickle(os.path.join(model_dir, model_name, 'comm_probs.pickle'))
     C = P.groupby('actual_comm').mean()
     C = C.T # transpose to (prob assigned, actual comm), as in the paper
     C = C.sort_index() # sort the rows alphabetically
@@ -139,6 +146,7 @@ for model_name in conditioned_models:
     Im = C[model_name].apply(ppl)
     Im.name = model_name
     I.append(Im)
+
 I = pd.concat(I, axis=1)
 
 for m1, m2 in combinations(conditioned_models, 2):
@@ -181,6 +189,8 @@ def cos_sim(v1, v2):
 
 def pairwise_sims(embedding, pairs):
     return pd.Series([cos_sim(*map(lambda x: embedding[list(comms).index(x)], pair)) for pair in pairs], index=pairs)
+
+pairs = list(combinations(comms, 2))
     
 df = pd.DataFrame([], index=pairs)
 for model in conditioned_models:
@@ -229,20 +239,8 @@ pivot_model_table(df_r)['r'].to_latex(os.path.join(floats_dir, 'comm_sim.tex'),
 
 ##### Compare LMCC and CCLM perplexity by community
 
-from scipy.special import entr
-
-def entropy(x):
[…]

### 2020-11-18 paper/paper.tex

--- 2020-11-18 paper/paper.tex
@@
 
-However we observe that the CCLM perplexity is only weakly correlated with the
-indiscernibility to the corresponding latent LMCC
-(\cref{fig:cclm_lmcc_ppl}).
+However, we observe that community-wise CCLM perplexity is only weakly (if at all) \bn{Changed the indiscernability (y-axis) from entropy to Ppl and the p-values all shot up above 0.1.
+However the r is still very weakly positive for every model.}
+correlated with community indiscernibility under the corresponding latent 
+LMCC (see \cref{fig:cclm_lmcc_ppl} and \cref{tab:cclm_lmcc_ppl}).
+In other words, the linguistic complexity of a given community
+is a poor predictor of how distinct its language is.
+
 \jp{actually, if we had many more communities, then the lines could be ``more blurred'' between various communities, and we may have a stronger correlation here.}
@@
   \caption{%
-    Mean linguistic perplexity (x-axis) and community indiscernibility
-    (y-axis) per community (Pearson's r = \num{0.30}, p =
-    \num{0.0444}).  Shown here for the LSTM model with $c = 1$, but
-    other models exhibit similar behavior.  }
+    Mean community-wise linguistic perplexity (x-axis) 
+    and indiscernibility (y-axis) 
+    (Pearson's r = \num{0.16}; p = \num{0.2991}).  
+    Shown here for the LSTM with $c = 1$.}
   \label{fig:cclm_lmcc_ppl}
@@
 
+\begin{table}
+  \centering
+  \input{floats/cclm_lmcc_ppl.tex}
+  \caption{
+  Pearson's r correlation between community-wise linguistic perplexity and indescernibility for each model.
+  }
+  \label{tab:cclm_lmcc_ppl}
+\end{table}
+
 \jp{We could also compute the entropy of rows of Cij = snap cosine similarity w/ softmax.}
+
+On the other hand, when we consider the \emph{information gain}\bn{define earlier}
+of the CCLM over the corresponding (same encoder architecture)
+un-conditioned
+
+\begin{figure}
+  \pplscatter{lstm-3-1_gain}{lstm-3-1_lmcc}
+  \caption{%
+    Mean community-wise information gain (x-axis) 
+    and indiscernibility (y-axis) 
+    %(Pearson's r = \num{0.16}; p = \num{0.2991}).  
+    Shown here for the LSTM with $c = 1$.}
+  \label{fig:cclm_lmcc_ppl}
+\end{figure}
+
+\begin{table}
+  \centering
+  \input{floats/lmcc_gain.tex}
+  \caption{
+    Pearson's r correlation between community-wise information gain (of the CCLM over the un-conditioned model) and indescernibility.
+  }
+  \label{tab:lmcc_gain}
+\end{table}
 

### 2020-11-19 analysis.py
partial refactor of tables

@@ -4,13 +4,22 @@ import numpy as np
 import torch
 from scipy.stats import pearsonr
 from scipy.special import entr
+from itertools import combinations
+from sklearn.decomposition import PCA
 
 model_dir = 'model/reddit2015'
 floats_dir = 'paper/floats'
 
-lcs = [None, 0, 1, 2, 3]
-architectures = ['LSTM', 'Transformer']
-lcs_str = [str(i) for i in lcs]
+def add_columns(df, new, suffix=''):
+    new = new.rename(lambda x: str(x) + suffix, axis=1)
+    index_name = df.index.name
+    for col in df.columns: # make the operation idempotent
+        if col in new.columns:
+            df = df.drop(col, axis=1)
+    df = pd.merge(df, new, how='outer', left_index=True, right_index=True, 
+            validate='one_to_one', sort=True)
+    df.index.name = index_name
+    return df
 
 def entropy(x):
     return entr(x).sum()
@@ -18,71 +27,19 @@ def entropy(x):
 def ppl(x):
     return np.exp(entropy(x))
 
-def model_params_to_name(encoder_arch, lc):
-    if encoder_arch == "Transformer":
-        prefix = 'transformer'
-    elif encoder_arch == "LSTM":
-        prefix = 'lstm'
-    if lc is None:
-        return f"{prefix}-3"
-    else:
-        return f"{prefix}-3-{lc}"
+def cos_sim(v1, v2):
+    return (v1 * v2).sum(axis=0) / (np.linalg.norm(v1, axis=0) * np.linalg.norm(v2, axis=0))
 
-def model_params_from_name(model_name):
-    if model_name.startswith('transformer'):
-        encoder_arch = 'Transformer' 
-    elif model_name.startswith('lstm'):
-        encoder_arch = 'LSTM'
-    layers_str = model_name[len(encoder_arch):]
-    if layers_str == '-3':
-        lc = 'None'
-    else:
-        lc = layers_str[-1]
-    return encoder_arch, lc 
 
-models = [model_params_to_name(arch, lc) for arch in architectures for lc in lcs] 
-conditioned_models = [model_params_to_name(arch, lc) for arch in architectures for lc in lcs if lc is not None]
+##### Create community dataframe (with communities as rows)
 
[…]

### 2020-11-19 paper/paper.tex
partial refactor of tables
discussion and info gain

@@
 \usepackage{pgfplots}
+\usepackage{pgfplotstable}
 \usepackage{todonotes}
 \usepackage{enumitem}
+\usepackage{array}
 \usepackage{cleveref}
@@
 \pgfplotsset{compat=1.16}
+\pgfplotstableset{
+  column type=r,
+  fixed zerofill,
+  clear infinite,
+  %dec sep align,
+  %empty cells with={\ensuremath{-}},
+}
 
@@
 For the LSTM, we trained the models on a simple left-to-right language modeling task
-with cross entropy loss over the vocabulary. 
+with cross-entropy loss over the vocabulary. 
 Because the transformer operates on all tokens in the sequence at once,
@@
 
+%\pgfplotstabletypeset[
+  %columns={best_epoch,lm_ppl,gain},
+  %columns/model/.style={string type,column type=l},
+  %columns/best_epoch/.style={column name={Epochs}, int detect},
+  %columns/lm_ppl/.style={column name={Ppl}},
+  %columns/gain/.style={column name={IG},precision=3,empty cells with={\ensuremath{-}}},
+  %every nth row={5}{before row=\midrule},
+  %]{floats/model.csv}
+
 \section{Results and analysis}
@@
 \subsection{Losses of CCLM}
+
+We measure the perplexity of each example in the test set. 
+For this, we use the negative exponential of the mean cross-entropy loss 
+for the sequence.
 
@@
 \end{table*}
+
+\subsection{Information gain of the CCLM}
+
+We consider the average information gain of each CCLM over its baseline
+un-conditioned counterpart with the same sequence encoder architecture.
+For a given message, the information gain is defined as the ratio of the
+perplexity of the baseline model to that of the CCLM:
+\[
+  IG_{CCLM}(m) = \frac{Ppl_{LM}(m)}{Ppl_{CCLM}(m)}
+\]
+
+\begin{table*}
+  \centering
+  \input{floats/info_gain.tex}
+  \caption{Mean information gain of the CCLM over the baseline model, for messages in the test set.}
+  \label{tab:model-info-gain}
+\end{table*}
+
+While the absolute performance of the LSTM models is better,
+we find the the transformer CCLMs have higher information gain 
+\cref{tab:model-info-gain}.
 
@@
 \begin{tikzpicture}
-  \confusiongroupplot{floats/lstm-3-1_comm_infer_confusion.csv}{floats/transformer-3-3_comm_infer_confusion.csv}
+  \confusiongroupplot{lstm-3-1}{transformer-3-3}
 \end{tikzpicture}
@@
 
-On the other hand, when we consider the \emph{information gain}\bn{define earlier}
-of the CCLM over the corresponding (same encoder architecture)
-un-conditioned
+On the other hand, community indiscernibility is correlated with
+the average information gain of the CCLM over the baseline model,
+for messages in its test set \cref{fig:comm-pca}.
+Communities for which the language model bennifits most from 
+the conditioned model are generally easier for the LMCC to classify.
 
@@
 We find that various communities exhibit widely varying levels of
-linguistic complexity (\cref{hyp:varying-complexity}), with perplexity
+linguistic complexity \cref{hyp:varying-complexity}, with perplexity
 per-word ranging from less than 53 to 104 for the best model
@@
 
-We find a weak correlation (TODO; r , p) between the perplexity of the
+We find a weak correlation \cref{tab:cclm_lmcc_ppl} between the perplexity of the \bn{I'm not sure if we can make this claim or not with the Ppl-Ppl correlations.
+However, given that *all* of them have positive correlation, even though the p-values are somewhat higher than .1 maybe we can?}
 CCLM and the indiscernibility of the corresponding community
@@
 albeit weakly. For LSTM models, the perplexity per word, averaged over
-all communities, varies between 78.64 and 80.70. It is steadily
-increasing when $l_c$ decreases.  For transformer models, it varies
-between 89.38 and 122.56. Here perplexity per-word is not a monotonous
+all communities, varies between 76.80 and 78.64. \bn{80.70 is the baseline model} 
+The perplexity per word is slightly higher for $l_c=0$, and has very
+little difference between the other models, with $l_c=1$ being the best.
+For transformer models, it varies
+between 89.38 and 122.56. Here perplexity per-word is clearly not a monotonous
 function of $l_c$: the maximum is reached at $l_c=2$. We did not
@@
 
-The representation of communities built by CCLM is exhibiting a clear
+The pattern of information gain by community is similar across architectures.
+Communities that bennifit most from the conditioned model behave that way for 
+both the LSTM and transformer.
+However, there are some differences.
+For example, /r/relationships has an information gain of \num{1.018} for
+the best CCLM LSTM but only \num{0.998} for the best transformer model. \bn{todo: community-wise CCLM info gain}
+Why was the LSTM able to take advantage of the community information for
+this community, where the transformer wasn't?
+It would be interesting to investigate these differences in future work,
+since it could reveal differences in the kind of linguistic variation the different
+model architectures capture.
+
+The representation of communities built by CCLM exhibits a clear
 correlation with user-coocurrence patterns
-(\cref{hyp:extra-linguistic-correlation}, TODO; r , p). A natural
+(\cref{hyp:extra-linguistic-correlation}). A natural
 question that may arise is how much this correlation can be explain by
@@
 
-- Comparison with topic modelling
-
-The pattern of how much the perplexity falls across different communities
-is similar across architectures, but with some differences.\jp{Either say something specific here or move the whole paragraph to future work.}
-It could be interesting to further investigate these differences in future work,
-since it could reveal differences in the kind of linguistic variation the different
-model architectures capture.
-
 \bibliography{paper}

### 2020-11-19 paper/paper.tex
deal with some comments and other minor changes

@@
 
+\jp{We want to identify a niche that we want to fill. A quick summary of the state of the art + what is missing.}
+
 In the present work, we investigate linguistic variation across online communities.
@@
 
-Second, we want to find out how, and how much, CCLMs can be attuned to
+Second, we want to find out how, and how much, CCLMs are attuned to
 the specificity of communities. We test whether 1. the linguistic
@@
 
-To sum up, we propose to test the following hypotheses:
+To sum up we aim at testing the following hypotheses:
 \begin{hypotheses}
@@
 
-We experiment with two kinds of model architecture: LSTM \citep{Hochreiter1997}\jp{Is it really a simple or a bi-lstm?
-Even If it's simple it may be worth specifying due to the prevalence of bilstm everywhere.}
-and Transformer \citep{Vaswani2017}.  In either case, the model
-is organised as a standard $n$-layer neural sequence encoder with
-token embedding (we set $n=3$ for all models). As is standard, words are predicted
-using a softmax projection layer over a pre-determined vocabulary.
-These encoders form the core of the community-conditioned language models (CCLM).
+We experiment with two kinds of model architecture: LSTM
+\citep{Hochreiter1997}\jp{Is it really a simple or a bi-lstm? Even If
+  it's simple it may be worth specifying due to the prevalence of
+  bilstm everywhere.}  and Transformer \citep{Vaswani2017}.  In either
+case, the model is organised as a standard $n$-layer neural sequence
+encoder (we set $n=3$ for all models). As is standard, word tokens are
+first converted to vectors using a trainable embedding layer. At the
+other end, word tokens are predicted using a softmax projection layer. The set of word tokens is a
+pre-determined vocabulary.  These encoders form the core of the
+community-conditioned language models (CCLM), and we call them
+unconditioned models.
 
@@
 We also note that the conditioned models perform differently across different communities---%
-for some communities, perplexity of the conditioned models is a lot lower,\bn{is this ok? took ``test'' out but made clear in the caption that it's the test set}%\jp{and training?}
-whereas for others it is the same or even slightly higher (\cref{tab:model-ppls}).\bn{talk about this in terms of mutual information instead?}
+for some communities, perplexity of the conditioned models is a lot lower than the unconditioned models,
+whereas for others it is the same or even slightly higher (\cref{tab:model-ppls}).\bn{talk about this in terms of mutual information instead?}\jp{What is that?}
 As an informal observation, we note 
@@
 communities where conditioning has the least effect
-tend to be more organized around more ``general interest'' topics,
-such as /r/relationships, where the subject matter is relevant to a broad
+tend to be more organized around more general interest topics,
+such as the \emph{relationships} subreddit, where the subject matter is relevant to a broad
 range of people.
-\bn{I really mean this as a quality of the topic, rather than the community itself, although the two are probably correlated}% \jp{Is there a way to quantify general interest topics?
-Community size in terms of number of users?}
 Conditioning the model on community has the most benefit
@@
 This makes sense intuitively, since communities with more niche topics
-would tend to have more specialized vocabulary and speech patterns.\bn{i think it's ok---i've seen this broader sense used in academic papers}%\jp{can you say speech for written language?}
+would tend to have more specialized vocabulary and other linguistic patterns.
 
@@
 albeit weakly. For LSTM models, the perplexity per word, averaged over
-all communities, varies between 76.80 and 78.64. \bn{80.70 is the baseline model} 
+all communities, varies between 76.80 and 78.64. (with 80.70 for the unconditioned model)
 The perplexity per word is slightly higher for $l_c=0$, and has very
@@
 However, there are some differences.
-For example, /r/relationships has an information gain of \num{1.018} for
+For example, the subreddit \emph{relationships} has an information gain of \num{1.018} for
 the best CCLM LSTM but only \num{0.998} for the best transformer model. \bn{todo: community-wise CCLM info gain}

### 2020-11-19 paper/paper.tex
comment out missing inputs
typography

@@
   \centering
-  \input{floats/info_gain.tex}
+  % \input{floats/info_gain.tex}
   \caption{Mean information gain of the CCLM over the baseline model, for messages in the test set.}
@@
   \centering
-  \input{floats/lmcc_gain.tex}
+  % \input{floats/lmcc_gain.tex}
   \caption{
@@
 We find that various communities exhibit widely varying levels of
-linguistic complexity \cref{hyp:varying-complexity}, with perplexity
+linguistic complexity \ref{hyp:varying-complexity}, with perplexity
 per-word ranging from less than 53 to 104 for the best model
@@
 we find that CCLMs can be used to recognize communities based on
-latent bayesian classification (\cref{hyp:LMCC-works}). Even the most
+latent bayesian classification \ref{hyp:LMCC-works}. Even the most
 generic communities appear to exhibit fairly specific linguistic
@@
 
-We find a weak correlation \cref{tab:cclm_lmcc_ppl} between the perplexity of the \bn{I'm not sure if we can make this claim or not with the Ppl-Ppl correlations.
+We find a weak correlation (\cref{tab:cclm_lmcc_ppl}) between the perplexity of the \bn{I'm not sure if we can make this claim or not with the Ppl-Ppl correlations.
 However, given that *all* of them have positive correlation, even though the p-values are somewhat higher than .1 maybe we can?}
 CCLM and the indiscernibility of the corresponding community
-(\cref{hyp:rich-harder-to-identify}). Because we have relatively few,
+\ref{hyp:rich-harder-to-identify}. Because we have relatively few,
 and only large communities in our test set, it is difficult to
@@
 We find that the layer at which the community embedding is taken into
-account in a CCLM influences its perplexity (\cref{hyp:layer-effect}),
+account in a CCLM influences its perplexity \ref{hyp:layer-effect},
 albeit weakly. For LSTM models, the perplexity per word, averaged over
@@
 For example, the subreddit \emph{relationships} has an information gain of \num{1.018} for
-the best CCLM LSTM but only \num{0.998} for the best transformer model. \bn{todo: community-wise CCLM info gain}
+the best CCLM LSTM but only \num{0.998} for the best transformer model.
 Why was the LSTM able to take advantage of the community information for
@@
 correlation with user-coocurrence patterns
-(\cref{hyp:extra-linguistic-correlation}). A natural
+(\ref{hyp:extra-linguistic-correlation}). A natural
 question that may arise is how much this correlation can be explain by

### 2020-11-20 paper/paper.tex
related work intro

@@
 
+In this work, we present results using a conditioned neural language models
+to model variation between speech communities.
+The architecture of these models concatenates a vector representation
+of the conditioned variable to the input of the sequence model.
+This approach has been applied in various conditioned text generation domains such as 
+image captioning \citep{Vinyals2015}, machine translation \citep{Kalchbrenner2013},
+but it has not, to our knowledeg, been used extensively to study linguistic variation.
+There are, however some related methods that have been used to study variation,
+and related applications of conditional neural language models.
+
+In the later category, conditoinal nerual language models have been used to study linguistic change.
+
 \paragraph{Vector representations of online communities encode semantic relationships}

### 2020-11-20 paper/paper.tex
2 paragraphs on related work

@@
 image captioning \citep{Vinyals2015}, machine translation \citep{Kalchbrenner2013},
-but it has not, to our knowledeg, been used extensively to study linguistic variation.
+but it has not, to our knowledge, been used extensively to study linguistic variation.
 There are, however some related methods that have been used to study variation,
@@
 
-In the later category, conditoinal nerual language models have been used to study linguistic change.
-
-\paragraph{Vector representations of online communities encode semantic relationships}
-https://www.aclweb.org/anthology/W17-2904.pdf
-
-\paragraph{Reddit Mining to Understand Gendered Movements}
-http://ceur-ws.org/Vol-2578/DARLIAP3.pdf
+In the later category, conditional neural language models have been
+used to study linguistic change.
 
@@
 
-\paragraph{Topic modelling}
-- Discovering Discrete Latent Topics with Neural Variational Inference https://arxiv.org/pdf/1706.00359.pdf
-- JeyHan Lau's work
+\cite{khan2020reddit}
 
 \paragraph{Pre-trained language models}
-- Unsupervised Domain Clusters in Pretrained Language Models https://www.aclweb.org/anthology/2020.acl-main.692/
+- Unsupervised Domain Clusters in Pretrained Language Models https://www.aclweb.org/anthology/2020.acl-main.692/ % 404?
+
+\paragraph{Topic modelling}
+Modelling communities is not unlike the classic task of \emph{topic
+  modelling}. In topic modelling, one most often want to
+\emph{discover} topics. In constrast, we take as a given that a given
+message belongs to a given community. However, deep-learning topic
+modelling is often built around the idea of a conditioned language
+model \citep{Lau2017a}, with a separate topic-discovery module.
+
+% - Discovering Discrete Latent Topics with Neural Variational Inference https://arxiv.org/pdf/1706.00359.pdf
+
+% \paragraph{Vector representations of online communities encode semantic relationships}
+% https://www.aclweb.org/anthology/W17-2904.pdf
 

### 2020-11-20 paper/paper.tex
cleanup

@@
 recognize messages coming from a given community $c_j$. We could use
-something like an F-score, but we prefer to use the entropy metric,
-which is has stronger information-theoretic grounds, and is a measure
-of the uncertainty of the LMCC, given messages from a particular 
-community. 
-Technically, the linguistic indiscernibility of community $c_j$ is defined 
-as the exponential of the entropy of the $j$ row in \(C\),
-which is a probability distribution 
-of predicted community for a random message from $c_j$.
-In sum, we have the formula: \bn{Took out $D$. is it needed elsewhere?}
+something like an F-score, but we prefer to use a perplexity-like metric,
+which is has stronger information-theoretic grounds.
+We define the linguistic indiscernibility of community $c_j$, noted $\Ind_j$, 
+as the exponential of the entropy of the $j$th row in \(C\),
+(which is a probability distribution
+of predicted community for a random message from $c_j$):
 \[\Ind_j = e^{-\sum_i C_{ij} log(C_{ij})}\]
+Intuititely, $\Ind_j$ is a measure of the uncertainty of the LMCC,
+given messages from a community $c_j$.
 
-Indiscernibility is a robust measure: it is almost perfectly
-correlated for various CCLM (transformer, LSTM and various $c$) $M$.
-In fact, we observe a correlation between the indiscernibility 
-measures by every pair of models with Pearson's r $\geq 0.98$ 
-and $p < 0.0001$ (see, for example, \cref{fig:lmcc_ppl}).
+Indiscernibility of a given community is a robust measure: it is
+almost perfectly correlated across CCLMs (such as transformers, LSTM
+and various values of $l_c$).  In fact, we observe a correlation
+between the indiscernibility measures by every pair of models with
+Pearson's $r \geq 0.98$ and $p < 0.0001$ (see, for example,
+\cref{fig:lmcc_ppl}).
 
@@
 
-\jp{We could also compute the entropy of rows of Cij = snap cosine similarity w/ softmax.}
+% We could also check if communities are well-separated in the SNAP community encodings
 

### 2020-11-20 paper/paper.tex
checking notes and consistency

@@
 for some communities, perplexity of the conditioned models is a lot lower than the unconditioned models,
-whereas for others it is the same or even slightly higher (\cref{tab:model-ppls}).\bn{talk about this in terms of mutual information instead?}\jp{What is that?}
+whereas for others it is the same or even slightly higher (\cref{tab:model-ppls}).\bn{talk about this in terms of mutual information instead?}\jp{This should make no difference (but I'm not sure on how you are taking averages; see other comment.)}
 As an informal observation, we note 
@@
 \]
-
+\jp{What we had before: average entropy gain is $H(m)-H_0(m)$, and
+  then taking the average over all messages:
+  $avg(H(m)-H_0(m)) = avg(H(m))-avg(H_0(m))$. Taking the exponential
+  gives a ratio of perplexity.  Here it seems that you have commuted
+  exponential and average. It is rather weird to do average in the
+  exponential space without normalising to a probability.}
 \begin{table*}
@@
 using data which spans January 2014 to April 2017. We refer the reader
-to \citet{Kumar2018}\jp{I actually cannot find in their paper what
-  they do.} for details, but the important point is that no linguistic
+to \citet{Kumar2018} % JP: I actually cannot find in their paper what they do.
+for details, but the important point is that no linguistic
 information is used to create these embeddings: they only depend on
@@
 probability of a message, which can be computed as
-\[P(m) = \frac 1 {\sum_i P(c=c_i) P(m|c=c_i ) }\].
+\[P(m) = \frac 1 {\sum_i P(c=c_i) P(m|c=c_i ) }\]
 
@@
 from community $c_j$.
-That is: \[C_{ij} = average_{m \in Messages(c_j)}(P(c=c_i | m))\].
+\[C_{ij} = average_{m \in Messages(c_j)}(P(c=c_i | m))\]
 As a sanity check, we observe that the diagonal dominates the matrix: on average the LMCC picks the right community for a message (\cref{fig:confusion}).
@@
 However, we observe that community-wise CCLM perplexity is only weakly (if at all) \bn{Changed the indiscernability (y-axis) from entropy to Ppl and the p-values all shot up above 0.1.
-However the r is still very weakly positive for every model.}
+However the r is still very weakly positive for every model.}\jp{Are indiscernability values correct?
+Seems inconsistent with the confusion matrix.}
 correlated with community indiscernibility under the corresponding latent 
@@
   \caption{
-  Pearson's r correlation between community-wise linguistic perplexity and indescernibility for each model.
+  Pearson's $r$ correlation coefficient between community-wise linguistic perplexity and indescernibility for each model.
   }
@@
   \centering
-  % \input{floats/lmcc_gain.tex}
+  \input{floats/lmcc_gain.tex}
   \caption{
-    Pearson's r correlation between community-wise information gain (of the CCLM over the un-conditioned model) and indescernibility.
+    Pearson's \(r\) correlation coefficient between community-wise information gain (of the CCLM over the un-conditioned model) and indescernibility.
   }
@@
 
-With indiscernibility values of 1.5 to 3.5 (out of a maximum of 46),
+With indiscernibility values of 1.5 to 3.5 (out of a maximum of 46)\jp{Inconsistent with data; to revisit},
 we find that CCLMs can be used to recognize communities based on
@@
 correlation with user-coocurrence patterns
-(\ref{hyp:extra-linguistic-correlation}). A natural
+\ref{hyp:extra-linguistic-correlation}. A natural
 question that may arise is how much this correlation can be explain by

### 2020-11-20 paper/paper.tex
todos discussion

@@
 
-We experiment with two kinds of model architecture: LSTM
-\citep{Hochreiter1997}\jp{Is it really a simple or a bi-lstm? Even If
-  it's simple it may be worth specifying due to the prevalence of
-  bilstm everywhere.}  and Transformer \citep{Vaswani2017}.  In either
+We experiment with two kinds of model architecture: a simple unidirectional LSTM
+\citep{Hochreiter1997} and a Transformer \citep{Vaswani2017}.  In either
 case, the model is organised as a standard $n$-layer neural sequence
@@
 the language modeling was achieved by masking and incrementally un-masking input tokens
-in a left-to-right fashion, following \citet{BERT}\jp{Or some other standard paper}.  
+in a left-to-right fashion (so we have an auto-regressive\jp{check} model).
 We used the AdamW \citep{Loshchilov2019} optimization algorithm,
-with an initial learning rate of \num{0.001}.\jp{Also for the transformer? No simulated annealing?}
+with an initial learning rate of \num{0.001}, without any extra control on the decay of learning rate.
 We used a batch size of \num{256} and a maximum sequence length of \num{64} tokens, 
@@
 for some communities, perplexity of the conditioned models is a lot lower than the unconditioned models,
-whereas for others it is the same or even slightly higher (\cref{tab:model-ppls}).\bn{talk about this in terms of mutual information instead?}\jp{This should make no difference (but I'm not sure on how you are taking averages; see other comment.)}
+whereas for others it is the same or even slightly higher (\cref{tab:model-ppls}).
+\bn{talk about this in terms of mutual information instead!}
 As an informal observation, we note 
@@
   then taking the average over all messages:
-  $avg(H(m)-H_0(m)) = avg(H(m))-avg(H_0(m))$. Taking the exponential
+  $avg(H(m)-H_0(m)) = avg(H(m))-avg(H_0(m)) = H - H₀$. Taking the exponential
   gives a ratio of perplexity.  Here it seems that you have commuted
   exponential and average. It is rather weird to do average in the
-  exponential space without normalising to a probability.}
+  exponential space without normalising to a probability.
+
+  Compare: $e^{avg(H(m))} / e^{avg(H_0(m))}$ with $avg \frac{e^{H(m)}}{e^{H₀(m)}}$.
+
+}
 \begin{table*}
@@
 using data which spans January 2014 to April 2017. We refer the reader
-to \citet{Kumar2018} % JP: I actually cannot find in their paper what they do.
+to \citet{Kumar2018}
+\jp{What they say they do (sec 6.2) More formally, we train user and community embeddings basedon the bipartite multigraph between users and communities, whereone user-community edge,(ui,cj), exists in the graph for each timea user,ui, posts on community,cj.
+Adapting the approach used ina number of recent successful works [27,53,62], we learn user andcommunity embeddings,ui∈Rd,cj∈Rd, from this graph using anegative-sampling optimization algorithm
+}
 for details, but the important point is that no linguistic
@@
 
-We find a weak correlation (\cref{tab:cclm_lmcc_ppl}) between the perplexity of the \bn{I'm not sure if we can make this claim or not with the Ppl-Ppl correlations.
+We cannot exclude a correlation (\cref{tab:cclm_lmcc_ppl}) between the perplexity of the
+% \bn{I'm not sure if we can make this claim or not with the Ppl-Ppl correlations.
 However, given that *all* of them have positive correlation, even though the p-values are somewhat higher than .1 maybe we can?}

### 2020-11-20 paper/paper.tex
address some todos

@@
 embedding with the community embedding created by \citet{Kumar2018},
-who generated them from the message-community co-occurence matrix,
-using data which spans January 2014 to April 2017. We refer the reader
+who generated them using using a negative-sampling optimization
+algorithm and the message-community co-occurence matrix as ground
+truth, using data which spans January 2014 to April 2017.
+We refer the reader
 to \citet{Kumar2018}
-\jp{What they say they do (sec 6.2) More formally, we train user and community embeddings basedon the bipartite multigraph between users and communities, whereone user-community edge,(ui,cj), exists in the graph for each timea user,ui, posts on community,cj.
-Adapting the approach used ina number of recent successful works [27,53,62], we learn user andcommunity embeddings,ui∈Rd,cj∈Rd, from this graph using anegative-sampling optimization algorithm
-}
 for details, but the important point is that no linguistic
 information is used to create these embeddings: they only depend on
-whether and how much a given user participates in a given
+whether a given user participates in a given
 community. In sharp contrasts, CCLM community embeddings depend in no
@@
 
-With indiscernibility values of 1.5 to 3.5 (out of a maximum of 46)\jp{Inconsistent with data; to revisit},
-we find that CCLMs can be used to recognize communities based on
-latent bayesian classification \ref{hyp:LMCC-works}. Even the most
-generic communities appear to exhibit fairly specific linguistic
-patterns. Furthermore, these patterns are just as recognizable by LSTM
-or Transformer models, when properly configured and trained.
+With indiscernibility values evenly ranging over the whole possible
+range (0 to 1), we find that CCLMs can be used to recognize certain
+communities based on latent bayesian classification
+\ref{hyp:LMCC-works}, but not all.  If generic communities exhibit
+distinctive linguistic patterns, none of our model is able to discern
+them. In contrast, the distinctive patterns of narrow-topic
+communities are just as recognizable by all our LSTM and Transformer
+models.
 
@@
 (perplexity per word above 100) have an indiscernibility of about 3
-(out of a maximum of 46), so they retain fairly specific patterns:
+(out of a maximum of 46)\jp{make consistent with data}, so they retain fairly specific patterns:
 they are not simply an even mixture of the kind of comments found in

### 2020-11-20 paper/paper.tex
more todos done

@@
 However, we observe that community-wise CCLM perplexity is only weakly (if at all) \bn{Changed the indiscernability (y-axis) from entropy to Ppl and the p-values all shot up above 0.1.
-However the r is still very weakly positive for every model.}\jp{Are indiscernability values correct?
-Seems inconsistent with the confusion matrix.}
+However the r is still very weakly positive for every model.}
 correlated with community indiscernibility under the corresponding latent 
@@
 
-\jp{actually, if we had many more communities, then the lines could be ``more blurred'' between various communities, and we may have a stronger correlation here.}
 \begin{figure}
@@
 (perplexity per word above 100) have an indiscernibility of about 3
-(out of a maximum of 46)\jp{make consistent with data}, so they retain fairly specific patterns:
+(out of a maximum of 46)\jp{make consistent with data once the plot is stablilized}, so they retain fairly specific patterns:
 they are not simply an even mixture of the kind of comments found in

### 2020-11-22 paper/paper.tex
fix information gain computation/definiton

@@
 \usepackage{latexsym}
+\usepackage{amsmath}
+\usepackage{amssymb}
 \usepackage{siunitx}
@@
 \usepackage{pgfplots}
-\usepackage{pgfplotstable}
 \usepackage{todonotes}
 \usepackage{enumitem}
+\usepackage{cleveref}
 \usepackage{array}
-\usepackage{cleveref}
+\usepackage{makecell}
+\usepackage{multirow}
 
@@
 \pgfplotsset{compat=1.16}
-\pgfplotstableset{
-  column type=r,
-  fixed zerofill,
-  clear infinite,
-  %dec sep align,
-  %empty cells with={\ensuremath{-}},
-}
 
@@
 \newcommand\Ppl{\mathsf{Ppl}}
+\newcommand\IG{\mathsf{IG}}
 \newcommand\Ind{I}
+\DeclareMathOperator*{\avg}{avg}
 \newcommand\BibTeX{B\textsc{ib}\TeX}
@@
 and used the weights from the epoch with the smallest validation loss for testing.
-In general, transformer models trained for about half as long as the LSTM models (\cref{tab:best-epoch}).
-
-\begin{table}
-  \centering
-  \input{floats/best_epoch.tex}
-  \caption{Epoch with the lowest validation loss.}
-  \label{tab:best-epoch}
-\end{table}
-
-%\pgfplotstabletypeset[
-  %columns={best_epoch,lm_ppl,gain},
-  %columns/model/.style={string type,column type=l},
-  %columns/best_epoch/.style={column name={Epochs}, int detect},
-  %columns/lm_ppl/.style={column name={Ppl}},
-  %columns/gain/.style={column name={IG},precision=3,empty cells with={\ensuremath{-}}},
-  %every nth row={5}{before row=\midrule},
-  %]{floats/model.csv}
+In general, transformer models trained for about half as long as the LSTM models (test epoch, \cref{tab:model-results}).
 
@@
 For this, we use the negative exponential of the mean cross-entropy loss 
-for the sequence.
-
+of the language model for a given message and average over a set of messages, $M$.
+\[\Ppl_M = \avg_{m\in M}(-e^{H(m)})\] \bn{Following the logic from IG, should we change Ppl to the exponential of the average cross entropy, rather than using the average of the message-level perplexity?}
 As one would expect, 
 the conditioned models have lower perplexity
-than their respective unconditioned baseline models,
+than their respective unconditioned baseline models (\cref{tab:model-results}),
 but this effect is different in magnitude
@@
 the best model ($l_c=1$) concatenates the community embedding after the first encoder layer,
-with two more LSTM layers after that.
+with two subsequent LSTM layers. 
 The best transformer model ($l_c=3$), on the other hand,
 incorporates the community information last,
-after all the sequence encoder layers
-(\cref{tab:model-ppls}).
+after all the sequence encoder layers.
 
-%TODO: For space reasons, it might be better to put the community-level table in the appendix and have the table in body just show the means (and perhaps medians/stddevs?) for each architecture.}
+\begin{table}
+  \footnotesize
+  \centering
+  \input{floats/model_results.tex}
+  \caption{
+    Performance of baseline (first row for each encoder architecture) and CCLM models. 
+    Perplexity and information gain are shown for the entire test set 
+    (\num{5000} messages for each community).}
+  \label{tab:model-results}
+\end{table}
 
-We also note that the conditioned models perform differently across different communities---%
-for some communities, perplexity of the conditioned models is a lot lower than the unconditioned models,
-whereas for others it is the same or even slightly higher (\cref{tab:model-ppls}).
-\bn{talk about this in terms of mutual information instead!}
+We also consider model performance stratified by community;
+that is, $\Ppl_{M_j}$, where $M_j$ is the test-set messages from community $c_j$.
+We observe a great deal of variation in baseline perplexity across communities,
+with $\Ppl_{M_j}$ ranging from \num{53.07} to \num{115.29} for the LSTM model
+(\cref{tab:community-results}).
+
+\subsection{Information gain of the CCLM}
+
+Next, we consider the \emph{information gain} of each CCLM over its baseline
+un-conditioned counterpart with the same sequence encoder architecture.
+For a given message, information gain is defined as the difference
+between the cross-entropy of the unconditioned model and the conditioned model:
+\[H_{LM}(m) - H_{CCLM}(m)\]
+For a set of messages, $M$, we consider the average information gain
+in exponential space (as a ratio of perplexities):
+\[\IG_M = \frac{e^{\avg_M(H_{LM}(m))}}{e^{\avg_M(H_{CCLM}(m))}}\]
+%\jp{What we had before: average entropy gain is $H(m)-H_0(m)$, and
+  %then taking the average over all messages:
+  %$avg(H(m)-H_0(m)) = avg(H(m))-avg(H_0(m)) = H - H₀$. Taking the exponential
+  %gives a ratio of perplexity.  Here it seems that you have commuted
+  %exponential and average. It is rather weird to do average in the
+  %exponential space without normalising to a probability.
+
+  %Compare: $e^{avg(H(m))} / e^{avg(H_0(m))}$ with $avg \frac{e^{H(m)}}{e^{H₀(m)}}$.
+%}
+
+While the absolute performance of the LSTM models is better,
+we find that the best transformer CCLM has higher information gain 
+than the best LSTM (\cref{tab:model-results}).
+
+The conditioned models also perform differently across different communities---%
+for some communities, $\IG_{M_j}$ around \num{1}, or even slightly lower,
+meaning that the CCLM has no advantage over the unconditioned baseline for that community.
+For others communities it is higher, meaning that the CCLM performs better than 
+the baseline for those communities (\cref{tab:community-results}).
+
+\begin{table*}
+  \normalsize
+  \centering
+  \input{floats/community_results.tex}
+  \caption{
+    Model performance by community. We show the absolute perplexity 
+    of the baseline models and the information gain of the CCLMs.
+    Note that the rows are sorted by the perplexity of the baseline
+    LSTM model.
+}
+  \label{tab:community-results}
+\end{table*}
+
 As an informal observation, we note 
@@
 tend to be more organized around more general interest topics,
-such as the \emph{relationships} subreddit, where the subject matter is relevant to a broad
+such as the \emph{relationships} and \emph{advice} subreddits, where the subject matter is relevant to a broad
 range of people.
 Conditioning the model on community has the most benefit
-for narrower special-interest subreddits such as those organized
+for narrower special-interest subreddits, such as those organized
 around a certain videogame, sports team, or subculture.
@@
 
-\begin{table*}
-  \footnotesize
-  \centering
-  \input{floats/model_ppls.tex}
-  \caption{Mean model perplexity for messages in the test set, broken down by community.}
-  \label{tab:model-ppls}
-\end{table*}
-
-\subsection{Information gain of the CCLM}
-
-We consider the average information gain of each CCLM over its baseline
-un-conditioned counterpart with the same sequence encoder architecture.
-For a given message, the information gain is defined as the ratio of the
-perplexity of the baseline model to that of the CCLM:
-\[
-  IG_{CCLM}(m) = \frac{Ppl_{LM}(m)}{Ppl_{CCLM}(m)}
-\]
-\jp{What we had before: average entropy gain is $H(m)-H_0(m)$, and
-  then taking the average over all messages:
-  $avg(H(m)-H_0(m)) = avg(H(m))-avg(H_0(m)) = H - H₀$. Taking the exponential
-  gives a ratio of perplexity.  Here it seems that you have commuted
-  exponential and average. It is rather weird to do average in the
-  exponential space without normalising to a probability.
-
-  Compare: $e^{avg(H(m))} / e^{avg(H_0(m))}$ with $avg \frac{e^{H(m)}}{e^{H₀(m)}}$.
-
-}
-\begin{table*}
-  \centering
-  % \input{floats/info_gain.tex}
-  \caption{Mean information gain of the CCLM over the baseline model, for messages in the test set.}
-  \label{tab:model-info-gain}
-\end{table*}
-
-While the absolute performance of the LSTM models is better,
-we find the the transformer CCLMs have higher information gain 
-\cref{tab:model-info-gain}.
-
 \subsection{Analysis of CCLM community embeddings}
 
-As a qualitative assesment of the community embeddings, 
+As a qualitative assessment of the community embeddings, 
 we manually identified topic categories into which at least three
@@
 \begin{figure*}
-\begin{tikzpicture}
-  \confusiongroupplot{lstm-3-1}{transformer-3-3}
-\end{tikzpicture}
+%\begin{tikzpicture}
+  %\confusiongroupplot{lstm-3-1}{transformer-3-3}
+%\end{tikzpicture}
 \caption{LMCC confusion matrices for the best LSTM (left, $c=1$) and transformer (right, $c=3$) models.}

### 2020-11-22 analysis.py

@@ -81,9 +81,11 @@ df_m = pd.DataFrame([], index=pd.Index(models, name='model'))
 df_mm = pd.DataFrame([],index=pd.MultiIndex.from_tuples(combinations(cond_models, 2), names=['model1', 'model2']))
 
 # Best model validation epoch (used for testing)
-best_epoch = pd.DataFrame([
+best_epoch = pd.Series([
     int(open(f'{model_dir}/{model}/saved-epoch.txt').read())
-    for model in models], index=models, columns=['best_epoch'])
+    for model in models], index=models)
+
+df_m['best_epoch'] = best_epoch
 
 
 ##### Language model perplexity (CCLM & baseline)
@@ -96,8 +98,6 @@ df_m['lm_ppl'] = lm_ppl[models].mean()
 
 ## Mean language model perplexity by community
 df_c = add_columns(df_c, lm_ppl.groupby('community').mean(), suffix='_lm_ppl')
-
-df_m = add_columns(df_m, best_epoch)
 
 
 #### Community embeddings
@@ -132,6 +132,7 @@ corr = pd.DataFrame([pearsonr(df_cc[f'{model}_cos_sim'], df_cc['snap_cos_sim'])
     for model in cond_models], 
     index=cond_models, 
     columns=['snap_cos_sim_corr_r', 'snap_cos_sim_corr_p'])
+
 df_m = add_columns(df_m, corr)
 
 
@@ -181,6 +182,7 @@ corr = pd.DataFrame([pearsonr(df_c[f'{model}_lmcc_ppl'], df_c[f'{model}_lm_ppl']
     for model in cond_models],
     index=cond_models,
     columns=['lmcc_cclm_corr_r', 'lmcc_cclm_corr_p'])
+
 df_m = add_columns(df_m, corr)
 
 ## Commuinty indiscrenibility correlation between two example models
@@ -191,31 +193,59 @@ print(f"r = {r:0.4f} p = {p:0.6f}")
 
 ##### Information gain (AKA mutual information)
 
-df_ig = pd.DataFrame([], index=lm_ppl.index)
-df_ig['community'] = lm_ppl['community']
-df_ig[cond_lstms] = lm_ppl[cond_lstms].apply(lambda x: lm_ppl['lstm-3']/x)
-df_ig[cond_transformers] = lm_ppl[cond_transformers].apply(lambda x: lm_ppl['transformer-3']/x)
+## Model entropy on test examples
[…]

### 2020-11-22 paper/paper.tex
scale indiscernibility by number of communities

@@
 \newcommand\IG{\mathsf{IG}}
-\newcommand\Ind{I}
-\DeclareMathOperator*{\avg}{avg}
+\newcommand\Ind{\mathsf{Ind}}
+\DeclareMathOperator*{\avg}{average}
 \newcommand\BibTeX{B\textsc{ib}\TeX}
@@
 for testing, \num{5000} for evaluation, and \num{40000} for training.
+
+In the remainder of the paper we use $M$ to refer to messages in the test set,
+and $M_j$ for the partition of the test set originating from community $c_j$.
 
@@
 belongs to a community $c_i$ is calculated as follows:
-\[P(c=c_i | m) = P(m | c_i)\frac {P(m)} {P(c=c_i)}\]
+\[P(c=c_i \mid m) = P(m | c_i)\frac {P(m)} {P(c=c_i)}\]
 In the above,
 $P(c=c_i)$ is the frequency of community $c_i$ in the dataset, and
-$P(m | c_i)$ (the probability that a message $m$
-belongs in a community $c_i$. ) is estimated by the CCLM. $P(m)$ is the absolute
+$P(m \mid c_i)$ (the probability of a message $m$
+in the context of community $c_i$) is estimated by the CCLM. 
+$P(m)$ is the absolute
 probability of a message, which can be computed as
-\[P(m) = \frac 1 {\sum_i P(c=c_i) P(m|c=c_i ) }\]
+\[P(m) = \sum_i P(c=c_i) P(m\mid c=c_i ). \]
+Since the test set is balanced across communities,
+we have an even prior for $P(c=c_i)$, giving us
+\[P(c=c_i\mid m) \propto  P(m\mid c_i).\]
 
 We can characterise the quality the LMCC with a confusion matrix
-$C$, defined by $C_{ij} = P(c=c_i | m)$ for a given (test set) message $m$ 
+$C$, defined by $C_{ij} = P(c=c_i \mid m)$ for a given (test set) message $m$ 
 from community $c_j$.
-\[C_{ij} = average_{m \in Messages(c_j)}(P(c=c_i | m))\]
+\[C_{ij} = \avg_{m\in M_j}P(c=c_i \mid m)\]
 As a sanity check, we observe that the diagonal dominates the matrix: on average the LMCC picks the right community for a message (\cref{fig:confusion}).
@@
 which is has stronger information-theoretic grounds.
-We define the linguistic indiscernibility of community $c_j$, noted $\Ind_j$, 
-as the exponential of the entropy of the $j$th row in \(C\),
-(which is a probability distribution
-of predicted community for a random message from $c_j$):
-\[\Ind_j = e^{-\sum_i C_{ij} log(C_{ij})}\]
-Intuititely, $\Ind_j$ is a measure of the uncertainty of the LMCC,
+To this end, we define the \emph{linguistic indiscernibility} of community $c_j$, 
+noted $\Ind_j$. 
+First, note that each row of $C$ is a probability distribution,
+since it is computed as the average of $P(c=c_i \mid m)$ for $m\in M_j$.
+The exponential of the entropy of the $j$th row of \(C\) is therefore a perplexity score.
+To compute $\Ind_j$, we scale this perplexity by the number of communities, 
+which makes $\Ind_j$ comparable across models trained on different
+numbers of communities.\footnote{Although we only consider one set of \num{46} communities in this paper.}
+\[\Ind_j = \frac{e^{-\sum_i C_{ij} log(C_{ij})}}{|C_j|}\]
+Intuitively, $\Ind_j$ is a measure of the uncertainty of the LMCC,
 given messages from a community $c_j$.
+It ranges from $0$ (for a completely deterministic $C_j$) 
+to $1$ (if $C_j$ gives equal weight to every community).
 
@@
 
-However, we observe that community-wise CCLM perplexity is only weakly (if at all) \bn{Changed the indiscernability (y-axis) from entropy to Ppl and the p-values all shot up above 0.1.
-However the r is still very weakly positive for every model.}
+However, we observe that community-wise perplexity is only weakly (if at all) 
 correlated with community indiscernibility under the corresponding latent 
-LMCC (see \cref{fig:cclm_lmcc_ppl} and \cref{tab:cclm_lmcc_ppl}).
+LMCC (see \cref{tab:ind-corrs} and, for example, \cref{fig:cclm_lmcc_ppl}).
 In other words, the linguistic complexity of a given community
 is a poor predictor of how distinct its language is.
+
+\begin{table}
+  \centering
+  \input{floats/ind_corrs.tex}
+  \caption{
+    Pearson's $r$ correlation coefficient between community indiscernibility ($\Ind_j$) and
+    two different predictors: CCLM perplexity on messages from community $c_j$ ($\Ppl_{M_j}$)
+    , and CCLM information gain on messages from $c_j$ ($\IG_{M_j}$).
+  }
+  \label{tab:ind-corrs}
+\end{table}
 
@@
 
-\begin{table}
-  \centering
-  \input{floats/cclm_lmcc_ppl.tex}
-  \caption{
-  Pearson's $r$ correlation coefficient between community-wise linguistic perplexity and indescernibility for each model.
-  }
-  \label{tab:cclm_lmcc_ppl}
-\end{table}
-
 % We could also check if communities are well-separated in the SNAP community encodings
@@
 On the other hand, community indiscernibility is correlated with
-the average information gain of the CCLM over the baseline model,
-for messages in its test set \cref{fig:comm-pca}.
+the information gain of the CCLM 
+on messages from that community's test set 
+(see \cref{tab:ind-corrs} and, for example, \cref{fig:cclm_lmcc_ppl}).
 Communities for which the language model bennifits most from 
@@
 \end{figure}
-
-\begin{table}
-  \centering
-  \input{floats/lmcc_gain.tex}
-  \caption{
-    Pearson's \(r\) correlation coefficient between community-wise information gain (of the CCLM over the un-conditioned model) and indescernibility.
-  }
-  \label{tab:lmcc_gain}
-\end{table}
 

### 2020-11-22 paper/paper.tex
colors for all the community-level plots

@@
 %
+
 % File acl2020.tex
@@
 \begin{table*}
-  \normalsize
+  \small
   \centering
@@
 \begin{tikzpicture}
-  \commpcagroupplot{floats/lstm-3-1_pca.csv}{floats/transformer-3-3_pca.csv}
+  \begin{groupplot}[
+    group style={group size=2 by 1}
+  ]
+    \nextgroupplot
+    \commscatter{lstm-3-1_pca1}{lstm-3-1_pca2}
+    \nextgroupplot
+    \commscatter{transformer-3-3_pca1}{transformer-3-3_pca2}
+  \end{groupplot}
 \end{tikzpicture}
@@
   Pairs from the same category are colored according to \cref{fig:comm-pca}.
-  We also mark pairs from different (\texttimes) and pairs where one or both
+  We also mark pairs from different categories (\texttimes) and pairs where one or both
   communities belong to the ``Other'' category (\textbigcircle).
@@
 \begin{figure*}
-%\begin{tikzpicture}
-  %\confusiongroupplot{lstm-3-1}{transformer-3-3}
-%\end{tikzpicture}
+\begin{tikzpicture}
+  \confusiongroupplot{lstm-3-1}{transformer-3-3}
+\end{tikzpicture}
 \caption{LMCC confusion matrices for the best LSTM (left, $c=1$) and transformer (right, $c=3$) models.}
@@
 \begin{figure}
-  \pplscatter{lstm-3-1_lmcc}{transformer-3-3_lmcc}
+  \begin{tikzpicture}
+    \begin{axis}
+    \commscatter{lstm-3-1_indisc}{transformer-3-3_indisc}
+    \end{axis}
+  \end{tikzpicture}
   \caption{%
-    Correlation between indiscernibility for the best LSTM ($c = 1$) and transformer ($c = 3$) models
+    Correlation between indiscernibility for 
+    the best LSTM ($c = 1$) and transformer ($c = 3$) models
     (Pearson's $r = 0.99$, $p < 0.0001$).
@@
 
-\begin{figure}
-  \pplscatter{lstm-3-1_cclm}{lstm-3-1_lmcc}
+\begin{figure*}
+  \begin{tikzpicture}
+  \begin{groupplot}[
+    group style={group size=2 by 1},
+    yticklabels={,,}, xticklabels={,,},
+    legend style={anchor=north, legend columns=1, font=\tiny},
+    every node near coord/.append style={font=\sffamily\tiny}%
+  ]
+    \nextgroupplot
+    \commscatter[]{lstm-3-1_lm_ppl}{lstm-3-1_indisc}
+    \nextgroupplot
+    \commscatter[]{lstm-3-1_info_gain}{lstm-3-1_indisc}
+    \legend{General interest,Videogames,Technology,Sports,Female-focused,Other}
+  \end{groupplot}
+  \end{tikzpicture}
   \caption{%
-    Mean community-wise linguistic perplexity (x-axis) 
-    and indiscernibility (y-axis) 
-    (Pearson's r = \num{0.16}; p = \num{0.2991}).  
-    Shown here for the LSTM with $c = 1$.}
+    Mean community indiscernibility (y-axis), 
+    in relationship to linguistic complexity (left) and 
+    CCLM information gain (right). Shown here for the LSTM
+    model with $l_c=1$. See \cref{tab:ind-corrs} for 
+    Pearson's r correlations for all models.
+  }
   \label{fig:cclm_lmcc_ppl}
-\end{figure}
+\end{figure*}
 
@@
 (see \cref{tab:ind-corrs} and, for example, \cref{fig:cclm_lmcc_ppl}).
-Communities for which the language model bennifits most from 
+Communities for which the language model benefits most from 
 the conditioned model are generally easier for the LMCC to classify.
-
-\begin{figure}
-  \pplscatter{lstm-3-1_gain}{lstm-3-1_lmcc}
-  \caption{%
-    Mean community-wise information gain (x-axis) 
-    and indiscernibility (y-axis) 
-    %(Pearson's r = \num{0.16}; p = \num{0.2991}).  
-    Shown here for the LSTM with $c = 1$.}
-  \label{fig:cclm_lmcc_ppl}
-\end{figure}
 

### 2020-11-22 paper/plot_macros.tex
info gain table/plot
partial refactor of tables
colors for all the community-level plots

@@
-\newcommand{\commpcaplot}[1]{%
+\newcommand{\commscatter}[3][\community]{%
   \addplot[
@@
   },
-    nodes near coords*={\community},
+  nodes near coords*={#1},
+  nodes near coords align={vertical},
+  every node near coord/.append style={font=\sffamily\tiny},
   visualization depends on={value \thisrow{community} \as \community}
-  ] table [x=pca1, y=pca2, meta=category] {#1};
-}
-
-\newcommand{\commpcagroupplot}[2]{%
-  \begin{groupplot}[
-    group style={group size=2 by 1}, 
-    yticklabels={,,}, xticklabels={,,},
-    legend style={anchor=north, legend columns=1, font=\tiny},
-    every node near coord/.append style={font=\sffamily\tiny}%
-    ]
-    \nextgroupplot
-    \commpcaplot{#1}
-    \nextgroupplot
-    \commpcaplot{#2}
-    \legend{General interest,Videogames,Technology,Sports,Female-focused,Other}
-  \end{groupplot}
+  ] table [x={#2}, y={#3}, meta=category] {floats/comm.csv};
 }
@@
   point meta=explicit,
-] table [x=actual_comm, y=confered_to, meta=avg_prob] {#1};
+  ] table [x=actual_comm, y=pred_comm, meta=#1] {floats/confusion.csv};
 }
@@
     yticklabels=none,
-  xticklabel style={font=\sffamily\tiny, rotate=90},
+    x tick label style={font=\sffamily\tiny,rotate=45,anchor=east},
     xtick=data,
@@
 }
-
-\newcommand{\pplscatter}[2]{%
-  \begin{tikzpicture}
-  \begin{axis}[]
-  \addplot[
-    only marks, black,
-  ] table [x=#1, y=#2] {floats/cclm_lmcc_ppl.csv};
-  \end{axis}
-  \end{tikzpicture}
-}

### 2020-11-23 paper/paper.tex
finish float refactor

@@
 \begin{table*}
-  \small
   \centering
@@
   \begin{groupplot}[
-    group style={group size=2 by 1}
+    group style={group size=2 by 1},
+    legend style={anchor=north, legend columns=1, font=\sffamily\tiny},
   ]
@@
     \commscatter{transformer-3-3_pca1}{transformer-3-3_pca2}
+    \legend{General interest,Videogames,Technology,Sports,Female-focused,Other}
   \end{groupplot}
@@
 \begin{figure}
-  \pairwisecommsim{lstm-3-1} 
+  \pairwisecommsim{snap_cos_sim}{lstm-3-1_cos_sim} 
   \caption{Pairwise cosine similarity between community vectors for 
@@
 Pearson's $r \geq 0.98$ and $p < 0.0001$ (see, for example,
-\cref{fig:lmcc_ppl}).
+\cref{fig:lmcc-ppl}).
 
@@
   }
-  \label{fig:lmcc_ppl}
+  \label{fig:lmcc-ppl}
 \end{figure}
@@
 correlated with community indiscernibility under the corresponding latent 
-LMCC (see \cref{tab:ind-corrs} and, for example, \cref{fig:cclm_lmcc_ppl}).
+LMCC (see \cref{tab:ind-corrs} and, for example, \cref{fig:cclm-lmcc-ppl}).
 In other words, the linguistic complexity of a given community
@@
   }
-  \label{fig:cclm_lmcc_ppl}
+  \label{fig:cclm-lmcc-ppl}
 \end{figure*}
@@
 on messages from that community's test set 
-(see \cref{tab:ind-corrs} and, for example, \cref{fig:cclm_lmcc_ppl}).
+(see \cref{tab:ind-corrs} and, for example, \cref{fig:cclm-lmcc-ppl}).
 Communities for which the language model benefits most from 
@@
 
-We cannot exclude a correlation (\cref{tab:cclm_lmcc_ppl}) between the perplexity of the
+We cannot exclude a correlation (\cref{tab:ind-corrs}) between the perplexity of the
 % \bn{I'm not sure if we can make this claim or not with the Ppl-Ppl correlations.

### 2020-11-23 paper/plot_macros.tex
finish float refactor

@@
+\usepackage{lmodern}
+
 \newcommand{\commscatter}[3][\community]{%
@@
     yticklabels=none,
-    x tick label style={font=\sffamily\tiny,rotate=45,anchor=east},
+    x tick label style={font=\sffamily\tiny,rotate=90,anchor=east},
     xtick=data,
@@
 
-\newcommand{\pairwisecommsim}[1]{%
+\newcommand{\pairwisecommsim}[2]{%
   \begin{tikzpicture}
@@
         },
-        ] table [x=snap, y=#1, meta=meta] {floats/comm_sim.csv};
+        ] table [x=#1, y=#2, meta=category] {floats/comm_comm.csv};
     \end{axis}

### 2020-11-23 analysis.py

@@ -170,18 +170,27 @@ df_confusion.columns = cond_models
 
 ## Linguistic indiscernibility 
 for model in cond_models:
-    df_c[f'{model}_lmcc_ppl'] = C[model].apply(ppl)
+    df_c[f'{model}_indisc'] = C[model].apply(lambda x: ppl(x) / len(comms))
 
 ## Pearson correlation of community-wise LMCC perplexity between pairs of models
 df_mm['lmcc_ppl_corr_r'], df_mm['lmcc_ppl_corr_p'] = zip(*[pearsonr(df_c[f'{m1}_lmcc_ppl'], df_c[f'{m2}_lmcc_ppl']) for m1, m2 in df_mm.index])
 
 
-##### Compare LMCC and CCLM perplexity by community
+##### Correlate indiscernibility (Ind) with info gain (IG) and perplexity (Ppl)
 
-corr = pd.DataFrame([pearsonr(df_c[f'{model}_lmcc_ppl'], df_c[f'{model}_lm_ppl']) 
+## Ppl
+corr = pd.DataFrame([pearsonr(df_c[f'{model}_indisc'], df_c[f'{model}_lm_ppl']) 
     for model in cond_models],
     index=cond_models,
-    columns=['lmcc_cclm_corr_r', 'lmcc_cclm_corr_p'])
+    columns=['ind_ppl_corr_r', 'ind_ppl_corr_p'])
+
+df_m = add_columns(df_m, corr)
+
+## IG
+corr = pd.DataFrame([pearsonr(df_c[f'{model}_info_gain'], df_c[f'{model}_lm_ppl']) 
+    for model in cond_models],
+    index=cond_models,
+    columns=['ind_ig_corr_r', 'ind_ig_corr_p'])
 
 df_m = add_columns(df_m, corr)
 
@@ -232,25 +241,24 @@ df_m[['best_epoch', 'lm_ppl', 'info_gain']].to_latex(os.path.join(floats_dir, 'm
 df_c[
         ['lstm-3_lm_ppl'] + [f'lstm-3-{i}_info_gain' for i in range(4)] +
         ['transformer-3_lm_ppl'] + [f'transformer-3-{i}_info_gain' for i in range(4)] 
-    ].to_latex(os.path.join(floats_dir, 'community_results.tex'),
+    ].sort_values('lstm-3_lm_ppl').to_latex(os.path.join(floats_dir, 'community_results.tex'),
[…]

### 2020-11-23 paper/paper.tex
related work & introduction

@@
 
-\jp{We want to identify a niche that we want to fill. A quick summary of the state of the art + what is missing.}
+Not only is linguistic variation an object of inquiry in its own right,
+but also important to consider in the development of robust, equitable language technology.
+Previous work has shown that by taking demographic information into account, 
+models can achieve better performance on various standard NLP tasks \citep{Hovy2015,Yang2017},
+but little is known about how variation at the level of speech community affects performance.
 
-In the present work, we investigate linguistic variation across online communities.
+There is a fair amount of previous work that investigates linguistic variation using computational methods.
+However, most of this work considers variation across macro-social categories,
+such as gender \citep{Burger2011,Ciot2013,Bamman2014}, age \citep{Nguyen2013}, 
+and geographic location \citep{Eisenstein2010,Bamman2014a}.
+An exception to this is \citet{DelTredici2017},
+who use the modified skip-gram model of \citet{Bamman2014a} to show that lexical semantic
+variation occurs across different communities organized around the same topic.
+
+In the present work, we continue this line of 
+investigating linguistic variation across online communities.
+In contrast to previous work, we investigate variation across speech communities
+(rather than macro-social categories), and use context-sensitive sequence models
+capable of capturing variation beyond the lexical level.
+
 Broadly construed, we want to investigate two questions.  First, we
@@
 but it has not, to our knowledge, been used extensively to study linguistic variation.
-There are, however some related methods that have been used to study variation,
-and related applications of conditional neural language models.
 
-In the later category, conditional neural language models have been
-used to study linguistic change.
+There are, however, related applications of conditional neural language models.
+\citet{Lau2017a} presents a neural language model that jointly learns to predict
+words on the sentence-level and represent topics on the document level.
+The topic representation is then fed back into the language model, 
+improving its performance on next word prediction.
+This is similar to how our model improves experiences improved performance
+by learning community representations. 
+Unlike our model, topics are inferred in an unsupervised way, 
+raising the question of whether communities could be identified from 
+unlabelled data as well.
 
-\paragraph{Language models for sociolinguistic variation}
-\cite{DelTredici2017}
-
-\cite{khan2020reddit}
-
-\paragraph{Pre-trained language models}
-- Unsupervised Domain Clusters in Pretrained Language Models https://www.aclweb.org/anthology/2020.acl-main.692/ % 404?
-
-\paragraph{Topic modelling}
-Modelling communities is not unlike the classic task of \emph{topic
-  modelling}. In topic modelling, one most often want to
-\emph{discover} topics. In constrast, we take as a given that a given
-message belongs to a given community. However, deep-learning topic
-modelling is often built around the idea of a conditioned language
-model \citep{Lau2017a}, with a separate topic-discovery module.
-
-% - Discovering Discrete Latent Topics with Neural Variational Inference https://arxiv.org/pdf/1706.00359.pdf
-
-% \paragraph{Vector representations of online communities encode semantic relationships}
-% https://www.aclweb.org/anthology/W17-2904.pdf
+Along these lines, \citep{OConnor2010} uses a Bayesian generative model to
+infer communities from variation in text data.
+In contrast to our work, this model treats words as
+independent events, ignoring the structure (and variation) in the
+construction of sequences.
+It does further suggest, however, that community-level variation
+could be modeled in an unsupervised way.
 

### 2020-11-23 paper/paper.tex
spell check

@@
 models are correlated with a model of communities based only on
-co-occurence of users between communities and 2. if latent community
+co-occurrence of users between communities and 2. if latent community
 classifiers based on the CCLM can accurately identify that a given
@@
 \item \label{hyp:LMCC-works} CCLM can be used to recognize communities based on
-  latent bayesian classification.
+  latent Bayesian classification.
 % \item \label{hyp:LMCC-works} The layer at which the community embedding is taken into account
@@
   This number of attention heads was chosen in part to give the LSTM and transformer
-  models a comparable number of paramters 
+  models a comparable number of parameters 
   (\num{22171203} and \num{21779523}, respectively).}
@@
     group style={group size=2 by 1},
+    yticklabels={,,}, xticklabels={,,},
     legend style={anchor=north, legend columns=1, font=\sffamily\tiny},
@@
 who generated them using using a negative-sampling optimization
-algorithm and the message-community co-occurence matrix as ground
+algorithm and the message-community co-occurrence matrix as ground
 truth, using data which spans January 2014 to April 2017.
@@
     group style={group size=2 by 1},
-    yticklabels={,,}, xticklabels={,,},
     legend style={anchor=north, legend columns=1, font=\tiny},
@@
 range (0 to 1), we find that CCLMs can be used to recognize certain
-communities based on latent bayesian classification
+communities based on latent Bayesian classification
 \ref{hyp:LMCC-works}, but not all.  If generic communities exhibit
@@
 function of $l_c$: the maximum is reached at $l_c=2$. We did not
-expect non-mononicity, and leave studying this effect to future work.
+expect non-monotonicity, and leave studying this effect to future work.
 
 The pattern of information gain by community is similar across architectures.
-Communities that bennifit most from the conditioned model behave that way for 
+Communities that benefit most from the conditioned model behave that way for 
 both the LSTM and transformer.
@@
 The representation of communities built by CCLM exhibits a clear
-correlation with user-coocurrence patterns
+correlation with user co-occurrence patterns
 \ref{hyp:extra-linguistic-correlation}. A natural
 question that may arise is how much this correlation can be explain by
-user-specifc linguistic patterns. We plan to control for this variable
+user-specific linguistic patterns. We plan to control for this variable
 in future work, for example by building author-conditioned language

### 2020-11-23 paper/paper.tex
intro tweaks

@@
 \maketitle
+\jp{Title suffers from wizwoz system syndrome.
+  http://www.cs.sjtu.edu.cn/~kzhu/writing-a-paper-slides.pdf (slide 10+)
+
+  Try to say what the paper brings to the state of the art rather than how it does it.
+}
 \begin{abstract}
+  Using Community-Conditioned Lanaguage Models (CCLMs), we investigate
+  the following questions: 1. Can the level of linguistic complexity
+  exhibited by different communities is captured. 2. Are CCLMs
+  well-tuned to various communities both by intrinsic (can we build a
+  community classifier based on CCLMs) and extrinsic methods (are
+  CCLMs representations correlated with other network-based
+  representations of communities).
 \end{abstract}
+\jp{Not so good abstract}
 
 \section{Introduction}
+\jp{I am seeing computer modern rather than times when building the
+  paper!?}  Linguistic communication requires that speakers share
+knowledge of certain linguistic conventions, such as syntactic
+structure, word meanings, and accepted patterns of interaction.
+Speakers assume that these conventions are \emph{common ground} among
+their interlocutors, based on joint membership in a community
+\cite{Stalnaker2002, Clark1996}.  Such a community may be of any size,
+from the very small, like members of a particular friend group, to the
+very large, like speakers of English.  Because it is the source of
+linguistic convention, the \emph{speech community} is an important
+concept in the study of linguistic variation \citep{Gumperz1972}.
 
-Linguistic communication requires that speakers share 
-knowledge of certain linguistic conventions, such as syntactic structure, 
-word meanings, and accepted patterns of interaction.
-Speakers assume that these conventions are \emph{common ground} among their interlocutors,
-based on joint membership in a community \cite{Stalnaker2002, Clark1996}.
-Such a community may be of any size,
-from the very small, like members of a particular friend group,
-to the very large, like speakers of English.
-Because it is the source of linguistic convention, the \emph{speech community} 
-is an important concept in the study of linguistic variation \citep{Gumperz1972}.
+Not only is linguistic variation an object of inquiry in its own
+right, but it is also important to consider it in the development of
+robust, equitable language technology.  Previous work has shown that
+by taking demographic information into account, models can achieve
+better performance on various standard NLP tasks
+\citep{Hovy2015,Yang2017}, but little is known about how variation at
+the level of speech community affects performance.
+%
+There is a fair amount of work that investigates linguistic variation
+using computational methods.  However, most of it considers variation
+across macro-social categories, such as gender
+\citep{Burger2011,Ciot2013,Bamman2014}, age \citep{Nguyen2013}, and
+geographic location \citep{Eisenstein2010,Bamman2014a}.
+\citet{DelTredici2017} provide an exception, by using the modified
+skip-gram model of \citet{Bamman2014a} to show that lexical semantic
+variation occurs across different communities organized around the
+same topic.
 
-Not only is linguistic variation an object of inquiry in its own right,
-but also important to consider in the development of robust, equitable language technology.
-Previous work has shown that by taking demographic information into account, 
-models can achieve better performance on various standard NLP tasks \citep{Hovy2015,Yang2017},
-but little is known about how variation at the level of speech community affects performance.
-
-There is a fair amount of previous work that investigates linguistic variation using computational methods.
-However, most of this work considers variation across macro-social categories,
-such as gender \citep{Burger2011,Ciot2013,Bamman2014}, age \citep{Nguyen2013}, 
-and geographic location \citep{Eisenstein2010,Bamman2014a}.
-An exception to this is \citet{DelTredici2017},
-who use the modified skip-gram model of \citet{Bamman2014a} to show that lexical semantic
-variation occurs across different communities organized around the same topic.
-
-In the present work, we continue this line of 
-investigating linguistic variation across online communities.
-In contrast to previous work, we investigate variation across speech communities
-(rather than macro-social categories), and use context-sensitive sequence models
+In the present work, we further investigate linguistic variation
+across online communities. In contrast to previous work, we
+investigate variation across speech communities (rather than
+macro-social categories), and use context-sensitive sequence models
 capable of capturing variation beyond the lexical level.
-
-Broadly construed, we want to investigate two questions.  First, we
-are interested in variation in the level of linguistic complexity
-exhibited by different communities.
-
+%
+We do so by investigate two questions.  First, we are interested in
+variation in the level of linguistic complexity exhibited by different
+communities.
 \jp{
@@
   of time/money and a big deal made out of nothing.  }
-
-It is folklore that certain communities exhibit more linguistic complexity than others.
-Here we want to quantify this phenomenon.
-More precisely, we investigate if the degree of complexity can be
-captured in a computational model.  For this purpose define various
-Community Conditioned Language Models (CCLM for short), which can be
-modeled as an estimator $M$ of the degree to which a given message $m$ 
-is acceptable in a given community $c$: \(P_M(m \mid c)\).
+%
+Indeed, it is folklore that certain communities exhibit more
+linguistic complexity than others.  Here we want to quantify this
+phenomenon.  More precisely, we investigate if the degree of
+complexity can be captured in a computational model.  For this purpose
+define various Community Conditioned Language Models (CCLM for short),
+which can be modeled as an estimator $M$ of the degree to which a
+given message $m$ is acceptable in a given community $c$:
+\(P_M(m \mid c)\).
 
@@
 recognize: the less linguistic diversity, the more specific the
-language become and the easier it is to identify.
+language becomes and the easier it is to identify.
 
-To sum up we aim at testing the following hypotheses:
+To sum up, we test the following hypotheses:
 \begin{hypotheses}

### 2020-11-23 paper/paper.tex
various minor revisions

@@
 but excluded four of them from further analysis: 
-two because they were primarily non-English and two with particularly short average message lengths.\footnote{
+two because they were primarily non-English and two with particularly short average message lengths.\\footnote{
 The macro-average message length was \num{39.3} ($\sigma=15.7$) and the two removed communities had average comment lengths of \num{10.7} and \num{18.1}.}
+% According to your footnote the removed communities are barely outliers. I wouldn't have removed them, but it is too late to do anything about that.
 Each community corpus consist of \num{50000} randomly selected messages from the year 2015.
@@
 
-We experiment with two kinds of model architecture: a simple unidirectional LSTM
-\citep{Hochreiter1997} and a Transformer \citep{Vaswani2017}.  In either
-case, the model is organised as a standard $n$-layer neural sequence
-encoder (we set $n=3$ for all models). As is standard, word tokens are
-first converted to vectors using a trainable embedding layer. At the
-other end, word tokens are predicted using a softmax projection layer. The set of word tokens is a
-pre-determined vocabulary.  These encoders form the core of the
-community-conditioned language models (CCLM), and we call them
+We experiment with two kinds of model architecture: a simple
+unidirectional LSTM \citep{Hochreiter1997} and a Transformer
+\citep{Vaswani2017}.  In either case, the model is organised as a
+standard $n$-layer neural sequence encoder (we set $n=3$ for all
+models). As conventional, word tokens are first converted to vectors
+using a trainable embedding layer. At the other end, word tokens are
+predicted using a softmax projection layer. The set of word tokens is
+a pre-determined vocabulary.  These encoders form the core of the
+of our models, as such do not take community into account, and we call them
 unconditioned models.
 
-To condition the model, we add a \emph{community embedding} parameter to the models.
-This parameter is concatenated with the hidden layer of the sequence encoder,
-for some layer $l_c \leq n$, and passed through a linear layer 
-which projects the resulting vector back to the original hidden layer size.
-For $l_c = n$, the output of this linear layer is passed directly to the softmax function,
-just as the final hidden layer of the sequence encoder is in other models.
-For $l_c=0$, the community embedding is concatenated with the token embedding.
-For this reason, we set the hidden size of the sequence encoder 
-and the size of the token to be embedding equal for all models.
+To condition the model (and obtain community-conditioned language
+models --- CCLM), we add a \emph{community embedding} parameter to the
+models.  This parameter is concatenated with the hidden layer of the
+sequence encoder, for some layer $l_c \leq n$, and passed through a
+linear layer which projects the resulting vector back to the original
+hidden layer size.  For $l_c = n$, the output of this linear layer is
+passed directly to the softmax function, just as the final hidden
+layer of the sequence encoder is in other models.  For $l_c=0$, the
+community embedding is concatenated with the token embedding.  For
+this reason, we set the hidden size of the sequence encoder and the
+size of the token to be embedding equal for all models.
 
@@
 
-In the remainder of the paper we use $M$ to refer to messages in the test set,
-and $M_j$ for the partition of the test set originating from community $c_j$.
+In the remainder of the paper we use $M$ to refer to messages in the
+test set, and $M_j$ for the partition of the test set originating from
+community $c_j$.
 
-We use a vocabulary size of \num{40000} tokens, including a special out-of-vocabulary token.
-The vocabulary consists of the most frequent tokens at the message level (across all communities),
-meaning that we only count each token once per message.
-The main reason for this choice is to avoid over-counting otherwise rare tokens that were spammed
-many times in a single message.
+We use a vocabulary size of \num{40000} tokens, including a special
+out-of-vocabulary token.  The vocabulary consists of the most frequent
+tokens at the message level (across all communities), meaning that we
+only count each token once per message.  The main reason for this
+choice is to avoid over-counting otherwise rare tokens that were
+repeated many times in a single message.
 
-For the LSTM, we trained the models on a simple left-to-right language modeling task
-with cross-entropy loss over the vocabulary. 
-Because the transformer operates on all tokens in the sequence at once,
-the language modeling was achieved by masking and incrementally un-masking input tokens
-in a left-to-right fashion (so we have an auto-regressive\jp{check} model).
-We used the AdamW \citep{Loshchilov2019} optimization algorithm,
-with an initial learning rate of \num{0.001}, without any extra control on the decay of learning rate.
-We used a batch size of \num{256} and a maximum sequence length of \num{64} tokens, 
-truncating longer messages (16.8\% of messages were longer than \num{64} tokens).
-During training, a dropout rate of $0.1$ was applied between encoder layers 
-and after each linear layer.
+For the LSTM, we trained the models on a simple left-to-right language
+modeling task with cross-entropy loss over the vocabulary.  Because
+the transformer operates on all tokens in the sequence at once, the
+language modeling was achieved by masking and incrementally un-masking
+input tokens in a left-to-right fashion (so we have an
+auto-regressive\jp{check} model).  We used the AdamW
+\citep{Loshchilov2019} optimization algorithm, with an initial
+learning rate of \num{0.001}, without any extra control on the decay
+of learning rate.
+%
+We used a batch size of \num{256} and a maximum sequence length of
+\num{64} tokens, truncating longer messages (16.8\% of messages were
+longer than \num{64} tokens).  During training, a dropout rate of
+$0.1$ was applied between encoder layers and after each linear layer.
 
@@
 
-We trained the models until the validation loss stopped decreasing 
-for two epochs in a row,
-and used the weights from the epoch with the smallest validation loss for testing.
-In general, transformer models trained for about half as long as the LSTM models (test epoch, \cref{tab:model-results}).
+We trained the models until the validation loss stopped decreasing for
+two epochs in a row, and used the weights from the epoch with the
+smallest validation loss for testing.  In general, transformer models
+trained for about half as long as the LSTM models (test epoch,
+\cref{tab:model-results}).
 
@@
 
-We measure the perplexity of each example in the test set. 
+We measure the perplexity of each example\jp{message? community?} in the test set. 
 For this, we use the negative exponential of the mean cross-entropy loss 
 of the language model for a given message and average over a set of messages, $M$.
-\[\Ppl_M = \avg_{m\in M}(-e^{H(m)})\] \bn{Following the logic from IG, should we change Ppl to the exponential of the average cross entropy, rather than using the average of the message-level perplexity?}
+\[\Ppl_M = \avg_{m\in M}(-e^{H(m)})\]
+\bn{Following the logic from IG, should we change Ppl to the exponential of the average cross entropy, rather than using the average of the message-level perplexity?}
+\jp{Yes}
 As one would expect, 
@@
 
-We also consider model performance stratified by community;
-that is, $\Ppl_{M_j}$, where $M_j$ is the test-set messages from community $c_j$.
-We observe a great deal of variation in baseline perplexity across communities,
-with $\Ppl_{M_j}$ ranging from \num{53.07} to \num{115.29} for the LSTM model
-(\cref{tab:community-results}).
+We also consider model performance stratified by community; that is,
+$\Ppl_{M_j}$, where $M_j$ is the test-set messages from community
+$c_j$.  We observe a great deal of variation in baseline perplexity
+across communities, with $\Ppl_{M_j}$ ranging from \num{53.07} to
+\num{115.29} for the LSTM model (\cref{tab:community-results}).
 
@@
 
-As an informal observation, we note 
-that across all the models we tested, 
-communities where conditioning has the least effect
-tend to be more organized around more general interest topics,
-such as the \emph{relationships} and \emph{advice} subreddits, where the subject matter is relevant to a broad
-range of people.
-Conditioning the model on community has the most benefit
-for narrower special-interest subreddits, such as those organized
-around a certain videogame, sports team, or subculture.
-This makes sense intuitively, since communities with more niche topics
-would tend to have more specialized vocabulary and other linguistic patterns.
+As an informal observation, we note that across all the models we
+tested, communities where conditioning has the least effect tend to be
+more organized around more general interest topics, such as the
+\emph{relationships} and \emph{advice} subreddits, where the subject
+matter is relevant to a broad range of people.  Conditioning the model
+on community has the most benefit for narrower special-interest
+subreddits, such as those organized around a certain videogame, sports
+team, or subculture.  This makes sense intuitively, since communities
+with more niche topics would tend to have more specialized vocabulary
+and other linguistic patterns.
 
@@
 community is. However, the CCLM induces a latent classifier,
-which we call a language model-based community classifier
+which we call a language-model-based community classifier
 (LMCC). Using the Bayes theorem, we can turn a
@@
 \[C_{ij} = \avg_{m\in M_j}P(c=c_i \mid m)\]
-As a sanity check, we observe that the diagonal dominates the matrix: on average the LMCC picks the right community for a message (\cref{fig:confusion}).
-
+We observe that the diagonal dominates the matrix: the LMCC generally finds the right community for a message (\cref{fig:confusion}).
+%
 \begin{figure*}
@@
 \end{figure*}
-
-However, we want to measure the ability of the LMCC to correctly
+%
+But we want a more quantitative measure, corresponding to the ability of the LMCC to correctly
 recognize messages coming from a given community $c_j$. We could use
@@
 since it is computed as the average of $P(c=c_i \mid m)$ for $m\in M_j$.
-The exponential of the entropy of the $j$th row of \(C\) is therefore a perplexity score.
-To compute $\Ind_j$, we scale this perplexity by the number of communities, 
-which makes $\Ind_j$ comparable across models trained on different
-numbers of communities.\footnote{Although we only consider one set of \num{46} communities in this paper.}
+The exponential of the entropy of the $j$th row of \(C\) is therefore a perplexity score,
+whose maximum is the total number of communities (46 for us).
+Thus we define $\Ind_j$ by scaling this perplexity by the number of communities:
 \[\Ind_j = \frac{e^{-\sum_i C_{ij} log(C_{ij})}}{|C_j|}\]
-Intuitively, $\Ind_j$ is a measure of the uncertainty of the LMCC,
-given messages from a community $c_j$.
-It ranges from $0$ (for a completely deterministic $C_j$) 
-to $1$ (if $C_j$ gives equal weight to every community).
+This way $Ind_j=0$ means that the LMCC is completely certain of its
+choice, while $Ind_j=1$ means that equal probability is given to all
+predictions.  In general $\Ind_j$ is a score of uncertainty of the
+LMCC, given messages from a community $c_j$.
 
@@
 In other words, the linguistic complexity of a given community
-is a poor predictor of how distinct its language is.
+is a poor predictor of how distinctive its language is.
 
@@
   \input{floats/ind_corrs.tex}
-  \caption{
-    Pearson's $r$ correlation coefficient between community indiscernibility ($\Ind_j$) and
-    two different predictors: CCLM perplexity on messages from community $c_j$ ($\Ppl_{M_j}$)
-    , and CCLM information gain on messages from $c_j$ ($\IG_{M_j}$).
-  }
+  \caption{ Pearson's $r$ correlation coefficient between community
+    indiscernibility ($\Ind_j$) and two different predictors: CCLM
+    perplexity on messages from community $c_j$ ($\Ppl_{M_j}$), and
+    CCLM information gain on messages from $c_j$ ($\IG_{M_j}$).  }
   \label{tab:ind-corrs}
@@
 improving its performance on next word prediction.
-This is similar to how our model improves experiences improved performance
+This is similar to how our model experiences improved performance
 by learning community representations. 
@@
 
-Along these lines, \citep{OConnor2010} uses a Bayesian generative model to
-infer communities from variation in text data.
-In contrast to our work, this model treats words as
-independent events, ignoring the structure (and variation) in the
-construction of sequences.
-It does further suggest, however, that community-level variation
-could be modeled in an unsupervised way.
+Along these lines, \citet{OConnor2010} uses a Bayesian generative
+model to infer communities from variation in text data.  In contrast
+to our work, this model treats words as independent events, ignoring
+the structure (and variation) in the construction of sequences.  It
+does further suggest, however, that community-level variation could be
+modeled in an unsupervised way.
 
@@
 
-We cannot exclude a correlation (\cref{tab:ind-corrs}) between the perplexity of the
+We cannot exclude a correlation (\cref{tab:ind-corrs}) between the
+perplexity of the
 % \bn{I'm not sure if we can make this claim or not with the Ppl-Ppl correlations.
@@
 CCLM and the indiscernibility of the corresponding community
-\ref{hyp:rich-harder-to-identify}. Because we have relatively few,
-and only large communities in our test set, it is difficult to
-provide a conclusive answer, but this kind of correlation would suggest that
+\ref{hyp:rich-harder-to-identify}. Because we have relatively few, and
+only large communities in our test set, it is difficult to provide a
+conclusive answer, but this kind of correlation would suggest that
 community-specific complexity can be partially attributed to using
-more varied patterns. However, even the most complex communities
-(perplexity per word above 100) have an indiscernibility of about 3
-(out of a maximum of 46)\jp{make consistent with data once the plot is stablilized}, so they retain fairly specific patterns:
-they are not simply an even mixture of the kind of comments found in
-other communities.
+more varied patterns. With an indiscernibility $> 0.9$, the most
+complex communities appear as as even mixture of the kind of language
+found in all communities, to our models.
 
@@
 albeit weakly. For LSTM models, the perplexity per word, averaged over
-all communities, varies between 76.80 and 78.64. (with 80.70 for the unconditioned model)
-The perplexity per word is slightly higher for $l_c=0$, and has very
-little difference between the other models, with $l_c=1$ being the best.
-For transformer models, it varies
-between 89.38 and 122.56. Here perplexity per-word is clearly not a monotonous
+all communities, varies between 76.80 and 78.64. (with 80.70 for the
+unconditioned model) The perplexity per word is slightly higher for
+$l_c=0$, and has very little difference between the other models, with
+$l_c=1$ being the best.  For transformer models, it varies between
+89.38 and 122.56. Here perplexity per-word is clearly not a monotonous
 function of $l_c$: the maximum is reached at $l_c=2$. We did not
-expect non-monotonicity, and leave studying this effect to future work.
+expect non-monotonicity, and leave studying this effect to future
+work.
 
-The pattern of information gain by community is similar across architectures.
-Communities that benefit most from the conditioned model behave that way for 
-both the LSTM and transformer.
-However, there are some differences.
-For example, the subreddit \emph{relationships} has an information gain of \num{1.018} for
-the best CCLM LSTM but only \num{0.998} for the best transformer model.
-Why was the LSTM able to take advantage of the community information for
-this community, where the transformer wasn't?
-It would be interesting to investigate these differences in future work,
-since it could reveal differences in the kind of linguistic variation the different
-model architectures capture.
+The pattern of information gain by community is similar across
+architectures.  Communities that benefit most from the conditioned
+model behave that way for both the LSTM and transformer.  However,
+there are some differences.  For example, the subreddit
+\emph{relationships} has an information gain of \num{1.018} for the
+best CCLM LSTM but only \num{0.998} for the best transformer model.
+Why was the LSTM able to take advantage of the community information
+for this community, where the transformer was not?  It would be
+interesting to investigate these differences in future work, since it
+could reveal differences in the kind of linguistic variation the
+different model architectures capture.
 

### 2020-11-23 paper/paper.tex
typography

@@
   \input{floats/comm_sim.tex}
-  \caption{Pearson's r between the pairwise similarity of community vectors
+  \caption{Pearson's $r$ between the pairwise similarity of community vectors
   in the CCLM models and the membership-based embedding of \citet{Kumar2018}.

### 2020-11-23 paper/paper.tex
typography
spellcheck

@@
 \maketitle
-\jp{Title suffers from wizwoz system syndrome.
-  http://www.cs.sjtu.edu.cn/~kzhu/writing-a-paper-slides.pdf (slide 10+)
-
-  Try to say what the paper brings to the state of the art rather than how it does it.
-}
 \begin{abstract}
-  Using Community-Conditioned Lanaguage Models (CCLMs), we investigate
+  Using Community-Conditioned Language Models (CCLMs), we investigate
   the following questions: 1. Can the level of linguistic complexity
@@
 \item \label{hyp:extra-linguistic-correlation} The CCLM representation of communities are correlated with
-  co-ocurrence of users between communities.
+  co-occurrence of users between communities.
 % \item Community stability is correlated with language complexity
@@
 unidirectional LSTM \citep{Hochreiter1997} and a Transformer
-\citep{Vaswani2017}.  In either case, the model is organised as a
+\citep{Vaswani2017}.  In either case, the model is organized as a
 standard $n$-layer neural sequence encoder (we set $n=3$ for all
@@
   \caption{
-    Model performance by community. We show the absolute perplexity 
+    Model perplexity and information gain by community. We show the absolute perplexity 
     of the baseline models and the information gain of the CCLMs.
-    Note that the rows are sorted by the perplexity of the baseline
+    Rows are sorted by the perplexity of the baseline
     LSTM model.
@@
 
-We can characterise the quality the LMCC with a confusion matrix
+We can characterize the quality the LMCC with a confusion matrix
 $C$, defined by $C_{ij} = P(c=c_i \mid m)$ for a given (test set) message $m$ 
@@
 raising the question of whether communities could be identified from 
-unlabelled data as well.
+unlabeled data as well.
 
@@
 more varied patterns. With an indiscernibility $> 0.9$, the most
-complex communities appear as as even mixture of the kind of language
+complex communities appear as an even mixture of the kind of language
 found in all communities, to our models.
@@
 \end{document}
+
+% LocalWords:  CCLMs Gumperz NLP Hovy Ciot Bamman DelTredici CFB SMU
+% LocalWords:  cancelled pre emptively CCLM Reddit subreddits dataset
+% LocalWords:  preprocessed tokenized SpaCy tokenizer Honnibal LSTM
+% LocalWords:  Hochreiter Vaswani softmax un AdamW Loshchilov IG Ppl
+% LocalWords:  LM videogame embeddings PCA Kumar lstm LMCC ij th Lau
+% LocalWords:  indiscernibility Vinyals Kalchbrenner OConnor
+% LocalWords:  monotonicity subreddit

### 2020-11-23 paper/paper.tex
tweak

@@
 linguistic complexity \ref{hyp:varying-complexity}, with perplexity
-per-word ranging from less than 53 to 104 for the best model
+per-word ranging from less than 53 to more than 104 for the best model
 architecture.
@@
 \ref{hyp:LMCC-works}, but not all.  If generic communities exhibit
-distinctive linguistic patterns, none of our model is able to discern
+distinctive linguistic patterns, none of our models is able to discern
 them. In contrast, the distinctive patterns of narrow-topic
@@
 
-We cannot exclude a correlation (\cref{tab:ind-corrs}) between the
+We find no statistically significant correlation (\cref{tab:ind-corrs}) between the
 perplexity of the
@@
 \ref{hyp:rich-harder-to-identify}. Because we have relatively few, and
-only large communities in our test set, it is difficult to provide a
-conclusive answer, but this kind of correlation would suggest that
-community-specific complexity can be partially attributed to using
-more varied patterns. With an indiscernibility $> 0.9$, the most
-complex communities appear as an even mixture of the kind of language
-found in all communities, to our models.
+only large communities in our test set, it is difficult to say if this
+finding generalizes. Regardless, this kind of correlation would suggest that
+community-specific complexity can be at least partially attributed to
+using more varied patterns. This is view is supported by the
+observation that, with an indiscernibility $> 0.9$, the most complex
+communities appear as an even mixture of the kind of language found in
+all communities, to our models.
 

### 2020-11-23 paper/paper.tex
formula for ppl

@@
 of the language model for a given message and average over a set of messages, $M$.
-\[\Ppl_M = \avg_{m\in M}(-e^{H(m)})\]
+\[\Ppl_M = \avg_{m\in M}(e^{-H(m)})\]
 \bn{Following the logic from IG, should we change Ppl to the exponential of the average cross entropy, rather than using the average of the message-level perplexity?}
 \jp{Yes}
+\[\Ppl_M = e^{\avg_{m\in M} H(m)}\]
 As one would expect, 

### 2020-11-23 analysis.py
ppl = e ^ (avg H), not avg(e^H)

@@ -93,15 +93,17 @@ df_m['best_epoch'] = best_epoch
 ## Model perplexity on test examples
 lm_ppl = pd.read_pickle(os.path.join(model_dir, 'test_ppl.pickle'))
 
+## Model entropy on test examples
+lm_entr = lm_ppl[models].apply(np.log)
+lm_entr['community'] = lm_ppl['community']
+
 ## Mean language model perplexity
-df_m['lm_ppl'] = lm_ppl[models].mean()
+df_m['lm_ppl'] = lm_entr[models].mean().apply(np.exp) 
 
 ## Mean language model perplexity by community
-df_c = add_columns(df_c, lm_ppl.groupby('community').mean(), suffix='_lm_ppl')
-
+df_c = add_columns(df_c, lm_entr.groupby('community').mean().apply(np.exp), suffix='_lm_ppl')
 
 #### Community embeddings
-
 comm_w = {model: np.load(os.path.join(model_dir, f'{model}/comm_embed.npy'))[1:] 
         for model in cond_models}
 
@@ -172,8 +174,34 @@ df_confusion.columns = cond_models
 for model in cond_models:
     df_c[f'{model}_indisc'] = C[model].apply(lambda x: ppl(x) / len(comms))
 
-## Pearson correlation of community-wise LMCC perplexity between pairs of models
-df_mm['lmcc_ppl_corr_r'], df_mm['lmcc_ppl_corr_p'] = zip(*[pearsonr(df_c[f'{m1}_lmcc_ppl'], df_c[f'{m2}_lmcc_ppl']) for m1, m2 in df_mm.index])
+## Pearson correlation of linguistic indiscernibility across models 
+df_mm['indisc_corr_r'], df_mm['indisc_corr_p'] = zip(*[pearsonr(df_c[f'{m1}_indisc'], df_c[f'{m2}_indisc']) for m1, m2 in df_mm.index])
+
+
+##### Information gain (AKA mutual information)
+
+## Model entropy on test examples
+
+lstm_baseline_ppl = np.exp(lm_entr['lstm-3'].mean())
+lstm_cond_ppl = lm_entr[cond_lstms].mean().apply(np.exp) 
+lstm_info_gain = lstm_cond_ppl.apply(lambda x: lstm_baseline_ppl / x)
+transformer_baseline_ppl = np.exp(lm_entr['transformer-3'].mean())
[…]

### 2020-11-24 paper/paper.tex
ppl = e ^ (avg H), not avg(e^H)
move big table to appendix

@@
 \usepgfplotslibrary{groupplots}
+\usepgfplotslibrary{statistics}
 \pgfplotsset{compat=1.16}
+\pgfplotsset{every tick label/.append style={font=\footnotesize}}
 
@@
 %
-We do so by investigate two questions.  First, we are interested in
+We do so by investigating two questions.  First, we are interested in
 variation in the level of linguistic complexity exhibited by different
@@
 
-We measure the perplexity of each example\jp{message? community?} in the test set. 
+We measure the perplexity of each message in the test set. 
 For this, we use the negative exponential of the mean cross-entropy loss 
 of the language model for a given message and average over a set of messages, $M$.
-\[\Ppl_M = \avg_{m\in M}(e^{-H(m)})\]
-\bn{Following the logic from IG, should we change Ppl to the exponential of the average cross entropy, rather than using the average of the message-level perplexity?}
-\jp{Yes}
 \[\Ppl_M = e^{\avg_{m\in M} H(m)}\]
@@
 \begin{table}
-  \footnotesize
+  \small
   \centering
@@
 $\Ppl_{M_j}$, where $M_j$ is the test-set messages from community
-$c_j$.  We observe a great deal of variation in baseline perplexity
-across communities, with $\Ppl_{M_j}$ ranging from \num{53.07} to
-\num{115.29} for the LSTM model (\cref{tab:community-results}).
-
-\subsection{Information gain of the CCLM}
-
-Next, we consider the \emph{information gain} of each CCLM over its baseline
-un-conditioned counterpart with the same sequence encoder architecture.
-For a given message, information gain is defined as the difference
-between the cross-entropy of the unconditioned model and the conditioned model:
-\[H_{LM}(m) - H_{CCLM}(m)\]
-For a set of messages, $M$, we consider the average information gain
-in exponential space (as a ratio of perplexities):
-\[\IG_M = \frac{e^{\avg_M(H_{LM}(m))}}{e^{\avg_M(H_{CCLM}(m))}}\]
-%\jp{What we had before: average entropy gain is $H(m)-H_0(m)$, and
-  %then taking the average over all messages:
-  %$avg(H(m)-H_0(m)) = avg(H(m))-avg(H_0(m)) = H - H₀$. Taking the exponential
-  %gives a ratio of perplexity.  Here it seems that you have commuted
-  %exponential and average. It is rather weird to do average in the
-  %exponential space without normalising to a probability.
-
-  %Compare: $e^{avg(H(m))} / e^{avg(H_0(m))}$ with $avg \frac{e^{H(m)}}{e^{H₀(m)}}$.
-%}
-
-While the absolute performance of the LSTM models is better,
-we find that the best transformer CCLM has higher information gain 
-than the best LSTM (\cref{tab:model-results}).
-
-The conditioned models also perform differently across different communities---%
-for some communities, $\IG_{M_j}$ around \num{1}, or even slightly lower,
-meaning that the CCLM has no advantage over the unconditioned baseline for that community.
-For others communities it is higher, meaning that the CCLM performs better than 
-the baseline for those communities (\cref{tab:community-results}).
-
-\begin{table*}
-  \centering
-  \input{floats/community_results.tex}
-  \caption{
-    Model perplexity and information gain by community. We show the absolute perplexity 
-    of the baseline models and the information gain of the CCLMs.
-    Rows are sorted by the perplexity of the baseline
-    LSTM model.
-}
-  \label{tab:community-results}
-\end{table*}
-
-As an informal observation, we note that across all the models we
-tested, communities where conditioning has the least effect tend to be
-more organized around more general interest topics, such as the
-\emph{relationships} and \emph{advice} subreddits, where the subject
-matter is relevant to a broad range of people.  Conditioning the model
-on community has the most benefit for narrower special-interest
-subreddits, such as those organized around a certain videogame, sports
-team, or subculture.  This makes sense intuitively, since communities
-with more niche topics would tend to have more specialized vocabulary
-and other linguistic patterns.
+$c_j$.  We observe a lot of variation in baseline perplexity
+across communities, with $\Ppl_{M_j}$ ranging from \num{34.47} to
+\num{66.43} for the LSTM model (see table 1 in the supplementary materials).
 
@@
 
-\begin{figure*}
+\begin{figure*}[t]
 \begin{tikzpicture}
@@
 \begin{table}
+  \small
   \centering
@@
 \end{table}
+
+\subsection{Information gain of the CCLM}
+
+Next, we consider the \emph{information gain} of each CCLM over its baseline
+un-conditioned counterpart with the same sequence encoder architecture.
+For a given message, information gain is defined as the difference
+between the cross-entropy of the unconditioned model and the conditioned model:
+\[H_{LM}(m) - H_{CCLM}(m)\]
+For a set of messages, $M$, we consider the average information gain
+in exponential space (as a ratio of perplexities):
+\[\IG_M = \frac{e^{\avg_M(H_{LM}(m))}}{e^{\avg_M(H_{CCLM}(m))}}\]
+%\jp{What we had before: average entropy gain is $H(m)-H_0(m)$, and
+  %then taking the average over all messages:
+  %$avg(H(m)-H_0(m)) = avg(H(m))-avg(H_0(m)) = H - H₀$. Taking the exponential
+  %gives a ratio of perplexity.  Here it seems that you have commuted
+  %exponential and average. It is rather weird to do average in the
+  %exponential space without normalising to a probability.
+
+  %Compare: $e^{avg(H(m))} / e^{avg(H_0(m))}$ with $avg \frac{e^{H(m)}}{e^{H₀(m)}}$.
+%}
+
+While the absolute performance of the LSTM models is better,
+we find that the best transformer CCLM has higher information gain 
+than the best LSTM (\cref{tab:model-results}).
+
+The conditioned models also perform differently across different communities---%
+for some communities, $\IG_{M_j}$ around \num{1}, or even slightly lower,
+meaning that the CCLM has no advantage over the unconditioned baseline for that community.
+For other communities it is higher, meaning that the CCLM performs better than 
+the baseline for those communities (\cref{fig:comm-info-gain}).
+
+\begin{figure}
+\begin{tikzpicture}
+  \infogainboxplot
+\end{tikzpicture}
+\caption{
+  Distribution of information gain across communities for each CCLM 
+  (top: LSTM models, bottom: transformer models).
+}
+\label{fig:comm-info-gain}
+\end{figure}
+
+This variation in information gain across communities is not 
+related to perplexity, however. 
+Some of the most linguistically complex communities benifit a lot
+from community conditioning, while others do not benefit at all. 
+Similarly, among the less linguistically complex communites 
+there some have high information gain, and some have information
+gain of \num{1} or less (\cref{fig:ppl-info-gain}).
+
+\begin{figure}
+\begin{tikzpicture}
+\begin{axis}
+\commscatter[]{lstm-3-1_lm_ppl}{lstm-3-1_info_gain}
+\end{axis}
+\end{tikzpicture}
+\caption{
+  Average community perplexity (x-axis) versus information gain (y-axis)
+  for the LSTM model with $l_c = 1$ .
+}
+\label{fig:ppl-info-gain}
+\end{figure}
+
+As an informal observation, we note that across all the models we
+tested, communities where conditioning has the least effect tend to be
+more organized around more general interest topics, such as the
+\emph{relationships} and \emph{advice} subreddits, where the subject
+matter is relevant to a broad range of people.  Conditioning the model
+on community has the most benefit for narrower special-interest
+subreddits, such as those organized around a certain videogame, sports
+team, or subculture.  This makes sense intuitively, since communities
+with more niche topics would tend to have more specialized vocabulary
+and other linguistic patterns.
 
@@
 Pearson's $r \geq 0.98$ and $p < 0.0001$ (see, for example,
-\cref{fig:lmcc-ppl}).
+\cref{fig:indisc-corr}).
 
@@
   }
-  \label{fig:lmcc-ppl}
+  \label{fig:indisc-corr}
 \end{figure}
@@
   \caption{%
-    Mean community indiscernibility (y-axis), 
+    Average community indiscernibility (y-axis), 
     in relationship to linguistic complexity (left) and 
@@
 the structure (and variation) in the construction of sequences.  It
-does further suggest, however, that community-level variation could be
+does further suggest, however, that community-level variation can be
 modeled in an unsupervised way.
@@
 linguistic complexity \ref{hyp:varying-complexity}, with perplexity
-per-word ranging from less than 53 to more than 104 for the best model
+per-word ranging from less than 53 to more than 104 for the best model \bn{update}
 architecture.
@@
 perplexity of the
-% \bn{I'm not sure if we can make this claim or not with the Ppl-Ppl correlations.
-However, given that *all* of them have positive correlation, even though the p-values are somewhat higher than .1 maybe we can?}
 CCLM and the indiscernibility of the corresponding community
@@
 $l_c=0$, and has very little difference between the other models, with
-$l_c=1$ being the best.  For transformer models, it varies between
+$l_c=1$ being the best.  For transformer models, it varies between \bn{update}
 89.38 and 122.56. Here perplexity per-word is clearly not a monotonous

### 2020-11-24 paper/plot_macros.tex
move big table to appendix

@@
   every node near coord/.append style={font=\sffamily\tiny},
-  visualization depends on={value \thisrow{community} \as \community}
+  visualization depends on={value \thisrow{community} \as \community},
   ] table [x={#2}, y={#3}, meta=category] {floats/comm.csv};
@@
 }
+
+\newcommand{\infogainboxplot}{
+  \begin{groupplot}[
+    group style={
+      group size=1 by 2,
+      every plot/.style={
+        xmin=0.65,
+        xmax=1.25,
+      },
+    },
+    height=4cm, width=8cm,
+    ytick={1,2,3,4,5,6,7,8},
+    yticklabels={0, 1, 2, 3, 0, 1, 2, 3},
+    ylabel={$l_c$},
+    ]
+  \nextgroupplot
+  \addplot+[boxplot,draw=blue,mark=*,solid]
+  table [y=lstm-3-0_info_gain] {floats/comm.csv};
+  \addplot+[boxplot,draw=blue,mark=*,solid]
+  table [y=lstm-3-1_info_gain] {floats/comm.csv};
+  \addplot+[boxplot,draw=blue,mark=*,solid]
+  table [y=lstm-3-2_info_gain] {floats/comm.csv};
+  \addplot+[boxplot,draw=blue,mark=*,solid]
+  table [y=lstm-3-3_info_gain] {floats/comm.csv};
+  \nextgroupplot
+  \addplot+[boxplot,draw=red,mark=*,solid]
+  table [y=transformer-3-0_info_gain] {floats/comm.csv};
+  \addplot+[boxplot,draw=red,mark=*,solid]
+  table [y=transformer-3-1_info_gain] {floats/comm.csv};
+  \addplot+[boxplot,draw=red,mark=*,solid]
+  table [y=transformer-3-2_info_gain] {floats/comm.csv};
+  \addplot+[boxplot,draw=red,mark=*,solid]
+  table [y=transformer-3-3_info_gain] {floats/comm.csv};
+  \end{groupplot}
+}

### 2020-11-24 paper/paper.tex
update ppl numbers in conclusion

@@
 \section{Introduction}
-\jp{I am seeing computer modern rather than times when building the
-  paper!?}  Linguistic communication requires that speakers share
+Linguistic communication requires that speakers share
 knowledge of certain linguistic conventions, such as syntactic
@@
 communities.
-\jp{
-  Some examples (perhaps include if there is space):
+%\jp{
+  %Some examples (perhaps include if there is space):
 
-  (CFB)
-  It was against SMU in the 47-48 Cotton bowl. They had cancelled a
-  game the year before against Miami for the same reason. There's
-  debate on if that's really the origin for the phrase but the story
-  itself is true.
+  %(CFB)
+  %It was against SMU in the 47-48 Cotton bowl. They had cancelled a
+  %game the year before against Miami for the same reason. There's
+  %debate on if that's really the origin for the phrase but the story
+  %itself is true.
 
-  (Videos)
-  Reminds me of Y2K computer bug. A lot of work went into
-  pre-emptively fixing it, and a lot of bugs were found and fixed,
-  some of which would have been catastrophic. Unfortunately, because
-  the work was successful, a lot of people now describe it as a waste
-  of time/money and a big deal made out of nothing.  }
+  %(Videos)
+  %Reminds me of Y2K computer bug. A lot of work went into
+  %pre-emptively fixing it, and a lot of bugs were found and fixed,
+  %some of which would have been catastrophic. Unfortunately, because
+  %the work was successful, a lot of people now describe it as a waste
+  %of time/money and a big deal made out of nothing.  }
 %
@@
 but excluded four of them from further analysis: 
-two because they were primarily non-English and two with particularly short average message lengths.\\footnote{
+two because they were primarily non-English and two with particularly short average message lengths.\footnote{
 The macro-average message length was \num{39.3} ($\sigma=15.7$) and the two removed communities had average comment lengths of \num{10.7} and \num{18.1}.}
@@
 input tokens in a left-to-right fashion (so we have an
-auto-regressive\jp{check} model).  We used the AdamW
+auto-regressive model).  We used the AdamW
 \citep{Loshchilov2019} optimization algorithm, with an initial
@@
 for models with different community embedding depths.
-For the LSTM encoder, 
-the best model ($l_c=1$) concatenates the community embedding after the first encoder layer,
-with two subsequent LSTM layers. 
-The best transformer model ($l_c=3$), on the other hand,
+The best transformer model ($l_c=3$)
 incorporates the community information last,
 after all the sequence encoder layers.
+For the LSTM encoder, 
+the model that concatenates the community embedding after the first encoder layer ($l_c=1$),
+performs almost as well as the model with $l_c=3$, but that model is still the best.
+There is much less variation in performance across different values of $l_c$ for the LSTM.
 
@@
 \end{tikzpicture}
-\caption{Community embedding PCA of the best LSTM (left, $c=1$) and transformer (right, $c=3$) models.}
+\caption{Community embedding PCA of an LSTM (left, $c=1$) and transformer (right, $c=3$) model.}
 \label{fig:comm-pca}
@@
   Average community perplexity (x-axis) versus information gain (y-axis)
-  for the LSTM model with $l_c = 1$ .
+  for the LSTM model with $l_c = 1$ (colors based on \cref{fig:comm-pca}).
 }
@@
     Correlation between indiscernibility for 
-    the best LSTM ($c = 1$) and transformer ($c = 3$) models
+    LSTM ($c = 1$) and transformer ($c = 3$) models
     (Pearson's $r = 0.99$, $p < 0.0001$).
@@
 
-In this work, we present results using a conditioned neural language models
+We have presented results using conditional neural language models
 to model variation between speech communities.
@@
 linguistic complexity \ref{hyp:varying-complexity}, with perplexity
-per-word ranging from less than 53 to more than 104 for the best model \bn{update}
+per-word ranging from less than 34 to more than 65 for the best model 
 architecture.
@@
 albeit weakly. For LSTM models, the perplexity per word, averaged over
-all communities, varies between 76.80 and 78.64. (with 80.70 for the
+messages from all communities, varies between \num{49.60} and \num{50.83}. (with \num{51.99} for the
 unconditioned model) The perplexity per word is slightly higher for
 $l_c=0$, and has very little difference between the other models, with
-$l_c=1$ being the best.  For transformer models, it varies between \bn{update}
-89.38 and 122.56. Here perplexity per-word is clearly not a monotonous
+$l_c=3$ being the best.  For transformer models, it varies between 
+\num{52.28} and \num{78.76}. Here perplexity per-word is clearly not a monotonous
 function of $l_c$: the maximum is reached at $l_c=2$. We did not

### 2020-11-24 paper/paper.tex
abstract; make space

@@
 \begin{abstract}
-  Using Community-Conditioned Language Models (CCLMs), we investigate
-  the following questions: 1. Can the level of linguistic complexity
-  exhibited by different communities is captured. 2. Are CCLMs
-  well-tuned to various communities both by intrinsic (can we build a
-  community classifier based on CCLMs) and extrinsic methods (are
-  CCLMs representations correlated with other network-based
-  representations of communities).
+  Community-Conditioned Language Models (CCLMs) are an under-researched tool for 
+  investigating community-level linguistic variation.
+  In this paper, experiment with different neural CCLM architectures and
+  report empirical results on data from 46 online communities.
+  We find that the community embeddings learned by these models
+  correlate with representations based on  non-linguistic features.
+  We define a metric of the linguistic \emph{indiscernability} of a community
+  and investigate the relationship between indiscernibility and other
+  linguistic properties of the community.
 \end{abstract}
-\jp{Not so good abstract}
 
@@
 
-Our dataset includes messages from \num{46} subreddits, 
+Our dataset includes messages from \num{46} English-language subreddits, 
 randomly selected from the set of all forums 
@@
 related to perplexity, however. 
-Some of the most linguistically complex communities benifit a lot
+Some of the most linguistically complex communities benefit a lot
 from community conditioning, while others do not benefit at all. 
-Similarly, among the less linguistically complex communites 
+Similarly, among the less linguistically complex communities 
 there some have high information gain, and some have information
@@
 
-\subsection{Latent language-model-based community classifiers}
+\subsection{Latent language model-based community classifiers}
 
@@
 community is. However, the CCLM induces a latent classifier,
-which we call a language-model-based community classifier
+which we call a language model-based community classifier
 (LMCC). Using the Bayes theorem, we can turn a
@@
 %
-\begin{figure*}
-%\begin{tikzpicture}
-  %\confusiongroupplot{lstm-3-1}{transformer-3-3}
-%\end{tikzpicture}
-\caption{LMCC confusion matrices for the best LSTM (left, $c=1$) and transformer (right, $c=3$) models.}
+\begin{figure}
+\begin{tikzpicture}
+  \confusionsingleplot{lstm-3-1}
+\end{tikzpicture}
+\caption{LMCC confusion matrices for the LSTM model with $c=1$.}
 \label{fig:confusion}
-\end{figure*}
+\end{figure}
 %
@@
 \cref{fig:indisc-corr}).
-
+%
 \begin{figure}
@@
 \end{figure}
-
+%
 However, we observe that community-wise perplexity is only weakly (if at all) 
@@
 is a poor predictor of how distinctive its language is.
-
+%
 \begin{table}
@@
 \end{table}
-
+%
 \begin{figure*}

### 2020-11-24 paper/plot_macros.tex
abstract; make space

@@
   ] table [x=actual_comm, y=pred_comm, meta=#1] {floats/confusion.csv};
+}
+
+\newcommand{\confusionsingleplot}[1]{
+  \begin{axis}[
+    group style={group size=2 by 1}, 
+    enlargelimits=false,
+    axis line style={draw=none},
+    %axis equal,
+    xmin=-1, xmax=46,
+    ymin=-1, ymax=46,
+    width=8cm,
+    xticklabels={Advice,AskWomen,BabyBumps,CFB,Drugs,EDH,EarthPorn,Fantasy,GameDeals,GlobalOffensive,Jokes,Kappa,KerbalSpaceProgram,KotakuInAction,LifeProTips,MLS,MMA,MaddenUltimateTeam,TwoXChromosomes,Warframe,airsoft,bodybuilding,breakingmom,cars,cringe,eu4,exjw,explainlikeimfive,femalefashionadvice,food,heroesofthestorm,jailbreak,justneckbeardthings,oculus,pcmasterrace,photography,reddevils,relationships,rupaulsdragrace,stopdrinking,streetwear,techsupport,todayilearned,toronto,videos,xxfitness},
+    yticklabels=none,
+    x tick label style={font=\sffamily\tiny,rotate=90,anchor=east},
+    xtick=data,
+    legend style={font=\sffamily\small},
+    colorbar right,
+    ]
+    \confusionplot{#1}
+  \end{axis}
 }

### 2021-02-18 data/collect_reddit_data.py
collect new reddit data

@@ -1,0 +1,122 @@
+import pandas as pd
+import os
+import google.auth
+from google.cloud import bigquery
+from google.cloud import bigquery_storage_v1beta1
+
+def remove_nul(filename):
+    with open(filename, 'r') as f:
+        data = f.read()
+    data = data.replace('\00', '')
+    with open(filename, 'w') as f:
+        f.write(data)
+
+def get_comment_counts(year, month, min_posts, bq_client, bqstorage_client):
+    query_string = f"""
+    SELECT subreddit, post_count
+    FROM (
+      SELECT subreddit, COUNT(id) as post_count
+      FROM `fh-bigquery.reddit_comments.{year}_{month:02d}`
+      GROUP BY subreddit
+    )
+    WHERE post_count > {min_posts}
+    ORDER BY post_count DESC
+    """
+    df = bq_client.query(query_string)\
+        .result()\
+	.to_dataframe(bqstorage_client=bqstorage_client)
+    df = df.set_index('subreddit')
+    df = df.rename({'post_count': f"{year}-{month:02d}"}, axis=1)
+    return df
+
+def sample_subreddit_comments(year, month, subs, n_comments,
+	gcp_project, gs_bucket, gs_dir, bq_dataset_id, bq_client):
+    # bq stratified sample https://stackoverflow.com/a/52901452
+    query_string = f"""
+    WITH table AS (
+      SELECT *, subreddit category
+      FROM `fh-bigquery.reddit_comments.{year}_{month:02d}` a
+    ), table_stats AS (
+      SELECT *, SUM(c) OVER() total
+      FROM (
+	SELECT category, COUNT(*) c
+	FROM table
+	GROUP BY 1
+	HAVING category in ({', '.join([f"'{sub}'" for sub in subs])})
+        )
+    )
+    SELECT r.id, r.link_id, r.parent_id, r.created_utc, r.subreddit, r.author, r.body
+    FROM (
+      SELECT ARRAY_AGG(a ORDER BY RAND() LIMIT {n_comments}) comment
+      FROM table a
+      JOIN table_stats b
+      USING(category)
+      GROUP BY category
+    ), UNNEST(comment) as r
+    """
+    # Query comments for the selecetd subs and save to a BQ table
+    table_id = f"{gcp_project}.{bq_dataset_id}.{year}_{month:02d}"
+    job_config = bigquery.QueryJobConfig(
+	allow_large_results=True,
[…]

### 2021-02-18 ngram_model.py
ngram stub

@@ -1,47 +1,6 @@
 import os
-import torch
-import random
-import data
 from torchtext.data.utils import ngrams_iterator
 from collections import Counter, defaultdict
 
-data_dir = "data/reddit2015"
-model_dir = "model/reddit2015"
-max_seq_len = 64
-file_limit = 50000
-
-dataset, fields = data.load_data_and_fields(data_dir, model_dir,
-        max_seq_len, file_limit)
-
-vocab_size = len(fields['text'].vocab.itos)
-comm_vocab_size = len(fields['community'].vocab.itos)
-comm_unk_idx = fields['community'].vocab.stoi['<unk>']
-text_pad_idx = fields['text'].vocab.stoi['<pad>']
-
-random.seed(42)
-random_state = random.getstate()
-train_data, val_data, test_data = dataset.split(split_ratio=[0.8,0.1,0.1], stratified=True, strata_field='community', random_state=random_state)
-
 def bigrams_iter(tokens):
     return zip(*[tokens[i:] for i in range(2)])
-
-counts = defaultdict(lambda: Counter())
-for i, example in enumerate(train_data):
-    if (i % 1000) == 0:
-        print(f"{i}/{len(train_data)}", end='\r')
-    tokens = ['<pad>'] + example.text + ['<eos>'] # use bad for BoS
-    tokens = [fields['text'].vocab.stoi[w] for w in tokens]
-    for w1,w2 in bigrams_iter(tokens):
-        counts[w1][w2] += 1
-
-model = {}
-for i, w1 in enumerate(counts):
-    print(f"{i}/{vocab_size}", end='\r')
-    denom = sum(counts[w1].values()) + vocab_size
-    model[w1] = defaultdict(lambda x: 1/denom)
-    for w2 in counts:
-        model[w2] = (counts[w1][w2] + 1) / denom 
-
-import pickle
-with open('model/reddit2015/bigram_model.pickle', 'wb') as f:
-    pickle.dump(model, f)

### 2021-02-18 data/dedup_reddit_comments.py
dedup comments

@@ -1,0 +1,33 @@
+from pathlib import Path
+
+def iter_comments(corpus_dir):
+    for comment_file in corpus_dir.glob('*.csv'):
+        print(comment_file)
+        with open(comment_file, 'r') as f:
+            reader = csv.DictReader(f)
+                for comment in reader:
+                    yield comment['id'], comment['body']
+
+
+def find_dupe_comments(comments):
+    min_length = 50
+    seen_hashes = set()
+    dupe_ids = []
+        for comment_id, body in comments: 
+            comment_len = len(body) 
+            if comment_len < min_length:
+                continue
+            # only consider the last 50 characters for duplicates (catches most bot-filled forms)
+            comment_hash = hash(body[-min_length:])
+            if comment_hash in seen_hashes:
+                dupe_ids.append(comment_id)
+            else:
+                seen_hashes.add(comment_hash) 
+    return dupe_ids
+
+corpus_dir = Path('synth_data')
+comment_iter = iter_comments(corpus_dir)
+dupe_ids = find_dupe_comments(coment_iter)
+with open(corpus_dir/'dupes.txt', 'w') as f:
+    f.write('\n'.join(dupe_ids))
+

### 2021-02-18 data/collect_reddit_data.py
drip

@@ -1,5 +1,6 @@
 import pandas as pd
 import os
+from pathlib import Path
 import google.auth
 from google.cloud import bigquery
 from google.cloud import bigquery_storage_v1beta1
@@ -120,3 +121,10 @@ if __name__ == '__main__':
         sample_subreddit_comments(2015, month, subs, n_sample_comments, 
             gcp_project, gs_bucket, gs_dir, bq_dataset_id, bq)
 
+    # copy the files locally from google storage
+    os.system("gsutil -m cp gs://bill-gu-research/CondLM/2015-*.csv reddit_sample/")
+
+    corpus_dir = Path('reddit_sample')
+    for comment_file in corpus_dir.glob('*.csv'):
+        print(comment_file)
+        remove_nul(comment_file)

### 2021-02-18 data/dedup_reddit_comments.py

@@ -1,33 +1,34 @@
 from pathlib import Path
+import csv
 
 def iter_comments(corpus_dir):
     for comment_file in corpus_dir.glob('*.csv'):
         print(comment_file)
         with open(comment_file, 'r') as f:
             reader = csv.DictReader(f)
-                for comment in reader:
-                    yield comment['id'], comment['body']
+            for comment in reader:
+                yield comment['id'], comment['body']
 
 
 def find_dupe_comments(comments):
     min_length = 50
     seen_hashes = set()
     dupe_ids = []
-        for comment_id, body in comments: 
-            comment_len = len(body) 
-            if comment_len < min_length:
-                continue
-            # only consider the last 50 characters for duplicates (catches most bot-filled forms)
-            comment_hash = hash(body[-min_length:])
-            if comment_hash in seen_hashes:
-                dupe_ids.append(comment_id)
-            else:
-                seen_hashes.add(comment_hash) 
+    for comment_id, body in comments: 
+        comment_len = len(body) 
+        if comment_len < min_length:
+            continue
+        # only consider last 50 chars (catches most bot-filled forms)
+        comment_hash = hash(body[-min_length:])
+        if comment_hash in seen_hashes:
+            dupe_ids.append(comment_id)
+        else:
+            seen_hashes.add(comment_hash) 
     return dupe_ids
 
-corpus_dir = Path('synth_data')
+corpus_dir = Path('reddit_sample')
 comment_iter = iter_comments(corpus_dir)
-dupe_ids = find_dupe_comments(coment_iter)
+dupe_ids = find_dupe_comments(comment_iter)
 with open(corpus_dir/'dupes.txt', 'w') as f:
     f.write('\n'.join(dupe_ids))
 

### 2021-02-18 data/dedup_reddit_comments.py
preprocessing pipeline

@@ -9,7 +9,6 @@ def iter_comments(corpus_dir):
             for comment in reader:
                 yield comment['id'], comment['body']
 
-
 def find_dupe_comments(comments):
     min_length = 50
     seen_hashes = set()

### 2021-02-18 data/tokenize_reddit_comments.py

@@ -1,0 +1,54 @@
+import re
+import mistletoe
+from bs4 import BeautifulSoup
+import spacy
+from spacy.lang.en import English
+from multiprocessing import Pool
+from pathlib import Path
+import csv
+
+def preprocess(text)
+    text = re.sub(r"\^", ' ', text)
+    # For some reason the pushshift.io comments have HTML escapes in the markdown
+    # we relpace them since the markdown parser doesn't recognie them.
+    text = text.replace('&amp;', '&')
+    text = text.replace('&lt;', '<')
+    text = text.replace('&gt;', '>')
+    # Render the markdown as an HTML string with mistletoe
+    html = mistletoe.markdown(text)
+    # Load HTML into BeautifulSoup 
+    soup = BeautifulSoup(html, 'lxml')
+    # Remove tables
+    (t.extract() for t in soup('table'))
+    # Remove code blocks
+    (t.extract() for t in soup('code'))
+    # Remove block quotes
+    (t.extract() for t in soup('blockquote'))
+    # Extract text
+    text = soup.text
+    # Replace whitespaces & other special chars with space 
+    text = re.sub(r"\s|\||\*", ' ', text)
+    # Normalize all sequences of more than 3 dots to 3 dots
+    text = re.sub(r"\.\.+(?=\.)", '..', text)
+    # Remove URLs
+    text = re.sub(r"https?:\/\/[^\s]+", '', text)
+    # Remove extra whitespace
+    text = re.sub(r"^\s+|\s+$|\s+(?=\s)", '', text)
+
+def preprocess_file(comment_file):
+    nlp = English()
+    spacy_tokenizer = nlp.Defaults.create_tokenizer(nlp)
+    # Use the SpaCy tokenizer
+    tokenized_comments = []
+    print(comment_file)
+    with open(comment_file, 'r') as f:
+        reader = csv.DictReader(f)
+        for comment in reader:
+            comment_text = preprocess(comment['body'])
+            tokenized = [token.text for token in spacy_tokenizer(comment_text)]
+            tokenized_comments.append((comment['subreddit'], comment_text))
+    return tokenized_comments
+
+corpus_dir = Path('reddit_sample')
+with Pool(processes=8) as p:
+    p.starmap(preprocess_file, corpus_dir.glob('*.csv'))

### 2021-02-19 ngram_model.py
start working on ngram

@@ -1,6 +1,65 @@
-import os
-from torchtext.data.utils import ngrams_iterator
-from collections import Counter, defaultdict
+from pathlib import Path
+import util
+import data
+import random
+import itertools
+import numpy as np
 
-def bigrams_iter(tokens):
-    return zip(*[tokens[i:] for i in range(2)])
+def iter_ngrams(text, n=3):
+    ts = itertools.tee(text, n)
+    for i, t in enumerate(ts[1:]):
+        for _ in range(i+1):
+            next(t)
+    return zip(*ts)
+
+# @click.argument('model_dir', type=click.Path(exists=True))
+# @click.argument('model_name', type=str)
+# @click.argument('data_dir', type=click.Path(exists=True))
+# @click.option('--rebuild-vocab/--no-rebuild-vocab', default=False)
+# @click.option('--vocab-size', default=40000)
+# @click.option('--max-seq-len', default=64)
+# @click.option('--file-limit', type=int, default=None,
+        # help="Number of examples per file (community).")
+
+model_dir = 'model/synth'
+model_name = 'bigram'
+data_dir = 'data/synth_data'
+rebuild_vocab = False
+vocab_size = 40000
+max_seq_len = 64
+file_limit = None
+
+model_dir = Path(model_dir)
+data_dir = Path(data_dir)
+save_dir = model_dir/model_name
+
+util.mkdir(model_dir)
+util.mkdir(save_dir)
+
+log = util.create_logger('ngram', save_dir/'ngram.log', True)
+
+dataset, fields = data.load_data_and_fields(data_dir, model_dir,
+        max_seq_len, file_limit, vocab_size, rebuild_vocab)
+
+vocab_size = len(fields['text'].vocab.itos)
+n_comms = len(fields['community'].vocab.itos)
+comm_unk_idx = fields['community'].vocab.stoi['<unk>']
+text_pad_idx = fields['text'].vocab.stoi['<pad>']
+log.info(f"Loaded {len(dataset)} examples.")
+
+random.seed(42)
+random_state = random.getstate()
+train_data, val_data, test_data = dataset.split(split_ratio=[0.8,0.1,0.1], stratified=True, strata_field='community', random_state=random_state)
+log.info(f"Splits: train: {len(train_data)} val: {len(val_data)} test: {len(test_data)} ")
+
+# bigram counts
+C = np.zeros((n_comms,vocab_size,vocab_size),int)
+
+for example in train_data:
[…]

### 2021-04-20 model.py
remove unused comm inference block

@@ -22,7 +22,6 @@ class CommunityConditionedLM(nn.Module):
         self.decoder = nn.Linear(hidden_size, n_tokens)
         self.encoder_before = encoder_before
         self.encoder_after = encoder_after
-        self.use_tunned_comm = False
         if use_community:
             self.comm_inference = nn.Embedding(n_comms, n_comms)
             self.comm_embed = WeightedEmbedding(n_comms, comm_emsize)
@@ -58,10 +57,7 @@ class CommunityConditionedLM(nn.Module):
         if self.encoder_before is not None:
             x = self.drop(self.encoder_before(x))
         if self.use_community:
-            if self.use_tunned_comm:
-                x_comm = self.comm_inference(comm).softmax(1)
-            else:
-                x_comm = F.one_hot(comm, num_classes=self.n_comms).type(torch.FloatTensor).to(device)
+            x_comm = F.one_hot(comm, num_classes=self.n_comms).type(torch.FloatTensor).to(device)
             x_comm = self.comm_embed(x_comm).repeat(text.shape[0],1,1)
             x = torch.cat((x, x_comm), 2)
             x = self.drop(self.comm_linear(x))

### 2021-04-20 paper/paper.tex
tweak vocab selection

@@
-%
+% This must be in the first 5 lines to tell arXiv to use pdfLaTeX, which is strongly recommended.
+%\pdfoutput=1
+% In particular, the hyperref package requires pdfLaTeX in order to break URLs across lines.
 
-% File acl2020.tex
-%
-%% Based on the style files for ACL 2020, which were
-%% Based on the style files for ACL 2018, NAACL 2018/19, which were
-%% Based on the style files for ACL-2015, with some improvements
-%%  taken from the NAACL-2016 style
-%% Based on the style files for ACL-2014, which were, in turn,
-%% based on ACL-2013, ACL-2012, ACL-2011, ACL-2010, ACL-IJCNLP-2009,
-%% EACL-2009, IJCNLP-2008...
-%% Based on the style files for EACL 2006 by 
-%%e.agirre@ehu.es or Sergi.Balari@uab.es
-%% and that of ACL 08 by Joakim Nivre and Noah Smith
+\documentclass[11pt]{article}
 
-\documentclass[11pt,a4paper]{article}
-\usepackage[hyperref]{acl2020}
-\usepackage[T1]{fontenc}
+% Remove the "review" option to generate the final version.
+\usepackage[review]{acl}
+
+% Standard package includes
 \usepackage{times}
 \usepackage{latexsym}
+
+% For proper rendering and hyphenation of words containing Latin characters (including in bib files)
+\usepackage[T1]{fontenc}
+% For Vietnamese characters
+% \usepackage[T5]{fontenc}
+% See https://www.latex-project.org/help/documentation/encguide.pdf for other character sets
+
+% This assumes your files are encoded as UTF8
+\usepackage[utf8]{inputenc}
+
+% This is not strictly necessary, and may be commented out,
+% but it will improve the layout of the manuscript,
+% and will typically save some space.
+\usepackage{microtype}
+
 \usepackage{amsmath}
@@
 
-% This is not strictly necessary, and may be commented out,
-% but it will improve the layout of the manuscript,
-% and will typically save some space.
-\usepackage{microtype}
-
-%\aclfinalcopy % Uncomment this line for the final submission
-%\def\aclpaperid{***} %  Enter the acl Paper ID here
-
-%\setlength\titlebox{5cm}
-% You can expand the titlebox if you need extra space
-% to show all the authors. Please do not make the titlebox
-% smaller than 5cm (the original size); we will check this
-% in the camera-ready version and ask you to change it back.
-
 \newcommand\Ppl{\mathsf{Ppl}}
@@
 
-\author{Anon.}
+% Author information can be set in various styles:
+% For several authors from the same institution:
+% \author{Author 1 \and ... \and Author n \\
+%         Address line \\ ... \\ Address line}
+% if the names do not fit well on one line use
+%         Author 1 \\ {\bf Author 2} \\ ... \\ {\bf Author n} \\
+% For authors from different institutions:
+% \author{Author 1 \\ Address line \\  ... \\ Address line
+%         \And  ... \And
+%         Author n \\ Address line \\ ... \\ Address line}
+% To start a seperate ``row'' of authors use \AND, as in
+% \author{Author 1 \\ Address line \\  ... \\ Address line
+%         \AND
+%         Author 2 \\ Address line \\ ... \\ Address line \And
+%         Author 3 \\ Address line \\ ... \\ Address line}
+
+\author{First Author \\
+  Affiliation / Address line 1 \\
+  Affiliation / Address line 2 \\
+  Affiliation / Address line 3 \\
+  \texttt{email@domain} \\\And
+  Second Author \\
+  Affiliation / Address line 1 \\
+  Affiliation / Address line 2 \\
+  Affiliation / Address line 3 \\
+  \texttt{email@domain} \\}
 
@@
   investigating community-level linguistic variation.
-  In this paper, experiment with different neural CCLM architectures and
+  In this paper, we experiment with different neural CCLM architectures and
   report empirical results on data from 46 online communities.
@@
 
-\jp{Perhaps: Outline of the paper.}
+%\jp{Perhaps: Outline of the paper.}
 
@@
 
-We use a vocabulary size of \num{40000} tokens, including a special
-out-of-vocabulary token.  The vocabulary consists of the most frequent
-tokens at the message level (across all communities), meaning that we
-only count each token once per message.  The main reason for this
-choice is to avoid over-counting otherwise rare tokens that were
-repeated many times in a single message.
+The vocabulary is drawn from the set of all words appearing with a frequency
+greater than \num{5e-5} (1 in \num{2000}) in at least one community.
+This results in a vocabulary size of \num{47403}, including a special
+out-of-vocablary token.
+
+%We use a vocabulary size of \num{40000} tokens, including a special
+%out-of-vocabulary token.  The vocabulary consists of the most frequent
+%tokens at the message level (across all communities), meaning that we
+%only count each token once per message.  The main reason for this
+%choice is to avoid over-counting otherwise rare tokens that were
+%repeated many times in a single message.
 

### 2021-04-20 data.py

@@ -7,28 +7,37 @@ import util
 
 def gen_examples(data_dir, fields, split, max_seq_len, file_limit):
     fields = fields.items()
-    for community, line in util.iter_data(data_dir, split, file_limit):
-        tokens = line.split(' ')[:max_seq_len]
-        yield tt.data.Example.fromlist([community, tokens], fields)
+    for community, text, i in util.iter_data(data_dir, split, file_limit):
+        tokens = text.lower().split(' ')[:max_seq_len]
+        yield tt.data.Example.fromlist([community, tokens, i], fields)
 
-def build_fields(data_dir, vocab_size):
+def build_fields(data_dir, min_freq, file_limit):
 
-    COMMUNITY = tt.data.Field(sequential=False, pad_token=None, unk_token=None)
-    TEXT = tt.data.Field(eos_token='<eos>', init_token='<bos>', tokenize=None)
-    fields = OrderedDict([('community', COMMUNITY), ('text', TEXT)])
+    fields = OrderedDict([
+        ('community', tt.data.Field(sequential=False, pad_token=None, unk_token=None)),
+        ('text', tt.data.Field(eos_token='<eos>', init_token='<bos>', tokenize=None)),
+        ('example_id', tt.data.Field(sequential=False, use_vocab=False))
+    ])
 
-    data = tt.data.Dataset(list(gen_examples(data_dir, fields, 'train', None, None)), fields=fields)
+    data = tt.data.Dataset(list(gen_examples(data_dir, fields, 'train', None, file_limit)),
+            fields=fields)
     fields['community'].build_vocab(data)
-
-    token_counts = Counter()
+    comms = fields['community'].vocab.itos
+    token_counts_comm = {comm: Counter() for comm in comms}
+    token_counts_all = Counter()
     for i, example in enumerate(data):
-        token_counts.update(set(example.text)) # avoid over-counting spammed tokens
-    fields['text'].vocab = tt.vocab.Vocab(token_counts, max_size=vocab_size,
+        token_counts_comm[example.community].update(example.text)
+    for comm in comms:
+        min_count = min_freq * sum(token_counts_comm[comm].values())
+        for w,c in token_counts_comm[comm].items():
+            if c >= min_count:
+                token_counts_all[w] += c
+    fields['text'].vocab = tt.vocab.Vocab(token_counts_all,
             specials=['<eos>', '<unk>', '<bos>', '<pad>'])
 
     return fields
 
-def load_fields(field_dir, data_dir=None, vocab_size=None):
[…]

### 2021-04-20 train.py

@@ -1,4 +1,3 @@
-import os
 import click
 import data
 import model
@@ -9,6 +8,8 @@ import torch.nn.functional as F
 import math
 import random
 import util
+import os
+from pathlib import Path
 
 def train(lm, batches, vocab_size, criterion, optimizer, log):
     lm.train()
@@ -55,15 +56,15 @@ def evaluate(lm, batches, vocab_size, criterion):
 @click.argument('data_dir', type=click.Path(exists=True))
 @click.option('--resume-training/--no-resume-training', default=False)
 @click.option('--rebuild-vocab/--no-rebuild-vocab', default=False)
-@click.option('--vocab-size', default=40000)
+@click.option('--min-token-freq', default=5e-5)
 @click.option('--encoder-layers', default=1)
 @click.option('--heads', default=8)
-@click.option('--hidden-size', default=256)
+@click.option('--hidden-size', default=128)
 @click.option('--condition-community/--no-condition-community', default=True)
 @click.option('--community-emsize', default=16)
 @click.option('--community-layer-no', default=0)
 @click.option('--dropout', default=0.1)
-@click.option('--batch-size', default=32)
+@click.option('--batch-size', default=128)
 @click.option('--max-seq-len', default=64)
 @click.option('--lr', default=0.001)
 @click.option('--max-epochs', type=int, default=None)
@@ -72,23 +73,24 @@ def evaluate(lm, batches, vocab_size, criterion):
 @click.option('--gpu-id', type=int, default=None,
         help="ID of the GPU, if traning with CUDA")
 def cli(architecture, model_dir, data_dir, resume_training, rebuild_vocab,
-        vocab_size, encoder_layers, heads, hidden_size,
+        min_token_freq, encoder_layers, heads, hidden_size,
         condition_community, community_emsize, community_layer_no, dropout,
         batch_size, max_seq_len, lr, max_epochs, file_limit, gpu_id):
 
+    model_dir = Path(model_dir)
     model_name = f"{architecture.lower()}-{encoder_layers}" + (f"-{community_layer_no}" if condition_community else "")
-    save_dir = os.path.join(model_dir, model_name)
+    save_dir = model_dir/model_name
     util.mkdir(save_dir)
-    log = util.create_logger('train', os.path.join(save_dir, 'training.log'), True)
[…]

### 2021-04-20 eval_model.py
eval_model just computes the nll matrix p(m|c) for each message, commpnity pair

@@ -1,0 +1,130 @@
+""" Use the trained models on test data, testing both CCLM and LMCC.
+    Produce results tabels that will be read by analysis.py to create plots, tables, etc.
+"""
+
+import util
+import data
+import model
+import torch
+import torch.nn as nn
+import torch.nn.functional as F
+import torchtext as tt
+import numpy as np
+import random
+import json
+import csv
+import os
+import click
+from pathlib import Path
+
+def exp_normalize(x, axis):
+    """ https://timvieira.github.io/blog/post/2014/02/11/exp-normalize-trick/ """
+    b = x.max(axis=axis)
+    y = np.exp(x - b)
+    return y / y.sum(axis)
+
+def batch_nll(lm, batch, pad_idx, comm=None):
+    """
+    Compute the negative log likelihood of the batch for the given community
+    If comm is None, then the actual communities for the batch will be used.
+    """
+    if comm is None:
+        comm = batch.community
+    text, lengths = batch.text
+    x_text = text[:-1]
+    y = text[1:]
+    y_hat = lm(x_text, comm)
+    vocab_size = y_hat.shape[-1]
+    nll_seq = F.nll_loss(y_hat.view(-1,vocab_size), y.view(-1), 
+            reduction='none', ignore_index=pad_idx).view(y.shape).sum(axis=0)
+    return nll_seq
+
+@click.command()
+@click.argument('model_family_dir', type=click.Path(exists=False))
+@click.argument('model_name', type=str)
+@click.argument('data_dir', type=click.Path(exists=True))
+@click.option('--batch-size', default=512)
+@click.option('--max-seq-len', default=64)
+@click.option('--file-limit', type=int, default=None,
+        help="Number of examples per file (community).")
+@click.option('--gpu-id', type=int, default=None,
+        help="ID of the GPU, if traning with CUDA")
+def cli(model_family_dir, model_name, data_dir, batch_size, max_seq_len, file_limit, gpu_id):
+
+    model_family_dir = Path(model_family_dir)
+    model_dir = model_family_dir/model_name
+    device = torch.device(f'cuda:{gpu_id}' if gpu_id is not None else 'cpu')
+    log = util.create_logger('test', os.path.join(model_dir, 'testing.log'), True)
+
+    log.info(f"Loading data from {data_dir}.")
+    fields = data.load_fields(model_family_dir)
+    fields['text'].include_lengths = True
[…]

### 2021-04-20 test_model.py

@@ -1,179 +1,0 @@
-""" Use the trained models on test data, testing both CCLM and LMCC.
-    Produce results tabels that will be read by analysis.py to create plots, tables, etc.
-"""
-
-from model import CommunityConditionedLM
-from data import load_data_and_fields
-import torch
-import torch.nn as nn
-import torch.nn.functional as F
-import torchtext as tt
-import pandas as pd
-import numpy as np
-import random
-import json
-import os
-import click
-from pathlib import Path
-
-def exp_normalize(x, axis):
-    """ https://timvieira.github.io/blog/post/2014/02/11/exp-normalize-trick/ """
-    b = x.max(axis=axis)
-    y = np.exp(x - b)
-    return y / y.sum(axis)
-
-def mask_seqs(max_seq_len, batch_size, lengths):
-    idxs = torch.arange(max_seq_len).expand(batch_size, max_seq_len).to(lengths.device)
-    mask = idxs < lengths.unsqueeze(1)
-    return mask.T.float()
-
-def batch_nll(model, batch, comm=None):
-    """
-    Compute the negative log likelihood of the batch for the given community
-    If comm is None, then the actual communities for the batch will be used.
-    """
-    if comm is None:
-        comm = batch.community
-    text, lengths = batch.text
-    max_seq_len = text.shape[0] - 1
-    x_text = text[:-1]
-    y = text[1:]
-    y_hat = model(x_text, comm)
-    vocab_size = y_hat.shape[-1]
-    nll_word = F.nll_loss(y_hat.view(-1,vocab_size), y.view(-1), reduction='none').view(y.shape)
-    mask = mask_seqs(max_seq_len, batch.batch_size, lengths-1)
-    nll_word = nll_word * mask
-    nll_seq = nll_word.sum(axis=0) 
-    return nll_seq
-
-@click.group()
-@click.argument('model_family_dir', type=click.Path(exists=False))
-@click.argument('data_dir', type=click.Path(exists=True))
-@click.option('--batch-size', default=512)
-@click.option('--max-seq-len', default=64)
-@click.option('--file-limit', type=int, default=None,
-        help="Number of examples per file (community).")
-@click.option('--gpu-id', type=int, default=None,
-        help="ID of the GPU, if traning with CUDA")
-@click.pass_context
-def cli(ctx, model_family_dir, data_dir, batch_size, max_seq_len, file_limit, gpu_id):
-    ctx.ensure_object(dict)
-
[…]

### 2021-04-21 paper/paper.tex
actually the current scheme isn't so bad... just make the non-lower-case explicit

@@
 
-The vocabulary is drawn from the set of all words appearing with a frequency
-greater than \num{5e-5} (1 in \num{2000}) in at least one community.
-This results in a vocabulary size of \num{47403}, including a special
-out-of-vocablary token.
-
-%We use a vocabulary size of \num{40000} tokens, including a special
-%out-of-vocabulary token.  The vocabulary consists of the most frequent
-%tokens at the message level (across all communities), meaning that we
-%only count each token once per message.  The main reason for this
-%choice is to avoid over-counting otherwise rare tokens that were
-%repeated many times in a single message.
+We use a vocabulary size of \num{40000} tokens, including a special
+out-of-vocabulary token.  The vocabulary consists of the most frequent
+tokens at the message level (across all communities), meaning that we
+only count each token once per message.  The main reason for this
+choice is to avoid over-counting otherwise rare tokens that were
+repeated many times in a single message.
 

### 2021-04-21 data.py

@@ -5,13 +5,15 @@ import torch
 import csv
 import util
 
-def gen_examples(data_dir, fields, split, max_seq_len, file_limit):
+def gen_examples(data_dir, fields, split, max_seq_len, file_limit, lower_case):
     fields = fields.items()
     for community, text, i in util.iter_data(data_dir, split, file_limit):
-        tokens = text.lower().split(' ')[:max_seq_len]
+        if lower_case:
+            text = text.lower()
+        tokens = text.split(' ')[:max_seq_len]
         yield tt.data.Example.fromlist([community, tokens, i], fields)
 
-def build_fields(data_dir, min_freq, file_limit):
+def build_fields(data_dir, vocab_size):
 
     fields = OrderedDict([
         ('community', tt.data.Field(sequential=False, pad_token=None, unk_token=None)),
@@ -19,25 +21,17 @@ def build_fields(data_dir, min_freq, file_limit):
         ('example_id', tt.data.Field(sequential=False, use_vocab=False))
     ])
 
-    data = tt.data.Dataset(list(gen_examples(data_dir, fields, 'train', None, file_limit)),
+    data = tt.data.Dataset(list(gen_examples(data_dir, fields, 'train', None, None)),
             fields=fields)
     fields['community'].build_vocab(data)
-    comms = fields['community'].vocab.itos
-    token_counts_comm = {comm: Counter() for comm in comms}
-    token_counts_all = Counter()
+    token_counts = Counter()
     for i, example in enumerate(data):
-        token_counts_comm[example.community].update(example.text)
-    for comm in comms:
-        min_count = min_freq * sum(token_counts_comm[comm].values())
-        for w,c in token_counts_comm[comm].items():
-            if c >= min_count:
-                token_counts_all[w] += c
-    fields['text'].vocab = tt.vocab.Vocab(token_counts_all,
+        token_counts.update(set(example.text)) # avoid over-counting spammed tokens
+    fields['text'].vocab = tt.vocab.Vocab(token_counts, max_size=vocab_size,
             specials=['<eos>', '<unk>', '<bos>', '<pad>'])
-
     return fields
 
-def load_fields(field_dir, data_dir=None, min_freq=0, file_limit=None):
+def load_fields(field_dir, data_dir=None, vocab_size):
     text_field_file = os.path.join(field_dir, 'text.field')
     comm_field_file = os.path.join(field_dir, 'community.field')
     print(text_field_file)
[…]

### 2021-04-21 train.py

@@ -56,7 +56,7 @@ def evaluate(lm, batches, vocab_size, criterion):
 @click.argument('data_dir', type=click.Path(exists=True))
 @click.option('--resume-training/--no-resume-training', default=False)
 @click.option('--rebuild-vocab/--no-rebuild-vocab', default=False)
-@click.option('--min-token-freq', default=5e-5)
+@click.option('--vocab-size', default=40000)
 @click.option('--encoder-layers', default=1)
 @click.option('--heads', default=8)
 @click.option('--hidden-size', default=128)
@@ -73,7 +73,7 @@ def evaluate(lm, batches, vocab_size, criterion):
 @click.option('--gpu-id', type=int, default=None,
         help="ID of the GPU, if traning with CUDA")
 def cli(architecture, model_dir, data_dir, resume_training, rebuild_vocab,
-        min_token_freq, encoder_layers, heads, hidden_size,
+        vocab_size, encoder_layers, heads, hidden_size,
         condition_community, community_emsize, community_layer_no, dropout,
         batch_size, max_seq_len, lr, max_epochs, file_limit, gpu_id):
 
@@ -85,7 +85,7 @@ def cli(architecture, model_dir, data_dir, resume_training, rebuild_vocab,
     log.info(f"Model will be saved to {save_dir}.")
 
     log.info(f"Loading data from {data_dir}.")
-    fields = data.load_fields(model_dir, data_dir, min_token_freq, file_limit)
+    fields = data.load_fields(model_dir, data_dir, vocab_size) 
     train_data = data.load_data(data_dir, fields, 'train', max_seq_len, file_limit)
     vocab_size = len(fields['text'].vocab.itos)
     comm_vocab_size = len(fields['community'].vocab.itos)

### 2021-04-21 data.py
fix arg bug

@@ -31,7 +31,7 @@ def build_fields(data_dir, vocab_size):
             specials=['<eos>', '<unk>', '<bos>', '<pad>'])
     return fields
 
-def load_fields(field_dir, data_dir=None, vocab_size):
+def load_fields(field_dir, data_dir=None, vocab_size=None):
     text_field_file = os.path.join(field_dir, 'text.field')
     comm_field_file = os.path.join(field_dir, 'community.field')
     print(text_field_file)

### 2021-05-03 paper/plot_macros.tex
updates to paper

@@
 }
+
+% https://tex.stackexchange.com/questions/183778/how-can-several-boxplots-be-combined-into-a-single-figure-as-groups/183856#183856
+\newcommand{\pplbycommbox}{
+  \begin{axis}
+    \addplot+[boxplot={draw position=-0.2, box extend=0.2}, draw=blue] table [y=lstm-3-0_ig] {floats/ig_by_comm.csv};
+    \addplot+[boxplot={draw position=0.2, box extend=0.2}, draw=red] table [y=transformer-3-0_ig] {floats/ig_by_comm.csv};
+    \addplot+[boxplot={draw position=.8, box extend=0.2}, draw=blue] table [y=lstm-3-1_ig] {floats/ig_by_comm.csv};
+    \addplot+[boxplot={draw position=1.2, box extend=0.2}, draw=red] table [y=transformer-3-1_ig] {floats/ig_by_comm.csv};
+    \addplot+[boxplot={draw position=1.8, box extend=0.2}, draw=blue] table [y=lstm-3-2_ig] {floats/ig_by_comm.csv};
+    \addplot+[boxplot={draw position=2.2, box extend=0.2}, draw=red] table [y=transformer-3-2_ig] {floats/ig_by_comm.csv};
+    \addplot+[boxplot={draw position=2.8, box extend=0.2}, draw=blue] table [y=lstm-3-3_ig] {floats/ig_by_comm.csv};
+    \addplot+[boxplot={draw position=3.2, box extend=0.2}, draw=red] table [y=transformer-3-3_ig] {floats/ig_by_comm.csv};
+  \end{axis}
+}

### 2021-05-03 analysis.py
results/analyisis notebooks

@@ -1,5 +1,6 @@
 import pandas as pd
 import os
+from pathlib import Path
 import numpy as np
 import torch
 from scipy.stats import pearsonr

### 2021-05-03 data/collect_reddit_data.py

@@ -86,9 +86,7 @@ if __name__ == '__main__':
     """
     We want to find subreddits that have at least 5000 comments per month 
     for every month of 2015.
-    We will sample of 250 subreddits from this list.
     And sample 5000 comments per month = 60000 comments per subreddit
-    Which is 60000 * 250 = 15,000,000 comments total
     """
 
     google_credentials_file = '/home/xnobwi/.google-cloud/bill-gu-research-177baaa5bef7.json'

### 2021-05-03 paper/paper.tex

@@
   In this paper, we experiment with different neural CCLM architectures and
-  report empirical results on data from 46 online communities.
+  report empirical results from CCLMs trained on 510 online communities.
   We find that the community embeddings learned by these models
@@
 better performance on various standard NLP tasks
-\citep{Hovy2015,Yang2017}, but little is known about how variation at
-the level of speech community affects performance.
+\citep{Hovy2015,Yang2017}, 
 %
-There is a fair amount of work that investigates linguistic variation
-using computational methods.  However, most of it considers variation
-across macro-social categories, such as gender
+However, most of the previous computational work on linguistic variation 
+has considered variation at the level of macro-social categories, such as gender
 \citep{Burger2011,Ciot2013,Bamman2014}, age \citep{Nguyen2013}, and
 geographic location \citep{Eisenstein2010,Bamman2014a}.
-\citet{DelTredici2017} provide an exception, by using the modified
+An exception is \citet{DelTredici2017} who use the modified
 skip-gram model of \citet{Bamman2014a} to show that lexical semantic
@@
 
-In the present work, we further investigate linguistic variation
-across online communities. In contrast to previous work, we
+In the present work, we investigate linguistic variation
+across different online communities. In contrast to previous work, we
 investigate variation across speech communities (rather than
@@
 Second, we want to find out how, and how much, CCLMs are attuned to
-the specificity of communities. We test whether 1. the linguistic
+the specificity of communities. We test whether 1.\ the linguistic
 models are correlated with a model of communities based only on
-co-occurrence of users between communities and 2. if latent community
+co-occurrence of users between communities and 2.\ if latent community
 classifiers based on the CCLM can accurately identify that a given
@@
 
-\section{Experimental setup}
-
-\subsection{Data sets}
-
-We investigate linguistic variation across various communities present
-on the social media website Reddit.\footnote{Comments were obtained
-  from the archive at \url{https://pushshift.io/}.
-  \cite{Baumgartner2020}.}
-%
-Reddit is divided into forums called \textit{subreddits}, 
-which are typically organized around a topic of interest. 
-Users create \textit{posts}, which consist of a link, image, 
-or text, along with a \emph{comment} section. 
-Comments are threaded: a comment can be made directly on a post,
-or appear as a reply to another comment.
-%
-Hereafter we refer to such comments as ``messages'', to match our
-convention in mathematical formulas: the letter $c$ stands for a
-community, and $m$ stands for a message.
-
-Our dataset includes messages from \num{46} English-language subreddits, 
-randomly selected from the set of all forums 
-with at least \num{15000} messages per month for each month
-in the three years between 2015 and 2017. 
-We initially selected \num{50} subreddits, 
-but excluded four of them from further analysis: 
-two because they were primarily non-English and two with particularly short average message lengths.\footnote{
-The macro-average message length was \num{39.3} ($\sigma=15.7$) and the two removed communities had average comment lengths of \num{10.7} and \num{18.1}.}
-% According to your footnote the removed communities are barely outliers. I wouldn't have removed them, but it is too late to do anything about that.
-Each community corpus consist of \num{50000} randomly selected messages from the year 2015.
-
-Messages were preprocessed as follows: 
-We excluded the content of block quotes, code blocks, and tables.
-We removed any markup (formatting) commands from the remaining text, extracting only rendered text.
-We tokenized the messages using the default English model for the SpaCy tokenizer Version 2.2.3 \citep{Honnibal2017}
-and lower-cased all tokens.
-
-\subsection{Models}
+\section{Community-conditioned language models}
 
@@
 size of the token to be embedding equal for all models.
+
+\subsection{Data sets}
+
+We investigate linguistic variation across various communities present
+on the social media website Reddit.\footnote{Comments were obtained
+  from the archive at \url{https://pushshift.io/}.
+  \cite{Baumgartner2020}.}
+%
+Reddit is divided into forums called \textit{subreddits}, 
+which are typically organized around a topic of interest. 
+Users create \textit{posts}, which consist of a link, image, 
+or text, along with a \emph{comment} section. 
+Comments are threaded: a comment can be made directly on a post,
+or appear as a reply to another comment.
+%
+Hereafter we refer to such comments as ``messages'', to match our
+convention in mathematical formulas: the letter $c$ stands for a
+community, and $m$ stands for a message.
+
+Our dataset includes messages from \num{510} subreddits, 
+sampled from the set of all forums 
+with at least \num{5000} messages per month for each month
+of the year 2015.
+Each community corpus consist of \num{42000} randomly selected messages from the year 2015.
+For each community, we reserve \num{1000} messages each for development and testing,
+leaving a total of \num{20.4}M messages for training.
+
+Messages were preprocessed as follows: 
+We excluded the content of block quotes, code blocks, and tables.
+We removed any markup (formatting) commands from the remaining text, extracting only rendered text.
+We tokenized the messages using the default English model for the SpaCy tokenizer Version 2.2.3 \citep{Honnibal2017}
+and lower-cased all tokens.
 
@@
 \begin{tikzpicture}
-  \infogainboxplot
+  \pplbycommbox
 \end{tikzpicture}
 \caption{
-  Distribution of information gain across communities for each CCLM 
-  (top: LSTM models, bottom: transformer models).
+  Distribution of $\Ppl_{M_j}$ for each model architecture.
 }

### 2021-05-04 results.py
results and analysis

@@ -1,0 +1,90 @@
+import pandas as pd
+from pathlib import Path
+import data
+import numpy as np
+import json
+
+def exp_normalize(x, axis):
+    """ https://timvieira.github.io/blog/post/2014/02/11/exp-normalize-trick/ """
+    b = x.max(axis=axis)
+    y = np.exp(x - b)
+    return y / y.sum(axis)
+
+results_agg_ppl = []
+results_comm_ppl = []
+
+model_family_dir = Path('model/reddit')
+fields = data.load_fields(model_family_dir)
+comms = fields['community'].vocab.itos
+
+for architecture in ('lstm', 'transformer'):
+    uncond_model_name = f'{architecture}-3'
+    uncond_model_dir = model_family_dir/uncond_model_name
+
+    # load the unconditioned model NLLs
+    uncond = pd.read_csv(uncond_model_dir/'nll.csv')\
+            .rename(columns={'nll': 'uncond_entr'})\
+            .drop('length', 1)
+
+    # repalce community indexes with names
+    uncond['community'] = uncond['community'].apply(lambda x:comms[x])
+
+    uncond_ppl = np.exp(uncond.uncond_entr.mean())
+    uncond_ppl_by_comm = uncond.groupby('community').uncond_entr.mean().apply(np.exp)
+    uncond_ppl_by_comm.name = uncond_model_name
+
+    results_comm_ppl.append(uncond_ppl_by_comm)
+    results_agg_ppl.append({
+            'model': uncond_model_name,
+            'test epoch': int(open(uncond_model_dir/'saved-epoch.txt').read()),
+            'Ppl': uncond_ppl,
+            'IG': None
+        })
+
+    for lc in range(4):
+        model_name = f'{architecture}-3-{lc}'
+        model_dir = model_family_dir/model_name
+        print(model_name)
+
+        # load conditioned NLLs (including off-community conditioning)
+        df = pd.read_csv(model_dir/'nll.csv')
+        # not sure why, but one row had missing values in lstm-3-0. possibly csv i/o problem?
+        # (example_id and length, unfortunately, so hard to replace...)
+        # drop it, but ensure it really is a limited problem
+        n = len(df)
+        df.dropna(inplace=True)
+        print(f"Dropped {n - len(df)} examples")
+        df.example_id = df.example_id.astype(int)
+        df.length = df.length.astype(int)
+        # repalce community indexes with names
[…]

### 2021-05-04 paper/paper.tex
new outline

@@
 \begin{abstract}
+  % Context (Optional)
+
+  % Problem?
+
+  - Can community-level linguistic variation be captured by statistical models?
+
+  - Is community-specific linguistic variation correlated with social
+    relationship between communities?
+
+  % Why it's interesting
+
+  % Results
+
+  % What follows from results? (Optional)
+
   Community-Conditioned Language Models (CCLMs) are an under-researched tool for 
@@
 \end{abstract}
+
+% Introduction
+
+% Data sets
+
+% Community Conditioned Language Models:
+%   Architecture (with emphasis on the embeddings)
+%   Training
+%   Results (loss and loss per community)
+
+% Latent bayesian community classifier
+%    Formula (no training!)
+%    Result 1: Confusion matrix
+%    Transition: too much data; can we extract a more useful metric?
+%    Indiscernibility: Formula, and table
+
+% Correlation between linguistic embedding with social embeddings
+%    Explain the problem (idea: comparing cosine similarities --> too much data, again)
+%    Sketch "procrustes" problem, explain what it yields. (Also what kind of similarity one gets with random input)
+%    Results:
+%      - Correlation between Kumar2018 and best LSTM
+%      - Correlation between Kumar2018 and best Transformer 
+%      - Correlation between Transformer and LSTM
+%      - Concatenations ???
 
@@
 \end{hypotheses}
-
-%\jp{Perhaps: Outline of the paper.}
 

### 2021-05-04 paper/paper.tex
restructure
some todos and outline of work

@@
 \begin{abstract}
+
+  TODO: fix this up.
   % Context (Optional)
@@
 
+% Community Conditioned Language Models:
 %   Data sets
-
-% Community Conditioned Language Models:
 %   Architecture (with emphasis on the embeddings)
@@
 % Correlation between linguistic embedding with social embeddings
+%    Kumar2018 Data set
 %    Explain the problem (idea: comparing cosine similarities --> too much data, again)
@@
 \end{hypotheses}
+
+TODO: outline of the paper.
 
@@
 
-\section{Results and analysis}
+\subsection{Results and analysis}
 
-\subsection{Losses of CCLM}
+\subsubsection{Losses of CCLM}
 
@@
 \num{66.43} for the LSTM model (see table 1 in the supplementary materials).
-
-\subsection{Analysis of CCLM community embeddings}
-
-As a qualitative assessment of the community embeddings, 
-we manually identified topic categories into which at least three
-of the selected communities fall.
-We plotted the two-component PCA projection of the embeddings,
-annotated with these categories (\cref{fig:comm-pca}).
-We see that there is some evidence of clustering according to
-the topic categories we identified,
-suggesting that the embedding space learned by the model is meaningful.
-
-\begin{figure*}[t]
-\begin{tikzpicture}
-  \begin{groupplot}[
-    group style={group size=2 by 1},
-    yticklabels={,,}, xticklabels={,,},
-    legend style={anchor=north, legend columns=1, font=\sffamily\tiny},
-  ]
-    \nextgroupplot
-    \commscatter{lstm-3-1_pca1}{lstm-3-1_pca2}
-    \nextgroupplot
-    \commscatter{transformer-3-3_pca1}{transformer-3-3_pca2}
-    \legend{General interest,Videogames,Technology,Sports,Female-focused,Other}
-  \end{groupplot}
-\end{tikzpicture}
-\caption{Community embedding PCA of an LSTM (left, $c=1$) and transformer (right, $c=3$) model.}
-\label{fig:comm-pca}
-\end{figure*}
-
-As a quantitative assessment, we compare the CCLM-learned community
-embedding with the community embedding created by \citet{Kumar2018},
-who generated them using using a negative-sampling optimization
-algorithm and the message-community co-occurrence matrix as ground
-truth, using data which spans January 2014 to April 2017.
-We refer the reader
-to \citet{Kumar2018}
-for details, but the important point is that no linguistic
-information is used to create these embeddings: they only depend on
-whether a given user participates in a given
-community. In sharp contrasts, CCLM community embeddings depend in no
-way on which user is the author of any given message: we only use the
-contents of messages, not authorship data.
-%
-We correlate the pairwise cosine similarity of the
-CCLM embeddings with the same metric on the membership embedding,
-restricted to the 46 communities chosen for our study
-(\cref{fig:pairwise-comm-sim}).
-
-\begin{figure}
-  \pairwisecommsim{snap_cos_sim}{lstm-3-1_cos_sim} 
-  \caption{Pairwise cosine similarity between community vectors for 
-    the membership-based embedding (x-axis) and
-  the LSTM CCLM embedding with $c=1$ (y-axis). 
-  Pairs from the same category are colored according to \cref{fig:comm-pca}.
-  We also mark pairs from different categories (\texttimes) and pairs where one or both
-  communities belong to the ``Other'' category (\textbigcircle).
-  }
-  \label{fig:pairwise-comm-sim}
-\end{figure}
-
-\begin{table}
-  \small
-  \centering
-  \input{floats/comm_sim.tex}
-  \caption{Pearson's $r$ between the pairwise similarity of community vectors
-  in the CCLM models and the membership-based embedding of \citet{Kumar2018}.
-  ($p<0.001$ for all models).
-  }
-  \label{tab:pairwise-comm-sim}
-\end{table}
 
@@
 
+\section{Social Analysis of CCLM community embeddings (TODO: better title)}
+
+- General idea; repeated.
+
+\subsection{Data set}
+We compare the CCLM-learned community
+embedding with the community embedding created by \citet{Kumar2018},
+who generated them using using a negative-sampling optimization
+algorithm and the message-community co-occurrence matrix as ground
+truth, using data which spans January 2014 to April 2017.
+We refer the reader
+to \citet{Kumar2018}
+for details, but the important point is that no linguistic
+information is used to create these embeddings: they only depend on
+whether a given user participates in a given
+community. In sharp contrasts, CCLM community embeddings depend in no
+way on which user is the author of any given message: we only use the
+contents of messages, not authorship data.
+%
+
+\subsection{Idea (JP)}
+% Explain the problem (idea: comparing cosine similarities --> too much data, again)
+
+correlate the pairwise cosine similarity of the
+CCLM embeddings with the same metric on the membership embedding,
+restricted to the 46 communities chosen for our study
+(\cref{fig:pairwise-comm-sim}).
+
+\subsection{Procrustes (JP)}
+
+%    Sketch "procrustes" problem, explain what it yields. (Also what kind of similarity one gets with random input)
+
+\subsection{Results}
+
+%    Results:
+%      - Correlation between Kumar2018 and best LSTM
+%      - Correlation between Kumar2018 and best Transformer 
+%      - Correlation between Transformer and LSTM
+%      - Concatenations ???
+
+\begin{figure*}[t]
+\begin{tikzpicture}
+  \begin{groupplot}[
+    group style={group size=2 by 1},
+    yticklabels={,,}, xticklabels={,,},
+    legend style={anchor=north, legend columns=1, font=\sffamily\tiny},
+  ]
+    \nextgroupplot
+    \commscatter{lstm-3-1_pca1}{lstm-3-1_pca2}
+    \nextgroupplot
+    \commscatter{transformer-3-3_pca1}{transformer-3-3_pca2}
+    \legend{General interest,Videogames,Technology,Sports,Female-focused,Other}
+  \end{groupplot}
+\end{tikzpicture}
+\caption{Plot of embeddings along the 2 principal dimensions of correlation between .... TODO of an LSTM (left, $c=1$) and transformer (right, $c=3$) model.
+Color correspond the 5-means clustering of .... TODO}
+\label{fig:comm-pca}
+\end{figure*}
+
 \section{Related work}

### 2021-05-04 paper/paper.tex
draft procrustes sectino

@@
 
-- General idea; repeated.
+In this section we investigate whether (and how much) the community
+embeddings obtained from the CCLMs correlate with the community-member
+relationship on reddit.
 
-\subsection{Data set}
-We compare the CCLM-learned community
+To this end, we compare the CCLM-learned community
 embedding with the community embedding created by \citet{Kumar2018},
@@
 
-\subsection{Idea (JP)}
+\subsection{Comparing embeddings: method}
 % Explain the problem (idea: comparing cosine similarities --> too much data, again)
 
-correlate the pairwise cosine similarity of the
-CCLM embeddings with the same metric on the membership embedding,
-restricted to the 46 communities chosen for our study
-(\cref{fig:pairwise-comm-sim}).
+Let us call \(L\) the matrix of (normed) linguistic embeddings and \(S\) the matrix
+of (normed) social embeddings. Thus \(L_i\) and \(S_i\) are the linguistic and
+social embeddings for community \(i\).
 
-\subsection{Procrustes (JP)}
+To compare embeddings, one might naively think that it suffices to
+calculate the distance between two embeddings index-wise, which is
+equal to the Frobenius distance between \(L\) and \(S\):
 
-%    Sketch "procrustes" problem, explain what it yields. (Also what kind of similarity one gets with random input)
+\[||L-S||_F = \sum_i (L_i - S_i)\]
+
+The problem with this approach is that the dimensions of \(L\) and \(S\) may
+not coincide. Applying a simple permutation of dimensions may widely
+change the correlation metric. To make the metric resilient to such
+effects, we can compute the minimum distance between \(L_i\) and \(S_i\), for
+any orthogonal matrix \(\Omega\) applied to \(L\):
+
+\[d(L,S) = \mathsf{argmin}_\Omega ||ΩL-S||_F\]
+
+Computing this distance is known as the Procrustes problem. Its
+solution is
+
+\[d(L,S) = \mathsf {Tr} Σ\], where the \(Σ\) is obtained by the singular value
+decomposition \(U^TΣV = LS^T\).  The vectors of \(U\) and \(V\) give
+the directions of correlation respectively of \(L\) and \(S\). The
+singular values \(\sigma_i\) (diagonal elements of \(Σ\)), say how much
+correlation there is between the directions \(U_i\) and \(V_i\).
+
+Further we can arrange the solution so that \(σ_i < σ_j\) iff \(i < j\). That
+is, the largest singular value \(σ_0\) corresponds to the principal
+direction of correlation (\(U_0\), \(V_0\)), \(σ_1\) to the second
+principal direction, etc.
+
+The \(d(L,S)\) metric varies between \(0\) (perfect independence) and
+\(n\) (perfect correlation, obtained for example if \(L=S\)).
+
+To get a sense of what value of \(d(L,S)\) indicates significant correlation,
+We have computed experimentally \(d(L,S)\) for 
+\(L\) and \(S\) chosen at random and 16 dimensions. We find
+a mean of TODO and a standard deviation of TODO.
+
+Thus \(d(L,S)\) above \(μ+3σ\) can be safely considered to indicate a
+statistically significant correlation between \(L\) and
+\(S\). However, below, we will also give a sense of \emph{how} the
+correlation is manifested by analysis of the two principal components
+of correlation in the linguistic embeddings, \(U_0\) and \(U_1\).
+
+(Another idea may be to correlate the cosine similarity between
+\((S_i,S_j)\) and \((L_i,L_j)\). But this would yield one number for
+each pair of communities (510 * 509 / 2): too much data to deal with,
+and it is hard to get a sense of the statistical significance of the
+resulting correlation coefficient.)
 
@@
 % LocalWords:  indiscernibility Vinyals Kalchbrenner OConnor
-% LocalWords:  monotonicity subreddit
+% LocalWords:  monotonicity subreddit Frobenius

### 2021-05-04 paper/paper.tex
fixes and tweaks

@@
 
-Computing this distance is known as the Procrustes problem. Its
+Computing this distance is known as the orthogonal Procrustes problem~\citep{Gower2004}. Its
 solution is
-
-\[d(L,S) = \mathsf {Tr} Σ\], where the \(Σ\) is obtained by the singular value
-decomposition \(U^TΣV = LS^T\).  The vectors of \(U\) and \(V\) give
+\[d(L,S) = \mathsf {Tr} Σ\] where the matrix \(Σ\) is obtained by the singular value
+decomposition \(U^TΣV = LS^T\). Additionally the vectors of \(U\) and \(V\) give
 the directions of correlation respectively of \(L\) and \(S\). The
@@
 
-Further we can arrange the solution so that \(σ_i < σ_j\) iff \(i < j\). That
+Further, we can arrange the solution so that \(σ_i < σ_j\) iff \(i < j\). That
 is, the largest singular value \(σ_0\) corresponds to the principal
-direction of correlation (\(U_0\), \(V_0\)), \(σ_1\) to the second
+directions of correlation (\(U_0\), \(V_0\)), \(σ_1\) to the second
 principal direction, etc.
@@
 The \(d(L,S)\) metric varies between \(0\) (perfect independence) and
-\(n\) (perfect correlation, obtained for example if \(L=S\)).
+\(n\) (perfect correlation, obtained for example if \(L=S\)), where
+$n$ is the number of community embeddings considered.
 
-To get a sense of what value of \(d(L,S)\) indicates significant correlation,
+To get a more precise sense of what value of \(d(L,S)\) indicates significant correlation,
 We have computed experimentally \(d(L,S)\) for 
@@
 
-Thus \(d(L,S)\) above \(μ+3σ\) can be safely considered to indicate a
+Thus \(d(L,S)\) below \(μ-3σ\) can be safely considered to indicate a
 statistically significant correlation between \(L\) and

### 2021-05-04 paper/paper.tex
notes

@@
 \newcommand\Ind{\mathsf{Ind}}
+\newcommand\LM{\mathrm{LM}}
+\newcommand\CCLM{\mathrm{CCLM}}
 \DeclareMathOperator*{\avg}{average}
@@
 
-TODO: outline of the paper.
+%TODO: outline of the paper.
 
@@
 standard $n$-layer neural sequence encoder (we set $n=3$ for all
-models). As conventional, word tokens are first converted to vectors
-using a trainable embedding layer. At the other end, word tokens are
-predicted using a softmax projection layer. The set of word tokens is
-a pre-determined vocabulary.  These encoders form the core of the
-of our models, as such do not take community into account, and we call them
-unconditioned models.
+models). As is convention, a sequence of word tokens is first converted to 
+vectors using a trainable embedding layer over a pre-deterimend vocabulary.
+At the other end, word tokens are predicted using a softmax projection 
+layer. What we have described so far does not take community into account
+and as such we call them \emph{uncodnitioned models}, but the same
+encoder architecture also forms the core of our conditioned models.
 
-To condition the model (and obtain community-conditioned language
-models --- CCLM), we add a \emph{community embedding} parameter to the
-models.  This parameter is concatenated with the hidden layer of the
-sequence encoder, for some layer $l_c \leq n$, and passed through a
+To condition the model , we add a \emph{community embedding} parameter.
+This parameter is concatenated with the hidden layer of the
+sequence encoder, at some layer $l_c \leq n$, and passed through a
 linear layer which projects the resulting vector back to the original
@@
 this reason, we set the hidden size of the sequence encoder and the
-size of the token to be embedding equal for all models.
+size of the token embedding to be equal for all models.
+
+%TODO: Diagram of model?
 
@@
 
-We investigate linguistic variation across various communities present
-on the social media website Reddit.\footnote{Comments were obtained
+We investigate linguistic variation across various communities 
+from the social media website Reddit.\footnote{Comments were obtained
   from the archive at \url{https://pushshift.io/}.
@@
 We excluded the content of block quotes, code blocks, and tables.
-We removed any markup (formatting) commands from the remaining text, extracting only rendered text.
-We tokenized the messages using the default English model for the SpaCy tokenizer Version 2.2.3 \citep{Honnibal2017}
-and lower-cased all tokens.
+We removed any markup (formatting) commands from the remaining text, 
+extracting only rendered text.
+We tokenized the messages using the default English model for the SpaCy tokenizer 
+Version 2.2.3 \citep{Honnibal2017}.
 
@@
 
-We combined the 46 community corpora and split the combined corpus,
-reserving 10\% for testing, 10\% for validation and the rest for training.
-Splits were stratified by community so that each community had \num{5000} examples
-for testing, \num{5000} for evaluation, and \num{40000} for training.
-
-In the remainder of the paper we use $M$ to refer to messages in the
-test set, and $M_j$ for the partition of the test set originating from
-community $c_j$.
-
 We use a vocabulary size of \num{40000} tokens, including a special
 out-of-vocabulary token.  The vocabulary consists of the most frequent
-tokens at the message level (across all communities), meaning that we
-only count each token once per message.  The main reason for this
-choice is to avoid over-counting otherwise rare tokens that were
-repeated many times in a single message.
+tokens across all communities.
 
@@
 smallest validation loss for testing.  In general, transformer models
-trained for about half as long as the LSTM models (test epoch,
-\cref{tab:model-results}).
+trained for about half as long as the LSTM models (\cref{tab:model-results}, 
+test epoch).
 
@@
 
-\subsubsection{Losses of CCLM}
+In this section, we report the performance of the conditioned and un-conditioned
+models on the test set.
+First, we define two performance metrics, perplexity and information gain.
+In the following, we use $M$ to refer to messages in the
+combined test set, and $M_j$ for the partition of the test set originating from 
+community $c_j$.
 
-We measure the perplexity of each message in the test set. 
-For this, we use the negative exponential of the mean cross-entropy loss 
-of the language model for a given message and average over a set of messages, $M$.
+\paragraph{Perplexity}
+
+For a given model let $H(m)$ be the model's cross-entropy loss,
+averaged over tokens in $m$.
+We define the perplexity on a set of messages, $M$,
+to be the exponential of the model's average cross-entropy loss:
 \[\Ppl_M = e^{\avg_{m\in M} H(m)}\]
-As one would expect, 
-the conditioned models have lower perplexity
-than their respective unconditioned baseline models (\cref{tab:model-results}),
-but this effect is different in magnitude
-for models with different community embedding depths.
-The best transformer model ($l_c=3$)
-incorporates the community information last,
-after all the sequence encoder layers.
+
+\paragraph{CCLM Information Gain}
+
+We also consider the information gain of each CCLM over its baseline
+un-conditioned counterpart with the same sequence encoder architecture.
+For a given message, information gain is defined as the difference
+between the cross-entropy of the unconditioned model and the conditioned model:
+\[H_{\mathrm{LM}}(m) - H_{\mathrm{CCLM}}(m)\]
+For a set of messages, $M$, we consider the average information gain
+in exponential space (as a ratio of perplexities):
+\[\IG_M = \frac{e^{\avg_M(H_{\mathrm{LM}}(m))}}{e^{\avg_M(H_{\mathrm{CCLM}}(m))}}\]
+%\jp{What we had before: average entropy gain is $H(m)-H_0(m)$, and
+  %then taking the average over all messages:
+  %$avg(H(m)-H_0(m)) = avg(H(m))-avg(H_0(m)) = H - H₀$. Taking the exponential
+  %gives a ratio of perplexity.  Here it seems that you have commuted
+  %exponential and average. It is rather weird to do average in the
+  %exponential space without normalising to a probability.
+  %Compare: $e^{avg(H(m))} / e^{avg(H_0(m))}$ with $avg \frac{e^{H(m)}}{e^{H₀(m)}}$.
+%}
+\bn{I think this is stated correctly now (not the average in exp space)}
+
+\paragraph{}
+
+Unsurprisingly, the conditioned models mostly have lower perplexity 
+than their respective unconditioned baseline models, 
+meaning information gain above one (\cref{tab:model-results}).
+Overall, the LSTM models perform better than the transformers, 
+although the best transformer has higher information gain than the best LSTM.
+
+The effect of $l_c$, the depth of the community embedding,
+is also different across architectures.
 For the LSTM encoder, 
-the model that concatenates the community embedding after the first encoder layer ($l_c=1$),
-performs almost as well as the model with $l_c=3$, but that model is still the best.
-There is much less variation in performance across different values of $l_c$ for the LSTM.
+the best model concatenates the community embedding after the first encoder layer ($l_c=1$),
+but all of the conditioned models perform similarly well.
+For the transformer, the best model incorporates the community information
+first, concatenating it directly to the word vectors ($l_c=0$).
+It performs similarly to the model that only integrates the community information
+after all all the transformer layers ($l_c=3$),
+but the two middle-layer models actually perform worse than the unconditioned model
+(with $\IG_M < 1$).
 
@@
 
-We also consider model performance stratified by community; that is,
-$\Ppl_{M_j}$, where $M_j$ is the test-set messages from community
-$c_j$.  We observe a lot of variation in baseline perplexity
-across communities, with $\Ppl_{M_j}$ ranging from \num{34.47} to
-\num{66.43} for the LSTM model (see table 1 in the supplementary materials).
+We also consider performance stratified by community; that is,
+$\Ppl_{M_j}$ and $\IG_{M_j}$, 
+where $M_j$ is the set of messages originating from community $c_j$ 
+(\cref{fig:comm-stratified-box}).
+We observe a lot of variation in baseline perplexity
+across communities, with $\Ppl_{M_j}$ ranging from \num{3.67} to
+\num{93.58} for the best LSTM model . %TODO make long table
+While the absolute performance ($\Ppl_{M_j}$) of the LSTM models is better,
+we jhat the best transformer models have somewhat higher information gain
+than their LSTM counterparts.
 
-\subsection{Information gain of the CCLM}
-
-Next, we consider the \emph{information gain} of each CCLM over its baseline
-un-conditioned counterpart with the same sequence encoder architecture.
-For a given message, information gain is defined as the difference
-between the cross-entropy of the unconditioned model and the conditioned model:
-\[H_{LM}(m) - H_{CCLM}(m)\]
-For a set of messages, $M$, we consider the average information gain
-in exponential space (as a ratio of perplexities):
-\[\IG_M = \frac{e^{\avg_M(H_{LM}(m))}}{e^{\avg_M(H_{CCLM}(m))}}\]
-%\jp{What we had before: average entropy gain is $H(m)-H_0(m)$, and
-  %then taking the average over all messages:
-  %$avg(H(m)-H_0(m)) = avg(H(m))-avg(H_0(m)) = H - H₀$. Taking the exponential
-  %gives a ratio of perplexity.  Here it seems that you have commuted
-  %exponential and average. It is rather weird to do average in the
-  %exponential space without normalising to a probability.
-
-  %Compare: $e^{avg(H(m))} / e^{avg(H_0(m))}$ with $avg \frac{e^{H(m)}}{e^{H₀(m)}}$.
-%}
-
-While the absolute performance of the LSTM models is better,
-we find that the best transformer CCLM has higher information gain 
-than the best LSTM (\cref{tab:model-results}).
+\newcommand{\modelboxplot}[3]{
+  \addplot+[
+    boxplot={draw position=#3, box extend=0.3}, 
+    draw=#2, mark=*, mark options={fill=#2, scale=0.5}, solid, fill=#2!10,
+    area legend] 
+    table [y=#1] {floats/comm.csv};
+}
+\begin{figure}[t]
+\begin{tikzpicture}
+  \begin{axis}[name=ig,
+    height=7cm,width=8cm,
+    boxplot/draw direction=x, axis x line*=bottom, axis y line*=left,y axis line style={opacity=0},
+    %xmax=1.6,
+    ylabel=$l_c$, xlabel=$\IG_{M_j}$,
+    legend style={font=\tiny,at={(1,1)},anchor=north east}, legend columns=2
+    ]
+    \addplot[thin, red, samples=2] coordinates {(1,-0.5)(1,3.5)};
+    \foreach \lc in {0, ..., 3}{
+      \modelboxplot{lstm-3-\lc_ig}{violet}{\lc+0.2}
+      \modelboxplot{transformer-3-\lc_ig}{teal}{\lc-0.2}
+    \legend{,LSTM, Transformer}
+    }
+  \end{axis}
+  \begin{axis}[name=ppl,
+    anchor=north west,at={($(ig.south west)-(0,1.1cm)$)},
+    width=8cm, height=2.5cm, 
+    boxplot/draw direction=x, axis x line*=bottom, hide y axis,
+    xlabel=$\Ppl_{M_j}$,
+    ]
+    \modelboxplot{lstm-3_ppl}{violet}{0.2}
+    \modelboxplot{transformer-3_ppl}{teal}{-.2}
+  \end{axis}
+\end{tikzpicture}
+\caption{
+   Average model performance by community. 
+   Also see table 1 in the supplementary materials
+   for a tabular version of these results with community names.
+}
+\label{fig:comm-stratified-box}
+\end{figure}
 
 The conditioned models also perform differently across different communities---%
-for some communities, $\IG_{M_j}$ around \num{1}, or even slightly lower,
-meaning that the CCLM has no advantage over the unconditioned baseline for that community.
-For other communities it is higher, meaning that the CCLM performs better than 
-the baseline for those communities (\cref{fig:comm-info-gain}).
+even among the best models, some communities have $\IG_{M_j} < 1$,
+meaning that the CCLM performs worse than the unconditioned baseline for 
+messages from that community.
+For other communities $\IG_{M_j}$ is much higher, meaning that the CCLM performs better  
+(\cref{fig:comm-info-gain}).\footnote{
+  Some of the communities with consistently high $IG_{M_j}$ 
+  across all models are primarily non-English, 
+  but surprisingly, not the three most extreme
+  outliers. There are /r/counting, /r/friendsafari, and /r/Fireteams, 
+  the later two of which are places where people coordinate to play video games
+  together. The coordination has highly strucutred and regular formats, which 
+  are presumably conventional to these communities.}
+
+The variation in information gain across communities is not 
+related to perplexity, however (\cref{fig:ppl-info-gain}).
 
@@
 \begin{tikzpicture}
-  \pplbycommbox
-\end{tikzpicture}
-\caption{
-  Distribution of $\Ppl_{M_j}$ for each model architecture.
-}
-\label{fig:comm-info-gain}
-\end{figure}
-
-This variation in information gain across communities is not 
-related to perplexity, however. 
-Some of the most linguistically complex communities benefit a lot
-from community conditioning, while others do not benefit at all. 
-Similarly, among the less linguistically complex communities 
-there some have high information gain, and some have information
-gain of \num{1} or less (\cref{fig:ppl-info-gain}).
-
-\begin{figure}
-\begin{tikzpicture}
-\begin{axis}
-\commscatter[]{lstm-3-1_lm_ppl}{lstm-3-1_info_gain}
+\begin{axis}[%
+    axis x line*=bottom, axis y line*=left,
+    xlabel=$\Ppl_{M_j}$, ylabel=$\IG_{M_j}$,
+    width=7cm,
+  ]
+  \addplot+[%
+    only marks,
+    mark=*,
+    mark options={scale=0.35, solid, color=violet!80, fill=violet},
+    font=\tiny,
+  ] table [x={lstm-3_ppl}, y={lstm-3-3_ig}] {floats/comm.csv};
 \end{axis}
@@
   Average community perplexity (x-axis) versus information gain (y-axis)
-  for the LSTM model with $l_c = 1$ (colors based on \cref{fig:comm-pca}).
+  for the best LSTM model ($l_c = 1$).
 }
@@
 
-\subsection{Latent language model-based community classifiers}
+\section{Latent language model-based community classifiers}
 
@@
 We observe that the diagonal dominates the matrix: the LMCC generally finds the right community for a message (\cref{fig:confusion}).
-%
+
 \begin{figure}
-\begin{tikzpicture}
-  \confusionsingleplot{lstm-3-1}
-\end{tikzpicture}
-\caption{LMCC confusion matrices for the LSTM model with $c=1$.}
+  \includegraphics[width=\linewidth]{floats/confusion_lstm-3-1}
+\caption{
+  Log-probability LMCC confusion matrices for the LSTM model with $l_c=1$. 
+  Communities are sorted by indiscernability.}
 \label{fig:confusion}
 \end{figure}
-%
+
 But we want a more quantitative measure, corresponding to the ability of the LMCC to correctly
@@
 between the indiscernibility measures by every pair of models with
-Pearson's $r \geq 0.98$ and $p < 0.0001$ (see, for example,
-\cref{fig:indisc-corr}).
-%
-\begin{figure}
-  \begin{tikzpicture}
-    \begin{axis}
-    \commscatter{lstm-3-1_indisc}{transformer-3-3_indisc}
-    \end{axis}
-  \end{tikzpicture}
-  \caption{%
-    Correlation between indiscernibility for 
-    LSTM ($c = 1$) and transformer ($c = 3$) models
-    (Pearson's $r = 0.99$, $p < 0.0001$).
-  }
-  \label{fig:indisc-corr}
-\end{figure}
-%
+Pearson's $r \geq 0.96$ and $p < 0.0001$.
+
 However, we observe that community-wise perplexity is only weakly (if at all) 
@@
 \end{table}
-%
+
 \begin{figure*}
@@
     legend style={anchor=north, legend columns=1, font=\tiny},
-    every node near coord/.append style={font=\sffamily\tiny}%
+    every node near coord/.append style={font=\sffamily\tiny},%
+    axis x line*=bottom, axis y line*=left,
   ]
-    \nextgroupplot
-    \commscatter[]{lstm-3-1_lm_ppl}{lstm-3-1_indisc}
-    \nextgroupplot
-    \commscatter[]{lstm-3-1_info_gain}{lstm-3-1_indisc}
-    \legend{General interest,Videogames,Technology,Sports,Female-focused,Other}
+  \nextgroupplot[xlabel=$\Ppl_{M_j}$, ylabel=$\Ind_j$]
+      \addplot+[%
+        only marks,
+        mark=*,
+        mark options={scale=0.35, solid, color=violet!80, fill=violet},
+        font=\tiny,
+      ] table [x={lstm-3-1_ppl}, y={lstm-3-1_indisc}] {floats/comm.csv};
+  \nextgroupplot[xlabel=$\IG_{M_j}$, ylabel=$\Ind_j$]
+      \addplot+[%
+        only marks,
+        mark=*,
+        mark options={scale=0.35, solid, color=violet!80, fill=violet},
+        font=\tiny,
+      ] table [x={lstm-3-1_ig}, y={lstm-3-1_indisc}] {floats/comm.csv};
   \end{groupplot}
@@
   \caption{%
-    Average community indiscernibility (y-axis), 
+    Community indiscernibility (y-axis), 
     in relationship to linguistic complexity (left) and 
@@
 
-Computing this distance is known as the orthogonal Procrustes problem~\citep{Gower2004}. Its
+Computing \(d(L,S)\) is known as the orthogonal Procrustes problem~\citep{Gower2004}. The
 solution is
@@
 decomposition \(U^TΣV = LS^T\). Additionally the vectors of \(U\) and \(V\) give
-the directions of correlation respectively of \(L\) and \(S\). The
-singular values \(\sigma_i\) (diagonal elements of \(Σ\)), say how much
+the directions of correlation respectively of \(L\) and \(S\). Each
+singular \(\sigma_i\) (elements of the diagonal matrix \(Σ\)), give a measure of how much
 correlation there is between the directions \(U_i\) and \(V_i\).
 
-Further, we can arrange the solution so that \(σ_i < σ_j\) iff \(i < j\). That
-is, the largest singular value \(σ_0\) corresponds to the principal
+Further, we can apply a permutation to \(U\),\(V\) and \(Σ\) such that \(σ_i < σ_j\) iff \(i < j\). That
+is, the largest singular value \(σ_0\) then corresponds to the principal
 directions of correlation (\(U_0\), \(V_0\)), \(σ_1\) to the second
@@
 
-The \(d(L,S)\) metric varies between \(0\) (perfect independence) and
-\(n\) (perfect correlation, obtained for example if \(L=S\)), where
-$n$ is the number of community embeddings considered.
+The \(d(L,S)\) metric range from \(0\) (corresponding to perfect
+independence) to \(n\) (corresponding to perfect correlation, obtained
+for example if \(L=S\)), where $n=510$ is the number of communities
+considered.
 
@@
 We have computed experimentally \(d(L,S)\) for 
-\(L\) and \(S\) chosen at random and 16 dimensions. We find
+\(L\) and \(S\) chosen as random 16 dimensions. We find
 a mean of TODO and a standard deviation of TODO.
 
-Thus \(d(L,S)\) below \(μ-3σ\) can be safely considered to indicate a
-statistically significant correlation between \(L\) and
+TODO: !!! we should take a 16 × 510 and a 300 × 510 matrix to test!!!
+
+Thus if \(d(L,S)\) is below \(μ-3σ\), we can safely assume that there
+is statistically significant correlation between \(L\) and
 \(S\). However, below, we will also give a sense of \emph{how} the
 correlation is manifested by analysis of the two principal components
-of correlation in the linguistic embeddings, \(U_0\) and \(U_1\).
+of correlation in the linguistic embeddings, \(U_0\) and \(U_1\). To
+do so we will plot the projection of embeddings on along \(U_0\) and
+\(U_1\), together with the corresponding singular values.
 
@@
 
-\begin{figure*}[t]
-\begin{tikzpicture}
-  \begin{groupplot}[
-    group style={group size=2 by 1},
-    yticklabels={,,}, xticklabels={,,},
-    legend style={anchor=north, legend columns=1, font=\sffamily\tiny},
-  ]
-    \nextgroupplot
-    \commscatter{lstm-3-1_pca1}{lstm-3-1_pca2}
-    \nextgroupplot
-    \commscatter{transformer-3-3_pca1}{transformer-3-3_pca2}
-    \legend{General interest,Videogames,Technology,Sports,Female-focused,Other}
-  \end{groupplot}
-\end{tikzpicture}
-\caption{Plot of embeddings along the 2 principal dimensions of correlation between .... TODO of an LSTM (left, $c=1$) and transformer (right, $c=3$) model.
-Color correspond the 5-means clustering of .... TODO}
-\label{fig:comm-pca}
-\end{figure*}
+%\begin{figure*}[t]
+%\begin{tikzpicture}
+  %\begin{groupplot}[
+    %group style={group size=2 by 1},
+    %yticklabels={,,}, xticklabels={,,},
+    %legend style={anchor=north, legend columns=1, font=\sffamily\tiny},
+  %]
+    %\nextgroupplot
+    %\commscatter{lstm-3-1_pca1}{lstm-3-1_pca2}
+    %\nextgroupplot
+    %\commscatter{transformer-3-3_pca1}{transformer-3-3_pca2}
+    %\legend{General interest,Videogames,Technology,Sports,Female-focused,Other}
+  %\end{groupplot}
+%\end{tikzpicture}
+%\caption{Plot of embeddings along the 2 principal dimensions of correlation between .... TODO of an LSTM (left, $c=1$) and transformer (right, $c=3$) model.
+%Color correspond the 5-means clustering of .... TODO}
+%\label{fig:comm-pca}
+%\end{figure*}
 

### 2021-05-13 paper/paper.tex
update CCLM results section

@@
 \num{93.58} for the best LSTM model . %TODO make long table
-While the absolute performance ($\Ppl_{M_j}$) of the LSTM models is better,
-we jhat the best transformer models have somewhat higher information gain
-than their LSTM counterparts.
+While the performance ($\Ppl_{M_j}$) of the LSTM models is better,
+we find that the best transformer CCLM has higher information gain 
+than the best LSTM.
 
@@
 }
+
 \begin{figure}[t]
@@
     boxplot/draw direction=x, axis x line*=bottom, axis y line*=left,y axis line style={opacity=0},
-    %xmax=1.6,
     ylabel=$l_c$, xlabel=$\IG_{M_j}$,
@@
   \begin{axis}[name=ppl,
-    anchor=north west,at={($(ig.south west)-(0,1.1cm)$)},
+    anchor=north west,at={($(ig.south west)-(0,1cm)$)},
     width=8cm, height=2.5cm, 
@@
 The conditioned models also perform differently across different communities---%
-even among the best models, some communities have $\IG_{M_j} < 1$,
-meaning that the CCLM performs worse than the unconditioned baseline for 
-messages from that community.
-For other communities $\IG_{M_j}$ is much higher, meaning that the CCLM performs better  
+for some communities, $\IG_{M_j}$ is around \num{1}, or even slightly lower,
+meaning that the CCLM has no advantage over the unconditioned baseline.
+For other communities it is much higher, meaning that the CCLM performs better 
 (\cref{fig:comm-info-gain}).\footnote{
-  Some of the communities with consistently high $IG_{M_j}$ 
-  across all models are primarily non-English, 
-  but surprisingly, not the three most extreme
-  outliers. There are /r/counting, /r/friendsafari, and /r/Fireteams, 
-  the later two of which are places where people coordinate to play video games
-  together. The coordination has highly strucutred and regular formats, which 
-  are presumably conventional to these communities.}
+  Some of the outlier communities with consistently high $IG_{M_j}$ 
+  across all models are primarily non-English.}
 
-The variation in information gain across communities is not 
-related to perplexity, however (\cref{fig:ppl-info-gain}).
+This variation in information gain across communities is not 
+related to perplexity, however. 
+Some of the most linguistically complex communities benefit a lot
+from community conditioning, while others do not benefit at all. 
+Similarly, among the less linguistically complex communities 
+there some have high information gain, and some have information
+gain of \num{1} or less (\cref{fig:ppl-info-gain}).
 
-\begin{figure}
-\begin{tikzpicture}
-\begin{axis}[%
-    axis x line*=bottom, axis y line*=left,
-    xlabel=$\Ppl_{M_j}$, ylabel=$\IG_{M_j}$,
-    width=7cm,
-  ]
-  \addplot+[%
-    only marks,
-    mark=*,
-    mark options={scale=0.35, solid, color=violet!80, fill=violet},
-    font=\tiny,
-  ] table [x={lstm-3_ppl}, y={lstm-3-3_ig}] {floats/comm.csv};
-\end{axis}
-\end{tikzpicture}
-\caption{
-  Average community perplexity (x-axis) versus information gain (y-axis)
-  for the best LSTM model ($l_c = 1$).
-}
-\label{fig:ppl-info-gain}
-\end{figure}
+%\begin{figure}
+%\begin{tikzpicture}
+%\begin{axis}
+%\commscatter[]{lstm-3-1_ppl}{lstm-3-1_ig}
+%\end{axis}
+%\end{tikzpicture}
+%\caption{
+  %Average community perplexity (x-axis) versus information gain (y-axis)
+  %for the LSTM model with $l_c = 1$ (colors based on \cref{fig:comm-pca}).
+%}
+%\label{fig:ppl-info-gain}
+%\end{figure}
 
@@
 
-\section{Latent language model-based community classifiers}
+\subsection{Latent language model-based community classifiers}
 
@@
 We observe that the diagonal dominates the matrix: the LMCC generally finds the right community for a message (\cref{fig:confusion}).
-
-\begin{figure}
-  \includegraphics[width=\linewidth]{floats/confusion_lstm-3-1}
-\caption{
-  Log-probability LMCC confusion matrices for the LSTM model with $l_c=1$. 
-  Communities are sorted by indiscernability.}
-\label{fig:confusion}
-\end{figure}
-
+% TODO: new results
+%\begin{figure}
+%\begin{tikzpicture}
+  %\confusionsingleplot{lstm-3-1}
+%\end{tikzpicture}
+%\caption{LMCC confusion matrices for the LSTM model with $c=1$.}
+%\label{fig:confusion}
+%\end{figure}
+%
 But we want a more quantitative measure, corresponding to the ability of the LMCC to correctly
@@
 between the indiscernibility measures by every pair of models with
-Pearson's $r \geq 0.96$ and $p < 0.0001$.
-
+Pearson's $r \geq 0.98$ and $p < 0.0001$ (see, for example,
+\cref{fig:indisc-corr}).
+%
+%\begin{figure}
+  %\begin{tikzpicture}
+    %\begin{axis}
+    %\commscatter{lstm-3-1_indisc}{transformer-3-3_indisc}
+    %\end{axis}
+  %\end{tikzpicture}
+  %\caption{%
+    %Correlation between indiscernibility for 
+    %LSTM ($c = 1$) and transformer ($c = 3$) models
+    %(Pearson's $r = 0.99$, $p < 0.0001$).
+  %}
+  %\label{fig:indisc-corr}
+%\end{figure}
+%
 However, we observe that community-wise perplexity is only weakly (if at all) 
@@
 \end{table}
-
-\begin{figure*}
-  \begin{tikzpicture}
-  \begin{groupplot}[
-    group style={group size=2 by 1},
-    legend style={anchor=north, legend columns=1, font=\tiny},
-    every node near coord/.append style={font=\sffamily\tiny},%
-    axis x line*=bottom, axis y line*=left,
-  ]
-  \nextgroupplot[xlabel=$\Ppl_{M_j}$, ylabel=$\Ind_j$]
-      \addplot+[%
-        only marks,
-        mark=*,
-        mark options={scale=0.35, solid, color=violet!80, fill=violet},
-        font=\tiny,
-      ] table [x={lstm-3-1_ppl}, y={lstm-3-1_indisc}] {floats/comm.csv};
-  \nextgroupplot[xlabel=$\IG_{M_j}$, ylabel=$\Ind_j$]
-      \addplot+[%
-        only marks,
-        mark=*,
-        mark options={scale=0.35, solid, color=violet!80, fill=violet},
-        font=\tiny,
-      ] table [x={lstm-3-1_ig}, y={lstm-3-1_indisc}] {floats/comm.csv};
-  \end{groupplot}
-  \end{tikzpicture}
-  \caption{%
-    Community indiscernibility (y-axis), 
-    in relationship to linguistic complexity (left) and 
-    CCLM information gain (right). Shown here for the LSTM
-    model with $l_c=1$. See \cref{tab:ind-corrs} for 
-    Pearson's r correlations for all models.
-  }
-  \label{fig:cclm-lmcc-ppl}
-\end{figure*}
+%
+%\begin{figure*}
+  %\begin{tikzpicture}
+  %\begin{groupplot}[
+    %group style={group size=2 by 1},
+    %legend style={anchor=north, legend columns=1, font=\tiny},
+    %every node near coord/.append style={font=\sffamily\tiny}%
+  %]
+    %\nextgroupplot
+    %\commscatter[]{lstm-3-1_ppl}{lstm-3-1_indisc}
+    %\nextgroupplot
+    %\commscatter[]{lstm-3-1_ig}{lstm-3-1_indisc}
+    %\legend{General interest,Videogames,Technology,Sports,Female-focused,Other}
+  %\end{groupplot}
+  %\end{tikzpicture}
+  %\caption{%
+    %Average community indiscernibility (y-axis), 
+    %in relationship to linguistic complexity (left) and 
+    %CCLM information gain (right). Shown here for the LSTM
+    %model with $l_c=1$. See \cref{tab:ind-corrs} for 
+    %Pearson's r correlations for all models.
+  %}
+  %\label{fig:cclm-lmcc-ppl}
+%\end{figure*}
 
@@
 
-Computing \(d(L,S)\) is known as the orthogonal Procrustes problem~\citep{Gower2004}. The
+Computing this distance is known as the orthogonal Procrustes problem~\citep{Gower2004}. Its
 solution is
@@
 decomposition \(U^TΣV = LS^T\). Additionally the vectors of \(U\) and \(V\) give
-the directions of correlation respectively of \(L\) and \(S\). Each
-singular \(\sigma_i\) (elements of the diagonal matrix \(Σ\)), give a measure of how much
+the directions of correlation respectively of \(L\) and \(S\). The
+singular values \(\sigma_i\) (diagonal elements of \(Σ\)), say how much
 correlation there is between the directions \(U_i\) and \(V_i\).
 
-Further, we can apply a permutation to \(U\),\(V\) and \(Σ\) such that \(σ_i < σ_j\) iff \(i < j\). That
-is, the largest singular value \(σ_0\) then corresponds to the principal
+Further, we can arrange the solution so that \(σ_i < σ_j\) iff \(i < j\). That
+is, the largest singular value \(σ_0\) corresponds to the principal
 directions of correlation (\(U_0\), \(V_0\)), \(σ_1\) to the second
@@
 
-The \(d(L,S)\) metric range from \(0\) (corresponding to perfect
-independence) to \(n\) (corresponding to perfect correlation, obtained
-for example if \(L=S\)), where $n=510$ is the number of communities
-considered.
+The \(d(L,S)\) metric varies between \(0\) (perfect independence) and
+\(n\) (perfect correlation, obtained for example if \(L=S\)), where
+$n$ is the number of community embeddings considered.
 
@@
 We have computed experimentally \(d(L,S)\) for 
-\(L\) and \(S\) chosen as random 16 dimensions. We find
+\(L\) and \(S\) chosen at random and 16 dimensions. We find
 a mean of TODO and a standard deviation of TODO.
 
-TODO: !!! we should take a 16 × 510 and a 300 × 510 matrix to test!!!
-
-Thus if \(d(L,S)\) is below \(μ-3σ\), we can safely assume that there
-is statistically significant correlation between \(L\) and
+Thus \(d(L,S)\) below \(μ-3σ\) can be safely considered to indicate a
+statistically significant correlation between \(L\) and
 \(S\). However, below, we will also give a sense of \emph{how} the
 correlation is manifested by analysis of the two principal components
-of correlation in the linguistic embeddings, \(U_0\) and \(U_1\). To
-do so we will plot the projection of embeddings on along \(U_0\) and
-\(U_1\), together with the corresponding singular values.
+of correlation in the linguistic embeddings, \(U_0\) and \(U_1\).
 

### 2021-05-13 paper/plot_macros.tex
update CCLM results section

--- 2020-11-24 paper/plot_macros.tex
@@
   scatter src=explicit symbolic,
-  scatter/classes={
-    general-interest={mark=*, blue},
-    videogames={mark=*, red},
-    technology={mark=*, green},
-    sports={mark=*, cyan},
-    female-focused={mark=*, magenta},
-    other={mark=o, black}%
-  },
   nodes near coords*={#1},
@@
   visualization depends on={value \thisrow{community} \as \community},
-  ] table [x={#2}, y={#3}, meta=category] {floats/comm.csv};
+  ] table [x={#2}, y={#3}] {floats/comm.csv};
 }
@@
         scatter src=explicit symbolic,
-        scatter/classes={
-          general-interest={mark=*, blue},
-          videogames={mark=*, red},
-          technology={mark=*, green},
-          sports={mark=*, cyan},
-          female-focused={mark=*, magenta},
-          other={mark=o, black},
-          different={mark=x}
-        },
-        ] table [x=#1, y=#2, meta=category] {floats/comm_comm.csv};
+        ] table [x=#1, y=#2] {floats/comm_comm.csv};
     \end{axis}
@@
   \addplot+[boxplot,draw=blue,mark=*,solid]
-  table [y=lstm-3-0_info_gain] {floats/comm.csv};
+  table [y=lstm-3-0_ig] {floats/comm.csv};
   \addplot+[boxplot,draw=blue,mark=*,solid]
-  table [y=lstm-3-1_info_gain] {floats/comm.csv};
+  table [y=lstm-3-1_ig] {floats/comm.csv};
   \addplot+[boxplot,draw=blue,mark=*,solid]
-  table [y=lstm-3-2_info_gain] {floats/comm.csv};
+  table [y=lstm-3-2_ig] {floats/comm.csv};
   \addplot+[boxplot,draw=blue,mark=*,solid]
-  table [y=lstm-3-3_info_gain] {floats/comm.csv};
+  table [y=lstm-3-3_ig] {floats/comm.csv};
   \nextgroupplot
   \addplot+[boxplot,draw=red,mark=*,solid]
-  table [y=transformer-3-0_info_gain] {floats/comm.csv};
+  table [y=transformer-3-0_ig] {floats/comm.csv};
   \addplot+[boxplot,draw=red,mark=*,solid]
-  table [y=transformer-3-1_info_gain] {floats/comm.csv};
+  table [y=transformer-3-1_ig] {floats/comm.csv};
   \addplot+[boxplot,draw=red,mark=*,solid]
-  table [y=transformer-3-2_info_gain] {floats/comm.csv};
+  table [y=transformer-3-2_ig] {floats/comm.csv};
   \addplot+[boxplot,draw=red,mark=*,solid]
-  table [y=transformer-3-3_info_gain] {floats/comm.csv};
+  table [y=transformer-3-3_ig] {floats/comm.csv};
   \end{groupplot}
 }
+
+% https://tex.stackexchange.com/questions/183778/how-can-several-boxplots-be-combined-into-a-single-figure-as-groups/183856#183856

### 2021-05-13 paper/paper.tex
some more work on results

@@
 \num{93.58} for the best LSTM model . %TODO make long table
-While the performance ($\Ppl_{M_j}$) of the LSTM models is better,
-we find that the best transformer CCLM has higher information gain 
-than the best LSTM.
+While the absolute performance ($\Ppl_{M_j}$) of the LSTM models is better,
+we jhat the best transformer models have somewhat higher information gain
+than their LSTM counterparts.
 
@@
 }
-
 \begin{figure}[t]
@@
     boxplot/draw direction=x, axis x line*=bottom, axis y line*=left,y axis line style={opacity=0},
+    %xmax=1.6,
     ylabel=$l_c$, xlabel=$\IG_{M_j}$,
@@
   \begin{axis}[name=ppl,
-    anchor=north west,at={($(ig.south west)-(0,1cm)$)},
+    anchor=north west,at={($(ig.south west)-(0,1.1cm)$)},
     width=8cm, height=2.5cm, 
@@
 The conditioned models also perform differently across different communities---%
-for some communities, $\IG_{M_j}$ is around \num{1}, or even slightly lower,
-meaning that the CCLM has no advantage over the unconditioned baseline.
-For other communities it is much higher, meaning that the CCLM performs better 
+even among the best models, some communities have $\IG_{M_j} < 1$,
+meaning that the CCLM performs worse than the unconditioned baseline for 
+messages from that community.
+For other communities $\IG_{M_j}$ is much higher, meaning that the CCLM performs better  
 (\cref{fig:comm-info-gain}).\footnote{
@@
 
-This variation in information gain across communities is not 
-related to perplexity, however. 
-Some of the most linguistically complex communities benefit a lot
-from community conditioning, while others do not benefit at all. 
-Similarly, among the less linguistically complex communities 
-there some have high information gain, and some have information
-gain of \num{1} or less (\cref{fig:ppl-info-gain}).
+The variation in information gain across communities is not 
+related to perplexity, however (\cref{fig:ppl-info-gain}).
 
-%\begin{figure}
-%\begin{tikzpicture}
-%\begin{axis}
-%\commscatter[]{lstm-3-1_ppl}{lstm-3-1_ig}
-%\end{axis}
-%\end{tikzpicture}
-%\caption{
-  %Average community perplexity (x-axis) versus information gain (y-axis)
-  %for the LSTM model with $l_c = 1$ (colors based on \cref{fig:comm-pca}).
-%}
-%\label{fig:ppl-info-gain}
-%\end{figure}
+\begin{figure}
+\begin{tikzpicture}
+\begin{axis}[%
+    axis x line*=bottom, axis y line*=left,
+    xlabel=$\Ppl_{M_j}$, ylabel=$\IG_{M_j}$,
+    width=7cm,
+  ]
+  \addplot+[%
+    only marks,
+    mark=*,
+    mark options={scale=0.35, solid, color=violet!80, fill=violet},
+    font=\tiny,
+  ] table [x={lstm-3_ppl}, y={lstm-3-3_ig}] {floats/comm.csv};
+\end{axis}
+\end{tikzpicture}
+\caption{
+  Average community perplexity (x-axis) versus information gain (y-axis)
+  for the best LSTM model ($l_c = 1$).
+}
+\label{fig:ppl-info-gain}
+\end{figure}
 

### 2021-05-13 paper/plot_macros.tex
some more work on results

@@
 \usepackage{lmodern}
-
-\newcommand{\commscatter}[3][\community]{%
-  \addplot[
-  scatter, only marks,
-  font=\tiny,
-  scatter src=explicit symbolic,
-  nodes near coords*={#1},
-  nodes near coords align={vertical},
-  every node near coord/.append style={font=\sffamily\tiny},
-  visualization depends on={value \thisrow{community} \as \community},
-  ] table [x={#2}, y={#3}] {floats/comm.csv};
-}
 

### 2021-05-14 paper/paper.tex
a couple of cclm embedding things

@@
 whether a given user participates in a given
-community. In sharp contrasts, CCLM community embeddings depend in no
+community. In sharp contrast, CCLM community embeddings depend in no
 way on which user is the author of any given message: we only use the
@@
 \(L\) and \(S\) chosen at random and 16 dimensions. We find
-a mean of TODO and a standard deviation of TODO.
+a mean of \num{73.83} and a standard deviation of \num{1.89}.
 
-Thus \(d(L,S)\) below \(μ-3σ\) can be safely considered to indicate a
+Thus \(d(L,S)\) below \(μ-3σ\) (\num{68.15}) can be safely considered to indicate a \bn{above  \(μ+3σ\) (\num{79.50})?}
 statistically significant correlation between \(L\) and

### 2021-05-14 paper/paper.tex

@@
 We have computed experimentally \(d(L,S)\) for 
-\(L\) and \(S\) chosen at random and 16 dimensions. We find
-a mean of \num{73.83} and a standard deviation of \num{1.89}.
+\(L\) and \(S\) chosen as random 16 dimensions. We find
+a mean of TODO and a standard deviation of TODO.
 
-Thus \(d(L,S)\) below \(μ-3σ\) (\num{68.15}) can be safely considered to indicate a \bn{above  \(μ+3σ\) (\num{79.50})?}
-statistically significant correlation between \(L\) and
+TODO: !!! we should take a 16 × 510 and a 300 × 510 matrix to test!!!
+
+Thus if \(d(L,S)\) is below \(μ-3σ\), we can safely assume that there \bn{above  \(μ+3σ\) (\num{79.50})?}
+is statistically significant correlation between \(L\) and
 \(S\). However, below, we will also give a sense of \emph{how} the
 correlation is manifested by analysis of the two principal components
-of correlation in the linguistic embeddings, \(U_0\) and \(U_1\).
+of correlation in the linguistic embeddings, \(U_0\) and \(U_1\). To
+do so we will plot the projection of embeddings on along \(U_0\) and
+\(U_1\), together with the corresponding singular values.
 

### 2021-05-14 paper/paper.tex
typography + minor fixes and questions

@@
 \usepackage{latexsym}
+\usepackage{newunicodechar}
+\newunicodechar{Σ}{\ensuremath{\Sigma}} % 
+\newunicodechar{σ}{\ensuremath{\mathnormal{\sigma}}} % 
+\newunicodechar{μ}{\ensuremath{\mathnormal{\mu}}} % 
+\newunicodechar{×}{\ensuremath{\times}} %
 
@@
 intuitively assume that linguistically poor communities are easier to
-recognize: the less linguistic diversity, the more specific the
-language becomes and the easier it is to identify.
+recognize: the less linguistic diversity, \jp{easiest it is to model computationally ...}
+% the more specific the language becomes and
+the easier it is to identify.
 
@@
 \caption{
-  Log-probability LMCC confusion matrices for the LSTM model with $l_c=1$. 
-  Communities are sorted by indiscernability.}
+  Log-probability LMCC confusion matrix for the LSTM model with $l_c=1$. 
+  Communities are sorted by indiscernability. (The first communities predominantly use a language other than English, and are thus very discernable.)}
 \label{fig:confusion}
@@
 solution is
-\[d(L,S) = \mathsf {Tr} Σ\] where the matrix \(Σ\) is obtained by the singular value
+\[d(L,S) = \mathsf {Tr} (Σ)\] where the matrix \(Σ\) is obtained by the singular value
 decomposition \(U^TΣV = LS^T\). Additionally the vectors of \(U\) and \(V\) give
 the directions of correlation respectively of \(L\) and \(S\). Each
-singular \(\sigma_i\) (elements of the diagonal matrix \(Σ\)), give a measure of how much
+singular value \(\sigma_i\) (the elements of the diagonal matrix \(Σ\)), give a measure of how much
 correlation there is between the directions \(U_i\) and \(V_i\).
 
-Further, we can apply a permutation to \(U\),\(V\) and \(Σ\) such that \(σ_i < σ_j\) iff \(i < j\). That
-is, the largest singular value \(σ_0\) then corresponds to the principal
+Further, we can apply a permutation to \(U\), \(V\) and \(Σ\) such that \(σ_i < σ_j\) iff \(i < j\). Doing so, the largest singular value \(σ_0\) corresponds to the principal
 directions of correlation (\(U_0\), \(V_0\)), \(σ_1\) to the second
@@
 
-The \(d(L,S)\) metric range from \(0\) (corresponding to perfect
+The \(d(L,S)\) metric ranges from \(0\) (corresponding to perfect
 independence) to \(n\) (corresponding to perfect correlation, obtained
@@
 \(L\) and \(S\) chosen as random 16 dimensions. We find
-a mean of TODO and a standard deviation of TODO.
+a mean \(μ_d\)=TODO and a standard deviation \(σ_d\)=TODO.\jp{Unfortunate overloading}
 
@@
 \((S_i,S_j)\) and \((L_i,L_j)\). But this would yield one number for
-each pair of communities (510 * 509 / 2): too much data to deal with,
+each pair of communities \((510 × 509 / 2)\): too much data to deal with,
 and it is hard to get a sense of the statistical significance of the

### 2021-05-14 paper/paper.tex
minor changes and fixes

@@
 \usepackage{newunicodechar}
+\newunicodechar{ }{{\,}}
+\newunicodechar{¬}{\ensuremath{\neg}} % 
+\newunicodechar{²}{^2}
+\newunicodechar{³}{^3}
+\newunicodechar{·}{\ensuremath{\cdot}}
+\newunicodechar{¹}{^1}
+\newunicodechar{×}{\ensuremath{\times}} % 
+\newunicodechar{÷}{\ensuremath{\div}} % 
+\newunicodechar{ʳ}{^r}
+\newunicodechar{ˡ}{^l}
+\newunicodechar{̷}{\not} % 
+\newunicodechar{Γ}{\ensuremath{\Gamma}}   %
+\newunicodechar{Δ}{\ensuremath{\Delta}} % 
+\newunicodechar{Η}{\ensuremath{\textrm{H}}} % 
+\newunicodechar{Θ}{\ensuremath{\Theta}} % 
+\newunicodechar{Λ}{\ensuremath{\Lambda}} % 
+\newunicodechar{Ξ}{\ensuremath{\Xi}} % 
+\newunicodechar{Π}{\ensuremath{\Pi}}   %
 \newunicodechar{Σ}{\ensuremath{\Sigma}} % 
+\newunicodechar{Φ}{\ensuremath{\Phi}} % 
+\newunicodechar{Ψ}{\ensuremath{\Psi}} % 
+\newunicodechar{Ω}{\ensuremath{\Omega}} % 
+\newunicodechar{α}{\ensuremath{\mathnormal{\alpha}}}
+\newunicodechar{α}{\ensuremath{\mathnormal{\alpha}}} % 
+\newunicodechar{β}{\ensuremath{\beta}} % 
+\newunicodechar{γ}{\ensuremath{\mathnormal{\gamma}}} % 
+\newunicodechar{δ}{\ensuremath{\mathnormal{\delta}}} % 
+\newunicodechar{ε}{\ensuremath{\mathnormal{\varepsilon}}} % 
+\newunicodechar{ζ}{\ensuremath{\mathnormal{\zeta}}} % 
+\newunicodechar{η}{\ensuremath{\mathnormal{\eta}}} % 
+\newunicodechar{θ}{\ensuremath{\mathnormal{\theta}}} % 
+\newunicodechar{ι}{\ensuremath{\mathnormal{\iota}}} % 
+\newunicodechar{κ}{\ensuremath{\mathnormal{\kappa}}} % 
+\newunicodechar{λ}{\ensuremath{\mathnormal{\lambda}}} % 
+\newunicodechar{μ}{\ensuremath{\mathnormal{\mu}}} % 
+\newunicodechar{ν}{\ensuremath{\mathnormal{\mu}}} % 
+\newunicodechar{ξ}{\ensuremath{\mathnormal{\xi}}} % 
+\newunicodechar{π}{\ensuremath{\mathnormal{\pi}}}
+\newunicodechar{π}{\ensuremath{\mathnormal{\pi}}} % 
+\newunicodechar{ρ}{\ensuremath{\mathnormal{\rho}}} % 
 \newunicodechar{σ}{\ensuremath{\mathnormal{\sigma}}} % 
-\newunicodechar{μ}{\ensuremath{\mathnormal{\mu}}} % 
-\newunicodechar{×}{\ensuremath{\times}} %
+\newunicodechar{τ}{\ensuremath{\mathnormal{\tau}}} % 
+\newunicodechar{φ}{\ensuremath{\mathnormal{\varphi}}} % 
+\newunicodechar{χ}{\ensuremath{\mathnormal{\chi}}} % 
+\newunicodechar{ψ}{\ensuremath{\mathnormal{\psi}}} % 
+\newunicodechar{ω}{\ensuremath{\mathnormal{\omega}}} %  
+\newunicodechar{ϕ}{\ensuremath{\mathnormal{\phi}}} % 
+\newunicodechar{ϵ}{\ensuremath{\mathnormal{\epsilon}}} % 
+\newunicodechar{ᵏ}{^k}
+\newunicodechar{ᵢ}{_i} % 
+\newunicodechar{ }{{\,}}
+\newunicodechar{′}{\ensuremath{^\prime}}  % 
+\newunicodechar{″}{\ensuremath{^\second}} % 
+\newunicodechar{‴}{\ensuremath{^\third}}  % 
+\newunicodechar{ⁱ}{^i}
+\newunicodechar{⁵}{\ensuremath{^5}}
+\newunicodechar{⁺}{\ensuremath{^+}} %% 
+\newunicodechar{⁺}{^+}
+\newunicodechar{⁻}{\ensuremath{^-}} %% 
+\newunicodechar{ⁿ}{^n}
+\newunicodechar{₀}{\ensuremath{_0}} % 
+\newunicodechar{₁}{\ensuremath{_1}}
+\newunicodechar{₁}{_1}
+\newunicodechar{₂}{\ensuremath{_2}}
+\newunicodechar{₂}{_2}
+\newunicodechar{₃}{\ensuremath{_3}}
+\newunicodechar{₊}{\ensuremath{_+}} %% 
+\newunicodechar{₋}{\ensuremath{_-}} %% 
+\newunicodechar{ℂ}{\ensuremath{\mathbb{C}}} % 
+\newunicodechar{ℕ}{\mathbb{N}} % 
+\newunicodechar{ℚ}{\ensuremath{\mathbb{Q}}}
+\newunicodechar{ℝ}{\ensuremath{\mathbb{R}}} % 
+\newunicodechar{ℤ}{\ensuremath{\mathbb{Z}}} % 
+\newunicodechar{ℳ}{\mathscr{M}}
+\newunicodechar{⅋}{\ensuremath{\parr}} % 
+\newunicodechar{←}{\ensuremath{\leftarrow}} % 
+\newunicodechar{↑}{\ensuremath{\uparrow}} % 
+\newunicodechar{→}{\ensuremath{\rightarrow}} %
+\newunicodechar{↔}{\ensuremath{\leftrightarrow}} % 
+\newunicodechar{↖}{\nwarrow} %
+\newunicodechar{↗}{\nearrow} %
+\newunicodechar{↝}{\ensuremath{\leadsto}}
+\newunicodechar{↦}{\ensuremath{\mapsto}}
+\newunicodechar{⇆}{\ensuremath{\leftrightarrows}} % 
+\newunicodechar{⇐}{\ensuremath{\Leftarrow}} % 
+\newunicodechar{⇒}{\ensuremath{\Rightarrow}} % 
+\newunicodechar{⇔}{\ensuremath{\Leftrightarrow}} % 
+\newunicodechar{∀}{\ensuremath{\forall}}   %
+\newunicodechar{∂}{\ensuremath{\partial}}
+\newunicodechar{∃}{\ensuremath{\exists}} % 
+\newunicodechar{∅}{\ensuremath{\varnothing}} % 
+\newunicodechar{∈}{\ensuremath{\in}}
+\newunicodechar{∉}{\ensuremath{\not\in}} % 
+\newunicodechar{∋}{\ensuremath{\ni}}  % 
+\newunicodechar{∎}{\ensuremath{\qed}} % end of proof. % or: {\blacksquare} 
+\newunicodechar{∏}{\prod}
+\newunicodechar{∑}{\sum}
+\newunicodechar{∗}{\ensuremath{\ast}} % 
+\newunicodechar{∘}{\ensuremath{\circ}} % 
+\newunicodechar{∙}{\ensuremath{\bullet}} % 
+\newunicodechar{∙}{\ensuremath{\cdot}}
+\newunicodechar{∞}{\ensuremath{\infty}} % 
+\newunicodechar{∣}{\ensuremath{\mid}} % 
+\newunicodechar{∧}{\ensuremath{\wedge}}% 
+\newunicodechar{∨}{\ensuremath{\vee}}% 
+\newunicodechar{∩}{\ensuremath{\cap}} % 
+\newunicodechar{∪}{\ensuremath{\cup}} % 
+\newunicodechar{∫}{\int}
+\newunicodechar{∷}{::} % proportion
+\newunicodechar{∼}{\ensuremath{\sim}} % 
+\newunicodechar{≃}{\ensuremath{\simeq}} % 
+\newunicodechar{≅}{\ensuremath{\cong}} % 
+\newunicodechar{≈}{\ensuremath{\approx}} % 
+\newunicodechar{≜}{\ensuremath{\stackrel{\scriptscriptstyle {\triangle}}{=}}} 
+\newunicodechar{≜}{\triangleq}
+\newunicodechar{≝}{\ensuremath{\stackrel{\scriptscriptstyle {\text{def}}}{=}}}
+\newunicodechar{≟}{\ensuremath{\stackrel {_\text{\textbf{?}}}{\text{\textbf{=}}\negthickspace\negthickspace\text{\textbf{=}}}}} % or {\ensuremath{\stackrel{\scriptscriptstyle ?}{=}}}
+\newunicodechar{≠}{\ensuremath{\neq}}% 
+\newunicodechar{≡}{\ensuremath{\equiv}}% 
+\newunicodechar{≤}{\ensuremath{\le}} % 
+\newunicodechar{≥}{\ensuremath{\ge}} % 
+\newunicodechar{⊂}{\ensuremath{\subset}} % 
+\newunicodechar{⊃}{\ensuremath{\supset}} %  
+\newunicodechar{⊆}{\ensuremath{\subseteq}} %  
+\newunicodechar{⊇}{\ensuremath{\supseteq}} % 
+\newunicodechar{⊎}{\ensuremath{\uplus}} %
+\newunicodechar{⊑}{\ensuremath{\sqsubseteq}} % 
+\newunicodechar{⊒}{\ensuremath{\sqsupseteq}} % 
+\newunicodechar{⊓}{\ensuremath{\sqcap}} % 
+\newunicodechar{⊔}{\ensuremath{\sqcup}} % 
+\newunicodechar{⊕}{\ensuremath{\oplus}} % 
+\newunicodechar{⊗}{\ensuremath{\otimes}} % 
+\newunicodechar{⊢}{\ensuremath{\vdash}} % 
+\newunicodechar{⊥}{\ensuremath{\bot}} % bottom
+\newunicodechar{⊧}{\models} % 
+\newunicodechar{⊨}{\models} % 
+\newunicodechar{⊩}{\Vdash}
+\newunicodechar{⊸}{\ensuremath{\multimap}} % 
+\newunicodechar{⋁}{\ensuremath{\bigvee}}
+\newunicodechar{⋃}{\ensuremath{\bigcup}} % 
+\newunicodechar{⋄}{\ensuremath{\diamond}} % 
+\newunicodechar{⋅}{\ensuremath{\cdot}}
+\newunicodechar{⋆}{\ensuremath{\star}} % 
+\newunicodechar{⋮}{\ensuremath{\vdots}} % 
+\newunicodechar{⋯}{\ensuremath{\cdots}} % 
+\newunicodechar{─}{---}
+\newunicodechar{■}{\ensuremath{\blacksquare}} % black square
+\newunicodechar{□}{\ensuremath{\square}} % 
+\newunicodechar{▹}{\ensuremath{\rhd}} % 
+\newunicodechar{◃}{\triangleleft{}}
+\newunicodechar{◅}{\ensuremath{\triangleleft{}}}
+\newunicodechar{◇}{\ensuremath{\diamond}} % 
+\newunicodechar{◽}{\ensuremath{\square}}
+\newunicodechar{★}{\ensuremath{\star}}   %
+\newunicodechar{♭}{\ensuremath{\flat}} % 
+\newunicodechar{♯}{\ensuremath{\sharp}} % 
+\newunicodechar{✓}{\ensuremath{\checkmark}} % 
+\newunicodechar{⟂}{\ensuremath{^\bot}} % PERPENDICULAR 
+\newunicodechar{⟦}{\ensuremath{\llbracket}}
+\newunicodechar{⟧}{\ensuremath{\rrbracket}}
+\newunicodechar{⟨}{\ensuremath{\langle}} % 
+\newunicodechar{⟩}{\ensuremath{\rangle}} % 
+\newunicodechar{⟶}{{\longrightarrow}}
+\newunicodechar{⟷} {\ensuremath{\leftrightarrow}}
+\newunicodechar{⟹}{\ensuremath{\Longrightarrow}} % 
+\newunicodechar{𝒟}{\ensuremath{\mathcal{D}}} % 
+\newunicodechar{𝒢}{\ensuremath{\mathcal{G}}}
+\newunicodechar{𝒦}{\ensuremath{\mathcal{K}}} % 
+\newunicodechar{𝒫}{\ensuremath{\mathcal{P}}}
+\newunicodechar{𝔸}{\ensuremath{\mathds{A}}} % 
+\newunicodechar{𝔹}{\ensuremath{\mathds{B}}} % 
+\newunicodechar{𝟙}{\ensuremath{\mathds{1}}} % 
+\newunicodechar{ℒ}{\ensuremath{\mathscr{L}}}
+\newunicodechar{⊤}{\ensuremath{\top}}
+\newunicodechar{ }{\quad}
+\newunicodechar{⊛}{\ensuremath{\circledast}}
 
@@
 intuitively assume that linguistically poor communities are easier to
-recognize: the less linguistic diversity, \jp{easiest it is to model computationally ...}
-% the more specific the language becomes and
-the easier it is to identify.
+recognize: that is the easiest it is to model computationally
+the easier it is to identify.\jp{Do we actually do this?}
 
@@
 \[\IG_M = \frac{e^{\avg_M(H_{\mathrm{LM}}(m))}}{e^{\avg_M(H_{\mathrm{CCLM}}(m))}}\]
-%\jp{What we had before: average entropy gain is $H(m)-H_0(m)$, and
-  %then taking the average over all messages:
-  %$avg(H(m)-H_0(m)) = avg(H(m))-avg(H_0(m)) = H - H₀$. Taking the exponential
-  %gives a ratio of perplexity.  Here it seems that you have commuted
-  %exponential and average. It is rather weird to do average in the
-  %exponential space without normalising to a probability.
-  %Compare: $e^{avg(H(m))} / e^{avg(H_0(m))}$ with $avg \frac{e^{H(m)}}{e^{H₀(m)}}$.
-%}
-\bn{I think this is stated correctly now (not the average in exp space)}
-
-\paragraph{}
-
+\[\IG_M = {e^{\avg_{m∈M}(H_{\mathrm{LM}}(m) - H_{\mathrm{CCLM}}(m))}}\]
+%
 Unsurprisingly, the conditioned models mostly have lower perplexity 
@@
   \caption{
-    Performance of baseline (first row for each encoder architecture) and CCLM models. 
-    Perplexity and information gain are shown for the entire test set 
-    (\num{5000} messages for each community).}
+    Performance of baseline (first row for each encoder architecture)
+    and CCLM models.  The scope of perplexity and information gain
+    ($M$) is the entire test set, i.e. $5000×510$ messages; \num{5000}
+    for each community.}
   \label{tab:model-results}
@@
   Average community perplexity (x-axis) versus information gain (y-axis)
-  for the best LSTM model ($l_c = 1$).
+  for the best LSTM model ($l_c = 1$). TODO: correlation factor
 }
@@
       ] table [x={lstm-3-1_ppl}, y={lstm-3-1_indisc}] {floats/comm.csv};
-  \nextgroupplot[xlabel=$\IG_{M_j}$, ylabel=$\Ind_j$]
+  \nextgroupplot[xlabel=$\IG_{M_j}$, ylabel=$\Ind_j$, xmax=1.6]
       \addplot+[%
@@
     Pearson's r correlations for all models.
+    (Communities with \(\IG > 1.6\) (3 outliers) are hidden to make the correlation easier to observe.)
   }

### 2021-05-14 paper/paper.tex
thinking about these numbers

@@
 solution is
-\[d(L,S) = \mathsf {Tr} (Σ)\] where the matrix \(Σ\) is obtained by the singular value
+\[d(L,S) = n - \mathsf {Tr} (Σ)\] where the matrix \(Σ\) is obtained by the singular value
 decomposition \(U^TΣV = LS^T\). Additionally the vectors of \(U\) and \(V\) give
@@
 
-TODO: !!! we should take a 16 × 510 and a 300 × 510 matrix to test!!!
-
-Thus if \(d(L,S)\) is below \(μ-3σ\), we can safely assume that there \bn{above  \(μ+3σ\) (\num{79.50})?}
+TODO: !!! we should take the correlation of a random 16 × 510 with the web embeddings !!!
+
+Thus if \(d(L,S)\) is below \(μ-3σ\), we can safely assume that there % below, because we have a distance rather than correlation in this convention.
 is statistically significant correlation between \(L\) and

### 2021-05-14 paper/paper.tex
minor updates

@@
 \newcommand\BibTeX{B\textsc{ib}\TeX}
-
+\newcommand\subreddit[1]{\texttt{/r/#1}}
 \newlist{hypotheses}{enumerate}{3}
@@
 We do so by investigating two questions.  First, we are interested in
-variation in the level of linguistic complexity exhibited by different
+variation in the level of linguistic complexity\jp{maybe revise terminology? Perhaps define linguistic complexity as perplexity?} exhibited by different
 communities.
@@
 
-%TODO: outline of the paper.
+\jp{outline of the paper is important here}
 
@@
 convention in mathematical formulas: the letter $c$ stands for a
-community, and $m$ stands for a message.
+community, and $m$ stands for a message. \jp{what does \(p\) stand for?}
 
 Our dataset includes messages from \num{510} subreddits, 
-sampled from the set of all forums 
+sampled\jp{what is the size of the whole population? why 510?} from the set of all forums 
 with at least \num{5000} messages per month for each month
@@
 
-Messages were preprocessed as follows: 
+Messages were preprocessed as follows. 
 We excluded the content of block quotes, code blocks, and tables.
@@
 
-For the LSTM, we trained the models on a simple left-to-right language
-modeling task with cross-entropy loss over the vocabulary.  Because
+For the LSTM architecture, we trained the models on a simple left-to-right language
+modeling task with cross-entropy loss.  Because
 the transformer operates on all tokens in the sequence at once, the
@@
 \citep{Loshchilov2019} optimization algorithm, with an initial
-learning rate of \num{0.001}, without any extra control on the decay
+learning rate of \num{0.001}, with no extra control on the decay
 of learning rate.
@@
 smallest validation loss for testing.  In general, transformer models
-trained for about half as long as the LSTM models (\cref{tab:model-results}, 
+were trained for about half as long as the LSTM models (\cref{tab:model-results}, 
 test epoch).
@@
 
-We also consider the information gain of each CCLM over its baseline
-un-conditioned counterpart with the same sequence encoder architecture.
+We also consider the average information gain per token of the CCLM over its baseline
+un-conditioned counterpart, with the same sequence encoder architecture.
 For a given message, information gain is defined as the difference
@@
 across communities, with $\Ppl_{M_j}$ ranging from \num{3.67} to
-\num{93.58} for the best LSTM model . %TODO make long table
+\num{93.58} for the best LSTM model\jp{Is this the cclm or the baseline?}. %TODO make long table
 While the absolute performance ($\Ppl_{M_j}$) of the LSTM models is better,
-we jhat the best transformer models have somewhat higher information gain
+the best transformer models have somewhat higher information gain
 than their LSTM counterparts.
@@
   but surprisingly, not the three most extreme
-  outliers. There are /r/counting, /r/friendsafari, and /r/Fireteams, 
+  outliers. There are \subreddit{counting}, \subreddit{friendsafari}, and \subreddit{Fireteams}, 
   the later two of which are places where people coordinate to play video games
-  together. The coordination has highly strucutred and regular formats, which 
+  together. The coordination has highly structured and regular formats, which 
   are presumably conventional to these communities.}
@@
 tested, communities where conditioning has the least effect tend to be
-more organized around more general interest topics, such as the
-\emph{relationships} and \emph{advice} subreddits, where the subject
+more organized around more general interest topics, such as 
+\subreddit{relationships} and \subreddit{advice}, where the subject
 matter is relevant to a broad range of people.  Conditioning the model
@@
 \section{Latent language model-based community classifiers}
-
-The perplexity of a CCLM gives us a measure of linguistic complexity,
+\jp{Perhaps use ``Latent Bayesian classifier'' instead.}
+The perplexity of a CCLM gives us a measure of linguistic complexity\jp{terminology?},
 but it is not a measure of how linguistically \emph{specific} the
@@
 which we call a language model-based community classifier
-(LMCC). Using the Bayes theorem, we can turn a
-CCLM into a LMCC: the probability that a given message $m$
+(LMCC).
+
+So far, we have only loaded the community embedding which corresponds
+to the actual community that the message originates from. By
+evaluating the model's loss (and thus probability) for each potential community,
+we can apply Bayesian reasoning to extract a latent community classifier for each message.
+The probability that a given message $m$
 belongs to a community $c_i$ is calculated as follows:
@@
 
-We can characterize the quality the LMCC with a confusion matrix
+We can characterize the quality\jp{performance} of the LMCC with a confusion matrix
 $C$, defined by $C_{ij} = P(c=c_i \mid m)$ for a given (test set) message $m$ 
@@
 between the indiscernibility measures by every pair of models with
-Pearson's $r \geq 0.96$ and $p < 0.0001$.
+Pearson's $r \geq 0.96$ and $Bp < 0.0001$.
 
@@
 
+\jp{Conclusion: perplexity (linguistic complexity) is not correlated
+  to how recognisable a community is. Presumably (in general) some very specific
+  lexical items are characteristic of a community.}
+
 \section{Social Analysis of CCLM community embeddings (TODO: better title)}
@@
 embeddings obtained from the CCLMs correlate with the community-member
-relationship on reddit.
+relationship on Reddit.
 
@@
 for details, but the important point is that no linguistic
-information is used to create these embeddings: they only depend on
-whether a given user participates in a given
-community. In sharp contrast, CCLM community embeddings depend in no
+information is used to create these embeddings: they only reflect the social
+relationship between communities, via 
+the user-community membership relation. In contrast, CCLM community embeddings depend in no
 way on which user is the author of any given message: we only use the
@@
 decomposition \(U^TΣV = LS^T\). Additionally the vectors of \(U\) and \(V\) give
-the directions of correlation respectively of \(L\) and \(S\). Each
+the directions of correlation respectively of \(L\) and \(S\). That is, each
 singular value \(\sigma_i\) (the elements of the diagonal matrix \(Σ\)), give a measure of how much
@@
 
-Further, we can apply a permutation to \(U\), \(V\) and \(Σ\) such that \(σ_i < σ_j\) iff \(i < j\). Doing so, the largest singular value \(σ_0\) corresponds to the principal
-directions of correlation (\(U_0\), \(V_0\)), \(σ_1\) to the second
-principal direction, etc.
+Further, we can apply a permutation to \(U\), \(V\) and \(Σ\) such
+that \(σ_i > σ_j\) iff \(i < j\). Doing so, the largest singular value
+\(σ_0\) corresponds to the principal directions of correlation
+(\(U_0\), \(V_0\)), \(σ_1\) to the second principal direction, etc.
 
@@
 
-Thus if \(d(L,S)\) is below \(μ-3σ\), we can safely assume that there % below, because we have a distance rather than correlation in this convention.
+Thus if \(d(L,S)\) is below \(μ_d-3σ_d\), we can safely assume that there % below, because we have a distance rather than correlation in this convention.
 is statistically significant correlation between \(L\) and
@@
 each pair of communities \((510 × 509 / 2)\): too much data to deal with,
-and it is hard to get a sense of the statistical significance of the
-resulting correlation coefficient.)
+and it is hard to get a sense of the meaning, let alone statistical significance of the
+resulting correlation coefficient.)\jp{Not sure if this discussion has its place here.}
 

### 2021-05-14 paper/paper.tex
text

@@
 
+Finally, we note that, even though community-conditioning affects
+perplexity very little overall, it provides large benefits in term of
+the underlying Bayesian classifier. Indiscernibility is on average
+TODO with a standard deviation of TODO. Roughly speaking, this means
+that, while the linguistic variations does not allow our model to
+pinpoint the originating community of a message, it can often
+eliminate TODO (a percentatge?) of communities as possible locations
+for it. Further more, the information gain provided by
+community-conditioning is a good predictor of how specific the
+community is.
+
 % Finally, we investigate the correlation between the stability of a
@@
 
+\jp{I propose that most of the following analysis should be spread between the relevant sections, and only have a tiny summary here + future work.}
 We find that various communities exhibit widely varying levels of

### 2021-05-14 paper/paper.tex
legend
reword

@@
   Average model performance by community.
-   Also see table 1 in the supplementary materials
+  The boxes indicate 90 percent confidence intervals. Bars indicate ... TODO:. The outliers are shown as dots.
+   See also table 1 in the supplementary materials
    for a tabular version of these results with community names.
@@
 \[C_{ij} = \avg_{m\in M_j}P(c=c_i \mid m)\]
-We observe that the diagonal dominates the matrix: the LMCC generally finds the right community for a message (\cref{fig:confusion}).
+We observe that the diagonal is clearly above the rest: on average the LMCC generally
+picks assigns the highest probability to the community where a message originates (\cref{fig:confusion}).
 

### 2021-05-18 paper/paper.tex

@@
 \caption{
-  Log-probability LMCC confusion matrix for the LSTM model with $l_c=1$. 
-  Communities are sorted by indiscernability. (The first communities predominantly use a language other than English, and are thus very discernable.)}
+  LMCC confusion matrix for the LSTM model with $l_c=1$, showing $\log(P(c_j\mid c_i))$, 
+  with $c_i$ on the x-axis and $c_j$ on the y-axis.
+  Communities are sorted by indiscernability.
+  (The first few communities predominantly use a language other than English, and are thus very discernable.)}
 \label{fig:confusion}
@@
 
-The \(d(L,S)\) metric ranges from \(0\) (corresponding to perfect
-independence) to \(n\) (corresponding to perfect correlation, obtained
-for example if \(L=S\)), where $n=510$ is the number of communities
-considered.
+The \(d(L,S)\) metric ranges from \(0\) 
+(corresponding to perfect correlation, 
+obtained for example if \(L=S\)) 
+to \(n\) 
+(corresponding to perfect independence), 
+where $n=510$ is the number of communities considered.
 
 To get a more precise sense of what value of \(d(L,S)\) indicates significant correlation,
-We have computed experimentally \(d(L,S)\) for 
-\(L\) and \(S\) chosen as random 16 dimensions. We find
-a mean \(μ_d\)=TODO and a standard deviation \(σ_d\)=TODO.\jp{Unfortunate overloading}
-
-TODO: !!! we should take the correlation of a random 16 × 510 with the web embeddings !!!
-
-Thus if \(d(L,S)\) is below \(μ_d-3σ_d\), we can safely assume that there % below, because we have a distance rather than correlation in this convention.
+We have experimentally computed \(d(L,S)\) for five random  ``embeddings'' 
+with the same shape as the CCLM embeddings.
+We observed a mean of \(μ_d=431.39\) and a (Bessel's-corrected) 
+standard deviation \(s_d=3.94\) in their distance from the social embedding, \(S\).
+%\jp{Unfortunate overloading} luckily it seems that sample stdev is often written with an $s$ to differentiate from population stdev
+
+Thus if \(d(L,S)\) is below \(419.56\)
+($\mu_d - 3s_d$), 
+we can safely assume that there 
 is statistically significant correlation between \(L\) and
-\(S\). However, below, we will also give a sense of \emph{how} the
+\(S\). 
+Indeed, this is very clearly the case (\cref{tab:embedding-correlations}),
+meaning that the CCLM embedding capture some of the same information as the
+social-network embeddings.
+We also give a sense of \emph{how} the
 correlation is manifested by analysis of the two principal components
 of correlation in the linguistic embeddings, \(U_0\) and \(U_1\). To
-do so we will plot the projection of embeddings on along \(U_0\) and
+do so we plot the projection of embeddings on along \(U_0\) and
 \(U_1\), together with the corresponding singular values.
+
+\begin{table}
+\centering
+\begin{tabular}{llcc}
+\toprule
+                        &   & LSTM     & Transformer \\
+\midrule
+  \multirow{4}{*}{$l_c$}  & 0 &         254.06  (44.99) &         239.41  (49.09)  \\
+                          & 1 &         245.14  (47.25) & \textbf{232.18} (50.38)   \\
+                          & 2 &         249.17  (46.23) &         233.47  (50.21)  \\
+                          & 3 & \textbf{241.13} (48.27) &         237.74   (49.13) \\
+\bottomrule
+\end{tabular}
+\caption{Distance between CCLM embeddings and the social network-based embedding
+of \citet{Kumar2018}, as measured by $d(L,S)$. In parentheses is the number of standard deviations
+from the mean distance of our random embedding sample.} \label{tab:embedding-correlations}
+\end{table}
 

### 2021-05-28 paper/paper.tex
with 10 random samples

@@
 We observed a mean of \(μ_d=431.39\) and a (Bessel's-corrected) 
-standard deviation \(s_d=3.94\) in their distance from the social embedding, \(S\).
+standard deviation \(s_d=2.90\) in their distance from the social embedding, \(S\).
 %\jp{Unfortunate overloading} luckily it seems that sample stdev is often written with an $s$ to differentiate from population stdev
 
-Thus if \(d(L,S)\) is below \(419.56\)
+Thus if \(d(L,S)\) is below \(422.79\)
 ($\mu_d - 3s_d$), 
@@
 \midrule
-  \multirow{4}{*}{$l_c$}  & 0 &         254.06  (44.99) &         239.41  (49.09)  \\
-                          & 1 &         245.14  (47.25) & \textbf{232.18} (50.38)   \\
-                          & 2 &         249.17  (46.23) &         233.47  (50.21)  \\
-                          & 3 & \textbf{241.13} (48.27) &         237.74   (49.13) \\
+  \multirow{4}{*}{$l_c$}  & 0 &         254.06  (61.21) &         239.41  (66.79)  \\
+                          & 1 &         245.14  (64.29) & \textbf{232.18} (68.54)   \\
+                          & 2 &         249.17  (62.90) &         233.47  (68.32)  \\
+                          & 3 & \textbf{241.13} (65.67) &         237.74   (66.84) \\
 \bottomrule
@@
 of \citet{Kumar2018}, as measured by $d(L,S)$. In parentheses is the number of standard deviations
-from the mean distance of our random embedding sample.} \label{tab:embedding-correlations}
+from the mean distance of our random embedding samples.} \label{tab:embedding-correlations}
 \end{table}

### 2021-05-28 paper/paper.tex
wibble

@@
 %
-Hereafter we refer to such comments as ``messages'', to match our
+Hereafter we refer to such comments as ``messages'', matching our
 convention in mathematical formulas: the letter $c$ stands for a
-community, and $m$ stands for a message. \jp{what does \(p\) stand for?}
+community, and $m$ stands for a message.
 
@@
 belongs to a community $c_i$ is calculated as follows:
-\[P(c=c_i \mid m) = P(m | c_i)\frac {P(m)} {P(c=c_i)}\]
+\[P(c=c_i \mid m) = P(m \mid c_i)\frac {P(m)} {P(c=c_i)}\]
 In the above,

### 2021-05-28 paper/paper.tex
discussion and update

@@
 
-Not only is linguistic variation an object of inquiry in its own
-right, but it is also important to consider it in the development of
-robust, equitable language technology.  Previous work has shown that
-by taking demographic information into account, models can achieve
-better performance on various standard NLP tasks
-\citep{Hovy2015,Yang2017}, 
-%
-However, most of the previous computational work on linguistic variation 
+% Not only is linguistic variation an object of inquiry in its own
+% right, but it is also important to consider it in the development of
+% robust, equitable language technology.  Previous work has shown that
+% by taking demographic information into account, models can achieve
+% better performance on various standard NLP tasks
+% \citep{Hovy2015,Yang2017}. \jp{Move this to related work so we can get to the point quicker.}
+=
+Most of the previous computational work on linguistic variation 
 has considered variation at the level of macro-social categories, such as gender
@@
 We do so by investigating two questions.  First, we are interested in
-variation in the level of linguistic complexity\jp{maybe revise terminology? Perhaps define linguistic complexity as perplexity?} exhibited by different
-communities.
+variation in the level of linguistic complexity\jp{maybe revise
+  terminology? Perhaps define linguistic complexity as perplexity?}
+exhibited by different communities.
 %\jp{
@@
 define various Community Conditioned Language Models (CCLM for short),
-which can be modeled as an estimator $M$ of the degree to which a
-given message $m$ is acceptable in a given community $c$:
-\(P_M(m \mid c)\).
+which can be modeled as the degree to which a
+given message $m$ is acceptable in a given community $c$, by estimating the probability
+\(P(m \mid c)\).
 
@@
 intuitively assume that linguistically poor communities are easier to
-recognize: that is the easiest it is to model computationally
-the easier it is to identify.\jp{Do we actually do this?}
+recognize: that is, the easiest it is to model computationally
+the easier it is to identify.
 
@@
 Our dataset includes messages from \num{510} subreddits, 
-sampled\jp{what is the size of the whole population? why 510?} from the set of all forums 
+which all subreddits 
 with at least \num{5000} messages per month for each month
@@
   \input{floats/ind_corrs.tex}
-  \caption{ Pearson's $r$ correlation coefficient between community
+  \caption{ (TODO: missing data) Pearson's $r$ correlation coefficient between community
     indiscernibility ($\Ind_j$) and two different predictors: CCLM
@@
 
-Thus if \(d(L,S)\) is below \(422.79\)
-($\mu_d - 3s_d$), 
-we can safely assume that there 
-is statistically significant correlation between \(L\) and
-\(S\). 
-Indeed, this is very clearly the case (\cref{tab:embedding-correlations}),
-meaning that the CCLM embedding capture some of the same information as the
-social-network embeddings.
-We also give a sense of \emph{how} the
-correlation is manifested by analysis of the two principal components
-of correlation in the linguistic embeddings, \(U_0\) and \(U_1\). To
-do so we plot the projection of embeddings on along \(U_0\) and
-\(U_1\), together with the corresponding singular values.
+Thus if \(d(L,S)\) is below the mean by several standard deviations, we can safely
+assume that there is statistically significant correlation between
+\(L\) and \(S\).  This is the case: we observe a difference between 61 and 68 standard deviations
+(\cref{tab:embedding-correlations}). Further, it is clear that the CCLM embedding
+capture some of the same information as the social-network embeddings.
+We also give a sense of \emph{how} the correlation is manifested by
+analysis of the two principal components of correlation in the
+linguistic embeddings, \(U_0\) and \(U_1\). To do so we plot the
+projection of embeddings on along \(U_0\) and \(U_1\), together with
+the corresponding singular values.\jp{TODO}
 
@@
 
-(Another idea may be to correlate the cosine similarity between
-\((S_i,S_j)\) and \((L_i,L_j)\). But this would yield one number for
-each pair of communities \((510 × 509 / 2)\): too much data to deal with,
-and it is hard to get a sense of the meaning, let alone statistical significance of the
-resulting correlation coefficient.)\jp{Not sure if this discussion has its place here.}
+% (Another idea may be to correlate the cosine similarity between
+% \((S_i,S_j)\) and \((L_i,L_j)\). But this would yield one number for
+% each pair of communities \((510 × 509 / 2)\): too much data to deal with,
+% and it is hard to get a sense of the meaning, let alone statistical significance of the
+% resulting correlation coefficient.)\jp{Not sure if this discussion has its place here.}
 
@@
 
-\jp{I propose that most of the following analysis should be spread between the relevant sections, and only have a tiny summary here + future work.}
-We find that various communities exhibit widely varying levels of
+% \jp{I propose that most of the following analysis should be spread between the relevant sections, and only have a tiny summary here + future work.}
+To sum up our findings, 
+we observed that various communities exhibit widely varying levels of
 linguistic complexity \ref{hyp:varying-complexity}, with perplexity
@@
 With indiscernibility values evenly ranging over the whole possible
-range (0 to 1), we find that CCLMs can be used to recognize certain
+range (1 to 510), we find that CCLMs can be used to recognize certain
 communities based on latent Bayesian classification
-\ref{hyp:LMCC-works}, but not all.  If generic communities exhibit
-distinctive linguistic patterns, none of our models is able to discern
-them. In contrast, the distinctive patterns of narrow-topic
-communities are just as recognizable by all our LSTM and Transformer
-models.
-
-We find no statistically significant correlation (\cref{tab:ind-corrs}) between the
-perplexity of the
-CCLM and the indiscernibility of the corresponding community
-\ref{hyp:rich-harder-to-identify}. Because we have relatively few, and
-only large communities in our test set, it is difficult to say if this
-finding generalizes. Regardless, this kind of correlation would suggest that
-community-specific complexity can be at least partially attributed to
-using more varied patterns. This is view is supported by the
-observation that, with an indiscernibility $> 0.9$, the most complex
-communities appear as an even mixture of the kind of language found in
-all communities, to our models.
+\ref{hyp:LMCC-works}, but not all.  We cannot rule out that generic
+communities exhibit distinctive linguistic patterns, but none of our
+CCLMs are able to discern them. In contrast, the distinctive patterns
+of narrow-topic communities are just as recognizable by all our LSTM
+and Transformer models.
+
+We find a mild correlation (\cref{tab:ind-corrs}) between the
+perplexity of the CCLM and the indiscernibility of the corresponding
+community \ref{hyp:rich-harder-to-identify}, with factors ranging from
+0.06 to 0.23. However, The trend is more clear when considering the
+correlation with information gain and indiscernibility (correlation
+from 0.39 to 0.46). So this means that, when trying to identify a
+community, it is not its absolute perplexity that matters, but rather
+the difference in perplexity with the baseline.
 
 We find that the layer at which the community embedding is taken into
-account in a CCLM influences its perplexity \ref{hyp:layer-effect},
+account in a CCLM influences its perplexity \ref{hyp:layer-effect},\jp{Delete this hypothesis?}
 albeit weakly. For LSTM models, the perplexity per word, averaged over
-messages from all communities, varies between \num{49.60} and \num{50.83}. (with \num{51.99} for the
+messages from all communities, varies between \num{49.60} and \num{50.83}. (with \num{51.99} for the\jp{update numbers?}
 unconditioned model) The perplexity per word is slightly higher for
@@
 could reveal differences in the kind of linguistic variation the
-different model architectures capture.
+different model architectures capture.\jp{update?}
 
@@
 \ref{hyp:extra-linguistic-correlation}. A natural
-question that may arise is how much this correlation can be explain by
+question that may arise is how much this correlation can be explained by
 user-specific linguistic patterns. We plan to control for this variable

### 2021-06-03 paper/paper.tex
remove p's from table columns
complexity terminology

@@
 \usepackage{newunicodechar}
-\newunicodechar{ }{{\,}}
-\newunicodechar{¬}{\ensuremath{\neg}} % 
-\newunicodechar{²}{^2}
-\newunicodechar{³}{^3}
-\newunicodechar{·}{\ensuremath{\cdot}}
-\newunicodechar{¹}{^1}
-\newunicodechar{×}{\ensuremath{\times}} % 
-\newunicodechar{÷}{\ensuremath{\div}} % 
-\newunicodechar{ʳ}{^r}
-\newunicodechar{ˡ}{^l}
-\newunicodechar{̷}{\not} % 
-\newunicodechar{Γ}{\ensuremath{\Gamma}}   %
-\newunicodechar{Δ}{\ensuremath{\Delta}} % 
-\newunicodechar{Η}{\ensuremath{\textrm{H}}} % 
-\newunicodechar{Θ}{\ensuremath{\Theta}} % 
-\newunicodechar{Λ}{\ensuremath{\Lambda}} % 
-\newunicodechar{Ξ}{\ensuremath{\Xi}} % 
-\newunicodechar{Π}{\ensuremath{\Pi}}   %
-\newunicodechar{Σ}{\ensuremath{\Sigma}} % 
-\newunicodechar{Φ}{\ensuremath{\Phi}} % 
-\newunicodechar{Ψ}{\ensuremath{\Psi}} % 
-\newunicodechar{Ω}{\ensuremath{\Omega}} % 
-\newunicodechar{α}{\ensuremath{\mathnormal{\alpha}}}
-\newunicodechar{α}{\ensuremath{\mathnormal{\alpha}}} % 
-\newunicodechar{β}{\ensuremath{\beta}} % 
-\newunicodechar{γ}{\ensuremath{\mathnormal{\gamma}}} % 
-\newunicodechar{δ}{\ensuremath{\mathnormal{\delta}}} % 
-\newunicodechar{ε}{\ensuremath{\mathnormal{\varepsilon}}} % 
-\newunicodechar{ζ}{\ensuremath{\mathnormal{\zeta}}} % 
-\newunicodechar{η}{\ensuremath{\mathnormal{\eta}}} % 
-\newunicodechar{θ}{\ensuremath{\mathnormal{\theta}}} % 
-\newunicodechar{ι}{\ensuremath{\mathnormal{\iota}}} % 
-\newunicodechar{κ}{\ensuremath{\mathnormal{\kappa}}} % 
-\newunicodechar{λ}{\ensuremath{\mathnormal{\lambda}}} % 
-\newunicodechar{μ}{\ensuremath{\mathnormal{\mu}}} % 
-\newunicodechar{ν}{\ensuremath{\mathnormal{\mu}}} % 
-\newunicodechar{ξ}{\ensuremath{\mathnormal{\xi}}} % 
-\newunicodechar{π}{\ensuremath{\mathnormal{\pi}}}
-\newunicodechar{π}{\ensuremath{\mathnormal{\pi}}} % 
-\newunicodechar{ρ}{\ensuremath{\mathnormal{\rho}}} % 
-\newunicodechar{σ}{\ensuremath{\mathnormal{\sigma}}} % 
-\newunicodechar{τ}{\ensuremath{\mathnormal{\tau}}} % 
-\newunicodechar{φ}{\ensuremath{\mathnormal{\varphi}}} % 
-\newunicodechar{χ}{\ensuremath{\mathnormal{\chi}}} % 
-\newunicodechar{ψ}{\ensuremath{\mathnormal{\psi}}} % 
-\newunicodechar{ω}{\ensuremath{\mathnormal{\omega}}} %  
-\newunicodechar{ϕ}{\ensuremath{\mathnormal{\phi}}} % 
-\newunicodechar{ϵ}{\ensuremath{\mathnormal{\epsilon}}} % 
-\newunicodechar{ᵏ}{^k}
-\newunicodechar{ᵢ}{_i} % 
-\newunicodechar{ }{{\,}}
-\newunicodechar{′}{\ensuremath{^\prime}}  % 
-\newunicodechar{″}{\ensuremath{^\second}} % 
-\newunicodechar{‴}{\ensuremath{^\third}}  % 
-\newunicodechar{ⁱ}{^i}
-\newunicodechar{⁵}{\ensuremath{^5}}
-\newunicodechar{⁺}{\ensuremath{^+}} %% 
-\newunicodechar{⁺}{^+}
-\newunicodechar{⁻}{\ensuremath{^-}} %% 
-\newunicodechar{ⁿ}{^n}
-\newunicodechar{₀}{\ensuremath{_0}} % 
-\newunicodechar{₁}{\ensuremath{_1}}
-\newunicodechar{₁}{_1}
-\newunicodechar{₂}{\ensuremath{_2}}
-\newunicodechar{₂}{_2}
-\newunicodechar{₃}{\ensuremath{_3}}
-\newunicodechar{₊}{\ensuremath{_+}} %% 
-\newunicodechar{₋}{\ensuremath{_-}} %% 
-\newunicodechar{ℂ}{\ensuremath{\mathbb{C}}} % 
-\newunicodechar{ℕ}{\mathbb{N}} % 
-\newunicodechar{ℚ}{\ensuremath{\mathbb{Q}}}
-\newunicodechar{ℝ}{\ensuremath{\mathbb{R}}} % 
-\newunicodechar{ℤ}{\ensuremath{\mathbb{Z}}} % 
-\newunicodechar{ℳ}{\mathscr{M}}
-\newunicodechar{⅋}{\ensuremath{\parr}} % 
-\newunicodechar{←}{\ensuremath{\leftarrow}} % 
-\newunicodechar{↑}{\ensuremath{\uparrow}} % 
-\newunicodechar{→}{\ensuremath{\rightarrow}} %
-\newunicodechar{↔}{\ensuremath{\leftrightarrow}} % 
-\newunicodechar{↖}{\nwarrow} %
-\newunicodechar{↗}{\nearrow} %
-\newunicodechar{↝}{\ensuremath{\leadsto}}
-\newunicodechar{↦}{\ensuremath{\mapsto}}
-\newunicodechar{⇆}{\ensuremath{\leftrightarrows}} % 
-\newunicodechar{⇐}{\ensuremath{\Leftarrow}} % 
-\newunicodechar{⇒}{\ensuremath{\Rightarrow}} % 
-\newunicodechar{⇔}{\ensuremath{\Leftrightarrow}} % 
-\newunicodechar{∀}{\ensuremath{\forall}}   %
-\newunicodechar{∂}{\ensuremath{\partial}}
-\newunicodechar{∃}{\ensuremath{\exists}} % 
-\newunicodechar{∅}{\ensuremath{\varnothing}} % 
-\newunicodechar{∈}{\ensuremath{\in}}
-\newunicodechar{∉}{\ensuremath{\not\in}} % 
-\newunicodechar{∋}{\ensuremath{\ni}}  % 
-\newunicodechar{∎}{\ensuremath{\qed}} % end of proof. % or: {\blacksquare} 
-\newunicodechar{∏}{\prod}
-\newunicodechar{∑}{\sum}
-\newunicodechar{∗}{\ensuremath{\ast}} % 
-\newunicodechar{∘}{\ensuremath{\circ}} % 
-\newunicodechar{∙}{\ensuremath{\bullet}} % 
-\newunicodechar{∙}{\ensuremath{\cdot}}
-\newunicodechar{∞}{\ensuremath{\infty}} % 
-\newunicodechar{∣}{\ensuremath{\mid}} % 
-\newunicodechar{∧}{\ensuremath{\wedge}}% 
-\newunicodechar{∨}{\ensuremath{\vee}}% 
-\newunicodechar{∩}{\ensuremath{\cap}} % 
-\newunicodechar{∪}{\ensuremath{\cup}} % 
-\newunicodechar{∫}{\int}
-\newunicodechar{∷}{::} % proportion
-\newunicodechar{∼}{\ensuremath{\sim}} % 
-\newunicodechar{≃}{\ensuremath{\simeq}} % 
-\newunicodechar{≅}{\ensuremath{\cong}} % 
-\newunicodechar{≈}{\ensuremath{\approx}} % 
-\newunicodechar{≜}{\ensuremath{\stackrel{\scriptscriptstyle {\triangle}}{=}}} 
-\newunicodechar{≜}{\triangleq}
-\newunicodechar{≝}{\ensuremath{\stackrel{\scriptscriptstyle {\text{def}}}{=}}}
-\newunicodechar{≟}{\ensuremath{\stackrel {_\text{\textbf{?}}}{\text{\textbf{=}}\negthickspace\negthickspace\text{\textbf{=}}}}} % or {\ensuremath{\stackrel{\scriptscriptstyle ?}{=}}}
-\newunicodechar{≠}{\ensuremath{\neq}}% 
-\newunicodechar{≡}{\ensuremath{\equiv}}% 
-\newunicodechar{≤}{\ensuremath{\le}} % 
-\newunicodechar{≥}{\ensuremath{\ge}} % 
-\newunicodechar{⊂}{\ensuremath{\subset}} % 
-\newunicodechar{⊃}{\ensuremath{\supset}} %  
-\newunicodechar{⊆}{\ensuremath{\subseteq}} %  
-\newunicodechar{⊇}{\ensuremath{\supseteq}} % 
-\newunicodechar{⊎}{\ensuremath{\uplus}} %
-\newunicodechar{⊑}{\ensuremath{\sqsubseteq}} % 
-\newunicodechar{⊒}{\ensuremath{\sqsupseteq}} % 
-\newunicodechar{⊓}{\ensuremath{\sqcap}} % 
-\newunicodechar{⊔}{\ensuremath{\sqcup}} % 
-\newunicodechar{⊕}{\ensuremath{\oplus}} % 
-\newunicodechar{⊗}{\ensuremath{\otimes}} % 
-\newunicodechar{⊢}{\ensuremath{\vdash}} % 
-\newunicodechar{⊥}{\ensuremath{\bot}} % bottom
-\newunicodechar{⊧}{\models} % 
-\newunicodechar{⊨}{\models} % 
-\newunicodechar{⊩}{\Vdash}
-\newunicodechar{⊸}{\ensuremath{\multimap}} % 
-\newunicodechar{⋁}{\ensuremath{\bigvee}}
-\newunicodechar{⋃}{\ensuremath{\bigcup}} % 
-\newunicodechar{⋄}{\ensuremath{\diamond}} % 
-\newunicodechar{⋅}{\ensuremath{\cdot}}
-\newunicodechar{⋆}{\ensuremath{\star}} % 
-\newunicodechar{⋮}{\ensuremath{\vdots}} % 
-\newunicodechar{⋯}{\ensuremath{\cdots}} % 
-\newunicodechar{─}{---}
-\newunicodechar{■}{\ensuremath{\blacksquare}} % black square
-\newunicodechar{□}{\ensuremath{\square}} % 
-\newunicodechar{▹}{\ensuremath{\rhd}} % 
-\newunicodechar{◃}{\triangleleft{}}
-\newunicodechar{◅}{\ensuremath{\triangleleft{}}}
-\newunicodechar{◇}{\ensuremath{\diamond}} % 
-\newunicodechar{◽}{\ensuremath{\square}}
-\newunicodechar{★}{\ensuremath{\star}}   %
-\newunicodechar{♭}{\ensuremath{\flat}} % 
-\newunicodechar{♯}{\ensuremath{\sharp}} % 
-\newunicodechar{✓}{\ensuremath{\checkmark}} % 
-\newunicodechar{⟂}{\ensuremath{^\bot}} % PERPENDICULAR 
-\newunicodechar{⟦}{\ensuremath{\llbracket}}
-\newunicodechar{⟧}{\ensuremath{\rrbracket}}
-\newunicodechar{⟨}{\ensuremath{\langle}} % 
-\newunicodechar{⟩}{\ensuremath{\rangle}} % 
-\newunicodechar{⟶}{{\longrightarrow}}
-\newunicodechar{⟷} {\ensuremath{\leftrightarrow}}
-\newunicodechar{⟹}{\ensuremath{\Longrightarrow}} % 
-\newunicodechar{𝒟}{\ensuremath{\mathcal{D}}} % 
-\newunicodechar{𝒢}{\ensuremath{\mathcal{G}}}
-\newunicodechar{𝒦}{\ensuremath{\mathcal{K}}} % 
-\newunicodechar{𝒫}{\ensuremath{\mathcal{P}}}
-\newunicodechar{𝔸}{\ensuremath{\mathds{A}}} % 
-\newunicodechar{𝔹}{\ensuremath{\mathds{B}}} % 
-\newunicodechar{𝟙}{\ensuremath{\mathds{1}}} % 
-\newunicodechar{ℒ}{\ensuremath{\mathscr{L}}}
-\newunicodechar{⊤}{\ensuremath{\top}}
-\newunicodechar{ }{\quad}
-\newunicodechar{⊛}{\ensuremath{\circledast}}
 
@@
 We do so by investigating two questions.  First, we are interested in
-variation in the level of linguistic complexity\jp{maybe revise
-  terminology? Perhaps define linguistic complexity as perplexity?}
-exhibited by different communities.
+variation in the level of \emph{linguistic complexity}
+exhibited by different communities.\footnote{
+  We will define linguistic complexity as language model perplexity.
+  We do not intend to make any claim about the relationhip
+  between language model perplexity and other notions of linguistic
+  complexity \citep[cf.][]{Hawkins2014}. Furthermore, we note that 
+  language \emph{exhibited} by a speech community \citep[i.e., \emph{parole},][]{deSaussure2011}
+  is  different from \emph{a language} (i.e., \emph{langue}).}
 %\jp{
@@
 %
-Indeed, it is folklore that certain communities exhibit more
-linguistic complexity than others.  Here we want to quantify this
+Indeed, it is folklore that certain communities tend to exhibit more predictable
+speech patterns than others.  Here we want to quantify this
 phenomenon.  More precisely, we investigate if the degree of
@@
 \jp{Perhaps use ``Latent Bayesian classifier'' instead.}
-The perplexity of a CCLM gives us a measure of linguistic complexity\jp{terminology?},
+The perplexity of a CCLM gives us a measure of linguistic complexity
 but it is not a measure of how linguistically \emph{specific} the
@@
   \centering
-  \input{floats/ind_corrs.tex}
+  \begin{tabular}{llrrrr}
+  \toprule
+      & $l_c$ & $\Ppl_{M_j}$ & $\IG_{M_j}$ \\
+  \midrule
+  \multirow{4}{*}{LSTM}         & 0 &            - &             - \\
+                                & 1 &         0.19 &         -0.40 \\
+                                & 2 &         0.19 &         -0.43 \\
+                                & 3 &         0.06 &         -0.45 \\
+  \midrule
+  \multirow{4}{*}{Transformer}  & 0 &         0.19 &         -0.41 \\
+                                & 1 &         0.23 &         -0.43 \\
+                                & 2 &         0.16 &         -0.39 \\
+                                & 3 &         0.12 &         -0.40 \\
+  \bottomrule
+  \end{tabular}
   \caption{ (TODO: missing data) Pearson's $r$ correlation coefficient between community
@@
     perplexity on messages from community $c_j$ ($\Ppl_{M_j}$), and
-    CCLM information gain on messages from $c_j$ ($\IG_{M_j}$).  }
+    CCLM information gain on messages from $c_j$ ($\IG_{M_j}$). $p < 0.001$
+    for all correlations except $\Ppl_{M_j}$ for the LSTM with 
+  $l_c=3$ ($p=0.218$) and Transfomer with $l_c=3$ ($p=0.010$).}
   \label{tab:ind-corrs}

### 2021-06-03 paper/paper.tex
summary paragraph and abstract

@@
 \begin{abstract}
-
-  TODO: fix this up.
-  % Context (Optional)
-
-  % Problem?
-
-  - Can community-level linguistic variation be captured by statistical models?
-
-  - Is community-specific linguistic variation correlated with social
-    relationship between communities?
-
-  % Why it's interesting
-
-  % Results
-
-  % What follows from results? (Optional)
-
-  Community-Conditioned Language Models (CCLMs) are an under-researched tool for 
-  investigating community-level linguistic variation.
-  In this paper, we experiment with different neural CCLM architectures and
-  report empirical results from CCLMs trained on 510 online communities.
-  We find that the community embeddings learned by these models
-  correlate with representations based on non-linguistic features.
-  We define a metric of the linguistic \emph{indiscernability} of a community
-  and investigate the relationship between indiscernibility and other
-  linguistic properties of the community.
+  Community-level linguistic variation is a core concept in sociolinguistics,
+  but more tools are needed to study variation from a computational perspective.
+  In this paper, we use conditional language models to investigate linguistic variation
+  across 510 online communities.
+  We characterise communities with several different language model-based metrics
+  and report on the correlations between them. 
+  We also investigate the community representations learned by the conditional language model
+  and find that they are highly correlated with a social network-based 
+  representation that doesn't use any linguistic input.
 \end{abstract}
@@
 % \citep{Hovy2015,Yang2017}. \jp{Move this to related work so we can get to the point quicker.}
-=
 Most of the previous computational work on linguistic variation 
@@
 investigate variation across speech communities (rather than
-macro-social categories), and use context-sensitive sequence models
+macro-social categories) and use context-sensitive sequence models
 capable of capturing variation beyond the lexical level.
-%
-We do so by investigating two questions.  First, we are interested in
+We do this in two ways:
+
+First, we are interested in
 variation in the level of \emph{linguistic complexity}
@@
   %Some examples (perhaps include if there is space):
-
+%  
   %(CFB)
@@
   %itself is true.
-
+%
   %(Videos)
@@
 speech patterns than others.  Here we want to quantify this
-phenomenon.  More precisely, we investigate if the degree of
-complexity can be captured in a computational model.  For this purpose
-define various Community Conditioned Language Models (CCLM for short),
-which can be modeled as the degree to which a
+phenomenon.  For this purpose, we define various Community Conditioned Language Models 
+(CCLM for short), model the degree to which a
 given message $m$ is acceptable in a given community $c$, by estimating the probability
@@
 the specificity of communities. We test whether 1.\ the linguistic
-models are correlated with a model of communities based only on
-co-occurrence of users between communities and 2.\ if latent community
-classifiers based on the CCLM can accurately identify that a given
-message comes from a given community. Additionally, one may
-intuitively assume that linguistically poor communities are easier to
-recognize: that is, the easiest it is to model computationally
-the easier it is to identify.
+models are correlated with a social network model of communities (based only on
+co-occurrence of users) and 2.\ if a latent community
+classifier based on the CCLM can accurately identify the community that a
+message comes from. Additionally, one may
+intuitively assume that linguistically complex communities are harder to
+recognize: that is, the easier it is to model the language produced by a community,
+the easier it is to identify messages coming from that community.
 
@@
   in a CCLM influences its perplexity.
-\item \label{hyp:LMCC-works} CCLM can be used to recognize communities based on
+\item \label{hyp:LMCC-works} CCLMs can be used to recognize communities based on
   latent Bayesian classification.
@@
 
-\jp{outline of the paper is important here}
+We proceed as follows: 
+First, we define the CCLM models used in the paper, as well as baseline un-conditioned madels.
+We describe how the models were trained and report on overall perplexity results, 
+as well as comparing the perplexity of the CCLM models to the baseline at the community level
+(\cref{sec:cclm}).
+Next, we define latent Bayesian community classifiers based on the CCLM. We define a measure
+of \emph{indiscernability} for communities and report on correlations between indiscernability and perplexity (\cref{sec:lmcc}).
+Then, we focus on the community embeddings learend by the CCLM,
+comparing them to a social network-based embedding (\cref{sec:embedding-analysis}).
+Finally, we give an overview of related work and discuss our findings in that context (\cref{sec:related-work,sec:discussion-conclusion}).
 
-\section{Community-conditioned language models}
+\section{Community-conditioned language models}\label{sec:cclm}
 
@@
 
-\section{Latent language model-based community classifiers}
+\section{Latent language model-based community classifiers} \label{sec:lmcc}
 \jp{Perhaps use ``Latent Bayesian classifier'' instead.}
@@
 
-\section{Social Analysis of CCLM community embeddings (TODO: better title)}
+  \section{Social Analysis of CCLM community embeddings (TODO: better title)}\label{sec:embedding-analysis}
 
@@
 
-\section{Related work}
+\section{Related work}\label{sec:related-work}
 
@@
 
-\section{Discussion and Conclusion}
+\section{Discussion and Conclusion}\label{sec:discussion-conclusion}
 

### 2021-06-03 paper/paper.tex
fixing some things

@@
 across communities, with $\Ppl_{M_j}$ ranging from \num{3.67} to
-\num{93.58} for the best LSTM model\jp{Is this the cclm or the baseline?}. %TODO make long table
+\num{93.58} for the best conditional LSTM model. %TODO make long table
 While the absolute performance ($\Ppl_{M_j}$) of the LSTM models is better,
@@
     boxplot/draw direction=x, axis x line*=bottom, axis y line*=left,y axis line style={opacity=0},
-    %xmax=1.6,
+    xmax=1.6,
     ylabel=$l_c$, xlabel=$\IG_{M_j}$,
@@
   Average model performance by community.
-  The boxes indicate 90 percent confidence intervals. Bars indicate ... TODO:. The outliers are shown as dots.
+  The boxes indicate upper and lower quartiles, while the whiskers are placed at the upper and lower maximum, with communities more than $1.5\times IQR$ (inter-quartile range) above the upper quartile considered outliers (represented as dots).
+The three most extreme outliers are excluded from this view (see \cref{fig:ppl-info-gain}).
    See also table 1 in the supplementary materials
@@
   Average community perplexity (x-axis) versus information gain (y-axis)
-  for the best LSTM model ($l_c = 1$). TODO: correlation factor
+  for the best LSTM model ($l_c = 1$). Pearson's r=$-0.61$ (p<0.001).
 }
@@
 
-\section{Latent language model-based community classifiers} \label{sec:lmcc}
-\jp{Perhaps use ``Latent Bayesian classifier'' instead.}
+\section{Latent Bayesian community classifiers} \label{sec:lmcc}
 The perplexity of a CCLM gives us a measure of linguistic complexity
 but it is not a measure of how linguistically \emph{specific} the
-community is. However, the CCLM induces a latent classifier,
+community is. However, the CCLM induces a latent Bayesian classifier,
 which we call a language model-based community classifier
@@
 
-We can characterize the quality\jp{performance} of the LMCC with a confusion matrix
+We can characterize the performance of the LMCC with a confusion matrix
 $C$, defined by $C_{ij} = P(c=c_i \mid m)$ for a given (test set) message $m$ 

### 2021-06-04 paper/paper.tex
improve presentation of embedding correlatino

@@
 \[d(L,S) = n - \mathsf {Tr} (Σ)\] where the matrix \(Σ\) is obtained by the singular value
-decomposition \(U^TΣV = LS^T\). Additionally the vectors of \(U\) and \(V\) give
+decomposition \(U^TΣV = LS^T\). This way, the vectors of \(U\) and \(V\) give
 the directions of correlation respectively of \(L\) and \(S\). That is, each
@@
 correlation there is between the directions \(U_i\) and \(V_i\).
+Additionally, assuming a test community \(c\) and a CCLM embedding $x_c$ for it, we can make a prediction
+of its social embedding to be $\hat y_c = V^TUx_c$ and expect a cosine similarity with the actual social embedding $y_c$ to be on average
+$1 - d(L,S) / n$.
 
@@
 
-The \(d(L,S)\) metric ranges from \(0\) 
-(corresponding to perfect correlation, 
-obtained for example if \(L=S\)) 
-to \(n\) 
-(corresponding to perfect independence), 
-where $n=510$ is the number of communities considered.
+The \(d(L,S)\) metric ranges from \(0\) (corresponding to perfect
+correlation, obtained for example if \(L=S\)) to \(n\) (corresponding
+to perfect independence), where $n=510$ is the number of communities
+considered.
+%
+For our various embiddings, we observe a distance \(d(L,S)\) between
+232 and 354, which corresponds to an expected cosine similarity beween
+0.5 and 0.54. This strongly suggests that much information is shared
+between social-network embeddings and linguistic embeddings.
 
-To get a more precise sense of what value of \(d(L,S)\) indicates significant correlation,
-We have experimentally computed \(d(L,S)\) for five random  ``embeddings'' 
-with the same shape as the CCLM embeddings.
-We observed a mean of \(μ_d=431.39\) and a (Bessel's-corrected) 
-standard deviation \(s_d=2.90\) in their distance from the social embedding, \(S\).
-%\jp{Unfortunate overloading} luckily it seems that sample stdev is often written with an $s$ to differentiate from population stdev
+However, to get a more precise sense of whether this correlation is
+statistically significant, we have experimentally computed \(d(L,S)\)
+for 10 random ``embeddings'' with the same shape as the CCLM
+embeddings.  We observed a mean of \(μ_d=431.39\) and a
+(Bessel's-corrected) standard deviation \(s_d=2.90\) in their distance
+from the social embedding,
+\(S\). %sigma vs. s notation: it seems that sample stdev is often written with an $s$ to differentiate from population stdev
 
-Thus if \(d(L,S)\) is below the mean by several standard deviations, we can safely
-assume that there is statistically significant correlation between
-\(L\) and \(S\).  This is the case: we observe a difference between 61 and 68 standard deviations
-(\cref{tab:embedding-correlations}). Further, it is clear that the CCLM embedding
-capture some of the same information as the social-network embeddings.
-We also give a sense of \emph{how} the correlation is manifested by
-analysis of the two principal components of correlation in the
-linguistic embeddings, \(U_0\) and \(U_1\). To do so we plot the
-projection of embeddings on along \(U_0\) and \(U_1\), together with
-the corresponding singular values.\jp{TODO}
+Because the value of \(d(L,S)\) that we get for actual embedding is
+between 61 and 68 standard deviations
+(\cref{tab:embedding-correlations}), we can safely conclude that there
+is statistically significant correlation between \(L\) and \(S\).
+
+% Further, it is clear that the CCLM embedding
+% capture some of the same information as the social-network embeddings.
+% We also give a sense of \emph{how} the correlation is manifested by
+% analysis of the two principal components of correlation in the
+% linguistic embeddings, \(U_0\) and \(U_1\). To do so we plot the
+% projection of embeddings on along \(U_0\) and \(U_1\), together with
+% the corresponding singular values.
 

### 2021-06-08 analysis.py
update results

@@ -8,8 +8,8 @@ from scipy.special import entr
 from itertools import combinations
 from sklearn.decomposition import PCA
 
-model_dir = 'model/reddit'
-floats_dir = 'paper/floats_new'
+model_dir = Path('model/reddit')
+floats_dir = Path('paper/floats_new')
 
 def add_columns(df, new, suffix=''):
     new = new.rename(lambda x: str(x) + suffix, axis=1)
@@ -34,42 +34,10 @@ def cos_sim(v1, v2):
 
 ##### Create community dataframe (with communities as rows)
 
-comms = torch.load(open(os.path.join(model_dir, 'community.field'), 'rb')).vocab.itos[1:]
+comms = torch.load(open(model_dir/'community.field', 'rb')).vocab.itos[1:]
 
 df_c = pd.DataFrame([], index=pd.Index(comms, name='community'))
 df_cc = pd.DataFrame([],index=pd.MultiIndex.from_tuples(combinations(comms, 2), names=['community1', 'community2']))
-
-# Manually assign communities to different types/subjects
-# NOTE this shouldn't really be a partition. Some communities clearly belong to multiple types, but we need a more sophisticated viz for that.
-comm_cats = {
-    'videogames': ['Warframe', 'eu4', 'GlobalOffensive', 'MaddenUltimateTeam', 'heroesofthestorm', 'EDH', 'KerbalSpaceProgram'],
-    'female-focused': ['xxfitness', 'femalefashionadvice', 'TwoXChromosomes', 'AskWomen', 'breakingmom', 'BabyBumps'],
-    'sports': ['MMA', 'reddevils', 'CFB', 'MLS'],
-    'general-interest': ['Advice', 'relationships', 'LifeProTips', 'explainlikeimfive', 'todayilearned'],
-    'technology': ['pcmasterrace', 'techsupport', 'jailbreak', 'oculus']
-    # Exclude categories with <3 comms
-    # 'gamergate': ['Kappa', 'KotakuInAction'],
-    # 'meme': ['justneckbeardthings', 'cringe'],
-    # 'photos': ['photography', 'EarthPorn'],
-    # 'support': ['stopdrinking', 'exjw'],
-    # fitness: ['xxfitness', 'bodybuilding']
-}
-comm_cats = {c:t for t in comm_cats for c in comm_cats[t]} # assumes one type/community
-comm_cats = pd.Series([comm_cats.get(c, 'other') for c in comms], index=comms)
-df_c['category'] = comm_cats
[…]

### 2021-06-08 results.py

@@ -18,6 +18,7 @@ fields = data.load_fields(model_family_dir)
 comms = fields['community'].vocab.itos
 
 for architecture in ('lstm', 'transformer'):
+
     uncond_model_name = f'{architecture}-3'
     uncond_model_dir = model_family_dir/uncond_model_name
 
@@ -42,6 +43,7 @@ for architecture in ('lstm', 'transformer'):
         })
 
     for lc in range(4):
+
         model_name = f'{architecture}-3-{lc}'
         model_dir = model_family_dir/model_name
         print(model_name)
@@ -51,6 +53,7 @@ for architecture in ('lstm', 'transformer'):
         # not sure why, but one row had missing values in lstm-3-0. possibly csv i/o problem?
         # (example_id and length, unfortunately, so hard to replace...)
         # drop it, but ensure it really is a limited problem
+
         n = len(df)
         df.dropna(inplace=True)
         print(f"Dropped {n - len(df)} examples")

### 2021-06-09 analysis.py
remove analysis.py. now we have results.py which produces stuff we analyze in the notebooks

@@ -1,239 +1,0 @@
-import pandas as pd
-import os
-from pathlib import Path
-import numpy as np
-import torch
-from scipy.stats import pearsonr
-from scipy.special import entr
-from itertools import combinations
-from sklearn.decomposition import PCA
-
-model_dir = Path('model/reddit')
-floats_dir = Path('paper/floats_new')
-
-def add_columns(df, new, suffix=''):
-    new = new.rename(lambda x: str(x) + suffix, axis=1)
-    index_name = df.index.name
-    for col in df.columns: # make the operation idempotent
-        if col in new.columns:
-            df = df.drop(col, axis=1)
-    df = pd.merge(df, new, how='outer', left_index=True, right_index=True, 
-            validate='one_to_one', sort=True)
-    df.index.name = index_name
-    return df
-
-def entropy(x):
-    return entr(x).sum()
-
-def ppl(x):
-    return np.exp(entropy(x))
-
-def cos_sim(v1, v2):
-    return (v1 * v2).sum(axis=0) / (np.linalg.norm(v1, axis=0) * np.linalg.norm(v2, axis=0))
-
-
-##### Create community dataframe (with communities as rows)
-
-comms = torch.load(open(model_dir/'community.field', 'rb')).vocab.itos[1:]
-
-df_c = pd.DataFrame([], index=pd.Index(comms, name='community'))
-df_cc = pd.DataFrame([],index=pd.MultiIndex.from_tuples(combinations(comms, 2), names=['community1', 'community2']))
-
-##### Create model dataframe (indexed by model architecture)
-
-cond_lstms = [f'lstm-3-{lc}' for lc in range(4)]
-cond_transformers = [f'transformer-3-{lc}' for lc in range(4)]
-cond_models = cond_lstms + cond_transformers
-models = cond_models + ['lstm-3', 'transformer-3']
-
-df_m = pd.DataFrame([], index=pd.Index(models, name='model'))
-df_mm = pd.DataFrame([],index=pd.MultiIndex.from_tuples(combinations(cond_models, 2), names=['model1', 'model2']))
-
-# Best model validation epoch (used for testing)
-best_epoch = pd.Series([
-    int(open(f'{model_dir}/{model}/saved-epoch.txt').read())
[…]

### 2021-06-09 eval_model.py

@@ -1,5 +1,5 @@
 """ Use the trained models on test data, testing both CCLM and LMCC.
-    Produce results tabels that will be read by analysis.py to create plots, tables, etc.
+    Produce results tabels that will be read by results.py to create plots, tables, etc.
 """
 
 import util

### 2021-06-09 results.py
consolidate analysis

@@ -7,8 +7,18 @@ import json
 def exp_normalize(x, axis):
     """ https://timvieira.github.io/blog/post/2014/02/11/exp-normalize-trick/ """
     b = x.max(axis=axis)
-    y = np.exp(x - b)
-    return y / y.sum(axis)
+    y = np.exp(x - np.expand_dims(b, axis=axis))
+    return y / np.expand_dims(y.sum(axis), axis=axis)
+
+def compute_confusion(nll_df):
+    nll_df = nll_df.sort_values(by=['community', 'example_id'])
+    nll = nll_df[comms].values * np.expand_dims(nll_df.length.values, axis=1)# -log(P(m|c)) // shape (m,c)
+    P_mc = -nll # -log(P(m|c)) // shape (m,c)
+    P_cm = exp_normalize(P_mc, axis=1) # P(c|m) // shape (m,c) 
+    P_cm_split = np.array(np.split(P_cm,510,axis=0)) # P(c_i|m_j from c_j) // shape (c_j, m_j, c_i)
+    P_cm_split_sum = P_cm_split.sum(axis=1) # sum_mj[P(c_i | m_j from c_j)] // shape (c_j, c_i)
+    confusion = P_cm_split_sum / P_cm_split_sum.sum(axis=1) # mean_mj[P(c_i | c_j)] // shape (c_j, c_i)
+    return confusion
 
 results_agg_ppl = []
 results_comm_ppl = []
@@ -50,19 +60,16 @@ for architecture in ('lstm', 'transformer'):
 
         # load conditioned NLLs (including off-community conditioning)
         df = pd.read_csv(model_dir/'nll.csv')
-        # not sure why, but one row had missing values in lstm-3-0. possibly csv i/o problem?
-        # (example_id and length, unfortunately, so hard to replace...)
-        # drop it, but ensure it really is a limited problem
 
-        n = len(df)
-        df.dropna(inplace=True)
-        print(f"Dropped {n - len(df)} examples")
-        df.example_id = df.example_id.astype(int)
-        df.length = df.length.astype(int)
         # repalce community indexes with names
         df['community'] = df['community'].apply(lambda x:comms[x])
         keys = ['community', 'example_id']
         df = pd.merge(df, uncond, left_on=keys, right_on=keys)
+
+        # compute the community confusion matrix 
+        confusion = pd.DataFrame(compute_confusion(df),
[…]

### 2021-06-09 paper/paper.tex
note on n-gram models
remove plot_macros (and also lmodern ... why did we load that??)
conll is using the emnlp template
edit discussion

--- 2021-06-03 paper/paper.tex
@@
 % Remove the "review" option to generate the final version.
-\usepackage[review]{acl}
+\usepackage[review]{emnlp2021}
 
@@
 \usepackage{latexsym}
-\usepackage{newunicodechar}
 
@@
 \usepackage{booktabs}
-\usepackage{pgfplots}
 \usepackage{todonotes}
@@
 
+\usepackage{newunicodechar}
+
+\usepackage{pgfplots}
 \usepgfplotslibrary{groupplots}
@@
 \setlist[hypotheses,1]{parsep=0pt,itemsep=1pt,font=\bfseries,label=(H\arabic*)}
-
-\input{plot_macros}
 
@@
   \centering
-  \input{floats/model_results.tex}
+  \begin{tabular}{llrrr}
+  \toprule
+                                 &       & \makecell{test \\ epoch} & $\Ppl_M$ & $\IG_M$     \\
+                                 & $l_c$ &         &                 &                   \\
+  \midrule
+  \multirow{5}{*}{LSTM}         & - &          12 &  68.74          &       -           \\
+                                & 0 &          13 &  66.16          &    1.039          \\
+                                & 1 &           7 &  \textbf{66.01} &    \textbf{1.041} \\
+                                & 2 &           4 &  66.19          &    1.039          \\
+                                & 3 &           4 &  66.35          &    1.036          \\
+  \midrule
+  \multirow{5}{*}{Transformer}  & - &           4 &  79.13          &        -          \\
+                                & 0 &           4 &  \textbf{75.66} &    \textbf{1.046} \\
+                                & 1 &           4 &  82.12          &    0.964          \\
+                                & 2 &           7 &  83.53          &    0.947          \\
+                                & 3 &           3 &  75.90          &    1.043          \\
+  \bottomrule
+  \end{tabular}
   \caption{
@@
 across communities, with $\Ppl_{M_j}$ ranging from \num{3.67} to
-\num{93.58} for the best conditional LSTM model. %TODO make long table
+\num{93.58} for the best conditional LSTM model 
+(see supplementary materials for perplexity by community). %TODO keep? 
 While the absolute performance ($\Ppl_{M_j}$) of the LSTM models is better,
@@
     \foreach \lc in {0, ..., 3}{
-      \modelboxplot{lstm-3-\lc_ig}{violet}{\lc+0.2}
-      \modelboxplot{transformer-3-\lc_ig}{teal}{\lc-0.2}
+      \modelboxplot{lstm-3-\lc-ig}{violet}{\lc+0.2}
+      \modelboxplot{transformer-3-\lc-ig}{teal}{\lc-0.2}
     \legend{,LSTM, Transformer}
@@
     ]
-    \modelboxplot{lstm-3_ppl}{violet}{0.2}
-    \modelboxplot{transformer-3_ppl}{teal}{-.2}
+    \modelboxplot{lstm-3-ppl}{violet}{0.2}
+    \modelboxplot{transformer-3-ppl}{teal}{-.2}
   \end{axis}
@@
     font=\tiny,
-  ] table [x={lstm-3_ppl}, y={lstm-3-3_ig}] {floats/comm.csv};
+  ] table [x={lstm-3-ppl}, y={lstm-3-3-ig}] {floats/comm.csv};
 \end{axis}
@@
 Indiscernibility of a given community is a robust measure: it is
-almost perfectly correlated across CCLMs (such as transformers, LSTM
+almost perfectly correlated across CCLMs (such as transformers, LSTMs
 and various values of $l_c$).  In fact, we observe a correlation
 between the indiscernibility measures by every pair of models with
-Pearson's $r \geq 0.96$ and $Bp < 0.0001$.
+Pearson's $r \geq 0.96$ and $p < 0.0001$.
 
@@
 is a poor predictor of how distinctive its language is.
+\jp{Conclusion: perplexity (linguistic complexity) is not correlated
+  to how recognisable a community is. Presumably (in general) some very specific
+  lexical items are characteristic of a community.}
 %
@@
   \midrule
-  \multirow{4}{*}{LSTM}         & 0 &            - &             - \\
+  \multirow{4}{*}{LSTM}         & 0 &         0.21 &         -0.41 \\
                                 & 1 &         0.19 &         -0.40 \\
@@
   \end{tabular}
-  \caption{ (TODO: missing data) Pearson's $r$ correlation coefficient between community
+  \caption{ Pearson's $r$ correlation coefficient between community
     indiscernibility ($\Ind_j$) and two different predictors: CCLM
@@
         font=\tiny,
-      ] table [x={lstm-3-1_ppl}, y={lstm-3-1_indisc}] {floats/comm.csv};
+      ] table [x={lstm-3-1-ppl}, y={lstm-3-1-indisc}] {floats/comm.csv};
   \nextgroupplot[xlabel=$\IG_{M_j}$, ylabel=$\Ind_j$, xmax=1.6]
@@
         font=\tiny,
-      ] table [x={lstm-3-1_ig}, y={lstm-3-1_indisc}] {floats/comm.csv};
+      ] table [x={lstm-3-1-ig}, y={lstm-3-1-indisc}] {floats/comm.csv};
   \end{groupplot}
@@
 the conditioned model are generally easier for the LMCC to classify.
+Intuitively this makes sense:
+the more distinctive a community's language is,
+the easier it is to guess where messages from that community originated
+and the easier it is to predict the content of the message if you know the community.
 
-Finally, we note that, even though community-conditioning affects
+Finally, we note that, even though community conditioning affects
 perplexity very little overall, it provides large benefits in term of
 the underlying Bayesian classifier. Indiscernibility is on average
-TODO with a standard deviation of TODO. Roughly speaking, this means
+\num{0.57} with a standard deviation of \num{0.22}. Roughly speaking, this means
 that, while the linguistic variations does not allow our model to
 pinpoint the originating community of a message, it can often
-eliminate TODO (a percentatge?) of communities as possible locations
+eliminate TODO (a percentatge?) of communities as possible locations \bn{43\%?}
 for it. Further more, the information gain provided by
@@
 
-\jp{Conclusion: perplexity (linguistic complexity) is not correlated
-  to how recognisable a community is. Presumably (in general) some very specific
-  lexical items are characteristic of a community.}
-
-  \section{Social Analysis of CCLM community embeddings (TODO: better title)}\label{sec:embedding-analysis}
+\section{Comparison of CCLM community embeddings with a social network embedding}\label{sec:embedding-analysis}
 
@@
 linguistic embeddings, \(U_0\) and \(U_1\). To do so we plot the
-projection of embeddings on along \(U_0\) and \(U_1\), together with
-the corresponding singular values.\jp{TODO}
+projection of each embedding along their first two principle components
+which, together with the corresponding singular values, gives an idea
+of how much and in what way they differ (see supplementary materials).
 
@@
     %\nextgroupplot
-    %\commscatter{lstm-3-1_pca1}{lstm-3-1_pca2}
+    %\commscatter{lstm-3-1-pca1}{lstm-3-1-pca2}
     %\nextgroupplot
@@
 we observed that various communities exhibit widely varying levels of
-linguistic complexity \ref{hyp:varying-complexity}, with perplexity
-per-word ranging from less than 34 to more than 65 for the best model 
-architecture.
+linguistic complexity \ref{hyp:varying-complexity}. For the best model architecture,
+perplexity per-word ranges from less than \num{4} to more than \num{93} 
+(but with only 25\% of communities under 63).
 
-With indiscernibility values evenly ranging over the whole possible
-range (1 to 510), we find that CCLMs can be used to recognize certain
+With indiscernibility, values are distributed evenly over the whole possible
+range (1 to 510). We find that CCLMs can be used to recognize certain
 communities based on latent Bayesian classification
-\ref{hyp:LMCC-works}, but not all.  We cannot rule out that generic
+\ref{hyp:LMCC-works}, but not all.  In general, we observe that the least
+discernible communities all have a very conversational, general-interest
+premise. We cannot rule out that these more generic
 communities exhibit distinctive linguistic patterns, but none of our
@@
 We find that the layer at which the community embedding is taken into
-account in a CCLM influences its perplexity \ref{hyp:layer-effect},\jp{Delete this hypothesis?}
-albeit weakly. For LSTM models, the perplexity per word, averaged over
-messages from all communities, varies between \num{49.60} and \num{50.83}. (with \num{51.99} for the\jp{update numbers?}
-unconditioned model) The perplexity per word is slightly higher for
-$l_c=0$, and has very little difference between the other models, with
-$l_c=3$ being the best.  For transformer models, it varies between 
-\num{52.28} and \num{78.76}. Here perplexity per-word is clearly not a monotonous
-function of $l_c$: the maximum is reached at $l_c=2$. We did not
-expect non-monotonicity, and leave studying this effect to future
-work.
+account ($l_c$) has a weak effect on the perplexity of the CCLM
+\ref{hyp:layer-effect}
+For LSTM models, the perplexity per word, averaged over
+messages from all communities, was between \num{66.01} and \num{66.35}
+(with \num{68.74} for the unconditioned model).
+For transformer models, it varies a bit more, between \num{75.66} and \num{83.53},
+but this seems to be mainly due to the poor performance of the models where
+the community embedding is inserted between transformer layers
+($l_c=2$ and $3$ both test above the unconditioned transformer's 
+average perplexity of \num{79.13}).
+This could have something to do with the community embedding and linear layer
+interrupting the delicate layer normalisation between transformer encoder layers.
 
@@
 model behave that way for both the LSTM and transformer.  However,
-there are some differences.  For example, the subreddit
-\emph{relationships} has an information gain of \num{1.018} for the
-best CCLM LSTM but only \num{0.998} for the best transformer model.
-Why was the LSTM able to take advantage of the community information
-for this community, where the transformer was not?  It would be
-interesting to investigate these differences in future work, since it
+there are some differences.  For example, many of the communities with the biggest 
+difference in information gain between the $l_c=0$ and $l_c=3$ LSTMs 
+seem to be organised around trading collectables or organising virtual meetups
+(\emph{friendsafari},
+\emph{Fireteams},
+\emph{RandomActsOfGaming},
+\emph{randomactsofcsgo},
+\emph{Pokemongiveaway},
+\emph{ACTrade},
+\emph{SVExchange})
+Why was the model where the community information was incorporated before
+the LSTM able to take better advantage of it for these communities in particular? 
+It would be interesting to investigate these differences in future work, since it
 could reveal differences in the kind of linguistic variation the
-different model architectures capture.\jp{update?}
+different model architectures capture.
 

### 2021-06-10 paper/paper.tex
tweaks

@@
 in exponential space (as a ratio of perplexities):
-\[\IG_M = \frac{e^{\avg_M(H_{\mathrm{LM}}(m))}}{e^{\avg_M(H_{\mathrm{CCLM}}(m))}}\]
-\[\IG_M = {e^{\avg_{m∈M}(H_{\mathrm{LM}}(m) - H_{\mathrm{CCLM}}(m))}}\]
+\[\IG_M = \frac{e^{\avg_{m \in M}(H_{\mathrm{LM}}(m))}}{e^{\avg_{m \in M}(H_{\mathrm{CCLM}}(m))}}\]
+\[\IG_M = {e^{\avg_{m \in M}(H_{\mathrm{LM}}(m) - H_{\mathrm{CCLM}}(m))}}\]
 %
@@
 For other communities $\IG_{M_j}$ is much higher, meaning that the CCLM performs better  
-(\cref{fig:comm-info-gain}).\footnote{
+(\cref{fig:comm-info-gain})\jp{missing label}.\footnote{
   Some of the communities with consistently high $IG_{M_j}$ 
@@
   are presumably conventional to these communities.}
-
+%
 The variation in information gain across communities is not 
@@
 
-As an informal observation, we note that across all the models we
+We observe empirically that across all the models we
 tested, communities where conditioning has the least effect tend to be
@@
 We can characterize the performance of the LMCC with a confusion matrix
-$C$, defined by $C_{ij} = P(c=c_i \mid m)$ for a given (test set) message $m$ 
+$C$, defined by $C_{ij} = P(c=c_i \mid m)$ for a given (test) set of messages $m$ 
 from community $c_j$.
 \[C_{ij} = \avg_{m\in M_j}P(c=c_i \mid m)\]
-We observe that the diagonal is clearly above the rest: on average the LMCC generally
+We observe that the diagonal is clearly above the rest: the LMCC generally
 picks assigns the highest probability to the community where a message originates (\cref{fig:confusion}).
@@
 The exponential of the entropy of the $j$th row of \(C\) is therefore a perplexity score,
-whose maximum is the total number of communities (46 for us).
+whose maximum is the total number of communities (510 for us).
 Thus we define $\Ind_j$ by scaling this perplexity by the number of communities:
@@
 \num{0.57} with a standard deviation of \num{0.22}. Roughly speaking, this means
-that, while the linguistic variations does not allow our model to
+that, while the linguistic variations do not allow our model to
 pinpoint the originating community of a message, it can often
-eliminate TODO (a percentatge?) of communities as possible locations \bn{43\%?}
-for it. Further more, the information gain provided by
+eliminate TODO (a percentatge?) of communities as possible origins \bn{43\%? JP: yes}
+for it. Furthermore, the information gain provided by
 community-conditioning is a good predictor of how specific the
-community is.
+community is.\jp{By empirical observation?}
 
@@
 
-The problem with this approach is that the dimensions of \(L\) and \(S\) may
-not coincide. Applying a simple permutation of dimensions may widely
-change the correlation metric. To make the metric resilient to such
-effects, we can compute the minimum distance between \(L_i\) and \(S_i\), for
+The problem with this approach is that even if several dimensions of
+\(L\) and \(S\) are capture similar information, they will not
+coincide in the \emph{representation}. That is, applying a simple
+rotation (orthogonal transformation) on either matrix widely changes
+the correlation metric. To make the metric resilient to such effects,
+we can compute the minimum distance between \(L_i\) and \(S_i\), for
 any orthogonal matrix \(\Omega\) applied to \(L\):
@@
 
-Computing \(d(L,S)\) is known as the orthogonal Procrustes problem~\citep{Gower2004}. The
+The problem of computing \(d(L,S)\) is known as the orthogonal Procrustes problem~\citep{Gower2004}. The
 solution is
 \[d(L,S) = n - \mathsf {Tr} (Σ)\] where the matrix \(Σ\) is obtained by the singular value
-decomposition \(U^TΣV = LS^T\). Additionally the vectors of \(U\) and \(V\) give
+decomposition (SVD) \(U^TΣV = LS^T\). Additionally the vectors of \(U\) and \(V\) give
 the directions of correlation respectively of \(L\) and \(S\). That is, each
@@
 
-Further, we can apply a permutation to \(U\), \(V\) and \(Σ\) such
+As common when doing SVD, we can arrange \(U\), \(V\) and \(Σ\) such
 that \(σ_i > σ_j\) iff \(i < j\). Doing so, the largest singular value
@@
 
-The \(d(L,S)\) metric ranges from \(0\) 
-(corresponding to perfect correlation, 
-obtained for example if \(L=S\)) 
-to \(n\) 
-(corresponding to perfect independence), 
-where $n=510$ is the number of communities considered.
+The \(d(L,S)\) metric ranges from \(0\) (corresponding to perfect
+correlation, obtained for example if \(L=S\)) to \(n\) (corresponding
+to perfect orthogonality), where $n=510$ is the number of communities
+considered.
 
-To get a more precise sense of what value of \(d(L,S)\) indicates significant correlation,
-We have experimentally computed \(d(L,S)\) for five random  ``embeddings'' 
+To get a more precise sense of what value of \(d(L,S)\) indicates a significant correlation,
+We have experimentally computed \(d(L,S)\) for ten random  ``embeddings'' 
 with the same shape as the CCLM embeddings.
 We observed a mean of \(μ_d=431.39\) and a (Bessel's-corrected) 
-standard deviation \(s_d=2.90\) in their distance from the social embedding, \(S\).
-%\jp{Unfortunate overloading} luckily it seems that sample stdev is often written with an $s$ to differentiate from population stdev
+standard deviation \(s_d=2.90\) in their distance from the social embedding, \(S\). % luckily it seems that sample stdev is often written with an $s$ to differentiate from population stdev
 
@@
 \(L\) and \(S\).  This is the case: we observe a difference between 61 and 68 standard deviations
-(\cref{tab:embedding-correlations}). Further, it is clear that the CCLM embedding
+(\cref{tab:embedding-correlations}). Further, it is clear that the CCLM embeddings
 capture some of the same information as the social-network embeddings.
@@
 
-\subsection{Results}
+% \subsection{Results}
 

### 2021-06-10 paper/paper.tex
compress

@@
 speech patterns than others.  Here we want to quantify this
-phenomenon.  For this purpose, we define various Community Conditioned Language Models 
+phenomenon.  For this purpose, we define (\cref{sec:cclm}) various Community Conditioned Language Models 
 (CCLM for short), model the degree to which a
@@
 models are correlated with a social network model of communities (based only on
-co-occurrence of users) and 2.\ if a latent community
+co-occurrence of users) (\cref{sec:embedding-analysis}) and 2.\ if a latent community
 classifier based on the CCLM can accurately identify the community that a
-message comes from. Additionally, one may
+message comes from. (\cref{sec:lmcc}) Additionally, one may
 intuitively assume that linguistically complex communities are harder to
@@
 
-To sum up, we test the following hypotheses:
+To sum up, we test the following hypotheses, which we review in our
+conclution (\cref{sec:discussion-conclusion}):
 \begin{hypotheses}
@@
 
-We proceed as follows: 
-First, we define the CCLM models used in the paper, as well as baseline un-conditioned madels.
-We describe how the models were trained and report on overall perplexity results, 
-as well as comparing the perplexity of the CCLM models to the baseline at the community level
-(\cref{sec:cclm}).
-Next, we define latent Bayesian community classifiers based on the CCLM. We define a measure
-of \emph{indiscernability} for communities and report on correlations between indiscernability and perplexity (\cref{sec:lmcc}).
-Then, we focus on the community embeddings learend by the CCLM,
-comparing them to a social network-based embedding (\cref{sec:embedding-analysis}).
-Finally, we give an overview of related work and discuss our findings in that context (\cref{sec:related-work,sec:discussion-conclusion}).
+% We proceed as follows: 
+% First, we define the CCLM models used in the paper, as well as baseline un-conditioned madels.
+% We describe how the models were trained and report on overall perplexity results, 
+% as well as comparing the perplexity of the CCLM models to the baseline at the community level
+% .
+% Next, we define latent Bayesian community classifiers based on the CCLM. We define a measure
+% of \emph{indiscernability} for communities and report on correlations between indiscernability and perplexity .
+% Then, we focus on the community embeddings learend by the CCLM,
+% comparing them to a social network-based embedding .
+% Finally, we give an overview of related work and discuss our findings in that context .
 

### 2021-06-10 paper/paper.tex
oops, unicode
move pcas to main; use same axis limits

@@
 \usepackage{newunicodechar}
-
+\input{unicode_symb}
 \usepackage{pgfplots}
+\usetikzlibrary{arrows}
 \usepgfplotslibrary{groupplots}
@@
 \pgfplotsset{compat=1.16}
-\pgfplotsset{every tick label/.append style={font=\footnotesize}}
+\pgfplotsset{every tick label/.append style={font=\tiny}}
 
@@
 which, together with the corresponding singular values, gives an idea
-of how much and in what way they differ (see supplementary materials).
+of how much and in what way they differ (see~\cref{fig:pca-aligned}).
+
+\begin{figure*}[t]
+    \centering
+  \begin{tikzpicture}
+    \begin{groupplot}[%
+      group style={
+        group size=2 by 1,
+        horizontal sep=0pt,
+      },
+      xmin=-1, xmax=1,
+      ymin=-1, ymax=1,
+      xtick={-0.5,0.5}, ytick={-0.5,0.5},
+      axis x line=middle,
+      axis y line=middle,
+      axis line style={(-)},
+      axis on top=true,
+      width=0.55\textwidth,
+      height=0.55\textwidth,
+      scatter/classes={
+        0={mark=x, blue},
+        1={mark=+, orange},
+        2={mark=o, green},
+        3={mark=triangle, gray},
+        4={mark=star, magenta},
+        5={mark=diamond, red}
+      },
+      ]
+      \nextgroupplot[
+        %xlabel=$U_0$, ylabel=$U_1$,
+        legend cell align=left,
+        legend style={font=\tiny},
+        legend entries = {
+          offmychest confession relationships relationship\_advice SuicideWatch, 
+          TrueReddit China europe nyc socialism,
+          frugalmalefashion running MechanicAdvice personalfinance Flipping,
+          darksouls skyrimmods dragonage Xcom MonsterHunter,
+          amiugly Rateme ladybonersgw SchoolIdolFestival GoneWildPlus,
+          facepalm CrappyDesign mildlyinfuriating cringepics oddlysatisfying,
+        },
+        %legend pos=outer south east
+        legend to name=pcalegend
+      ]
+      \addplot[%
+        scatter, only marks,
+        mark options={thick, scale=0.75},
+       scatter src=explicit symbolic,
+        ] table [x={web-PCA0}, y={web-PCA1}, meta={web-cluster}] {floats/comm.csv};
+      \nextgroupplot[
+      %xlabel=$V_0$, ylabel=$V_0$
+      ]
+      \addplot[%
+        scatter, only marks,
+        mark options={thick, scale=0.75},
+       scatter src=explicit symbolic,
+        ] table [x={lstm-3-1-PCA0}, y={lstm-3-1-PCA1}, meta={web-cluster}] {floats/comm.csv};
+    \end{groupplot}
+  \end{tikzpicture}
+  \captionof{figure}{First two components of the aligned social (top) and linguistic (bottom) embeddings,
+    where the lingusitic embedding is taken from the LTSM with $l_c=1$.
+    Correlation between these directions is given by $\sigma_0 = 53.4$ and $\sigma_1 = 35.6$.
+    Colors are assigned by k-means clustering of the social embedding. The legend shows the
+    closest 5 communites to each cluster centroid.}
+  \label{fig:pca-aligned}
+\end{figure*}
 

### 2021-06-11 paper/paper.tex
fixe

@@
   \end{tikzpicture}
-  \captionof{figure}{First two components of the aligned social (top) and linguistic (bottom) embeddings,
+  \captionof{figure}{First two components of the aligned social (left) and linguistic (right) embeddings,
     where the lingusitic embedding is taken from the LTSM with $l_c=1$.
     Correlation between these directions is given by $\sigma_0 = 53.4$ and $\sigma_1 = 35.6$.
-    Colors are assigned by k-means clustering of the social embedding. The legend shows the
-    closest 5 communites to each cluster centroid.}
+    Colors are assigned by k-means clustering of the social embedding.
+    % The legend shows the closest 5 communites to each cluster centroid.
+    % JP: commented out this sentence which does not seem to refer to something visible.
+    % JP: If the legend shows in the supplementary material, then the whole figure should be repeated there.
+
+  }
   \label{fig:pca-aligned}

### 2021-06-11 paper/paper.tex
small fixes

@@
 For other communities $\IG_{M_j}$ is much higher, meaning that the CCLM performs better  
-(\cref{fig:comm-info-gain})\jp{missing label}.\footnote{
+(\cref{fig:comm-stratified-box}).\footnote{
   Some of the communities with consistently high $IG_{M_j}$ 
@@
 is a poor predictor of how distinctive its language is.
-\jp{Conclusion: perplexity (linguistic complexity) is not correlated
-  to how recognisable a community is. Presumably (in general) some very specific
-  lexical items are characteristic of a community.}
+%\jp{Conclusion: perplexity (linguistic complexity) is not correlated
+  %to how recognisable a community is. Presumably (in general) some very specific
+  %lexical items are characteristic of a community.}
 %
@@
 the underlying Bayesian classifier. Indiscernibility is on average
-\num{0.57} with a standard deviation of \num{0.22}. Roughly speaking, this means
+\num{0.57} with a standard deviation of \num{0.22}. This means
 that, while the linguistic variations do not allow our model to
-pinpoint the originating community of a message, it can often
-eliminate TODO (a percentatge?) of communities as possible origins \bn{43\%? JP: yes}
+pinpoint the originating community of a message, it can on average
+eliminate 43\% of communities as possible origins 
 for it. Furthermore, the information gain provided by
-community-conditioning is a good predictor of how specific the
-community is.\jp{By empirical observation?}
+community-conditioning is a good predictor of how discernable the
+community is.
 
@@
   \end{tikzpicture}
-  \captionof{figure}{First two components of the aligned social (left) and linguistic (right) embeddings,
+  \caption{First two components of the aligned social (left) and linguistic (right) embeddings,
     where the lingusitic embedding is taken from the LTSM with $l_c=1$.
@@
     % JP: If the legend shows in the supplementary material, then the whole figure should be repeated there.
-
   }

### 2021-06-11 paper/paper.tex
add pca legend to appendix

@@
 
+\input{floats/pca_aligned}
 \begin{figure*}[t]
     \centering
-  \begin{tikzpicture}
-    \begin{groupplot}[%
-      group style={
-        group size=2 by 1,
-        horizontal sep=0pt,
-      },
-      xmin=-1, xmax=1,
-      ymin=-1, ymax=1,
-      xtick={-0.5,0.5}, ytick={-0.5,0.5},
-      axis x line=middle,
-      axis y line=middle,
-      axis line style={(-)},
-      axis on top=true,
-      width=0.55\textwidth,
-      height=0.55\textwidth,
-      scatter/classes={
-        0={mark=x, blue},
-        1={mark=+, orange},
-        2={mark=o, green},
-        3={mark=triangle, gray},
-        4={mark=star, magenta},
-        5={mark=diamond, red}
-      },
-      ]
-      \nextgroupplot[
-        %xlabel=$U_0$, ylabel=$U_1$,
-        legend cell align=left,
-        legend style={font=\tiny},
-        legend entries = {
-          offmychest confession relationships relationship\_advice SuicideWatch, 
-          TrueReddit China europe nyc socialism,
-          frugalmalefashion running MechanicAdvice personalfinance Flipping,
-          darksouls skyrimmods dragonage Xcom MonsterHunter,
-          amiugly Rateme ladybonersgw SchoolIdolFestival GoneWildPlus,
-          facepalm CrappyDesign mildlyinfuriating cringepics oddlysatisfying,
-        },
-        %legend pos=outer south east
-        legend to name=pcalegend
-      ]
-      \addplot[%
-        scatter, only marks,
-        mark options={thick, scale=0.75},
-       scatter src=explicit symbolic,
-        ] table [x={web-PCA0}, y={web-PCA1}, meta={web-cluster}] {floats/comm.csv};
-      \nextgroupplot[
-      %xlabel=$V_0$, ylabel=$V_0$
-      ]
-      \addplot[%
-        scatter, only marks,
-        mark options={thick, scale=0.75},
-       scatter src=explicit symbolic,
-        ] table [x={lstm-3-1-PCA0}, y={lstm-3-1-PCA1}, meta={web-cluster}] {floats/comm.csv};
-    \end{groupplot}
-  \end{tikzpicture}
+    \PCAAligned 
   \caption{First two components of the aligned social (left) and linguistic (right) embeddings,
@@
     Colors are assigned by k-means clustering of the social embedding. 
-    % The legend shows the closest 5 communites to each cluster centroid.
-    % JP: commented out this sentence which does not seem to refer to something visible.
-    % JP: If the legend shows in the supplementary material, then the whole figure should be repeated there.
+    This figure is reproduced in the supplementary materials with a legend that helps 
+    characterise the clusters.
   }

### 2021-10-05 paper/paper.tex
unified preample

@@
 
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{siunitx}
-\usepackage{booktabs}
-\usepackage{todonotes}
-\usepackage{enumitem}
-\usepackage{cleveref}
-\usepackage{array}
-\usepackage{makecell}
-\usepackage{multirow}
-
-\usepackage{newunicodechar}
-\input{unicode_symb}
-\usepackage{pgfplots}
-\usetikzlibrary{arrows}
-\usepgfplotslibrary{groupplots}
-\usepgfplotslibrary{statistics}
-\pgfplotsset{compat=1.16}
-\pgfplotsset{every tick label/.append style={font=\tiny}}
-
-\renewcommand{\UrlFont}{\ttfamily\small}
-
-\newcommand\jp[1]{\todo[backgroundcolor=blue!10]{JP: #1}}
-\newcommand\bn[1]{\todo[backgroundcolor=green!10]{BN: #1}}
-
-\newcommand\Ppl{\mathsf{Ppl}}
-\newcommand\IG{\mathsf{IG}}
-\newcommand\Ind{\mathsf{Ind}}
-\newcommand\LM{\mathrm{LM}}
-\newcommand\CCLM{\mathrm{CCLM}}
-\DeclareMathOperator*{\avg}{average}
-\newcommand\BibTeX{B\textsc{ib}\TeX}
-\newcommand\subreddit[1]{\texttt{/r/#1}}
-\newlist{hypotheses}{enumerate}{3}
-\setlist[hypotheses,1]{parsep=0pt,itemsep=1pt,font=\bfseries,label=(H\arabic*)}
+\input{preamble}
 
@@
 geographic location \citep{Eisenstein2010,Bamman2014a}.
-An exception is \citet{DelTredici2017} who use the modified
-skip-gram model of \citet{Bamman2014a} to show that lexical semantic
-variation occurs across different communities organized around the
-same topic.
+Some recent exceptions are \citet{DelTredici2017} who use a modified
+skip-gram model to show that lexical semantic variation occurs across 
+different communities organised around the same topic and 
+\citet{Lucy2021}, who use contextualised word representations to
+investigate community-level variation in word sense distributions.
 
@@
 exhibited by different communities.\footnote{
-  We will define linguistic complexity as language model perplexity.
-  We do not intend to make any claim about the relationhip
+  We define linguistic complexity as language model perplexity
+  (i.e., as the inverse of predictability, for a given language model).
+  We do not intend to make any claim about the relationship
   between language model perplexity and other notions of linguistic
   complexity \citep[cf.][]{Hawkins2014}. Furthermore, we note that 
-  language \emph{exhibited} by a speech community \citep[i.e., \emph{parole},][]{deSaussure2011}
+  language \emph{exhibited} by a speech community \citep[i.e., \emph{parole},][]{deSaussure2011},
+  which is at issue here,
   is  different from \emph{a language} (i.e., \emph{langue}).}
@@
 speech patterns than others.  Here we want to quantify this
-phenomenon.  For this purpose, we define (\cref{sec:cclm}) various Community Conditioned Language Models 
-(CCLM for short), model the degree to which a
+phenomenon.  For this purpose, we define (\cref{sec:cclm}) 
+various Community Conditioned Language Models 
+(CCLM for short), and model the degree to which a
 given message $m$ is acceptable in a given community $c$, by estimating the probability
@@
 intuitively assume that linguistically complex communities are harder to
-recognize: that is, the easier it is to model the language produced by a community,
+recognise: that is, the easier it is to model the language produced by a community,
 the easier it is to identify messages coming from that community.
@@
 To sum up, we test the following hypotheses, which we review in our
-conclution (\cref{sec:discussion-conclusion}):
+conclusion (\cref{sec:discussion-conclusion}):
 \begin{hypotheses}
@@
   in a CCLM influences its perplexity.
-\item \label{hyp:LMCC-works} CCLMs can be used to recognize communities based on
+\item \label{hyp:LMCC-works} CCLMs can be used to recognise communities based on
   latent Bayesian classification.
@@
 unidirectional LSTM \citep{Hochreiter1997} and a Transformer
-\citep{Vaswani2017}.  In either case, the model is organized as a
+\citep{Vaswani2017}.  In either case, the model is organised as a
 standard $n$-layer neural sequence encoder (we set $n=3$ for all
 models). As is convention, a sequence of word tokens is first converted to 
-vectors using a trainable embedding layer over a pre-deterimend vocabulary.
+vectors using a trainable embedding layer over a pre-determined vocabulary.
 At the other end, word tokens are predicted using a softmax projection 
@@
 Reddit is divided into forums called \textit{subreddits}, 
-which are typically organized around a topic of interest. 
+which are typically organised around a topic of interest. 
 Users create \textit{posts}, which consist of a link, image, 
@@
 Our dataset includes messages from \num{510} subreddits, 
-which all subreddits 
+the set of all subreddits 
 with at least \num{5000} messages per month for each month
 of the year 2015.
+Most but not all of these communities (with about 10 exceptions) 
+are mostly English-language.
 Each community corpus consist of \num{42000} randomly selected messages from the year 2015.
@@
 auto-regressive model).  We used the AdamW
-\citep{Loshchilov2019} optimization algorithm, with an initial
+\citep{Loshchilov2019} optimisation algorithm, with an initial
 learning rate of \num{0.001}, with no extra control on the decay
@@
 The three most extreme outliers are excluded from this view (see \cref{fig:ppl-info-gain}).
-   See also table 1 in the supplementary materials
-   for a tabular version of these results with community names.
+   %See also table 1 in the supplementary materials % not very useful for this actually since we sort by indisc and it's only one model anyway.
+   %for a tabular version of these results with community names.
 }
@@
 tested, communities where conditioning has the least effect tend to be
-more organized around more general interest topics, such as 
+more organised around more general interest topics, such as 
 \subreddit{relationships} and \subreddit{advice}, where the subject
@@
 on community has the most benefit for narrower special-interest
-subreddits, such as those organized around a certain videogame, sports
+subreddits, such as those organised around a certain videogame, sports
 team, or subculture.  This makes sense intuitively, since communities
-with more niche topics would tend to have more specialized vocabulary
+with more niche topics would tend to have more specialised vocabulary
 and other linguistic patterns.
@@
 
-We can characterize the performance of the LMCC with a confusion matrix
+We can characterise the performance of the LMCC with a confusion matrix
 $C$, defined by $C_{ij} = P(c=c_i \mid m)$ for a given (test) set of messages $m$ 
@@
 But we want a more quantitative measure, corresponding to the ability of the LMCC to correctly
-recognize messages coming from a given community $c_j$. We could use
+recognise messages coming from a given community $c_j$. We could use
 something like an F-score, but we prefer to use a perplexity-like metric,
@@
 With indiscernibility, values are distributed evenly over the whole possible
-range (1 to 510). We find that CCLMs can be used to recognize certain
+range (1 to 510). We find that CCLMs can be used to recognise certain
 communities based on latent Bayesian classification

### 2021-10-07 paper/paper1.tex
short paper, with CCLM embeddings

--- 2021-10-05 paper/paper.tex
@@
 %
-The variation in information gain across communities is not 
-related to perplexity, however (\cref{fig:ppl-info-gain}).
-
-\begin{figure}
-\begin{tikzpicture}
-\begin{axis}[%
-    axis x line*=bottom, axis y line*=left,
-    xlabel=$\Ppl_{M_j}$, ylabel=$\IG_{M_j}$,
-    width=7cm,
-  ]
-  \addplot+[%
-    only marks,
-    mark=*,
-    mark options={scale=0.35, solid, color=violet!80, fill=violet},
-    font=\tiny,
-  ] table [x={lstm-3-ppl}, y={lstm-3-3-ig}] {floats/comm.csv};
-\end{axis}
-\end{tikzpicture}
-\caption{
-  Average community perplexity (x-axis) versus information gain (y-axis)
-  for the best LSTM model ($l_c = 1$). Pearson's r=$-0.61$ (p<0.001).
-}
-\label{fig:ppl-info-gain}
-\end{figure}
+%The variation in information gain across communities is not 
+%related to perplexity, however (\cref{fig:ppl-info-gain}).
 
@@
 and other linguistic patterns.
-
-\section{Latent Bayesian community classifiers} \label{sec:lmcc}
-The perplexity of a CCLM gives us a measure of linguistic complexity
-but it is not a measure of how linguistically \emph{specific} the
-community is. However, the CCLM induces a latent Bayesian classifier,
-which we call a language model-based community classifier
-(LMCC).
-
-So far, we have only loaded the community embedding which corresponds
-to the actual community that the message originates from. By
-evaluating the model's loss (and thus probability) for each potential community,
-we can apply Bayesian reasoning to extract a latent community classifier for each message.
-The probability that a given message $m$
-belongs to a community $c_i$ is calculated as follows:
-\[P(c=c_i \mid m) = P(m \mid c_i)\frac {P(m)} {P(c=c_i)}\]
-In the above,
-$P(c=c_i)$ is the frequency of community $c_i$ in the dataset, and
-$P(m \mid c_i)$ (the probability of a message $m$
-in the context of community $c_i$) is estimated by the CCLM. 
-$P(m)$ is the absolute
-probability of a message, which can be computed as
-\[P(m) = \sum_i P(c=c_i) P(m\mid c=c_i ). \]
-Since the test set is balanced across communities,
-we have an even prior for $P(c=c_i)$, giving us
-\[P(c=c_i\mid m) \propto  P(m\mid c_i).\]
-
-We can characterise the performance of the LMCC with a confusion matrix
-$C$, defined by $C_{ij} = P(c=c_i \mid m)$ for a given (test) set of messages $m$ 
-from community $c_j$.
-\[C_{ij} = \avg_{m\in M_j}P(c=c_i \mid m)\]
-We observe that the diagonal is clearly above the rest: the LMCC generally
-picks assigns the highest probability to the community where a message originates (\cref{fig:confusion}).
-
-\begin{figure}
-  \includegraphics[width=\linewidth]{floats/confusion_lstm-3-1}
-\caption{
-  LMCC confusion matrix for the LSTM model with $l_c=1$, showing $\log(P(c_j\mid c_i))$, 
-  with $c_i$ on the x-axis and $c_j$ on the y-axis.
-  Communities are sorted by indiscernability.
-  (The first few communities predominantly use a language other than English, and are thus very discernable.)}
-\label{fig:confusion}
-\end{figure}
-
-But we want a more quantitative measure, corresponding to the ability of the LMCC to correctly
-recognise messages coming from a given community $c_j$. We could use
-something like an F-score, but we prefer to use a perplexity-like metric,
-which is has stronger information-theoretic grounds.
-To this end, we define the \emph{linguistic indiscernibility} of community $c_j$, 
-noted $\Ind_j$. 
-First, note that each row of $C$ is a probability distribution,
-since it is computed as the average of $P(c=c_i \mid m)$ for $m\in M_j$.
-The exponential of the entropy of the $j$th row of \(C\) is therefore a perplexity score,
-whose maximum is the total number of communities (510 for us).
-Thus we define $\Ind_j$ by scaling this perplexity by the number of communities:
-\[\Ind_j = \frac{e^{-\sum_i C_{ij} log(C_{ij})}}{|C_j|}\]
-This way $Ind_j=0$ means that the LMCC is completely certain of its
-choice, while $Ind_j=1$ means that equal probability is given to all
-predictions.  In general $\Ind_j$ is a score of uncertainty of the
-LMCC, given messages from a community $c_j$.
-
-Indiscernibility of a given community is a robust measure: it is
-almost perfectly correlated across CCLMs (such as transformers, LSTMs
-and various values of $l_c$).  In fact, we observe a correlation
-between the indiscernibility measures by every pair of models with
-Pearson's $r \geq 0.96$ and $p < 0.0001$.
-
-However, we observe that community-wise perplexity is only weakly (if at all) 
-correlated with community indiscernibility under the corresponding latent 
-LMCC (see \cref{tab:ind-corrs} and, for example, \cref{fig:cclm-lmcc-ppl}).
-In other words, the linguistic complexity of a given community
-is a poor predictor of how distinctive its language is.
-%\jp{Conclusion: perplexity (linguistic complexity) is not correlated
-  %to how recognisable a community is. Presumably (in general) some very specific
-  %lexical items are characteristic of a community.}
-%
-\begin{table}
-  \small
-  \centering
-  \begin{tabular}{llrrrr}
-  \toprule
-      & $l_c$ & $\Ppl_{M_j}$ & $\IG_{M_j}$ \\
-  \midrule
-  \multirow{4}{*}{LSTM}         & 0 &         0.21 &         -0.41 \\
-                                & 1 &         0.19 &         -0.40 \\
-                                & 2 &         0.19 &         -0.43 \\
-                                & 3 &         0.06 &         -0.45 \\
-  \midrule
-  \multirow{4}{*}{Transformer}  & 0 &         0.19 &         -0.41 \\
-                                & 1 &         0.23 &         -0.43 \\
-                                & 2 &         0.16 &         -0.39 \\
-                                & 3 &         0.12 &         -0.40 \\
-  \bottomrule
-  \end{tabular}
-  \caption{ Pearson's $r$ correlation coefficient between community
-    indiscernibility ($\Ind_j$) and two different predictors: CCLM
-    perplexity on messages from community $c_j$ ($\Ppl_{M_j}$), and
-    CCLM information gain on messages from $c_j$ ($\IG_{M_j}$). $p < 0.001$
-    for all correlations except $\Ppl_{M_j}$ for the LSTM with 
-  $l_c=3$ ($p=0.218$) and Transfomer with $l_c=3$ ($p=0.010$).}
-  \label{tab:ind-corrs}
-\end{table}
-
-\begin{figure*}
-  \begin{tikzpicture}
-  \begin{groupplot}[
-    group style={group size=2 by 1},
-    legend style={anchor=north, legend columns=1, font=\tiny},
-    every node near coord/.append style={font=\sffamily\tiny},%
-    axis x line*=bottom, axis y line*=left,
-  ]
-  \nextgroupplot[xlabel=$\Ppl_{M_j}$, ylabel=$\Ind_j$]
-      \addplot+[%
-        only marks,
-        mark=*,
-        mark options={scale=0.35, solid, color=violet!80, fill=violet},
-        font=\tiny,
-      ] table [x={lstm-3-1-ppl}, y={lstm-3-1-indisc}] {floats/comm.csv};
-  \nextgroupplot[xlabel=$\IG_{M_j}$, ylabel=$\Ind_j$, xmax=1.6]
-      \addplot+[%
-        only marks,
-        mark=*,
-        mark options={scale=0.35, solid, color=violet!80, fill=violet},
-        font=\tiny,
-      ] table [x={lstm-3-1-ig}, y={lstm-3-1-indisc}] {floats/comm.csv};
-  \end{groupplot}
-  \end{tikzpicture}
-  \caption{%
-    Community indiscernibility (y-axis), 
-    in relationship to linguistic complexity (left) and 
-    CCLM information gain (right). Shown here for the LSTM
-    model with $l_c=1$. See \cref{tab:ind-corrs} for 
-    Pearson's r correlations for all models.
-    (Communities with \(\IG > 1.6\) (3 outliers) are hidden to make the correlation easier to observe.)
-  }
-  \label{fig:cclm-lmcc-ppl}
-\end{figure*}
-
-% We could also check if communities are well-separated in the SNAP community encodings
-
-On the other hand, community indiscernibility is correlated with
-the information gain of the CCLM
-on messages from that community's test set
-(see \cref{tab:ind-corrs} and, for example, \cref{fig:cclm-lmcc-ppl}).
-Communities for which the language model benefits most from 
-the conditioned model are generally easier for the LMCC to classify.
-Intuitively this makes sense:
-the more distinctive a community's language is,
-the easier it is to guess where messages from that community originated
-and the easier it is to predict the content of the message if you know the community.
-
-Finally, we note that, even though community conditioning affects
-perplexity very little overall, it provides large benefits in term of
-the underlying Bayesian classifier. Indiscernibility is on average
-\num{0.57} with a standard deviation of \num{0.22}. This means
-that, while the linguistic variations do not allow our model to
-pinpoint the originating community of a message, it can on average
-eliminate 43\% of communities as possible origins 
-for it. Furthermore, the information gain provided by
-community-conditioning is a good predictor of how discernable the
-community is.
-
-% Finally, we investigate the correlation between the stability of a
-% community and the perplexity of the LMCC for it. Stability is defined as TODO.
-
-% - compute pearson correlation coefficient
-
-% We see a positive correlation: this suggests that the more stable a
-% community is, the more difficult it is to identify message as coming
-% from it. What this suggests is that stable communities tend to use a
-% varied and standard, subset of English, while unstable communities
-% resort to more formulaic language. This may suggests that to 'fit' in an
-% unstable community, one must use more obvious linguistic cues than in
-% a stable community. In fact, it is reasonable to think that a new
-% member will ostensibly use community-specifc language, but once
-% established in a community, a member would tend not to do so.
 

### 2021-10-07 paper/paper1.tex
remove irrelevant things

@@
   across 510 online communities.
-  We characterise communities with several different language model-based metrics
-  and report on the correlations between them. 
-  We also investigate the community representations learned by the conditional language model
+  We investigate the community representations learned by the conditional language model
   and find that they are highly correlated with a social network-based 
@@
 investigate variation across speech communities (rather than
-macro-social categories) and use context-sensitive sequence models
-capable of capturing variation beyond the lexical level.
-We do this in two ways:
+macro-social categories).
 
-First, we are interested in
-variation in the level of \emph{linguistic complexity}
-exhibited by different communities.\footnote{
-  We define linguistic complexity as language model perplexity
-  (i.e., as the inverse of predictability, for a given language model).
-  We do not intend to make any claim about the relationship
-  between language model perplexity and other notions of linguistic
-  complexity \citep[cf.][]{Hawkins2014}. Furthermore, we note that 
-  language \emph{exhibited} by a speech community \citep[i.e., \emph{parole},][]{deSaussure2011},
-  which is at issue here,
-  is  different from \emph{a language} (i.e., \emph{langue}).}
 %\jp{
@@
 %
-Indeed, it is folklore that certain communities tend to exhibit more predictable
-speech patterns than others.  Here we want to quantify this
-phenomenon.  For this purpose, we define (\cref{sec:cclm}) 
+% Indeed, it is folklore that certain communities tend to exhibit more predictable
+% speech patterns than others.  Here we want to quantify this
+% phenomenon.
+
+For this purpose, we define (\cref{sec:cclm}) 
 various Community Conditioned Language Models 
-(CCLM for short), and model the degree to which a
-given message $m$ is acceptable in a given community $c$, by estimating the probability
-\(P(m \mid c)\).
+(CCLM for short).
+% , and model the degree to which a
+% given message $m$ is acceptable in a given community $c$, by estimating the probability
+% \(P(m \mid c)\).
 
-Second, we want to find out how, and how much, CCLMs are attuned to
-the specificity of communities. We test whether 1.\ the linguistic
+We want to find out how, and how much, CCLMs are attuned to
+In practice, we test whether the linguistic
 models are correlated with a social network model of communities (based only on
-co-occurrence of users) (\cref{sec:embedding-analysis}) and 2.\ if a latent community
-classifier based on the CCLM can accurately identify the community that a
-message comes from. (\cref{sec:lmcc}) Additionally, one may
-intuitively assume that linguistically complex communities are harder to
-recognise: that is, the easier it is to model the language produced by a community,
-the easier it is to identify messages coming from that community.
+co-occurrence of users) (\cref{sec:embedding-analysis})
 
-To sum up, we test the following hypotheses, which we review in our
-conclusion (\cref{sec:discussion-conclusion}):
-\begin{hypotheses}
-\item \label{hyp:varying-complexity} Different communities have different levels of linguistic
-  complexity
-\item \label{hyp:layer-effect} The layer at which the community embedding is taken into account
-  in a CCLM influences its perplexity.
-\item \label{hyp:LMCC-works} CCLMs can be used to recognise communities based on
-  latent Bayesian classification.
-% \item \label{hyp:LMCC-works} The layer at which the community embedding is taken into account
-%   influences CCLM and its associated latent classifier.
-  % This is more of a justification for studying the latent classifier than a real "user-facing" hypothesis.
-\item \label{hyp:extra-linguistic-correlation} The CCLM representation of communities are correlated with
-  co-occurrence of users between communities.
-% \item Community stability is correlated with language complexity
-\item \label{hyp:rich-harder-to-identify} Linguistically poor communities are easier to identify than
-  linguistically rich communities.
-\end{hypotheses}
+% To sum up, we test the following hypotheses, which we review in our
+% conclusion (\cref{sec:discussion-conclusion}):
+% \begin{hypotheses}
+% \item \label{hyp:varying-complexity} Different communities have different levels of linguistic
+%   complexity
+% \item \label{hyp:layer-effect} The layer at which the community embedding is taken into account
+%   in a CCLM influences its perplexity.
+% \item \label{hyp:LMCC-works} CCLMs can be used to recognise communities based on
+%   latent Bayesian classification.
+% % \item \label{hyp:LMCC-works} The layer at which the community embedding is taken into account
+% %   influences CCLM and its associated latent classifier.
+%   % This is more of a justification for studying the latent classifier than a real "user-facing" hypothesis.
+% \item \label{hyp:extra-linguistic-correlation} The CCLM representation of communities are correlated with
+%   co-occurrence of users between communities.
+% % \item Community stability is correlated with language complexity
+% \item \label{hyp:rich-harder-to-identify} Linguistically poor communities are easier to identify than
+%   linguistically rich communities.
+% \end{hypotheses}
 
@@
 % \jp{I propose that most of the following analysis should be spread between the relevant sections, and only have a tiny summary here + future work.}
-To sum up our findings, 
-we observed that various communities exhibit widely varying levels of
-linguistic complexity \ref{hyp:varying-complexity}. For the best model architecture,
-perplexity per-word ranges from less than \num{4} to more than \num{93} 
-(but with only 25\% of communities under 63).
 
-With indiscernibility, values are distributed evenly over the whole possible
-range (1 to 510). We find that CCLMs can be used to recognise certain
-communities based on latent Bayesian classification
-\ref{hyp:LMCC-works}, but not all.  In general, we observe that the least
-discernible communities all have a very conversational, general-interest
-premise. We cannot rule out that these more generic
-communities exhibit distinctive linguistic patterns, but none of our
-CCLMs are able to discern them. In contrast, the distinctive patterns
-of narrow-topic communities are just as recognizable by all our LSTM
-and Transformer models.
+% To sum up our findings, 
+% we observed that various communities exhibit widely varying levels of
+% linguistic complexity \ref{hyp:varying-complexity}. For the best model architecture,
+% perplexity per-word ranges from less than \num{4} to more than \num{93} 
+% (but with only 25\% of communities under 63).
 
-We find a mild correlation (\cref{tab:ind-corrs}) between the
-perplexity of the CCLM and the indiscernibility of the corresponding
-community \ref{hyp:rich-harder-to-identify}, with factors ranging from
-0.06 to 0.23. However, The trend is more clear when considering the
-correlation with information gain and indiscernibility (correlation
-from 0.39 to 0.46). So this means that, when trying to identify a
-community, it is not its absolute perplexity that matters, but rather
-the difference in perplexity with the baseline.
+% With indiscernibility, values are distributed evenly over the whole possible
+% range (1 to 510). We find that CCLMs can be used to recognise certain
+% communities based on latent Bayesian classification
+% \ref{hyp:LMCC-works}, but not all.  In general, we observe that the least
+% discernible communities all have a very conversational, general-interest
+% premise. We cannot rule out that these more generic
+% communities exhibit distinctive linguistic patterns, but none of our
+% CCLMs are able to discern them. In contrast, the distinctive patterns
+% of narrow-topic communities are just as recognizable by all our LSTM
+% and Transformer models.
+
+% We find a mild correlation (\cref{tab:ind-corrs}) between the
+% perplexity of the CCLM and the indiscernibility of the corresponding
+% community \ref{hyp:rich-harder-to-identify}, with factors ranging from
+% 0.06 to 0.23. However, The trend is more clear when considering the
+% correlation with information gain and indiscernibility (correlation
+% from 0.39 to 0.46). So this means that, when trying to identify a
+% community, it is not its absolute perplexity that matters, but rather
+% the difference in perplexity with the baseline.
 

### 2021-10-08 paper/paper1.tex
text

@@
 
-\subsection{Comparing embeddings: method}
+\subsection{Comparing embeddings: anecdotal evidence}
+\label{sec:storytelling}
+TODO: BILL
+
+\subsection{Comparing embeddings: statistical evidence}
+
+The anecdotal evidence suggests that for some communities, proximity
+in linguistic patterns do match social proximity… and for others it
+does not. Unfortunately, it seems hard, if not impossible, to draw a
+general conclusion from this way. In this section, we propose a
+systematic approach thanks to which we can quantify the correlation
+between social proximity and linguistic proximity, and measure its
+statistical significance.
+
+Let us call \(L\) the matrix of (normed) linguistic embeddings and
+\(S\) the matrix of (normed) social embeddings. Thus \(L_i\) and
+\(S_i\) are the linguistic and social embeddings for community \(i\).
+%
+A straightforward, but ultimately flawed, generalization of the
+methodology used to collect data in \cref{sec:storytelling} is to
+consider each pair of communities \((i,j)\), and compute the
+correlation between the cosine similarities of both embeddings. That
+is, compute the Pearson correlation factor of the data set:
+
+\[\C = {(x = L_i · L_j, y = S_i · S_j) \text {~for~} i,j ∈ [1,512]\}\]
+
+One could then use a standard statistical test to verify to obtain a
+p-value for the correlation [TODO: what do we get?].  The flaw of
+this methodology is that the number of pairs grows with the square
+of the number of communities. With 510 communities, we have 129795
+pairs, and standard statistical tests on Pearson correlation will
+assure that we have a statistically significant correlation in all
+but the weakest correlation. A further flaw is that the distribution of data points
+of \(C\) are \emph{not} independent. Far from it, the are generated
+from 1/510 independent variables. We consider this last flaw fatal, and take the following
+approach instead.
+
 % Explain the problem (idea: comparing cosine similarities --> too much data, again)
 
-Let us call \(L\) the matrix of (normed) linguistic embeddings and \(S\) the matrix
-of (normed) social embeddings. Thus \(L_i\) and \(S_i\) are the linguistic and
-social embeddings for community \(i\).
+Instead of comparing embedding pairs, we will compare embeddings
+community by community.
 
-To compare embeddings, one might naively think that it suffices to
+At this stage, one might naively think that it would suffice to
 calculate the distance between two embeddings index-wise, which is
@@
 The problem with this approach is that even if several dimensions of
-\(L\) and \(S\) are capture similar information, they will not
-coincide in the \emph{representation}. That is, applying a simple
-rotation (orthogonal transformation) on either matrix widely changes
-the correlation metric. To make the metric resilient to such effects,
-we can compute the minimum distance between \(L_i\) and \(S_i\), for
-any orthogonal matrix \(\Omega\) applied to \(L\):
+\(L\) and \(S\) are correlated, they will not coincide in the
+\emph{representation}. That is, applying a simple rotation (orthogonal
+transformation) on either matrix widely changes the \(||L-S||_F\)
+correlation metric. Another way to consider the same problem is that,
+applying such a rotation does not change the cosine similarities
+considered in the previous section.
+
+To make the metric independent of the representation (up to orthogonal
+transformations), we compute the \emph{minimum} distance between
+\(L_i\) and \(S_i\), for any orthogonal matrix \(\Omega\) applied to
+\(L\):
 
@@
 
+Here, the orthogonal matrix \(Ω\) gives a map from linguistic embeddings to social embeddings.
 The problem of computing \(d(L,S)\) is known as the orthogonal Procrustes problem~\citep{Gower2004}. The
@@
 
-To get a more precise sense of what value of \(d(L,S)\) indicates a significant correlation,
-We have experimentally computed \(d(L,S)\) for ten random  ``embeddings'' 
-with the same shape as the CCLM embeddings.
-We observed a mean of \(μ_d=431.39\) and a (Bessel's-corrected) 
-standard deviation \(s_d=2.90\) in their distance from the social embedding, \(S\). % luckily it seems that sample stdev is often written with an $s$ to differentiate from population stdev
+Now, to test if \(d(L,S)\) corresponds to a significant correlation,
+it suffices to check if its value is significantly larger than it
+value for random linguistic embeddings \(L\). Even though this is the
+distribution of \(d(L,S)\) for random embeddings is difficult to
+compute analytically, one can instead evaluate it using a Monte Carlo
+method.
 
-Thus if \(d(L,S)\) is below the mean by several standard deviations, we can safely
-assume that there is statistically significant correlation between
-\(L\) and \(S\).  This is the case: we observe a difference between 61 and 68 standard deviations
-(\cref{tab:embedding-correlations}). Further, it is clear that the CCLM embeddings
-capture some of the same information as the social-network embeddings.
-We also give a sense of \emph{how} the correlation is manifested by
+Doing so, we observed that \(d(L,S)\) exhibits a mean of
+\(μ_d=431.39\) and a (Bessel's-corrected) standard deviation
+\(s_d=2.90\) in their distance from the social embedding,
+\(S\). % luckily it seems that sample stdev is often written with an $s$ to differentiate from population stdev
+
+Thus if the real \(d(L,S)\) is below the mean by several standard
+deviations, we can safely assume that there is statistically
+significant correlation between \(L\) and \(S\). A 4-sigma difference
+has less than one percent chance of occurring randomly. In our case:
+we observe a difference between 61 and 68 standard deviations
+(\cref{tab:embedding-correlations}). This definitely indicates a
+significant correlation.  Furthermore, by coming back to the
+definition of \(d(L,S)\), we know that, on average, the cosine
+similarity between \(ΩL\) and \(S\) is \(510 - 232.18 / 510\).  It
+further means that if we obtain a linguistic embedding \(L_k\) for a
+new community $k$, we can estimate its social embedding by \(ΩL_k\),
+and the cosine similarity with its true social embedding \(S_k\) is
+expected to be \((431.39 - 232.18) / 510\) --- accounting for
+over-fitting effects by taking the average distance rather than the
+maximum. % JP: TODO is the number from the table. this should probably be rewritten
+In sum, it is clear that the CCLM embeddings predict some aspect the social-network embeddings--- but far from all of it.
+
+To finish, we also give a sense of \emph{how} the correlation is manifested overall, by
 analysis of the two principal components of correlation in the

### 2021-10-12 results.py
unigram analysis fixes

@@ -15,26 +15,28 @@ def exp_normalize(x, axis):
     y = np.exp(x - np.expand_dims(b, axis=axis))
     return y / np.expand_dims(y.sum(axis), axis=axis)
 
-def compute_confusion(nll_df): # TODO per_word switch
-    nll_df = nll_df.sort_values(by=['community', 'example_id'])
-    nll = nll_df[comms].values * np.expand_dims(nll_df.length.values, axis=1)# -log(P(m|c)) // shape (m,c)
+def lmcc_probs(nll_df): 
+    nll = nll_df[comms].values # -log(P(m|c)) // shape (m,c)
     P_mc = -nll # -log(P(m|c)) // shape (m,c)
     P_cm = exp_normalize(P_mc, axis=1) # P(c|m) // shape (m,c) 
-    P_cm_split = np.array(np.split(P_cm,510,axis=0)) # P(c_i|m_j from c_j) // shape (c_j, m_j, c_i)
-    P_cm_split_sum = P_cm_split.sum(axis=1) # sum_mj[P(c_i | m_j from c_j)] // shape (c_j, c_i)
-    confusion = P_cm_split_sum / P_cm_split_sum.sum(axis=1) # mean_mj[P(c_i | c_j)] // shape (c_j, c_i)
-    return confusion
+    lmcc_df = pd.DataFrame(P_cm, columns=comms, index=nll_df.index)
+    lmcc_df = pd.merge(nll_df[['community', 'example_id']], lmcc_df, left_index=True, right_index=True)
+    return lmcc_df
+
+def lmcc_confusion(lmcc_df):
+    total_probs = lmcc_df.groupby('community')[comms].sum()
+    return total_probs / total_probs.sum(axis=1)
 
 if __name__ == '__main__':
 
     results_agg_ppl = []
     results_comm_ppl = []
 
-    uncond_models = ['lstm-3', 'transformer-3', 'unigram' ]
+    uncond_models = ['unigram', 'lstm-3', 'transformer-3']
     cond_models = [
-            [f'lstm-3-{i}' for i in range(1)], 
-            [f'transformer-3-{i}' for i in range(1)],
-            ['unigram-cond']
+            ['unigram-cond'],
+            [f'lstm-3-{i}' for i in range(4)], 
+            [f'transformer-3-{i}' for i in range(4)]
     ]
 
     for uncond_model_name, cond_models in zip(uncond_models, cond_models): 
@@ -45,18 +47,16 @@ if __name__ == '__main__':
 
[…]

### 2021-10-12 unigram_model.py

@@ -49,9 +49,18 @@ def cross_entropy(p, q):
     sum_ = 0
     for word in p:
         if not word in q:
-            return -np.inf
+            return np.inf
         else:
             sum_ += p[word] * np.log(q[word])
+    return -sum_
+
+def cross_entropy_message(m, q):
+    sum_ = 0
+    for word in m:
+        if not word in q:
+            return np.inf
+        else:
+            sum_ += np.log(q[word])
     return -sum_
 
 def word_freq_in_message(m):
@@ -75,12 +84,12 @@ with open(cond_file, 'w') as f_cond, open(ucond_file, 'w') as f_ucond:
             print(f'{i}/{len(test_data)}', end='\r')
         m = [w if w in vocab else '<unk>' for w in ex.text]
         message_len = len(m)
-        m_word_freq = word_freq_in_message(m)
+        # m_word_freq = word_freq_in_message(m)
         row = {'community': ex.community, 'example_id': ex.example_id, 'length': message_len}
-        row['nll'] = cross_entropy(m_word_freq, unigram_freqs['<comm_totals>']) 
+        row['nll'] = cross_entropy_message(m, unigram_freqs['<comm_totals>']) 
         out = writer_ucond.writerow(row)
         del row['nll']
         for c in comms:
-            row[c] = cross_entropy(m_word_freq, unigram_freqs[c]) 
+            row[c] = cross_entropy_message(m, unigram_freqs[c]) 
         out = writer_cond.writerow(row)
 

### 2021-10-12 util.py

@@ -41,7 +41,7 @@ def iter_data(data_dir, split, file_limit=None):
     communities = get_communities(data_dir)
     for community in communities:
         filename = data_filename(data_dir, split, community)
-        print(filename)
+        # print(filename)
         for i, line in enumerate(iter_file(filename)):
             yield community, line, i
             if file_limit and i >= file_limit:

### 2021-10-12 paper/paper1.tex
various small changes which are essentially procrastination

@@
 
-\title{Community-Conditioned Language Models}
+\title{Community-conditioned language models for linguistic variation}
 
@@
   across 510 online communities.
-  We investigate the community representations learned by the conditional language model
+  We probe the community representations learned by the conditional language model
   and find that they are highly correlated with a social network-based 
@@
 Linguistic communication requires that speakers share
-knowledge of certain linguistic conventions, such as syntactic
-structure, word meanings, and accepted patterns of interaction.
+certain linguistic conventions, such as syntactic
+structure, word meanings, and patterns of interaction.
 Speakers assume that these conventions are \emph{common ground} among
@@
 very large, like speakers of English.  Because it is the source of
-linguistic convention, the \emph{speech community} is an important
-concept in the study of linguistic variation \citep{Gumperz1972}.
+linguistic convention, the \emph{speech community} \citep{Gumperz1972}
+is an important concept in the study of linguistic variation.
 
@@
 % \citep{Hovy2015,Yang2017}. \jp{Move this to related work so we can get to the point quicker.}
-Most of the previous computational work on linguistic variation 
+Most previous computational work on linguistic variation 
 has considered variation at the level of macro-social categories, such as gender
@@
 geographic location \citep{Eisenstein2010,Bamman2014a}.
-Some recent exceptions are \citet{DelTredici2017} who use a modified
-skip-gram model to show that lexical semantic variation occurs across 
+Recent exceptions are \citet{DelTredici2017} who use a modified
+skip-gram model to show that lexical semantic variation occurs even across 
 different communities organised around the same topic and 
@@
 In the present work, we investigate linguistic variation
-across different online communities. In contrast to previous work, we
-investigate variation across speech communities (rather than
-macro-social categories).
+in online communities on the social media website Reddit. 
+In contrast to previous work, we investigate variation across 
+speech communities, rather than macro-social categories.
 
@@
 
-For this purpose, we define (\cref{sec:cclm}) 
-various Community Conditioned Language Models 
-(CCLM for short).
+For this purpose, we introduce (\cref{sec:cclm}) various 
+Community-Conditioned Language Models (CCLMs for short),
+which learnan embedding of the communites they are conditioned on 
+as a side effect of their language modeling objective.
+We test which architectures best make use of the community information and 
+how and in what ways the resulting embeddings are
+correlated with a social network-based representation of communities(\cref{sec:embedding-analysis}).
 % , and model the degree to which a
@@
 % \(P(m \mid c)\).
-
-We want to find out how, and how much, CCLMs are attuned to
-In practice, we test whether the linguistic
-models are correlated with a social network model of communities (based only on
-co-occurrence of users) (\cref{sec:embedding-analysis})
 
@@
 
-\section{Community-conditioned language models}\label{sec:cclm}
+\section{Community-conditioned language models (CCLMs)}\label{sec:cclm}
 
@@
 \citep{Vaswani2017}.  In either case, the model is organised as a
-standard $n$-layer neural sequence encoder (we set $n=3$ for all
-models). As is convention, a sequence of word tokens is first converted to 
-vectors using a trainable embedding layer over a pre-determined vocabulary.
+standard $3$-layer neural sequence encoder, where the input for the
+$t$th timestep of the $n+1$st layer is the $t$th hidden state of the 
+$n$th layer.
+As usual, the input to the first layer, is a sequence of tokens,
+encoded with a trainable embedding layer over a pre-determined vocabulary.
 At the other end, word tokens are predicted using a softmax projection 
@@
 
-To condition the model , we add a \emph{community embedding} parameter.
-This parameter is concatenated with the hidden layer of the
+In the conditional models, we add a \emph{community embedding} parameter.
+This parameter is concatenated (at each time step) with the hidden layer of the
 sequence encoder, at some layer $l_c \leq n$, and passed through a
@@
 
-\subsection{Results and analysis}
+\section{CCLM Performance}
 
 In this section, we report the performance of the conditioned and un-conditioned
-models on the test set.
-First, we define two performance metrics, perplexity and information gain.
+models on the held out test set.
+First, we define two performance metrics: perplexity and information gain.
 In the following, we use $M$ to refer to messages in the
@@
 
-For a given model let $H(m)$ be the model's cross-entropy loss,
+For a given model, let $H(m)$ be the model's cross-entropy loss,
 averaged over tokens in $m$.
@@
 meaning information gain above one (\cref{tab:model-results}).
-Overall, the LSTM models perform better than the transformers, 
-although the best transformer has higher information gain than the best LSTM.
+While the absolute performance ($\Ppl_{M_j}$) of the LSTM models is better,
+the best transformer models have somewhat higher information gain
+than their LSTM counterparts.
 
@@
 \num{93.58} for the best conditional LSTM model 
-(see supplementary materials for perplexity by community). %TODO keep? 
-While the absolute performance ($\Ppl_{M_j}$) of the LSTM models is better,
-the best transformer models have somewhat higher information gain
-than their LSTM counterparts.
+(\cref{fig:comm-stratified-box}; also see supplementary materials for perplexity by community). %TODO keep? 
+The conditioned models also perform differently across different communities---%
+even among the best models, some communities have $\IG_{M_j} < 1$,
+meaning that the CCLM performs worse than the unconditioned baseline for 
+messages from that community.
+For other communities $\IG_{M_j}$ is much higher, meaning that the CCLM performs better  
+(\cref{fig:comm-stratified-box}).\footnote{
+  Some of the communities with consistently high $IG_{M_j}$ 
+  across all models are primarily non-English, 
+  but surprisingly, not the three most extreme
+  outliers. There are \subreddit{counting}, \subreddit{friendsafari}, and \subreddit{Fireteams}, 
+  the later two of which are places where people coordinate to play video games
+  together. The messages in these communites adhere to highly regular formats, 
+  which are presumably conventional to the communities.}
+%
+%The variation in information gain across communities is not 
+%related to perplexity, however (\cref{fig:ppl-info-gain}).
 
@@
 
-The conditioned models also perform differently across different communities---%
-even among the best models, some communities have $\IG_{M_j} < 1$,
-meaning that the CCLM performs worse than the unconditioned baseline for 
-messages from that community.
-For other communities $\IG_{M_j}$ is much higher, meaning that the CCLM performs better  
-(\cref{fig:comm-stratified-box}).\footnote{
-  Some of the communities with consistently high $IG_{M_j}$ 
-  across all models are primarily non-English, 
-  but surprisingly, not the three most extreme
-  outliers. There are \subreddit{counting}, \subreddit{friendsafari}, and \subreddit{Fireteams}, 
-  the later two of which are places where people coordinate to play video games
-  together. The coordination has highly structured and regular formats, which 
-  are presumably conventional to these communities.}
-%
-%The variation in information gain across communities is not 
-%related to perplexity, however (\cref{fig:ppl-info-gain}).
-
 We observe empirically that across all the models we
 tested, communities where conditioning has the least effect tend to be
-more organised around more general interest topics, such as 
+organised around more general interest topics, such as 
 \subreddit{relationships} and \subreddit{advice}, where the subject
@@
 
-In this section we investigate whether (and how much) the community
+In this section we investigate whether (and how much) community
 embeddings obtained from the CCLMs correlate with the community-member
@@
 embedding with the community embedding created by \citet{Kumar2018},
-who generated them using using a negative-sampling optimization
-algorithm and the message-community co-occurrence matrix as ground
-truth, using data which spans January 2014 to April 2017.
+which were generated using using a negative-sampling optimization
+algorithm, with the message author-community co-occurrence matrix as ground
+truth, using data from January 2014 to April 2017.
 We refer the reader
@@
 \label{sec:storytelling}
-TODO: BILL
+
+To get a handle on the relationship between the linguistic and social embeddigs
 
@@
 
-\[\C = {(x = L_i · L_j, y = S_i · S_j) \text {~for~} i,j ∈ [1,512]\}\]
+\[C = \{(x = L_i · L_j, y = S_i · S_j) \text {~for~} i,j ∈ [1,512]\}\] % BN: changed \C to C. Definition for \C may be missing
 
@@
 Now, to test if \(d(L,S)\) corresponds to a significant correlation,
-it suffices to check if its value is significantly larger than it
+it suffices to check if its value is significantly larger than the same
 value for random linguistic embeddings \(L\). Even though this is the

### 2021-10-13 paper/paper1.tex
anicdotal evidence figure & discussion

@@
 To this end, we compare the CCLM-learned community
-embedding with the community embedding created by \citet{Kumar2018},
+embeddings\footnote{
+  In this section, we only consider the embeddings from 
+  the \emph{best} (highest informaiton gain) CCLM from each architecture family; 
+  that is, the LSTM with $l_c=1$ and the Transformer with $l_c=0$, however we
+  observed similar results for other values of $l_c$.}
+with the community embedding created by \citet{Kumar2018},
 which were generated using using a negative-sampling optimization
@@
 
-To get a handle on the relationship between the linguistic and social embeddigs
+\begin{figure*}[ht]
+\begin{tikzpicture}
+  \begin{groupplot}[group style={
+      group size=2 by 1,
+      horizontal sep=1cm,
+    },
+    height=5cm,width=8cm,
+    axis x line*=bottom, axis y line*=left,x axis line style={opacity=0},
+    ymin=-0.35, ymax=1.00,
+    xtick=data,
+    xticklabel style={align=left, anchor=east, rotate=60},
+    minor tick length=1ex,
+    ]
+    \nextgroupplot[ybar,
+      bar width=0.25cm,
+      xticklabels from table={floats/low_soc_high_ling.csv}{comms-label},
+      ylabel={cosine similarity},
+    ]
+    \addplot[draw=violet,fill=violet!10] table [y=lstm-sim] {floats/low_soc_high_ling.csv};
+    \addplot[draw=teal,fill=teal!10] table [y=transformer-sim] {floats/low_soc_high_ling.csv};
+    \addplot[draw=red,fill=red!10] table [y=social-sim] {floats/low_soc_high_ling.csv};
+    \nextgroupplot[ybar,
+      bar width=0.25cm,
+      xticklabels from table={floats/low_ling_high_soc.csv}{comms-label},
+      legend style={font=\tiny,at={(1,1)},anchor=south east}, legend columns=3
+    ]
+    \addplot[draw=violet,fill=violet!10] table [y=lstm-sim] {floats/low_ling_high_soc.csv};
+    \addplot[draw=teal,fill=teal!10] table [y=transformer-sim] {floats/low_ling_high_soc.csv};
+    \addplot[draw=red,fill=red!10] table [y=social-sim] {floats/low_ling_high_soc.csv};
+    \legend{LSTM, Transformer, Social}
+  \end{groupplot}
+\end{tikzpicture}
+\caption{
+  Communities with high linguistic similarity but low social similarity (left) and \textit{vice versa} (right). 
+  Linguistic similarity is measured with the embeddings from the best LSTM ($l_c=1$) and Transformer ($l_c=0$)
+  models and social similarity is measured with the embedding from \citet{Kumar2018}.
+}
+\label{fig:embedding-differences-examples}
+\end{figure*}
+
+Let us call \(L\) the matrix of (normed) linguistic embeddings and
+\(S\) the matrix of (normed) social embeddings. Thus \(L_i\) and
+\(S_i\) are the linguistic and social embeddings for community \(i\).
+
+A straightforward, but ultimately flawed, way to measure 
+how similar the two spaces are is to
+consider each pair of communities \((i,j)\), and compute the
+correlation between the cosine similarities of both embeddings. That
+is, compute the Pearson correlation factor of the data set:
+\[C = \{(x = L_i · L_j, y = S_i · S_j) \text {~for~} i,j ∈ [1,510]\}\] % BN: changed \C to C. Definition for \C may be missing
+
+Indeed, $C$ shows positive correlation for both the LSTM ($r=0.438$)
+and transformer ($r=0.452$) linguistic embeddings.\footnote{
+  By comparison, the correlation between the two linguistic embeddings is \num{0.759}.}
+The correlations are significant with $p<0.001$ in all cases;
+however, we note that the number of pairs grows with the square
+of the number of communities (with 510 communities, we have 129795)
+pairs), meaning that standard statistical tests on Pearson correlation will
+assure us of statistical significance in all
+but the weakest of correlations. A further flaw is that the data points in 
+\(C\) are \emph{not} distributed independently---far from it since each data point is generated 
+from \num{2} of \num{510} independent variables. We consider this last flaw fatal, 
+and take a different approach for computing the correlation between community
+embeddings (\cref{sec:statistical-evidence}).
+
+We can, however, use cosine similarity to find pairs of communities
+that are similar in one embedding but different in the other. 
+This gives us a way of characterizing the differences between the two kinds of
+community representation.
+The examples shown in \cref{fig:embedding-differences-examples} were selected by taking the \num{5} 
+most linguistically similar (in the LSTM embedding) pairs of communities that had below median social similarity 
+(and \textit{vice versa}).\footnote{
+Median similarity among pairs of communities was \num{0.177} for the social embedding and
+\num{0.010} and \num{0.012} for the LSTM and transformer linguistic embeddings, respectively.}
+
+As one would expect, it seems that the communities with similar linguistic embeddings
+but dissimilar social embeddings (\cref{fig:embedding-differences-examples}, left) share a similar
+topic, mode of interaction, or language variety, but for various reasons, may attract a quite
+different set of members.
+For example, \subreddit{hiphopheads} and \subreddit{Monstercat} are both topically related to music,
+but the music genres are different, and the later has a more geographically local focus (Monstercat is an 
+independent electronic music label based in Vancouver).
+While both \subreddit{MLPLounge} and \subreddit{CasualConversation} are oriented towards casual general-topic conversation,
+the former is geared towards members of a niche internet sub-culture.
+% adult fans of the television show My Little Pony.
+There are other subreddits (mentioned in r/MLPLounge's "about") that are for on-topic discussion of MLP
+Both \subreddit{rocketbeans} and \subreddit{de} are primarily German-language subreddits, but the former is comprised of 
+fans of a computer gaming-related YouTube channel, while the later is more general-interest.
+
+Differences at the other end of the spectrum (\cref{fig:embedding-differences-examples}, right)
+are somewhat harder to interpret.
+It is mostly easy to see why these communities would have different linguistic embeddings---%
+in all cases the topics are quite different.
+The reason they have similar social embeddings is less obvious,
+but we can discern some trends in how the communities are premised.
+The \subreddit{progresspics} and \subreddit{TalesFromRetail} are premised, in part,
+on seeking support from other people with similar experiences;
+\subreddit{legaladvice}, \subreddit{Cooking}, and \subreddit{loseit}
+all have a knowledge-sharing component to them;
+\subreddit{running} and \subreddit{Coffee} are hobby-focused;
+and \subreddit{self} (often) and \subreddit{askscince} (by premise) are
+places people ask and answer questions.
+It may be that there are different patterns in the \emph{social function}
+that people attribute to this particular social media website---%
+people who use Reddit in one way are more likely
+to belong to communities that are premised on the same kind of social function,
+even if the topics (and indeed language) of those communities are quite different.
+Testing this hypothesis would require a more focused study design and 
+ideally consider communities from multiple social networks (online or otherwise).
 
 \subsection{Comparing embeddings: statistical evidence}
+\label{sec:statistical-evidence}
 
@@
 
-Let us call \(L\) the matrix of (normed) linguistic embeddings and
-\(S\) the matrix of (normed) social embeddings. Thus \(L_i\) and
-\(S_i\) are the linguistic and social embeddings for community \(i\).
-%
-A straightforward, but ultimately flawed, generalization of the
-methodology used to collect data in \cref{sec:storytelling} is to
-consider each pair of communities \((i,j)\), and compute the
-correlation between the cosine similarities of both embeddings. That
-is, compute the Pearson correlation factor of the data set:
-
-\[C = \{(x = L_i · L_j, y = S_i · S_j) \text {~for~} i,j ∈ [1,512]\}\] % BN: changed \C to C. Definition for \C may be missing
-
-One could then use a standard statistical test to verify to obtain a
-p-value for the correlation [TODO: what do we get?].  The flaw of
-this methodology is that the number of pairs grows with the square
-of the number of communities. With 510 communities, we have 129795
-pairs, and standard statistical tests on Pearson correlation will
-assure that we have a statistically significant correlation in all
-but the weakest correlation. A further flaw is that the distribution of data points
-of \(C\) are \emph{not} independent. Far from it, the are generated
-from 1/510 independent variables. We consider this last flaw fatal, and take the following
-approach instead.
-
 % Explain the problem (idea: comparing cosine similarities --> too much data, again)
 
-Instead of comparing embedding pairs, we will compare embeddings
-community by community.
-
-At this stage, one might naively think that it would suffice to
+Instead of comparing embedding pairs, as in \cref{sec:storytelling}, 
+we will compare embeddings community by community.
+A naive approach would be to 
 calculate the distance between two embeddings index-wise, which is
@@
 correlation metric. Another way to consider the same problem is that,
-applying such a rotation does not change the cosine similarities
+applying such a rotation does \emph{not} change the cosine similarities
 considered in the previous section.
@@
 Here, the orthogonal matrix \(Ω\) gives a map from linguistic embeddings to social embeddings.
-The problem of computing \(d(L,S)\) is known as the orthogonal Procrustes problem~\citep{Gower2004}. The
-solution is
+The problem of computing \(d(L,S)\) is known as the orthogonal Procrustes problem~\citep{Gower2004}.\footnote{
+This approach has also been used to compare word embeddings across representations \citep[e.g.,][]{Hamilton2016}.}
+The solution is
 \[d(L,S) = n - \mathsf {Tr} (Σ)\] where the matrix \(Σ\) is obtained by the singular value
@@
     This figure is reproduced in the supplementary materials with a legend that helps 
-    characterise the clusters.
+    to characterise the clusters.
   }

### 2021-10-14 paper/paper1.tex
revision in sec 4

@@
 
-\subsection{Comparing embeddings: anecdotal evidence}
+\subsection{Comparing embeddings: cosine similarities}
 \label{sec:storytelling}
@@
 
-Let us call \(L\) the matrix of (normed) linguistic embeddings and
-\(S\) the matrix of (normed) social embeddings. Thus \(L_i\) and
-\(S_i\) are the linguistic and social embeddings for community \(i\).
+A difficulty is that the social embeddings and linguistic embeddings
+range over completely unrelated spaces, so one cannot use the usual
+cosine similarity metric \emph{across} this boundary. One can,
+however, use cosine similarity between \emph{pairs} of communities,
+and verify the similarities correlated between linguistic to social
+embeddings.  This gives us a way of characterizing the differences
+between the two kinds of community representation. To get a more
+concrete sense of what this method yields, we first survey some of the
+most salient community pairs.
 
-A straightforward, but ultimately flawed, way to measure 
-how similar the two spaces are is to
-consider each pair of communities \((i,j)\), and compute the
-correlation between the cosine similarities of both embeddings. That
-is, compute the Pearson correlation factor of the data set:
-\[C = \{(x = L_i · L_j, y = S_i · S_j) \text {~for~} i,j ∈ [1,510]\}\] % BN: changed \C to C. Definition for \C may be missing
+\todo{insert some examples with very high correlation}
 
-Indeed, $C$ shows positive correlation for both the LSTM ($r=0.438$)
-and transformer ($r=0.452$) linguistic embeddings.\footnote{
-  By comparison, the correlation between the two linguistic embeddings is \num{0.759}.}
-The correlations are significant with $p<0.001$ in all cases;
-however, we note that the number of pairs grows with the square
-of the number of communities (with 510 communities, we have 129795)
-pairs), meaning that standard statistical tests on Pearson correlation will
-assure us of statistical significance in all
-but the weakest of correlations. A further flaw is that the data points in 
-\(C\) are \emph{not} distributed independently---far from it since each data point is generated 
-from \num{2} of \num{510} independent variables. We consider this last flaw fatal, 
-and take a different approach for computing the correlation between community
-embeddings (\cref{sec:statistical-evidence}).
+The examples shown in \cref{fig:embedding-differences-examples} were
+selected by taking the \num{5} most linguistically similar (in the
+LSTM embedding) pairs of communities that had below median social
+similarity (and \textit{vice versa}).\footnote{ Median similarity
+  among pairs of communities was \num{0.177} for the social embedding
+  and \num{0.010} and \num{0.012} for the LSTM and transformer
+  linguistic embeddings, respectively.}
 
-We can, however, use cosine similarity to find pairs of communities
-that are similar in one embedding but different in the other. 
-This gives us a way of characterizing the differences between the two kinds of
-community representation.
-The examples shown in \cref{fig:embedding-differences-examples} were selected by taking the \num{5} 
-most linguistically similar (in the LSTM embedding) pairs of communities that had below median social similarity 
-(and \textit{vice versa}).\footnote{
-Median similarity among pairs of communities was \num{0.177} for the social embedding and
-\num{0.010} and \num{0.012} for the LSTM and transformer linguistic embeddings, respectively.}
+As one would expect, it seems that the communities with similar
+linguistic embeddings but dissimilar social embeddings
+(\cref{fig:embedding-differences-examples}, left) share a similar
+topic, mode of interaction, or language variety, but for various
+reasons, may attract a quite different set of members.  For example,
+\subreddit{hiphopheads} and \subreddit{Monstercat} are both topically
+related to music, but the music genres are different, and the later
+has a more geographically local focus (Monstercat is an independent
+electronic music label based in Vancouver).  While both
+\subreddit{MLPLounge} and \subreddit{CasualConversation} are oriented
+towards casual general-topic conversation, the former is geared
+towards members of a niche internet
+sub-culture. % adult fans of the television show My Little Pony. There are other subreddits (mentioned in r/MLPLounge's "about") that are for on-topic discussion of MLP
+Both \subreddit{rocketbeans} and \subreddit{de} are primarily
+German-language subreddits, but the former is comprised of fans of a
+computer gaming-related YouTube channel, while the later is more
+general-interest.
 
-As one would expect, it seems that the communities with similar linguistic embeddings
-but dissimilar social embeddings (\cref{fig:embedding-differences-examples}, left) share a similar
-topic, mode of interaction, or language variety, but for various reasons, may attract a quite
-different set of members.
-For example, \subreddit{hiphopheads} and \subreddit{Monstercat} are both topically related to music,
-but the music genres are different, and the later has a more geographically local focus (Monstercat is an 
-independent electronic music label based in Vancouver).
-While both \subreddit{MLPLounge} and \subreddit{CasualConversation} are oriented towards casual general-topic conversation,
-the former is geared towards members of a niche internet sub-culture.
-% adult fans of the television show My Little Pony.
-There are other subreddits (mentioned in r/MLPLounge's "about") that are for on-topic discussion of MLP
-Both \subreddit{rocketbeans} and \subreddit{de} are primarily German-language subreddits, but the former is comprised of 
-fans of a computer gaming-related YouTube channel, while the later is more general-interest.
+Differences at the other end of the spectrum
+(\cref{fig:embedding-differences-examples}, right) are somewhat harder
+to interpret.  It is mostly easy to see why these communities would
+have different linguistic embeddings---%
+in all cases the topics are quite different.  The reason they have
+similar social embeddings is less obvious, but we can discern some
+trends in how the communities are premised.  The
+\subreddit{progresspics} and \subreddit{TalesFromRetail} are premised,
+in part, on seeking support from other people with similar
+experiences; \subreddit{legaladvice}, \subreddit{Cooking}, and
+\subreddit{loseit} all have a knowledge-sharing component to them;
+\subreddit{running} and \subreddit{Coffee} are hobby-focused; and
+\subreddit{self} (often) and \subreddit{askscience} (by premise) are
+places people ask and answer questions.  It may be that there are
+different patterns in the \emph{social function} that people attribute
+to this particular social media website---%
+people who use Reddit in one way are more likely to belong to
+communities that are premised on the same kind of social function,
+even if the topics (and indeed language) of those communities are
+quite different.  Testing this hypothesis would require a more focused
+study design and ideally consider communities from multiple social
+networks (online or otherwise).
 
-Differences at the other end of the spectrum (\cref{fig:embedding-differences-examples}, right)
-are somewhat harder to interpret.
-It is mostly easy to see why these communities would have different linguistic embeddings---%
-in all cases the topics are quite different.
-The reason they have similar social embeddings is less obvious,
-but we can discern some trends in how the communities are premised.
-The \subreddit{progresspics} and \subreddit{TalesFromRetail} are premised, in part,
-on seeking support from other people with similar experiences;
-\subreddit{legaladvice}, \subreddit{Cooking}, and \subreddit{loseit}
-all have a knowledge-sharing component to them;
-\subreddit{running} and \subreddit{Coffee} are hobby-focused;
-and \subreddit{self} (often) and \subreddit{askscince} (by premise) are
-places people ask and answer questions.
-It may be that there are different patterns in the \emph{social function}
-that people attribute to this particular social media website---%
-people who use Reddit in one way are more likely
-to belong to communities that are premised on the same kind of social function,
-even if the topics (and indeed language) of those communities are quite different.
-Testing this hypothesis would require a more focused study design and 
-ideally consider communities from multiple social networks (online or otherwise).
+We simultaneously have examples of high correlation between social and
+linguistic embeddings, but also of low correlation. This is
+essentially anecdotal evidence: we must resort to statistical tools to
+extract the general trends.
+A straightforward, but ultimately flawed,
+way to measure how similar the two spaces are is to consider each pair
+of communities \((i,j)\), and compute the correlation between the
+cosine similarities of both embeddings.
 
-\subsection{Comparing embeddings: statistical evidence}
+That is, we can compute the Pearson correlation factor of the data set:
+\[C = \{(x = L_i · L_j, y = S_i · S_j) \text {~for~} i,j ∈
+  [1,510]\}\] % BN: changed \C to C. Definition for \C may be missing
+where \(L_i\) and \(S_i\) are the linguistic and social embeddings for
+community \(i\).  (Thus \(L\) is the matrix of (normed) linguistic
+embeddings and \(S\) the matrix of (normed) social embeddings.)
+
+The analysis shows positive correlation for both the LSTM ($r=0.438$)
+and transformer ($r=0.452$) linguistic embeddings.\footnote{ By
+  comparison, the correlation between the two linguistic embeddings is
+  \num{0.759}.}  The correlations are significant with $p<0.001$ in
+all cases.  However, we note that the number of pairs grows with the
+square of the number of communities (with 510 communities, we have
+129795) pairs), meaning that standard statistical tests on Pearson
+correlation will assure us of statistical significance in all but the
+weakest of correlations. A further flaw is that the data points in
+\(C\) are \emph{not} distributed independently---far from it in fact,
+since each data point is generated from \num{2} of \num{510}
+independent variables. We consider this last flaw fatal, and take a
+different approach for computing the correlation between community
+embeddings in the next section.
+
+\subsection{Comparing embeddings: procrustes method}
 \label{sec:statistical-evidence}

### 2021-10-14 paper/paper1.tex
update conclusion

@@
 
-% \jp{I propose that most of the following analysis should be spread between the relevant sections, and only have a tiny summary here + future work.}
-
-% To sum up our findings, 
-% we observed that various communities exhibit widely varying levels of
-% linguistic complexity \ref{hyp:varying-complexity}. For the best model architecture,
-% perplexity per-word ranges from less than \num{4} to more than \num{93} 
-% (but with only 25\% of communities under 63).
-
-% With indiscernibility, values are distributed evenly over the whole possible
-% range (1 to 510). We find that CCLMs can be used to recognise certain
-% communities based on latent Bayesian classification
-% \ref{hyp:LMCC-works}, but not all.  In general, we observe that the least
-% discernible communities all have a very conversational, general-interest
-% premise. We cannot rule out that these more generic
-% communities exhibit distinctive linguistic patterns, but none of our
-% CCLMs are able to discern them. In contrast, the distinctive patterns
-% of narrow-topic communities are just as recognizable by all our LSTM
-% and Transformer models.
-
-% We find a mild correlation (\cref{tab:ind-corrs}) between the
-% perplexity of the CCLM and the indiscernibility of the corresponding
-% community \ref{hyp:rich-harder-to-identify}, with factors ranging from
-% 0.06 to 0.23. However, The trend is more clear when considering the
-% correlation with information gain and indiscernibility (correlation
-% from 0.39 to 0.46). So this means that, when trying to identify a
-% community, it is not its absolute perplexity that matters, but rather
-% the difference in perplexity with the baseline.
+To sum up our findings, we have defined community-conditioned language
+models (CCLMs). These models are generally able to attune to
+community-specific language, as witnessed by the information gain that
+they exhibit over a baseline, non-conditioned models.
 
 We find that the layer at which the community embedding is taken into
-account ($l_c$) has a weak effect on the perplexity of the CCLM
-\ref{hyp:layer-effect}
+account ($l_c$) has a weak effect on the information gain and the
+perplexity of the CCLMs.
+
 For LSTM models, the perplexity per word, averaged over
@@
 
-The representation of communities built by CCLM exhibits a clear
-correlation with user co-occurrence patterns
-\ref{hyp:extra-linguistic-correlation}. A natural
-question that may arise is how much this correlation can be explained by
-user-specific linguistic patterns. We plan to control for this variable
-in future work, for example by building author-conditioned language
-models and testing if community vector embeddings can be determined by
-the set of the vector embeddings of its members.
+% we observed that various communities exhibit widely varying levels of
+% linguistic complexity \ref{hyp:varying-complexity}. For the best model architecture,
+% perplexity per-word ranges from less than \num{4} to more than \num{93} 
+% (but with only 25\% of communities under 63).
+
+% With indiscernibility, values are distributed evenly over the whole possible
+% range (1 to 510). We find that CCLMs can be used to recognise certain
+% communities based on latent Bayesian classification
+% \ref{hyp:LMCC-works}, but not all.  In general, we observe that the least
+% discernible communities all have a very conversational, general-interest
+% premise. We cannot rule out that these more generic
+% communities exhibit distinctive linguistic patterns, but none of our
+% CCLMs are able to discern them. In contrast, the distinctive patterns
+% of narrow-topic communities are just as recognizable by all our LSTM
+% and Transformer models.
+
+% We find a mild correlation (\cref{tab:ind-corrs}) between the
+% perplexity of the CCLM and the indiscernibility of the corresponding
+% community \ref{hyp:rich-harder-to-identify}, with factors ranging from
+% 0.06 to 0.23. However, The trend is more clear when considering the
+% correlation with information gain and indiscernibility (correlation
+% from 0.39 to 0.46). So this means that, when trying to identify a
+% community, it is not its absolute perplexity that matters, but rather
+% the difference in perplexity with the baseline.
+
+Our main result is that the the representation of communities built by
+CCLM has a positive correlation with user co-occurrence patterns.
+Even though kind of correlation is a core hypothesis of
+socio-linguistics,\todo{what was this effect called?}  we believe that
+this study is the first to test it on topic-organised
+communities. Furthermore, it appears that our method (correlating
+linguistic embeddings and social embeddings) is novel. Indeed, even
+though the procrustes method has been used to correlate two sets of
+linguistic emdeddings \emph{for the same model}, we find no evidence
+of the method being applied to embeddings for widely different models,
+as we have done.
+
+% A natural
+% question that may arise is how much this correlation can be explained by
+% user-specific linguistic patterns. We plan to control for this variable
+% in future work, for example by building author-conditioned language
+% models and testing if community vector embeddings can be determined by
+% the set of the vector embeddings of its members.
 

### 2021-10-14 paper/paper1.tex
typo

@@
 Community-Conditioned Language Models (CCLMs for short),
-which learnan embedding of the communites they are conditioned on 
+which learn embedding of the communites they are conditioned on 
 as a side effect of their language modeling objective.

### 2021-10-14 paper/paper1.tex
flyspell
wibble

@@
+% -*- eval: (flyspell-mode 1); -*-
+
 % This must be in the first 5 lines to tell arXiv to use pdfLaTeX, which is strongly recommended.
@@
 definition of \(d(L,S)\), we know that, on average, the cosine
-similarity between \(ΩL\) and \(S\) is \(510 - 232.18 / 510\).  It
+similarity between \(ΩL\) and \(S\) is \(0.45 = \) (\(510 - 232.18 / 510\)).  It
 further means that if we obtain a linguistic embedding \(L_k\) for a
@@
 and the cosine similarity with its true social embedding \(S_k\) is
-expected to be \((431.39 - 232.18) / 510\) --- accounting for
+expected to be \(0.39\) = \((431.39 - 232.18) / 510\) --- accounting for
 over-fitting effects by taking the average distance rather than the

### 2021-10-16 paper/paper1.tex
final edits for aclrr. add ling. sim/soc sim examples to fig 2; add some related work; ethics statement; small changes to intrudction, conclusion

@@
 % Remove the "review" option to generate the final version.
-\usepackage[review]{emnlp2021}
+\usepackage[review]{acl}
 
@@
 \usepackage{microtype}
+
+% If the title and author information does not fit in the area allocated, uncomment the following
+%
+%\setlength\titlebox{<dim>}
+%
+% and set <dim> to something 5cm or larger.
 
@@
 \section{Introduction}
+
 Linguistic communication requires that speakers share
@@
 their interlocutors, based on joint membership in a community
-\cite{Stalnaker2002, Clark1996}.  Such a community may be of any size,
-from the very small, like members of a particular friend group, to the
-very large, like speakers of English.  Because it is the source of
-linguistic convention, the \emph{speech community} \citep{Gumperz1972}
-is an important concept in the study of linguistic variation.
+\cite{Stalnaker2002, Clark1996}.  Such a \emph{speech community} \citep{Gumperz1972} 
+may be of any size, from the very small, like members of a particular friend group, 
+to the very large, like speakers of English; however as \citet{Eckert1992}
+point out, it is \emph{communities of practice}---defined by mutual social engagement
+in some common activity---that are the primary locus of linguistic variation.
 
@@
 
-\paragraph{Perplexity}
+\subsection{Perplexity}\label{sec:ppl}
 
@@
 
-\paragraph{CCLM Information Gain}
+\paragraph{CCLM Information Gain}\label{sec:ig}
 
@@
 The boxes indicate upper and lower quartiles, while the whiskers are placed at the upper and lower maximum, with communities more than $1.5\times IQR$ (inter-quartile range) above the upper quartile considered outliers (represented as dots).
-The three most extreme outliers are excluded from this view (see \cref{fig:ppl-info-gain}).
+The three most extreme outliers are excluded from this view.}.% \cref{fig:ppl-info-gain}).
    %See also table 1 in the supplementary materials % not very useful for this actually since we sort by indisc and it's only one model anyway.
    %for a tabular version of these results with community names.
-}
 \label{fig:comm-stratified-box}
@@
   In this section, we only consider the embeddings from 
-  the \emph{best} (highest informaiton gain) CCLM from each architecture family; 
+  the \emph{best} (highest information gain) CCLM from each architecture family; 
   that is, the LSTM with $l_c=1$ and the Transformer with $l_c=0$, however we
@@
 
+\newcommand{\pairwisecossimbarplot}[1]{
+    \nextgroupplot[ybar=0pt,
+      bar width=0.25cm,
+      xticklabels from table={floats/#1}{comms-label},
+      enlargelimits=0.15,
+      yticklabels={,,}
+    ]
+    \addplot[draw=violet,fill=violet!10] table [y=lstm-sim] {floats/#1};
+    \addplot[draw=teal,fill=teal!10] table [y=transformer-sim] {floats/#1};
+    \addplot[draw=red,fill=red!10] table [y=social-sim] {floats/#1};
+}
+
 \begin{figure*}[ht]
@@
   \begin{groupplot}[group style={
-      group size=2 by 1,
-      horizontal sep=1cm,
+      group size=3 by 1,
+      horizontal sep=0.5cm,
     },
-    height=5cm,width=8cm,
+    height=5cm,width=0.39\textwidth,
     axis x line*=bottom, axis y line*=left,x axis line style={opacity=0},
@@
     xtick=data,
-    xticklabel style={align=left, anchor=east, rotate=60},
+    xticklabel style={align=right, anchor=east, rotate=60},
     minor tick length=1ex,
+    legend style={font=\tiny,at={(1,1)},anchor=north east}, legend columns=3
     ]
-    \nextgroupplot[ybar,
+    \nextgroupplot[ybar=0pt,
       bar width=0.25cm,
-      xticklabels from table={floats/low_soc_high_ling.csv}{comms-label},
-      ylabel={cosine similarity},
+      xticklabels from table={floats/high_ling_high_soc.csv}{comms-label},
+      enlargelimits=0.15,
     ]
-    \addplot[draw=violet,fill=violet!10] table [y=lstm-sim] {floats/low_soc_high_ling.csv};
-    \addplot[draw=teal,fill=teal!10] table [y=transformer-sim] {floats/low_soc_high_ling.csv};
-    \addplot[draw=red,fill=red!10] table [y=social-sim] {floats/low_soc_high_ling.csv};
-    \nextgroupplot[ybar,
-      bar width=0.25cm,
-      xticklabels from table={floats/low_ling_high_soc.csv}{comms-label},
-      legend style={font=\tiny,at={(1,1)},anchor=south east}, legend columns=3
-    ]
-    \addplot[draw=violet,fill=violet!10] table [y=lstm-sim] {floats/low_ling_high_soc.csv};
-    \addplot[draw=teal,fill=teal!10] table [y=transformer-sim] {floats/low_ling_high_soc.csv};
-    \addplot[draw=red,fill=red!10] table [y=social-sim] {floats/low_ling_high_soc.csv};
+    \addplot[draw=violet,fill=violet!10] table [y=lstm-sim] {floats/high_ling_high_soc.csv};
+    \addplot[draw=teal,fill=teal!10] table [y=transformer-sim] {floats/high_ling_high_soc.csv};
+    \addplot[draw=red,fill=red!10] table [y=social-sim] {floats/high_ling_high_soc.csv};
+    \pairwisecossimbarplot{high_ling_low_soc.csv}{}{}
+    \pairwisecossimbarplot{low_ling_high_soc.csv}{}{}
     \legend{LSTM, Transformer, Social}
@@
 \caption{
-  Communities with high linguistic similarity but low social similarity (left) and \textit{vice versa} (right). 
-  Linguistic similarity is measured with the embeddings from the best LSTM ($l_c=1$) and Transformer ($l_c=0$)
-  models and social similarity is measured with the embedding from \citet{Kumar2018}.
-}
+  Cosine similarity between pairs of communities, computed for vectors from the
+  best CCLM embeddings (LSTM: $l_c=1$, Transformer: $l_c=0$)
+  and the social embedding from \citet{Kumar2018}.
+  Communities with high linguistic and social similarity \textbf{(left)},
+  high linguistic but low social similarity \textbf{(center)}, and
+  low linguistic but high social similarity \textbf{(right)}.
+  See text for details on the selection criteria.}
   \label{fig:embedding-differences-examples}
@@
 
-\todo{insert some examples with very high correlation}
-
-The examples shown in \cref{fig:embedding-differences-examples} were
-selected by taking the \num{5} most linguistically similar (in the
-LSTM embedding) pairs of communities that had below median social
-similarity (and \textit{vice versa}).\footnote{ Median similarity
+We consider communities from three different selection criteria: 
+Those with high social \emph{and} linguistic similarity 
+(where the sum of the two is highest),
+those with high linguistic and low social similarity
+(where social similarity is below the median and linguistic similarity is highest), and
+those with low linguistic and high social similarity
+(where linguistic similarity is below these median and linguistic similarity is highest).\footnote{We use the LSTM ($l_c=1$) community vectors for these purposes, 
+  but results attain with the best transformer model.}\footnote{Median similarity
   among pairs of communities was \num{0.177} for the social embedding
@@
   linguistic embeddings, respectively.}
+We do not consider pairs of communities that are different in both ways,
+since these don't offer much in the way of understanding the respective embeddings.
 
-As one would expect, it seems that the communities with similar
-linguistic embeddings but dissimilar social embeddings
-(\cref{fig:embedding-differences-examples}, left) share a similar
-topic, mode of interaction, or language variety, but for various
-reasons, may attract a quite different set of members.  For example,
+Unsuprisingly, the first category (\cref{fig:embedding-differences-examples}, left)yields communities that are qualitatively very similar. 
+The \subreddit{SSBPM} and \subreddit{darksouls} communities are focused
+around discussion of a particular videogame, and are paired
+with communities that discuss a variation of the same game.
+The \subreddit{amiugly} and \subreddit{Rateme} communities are both 
+forums where the posts are selfies and the comments are mostly 
+comments on the person's appearance.
+The two communities paired with \subreddit{reddevils} are 
+likewise comprised of fans of a particular English football club.
+That the three football club communities are so closely related points
+to an important point about the social embedding: Because of the way
+it was trained, it does not have to be the case that communities with
+a similar representation have a high overlap in their membership,
+but rather that the \emph{kinds of members} of the first community
+are also the \emph{kinds of members} who would belong to the second, 
+based on their membership in other communities.
+
+Communities with similar linguistic embeddings but dissimilar social embeddings
+(\cref{fig:embedding-differences-examples}, left) tend to share a similar
+topic, mode of interaction, or language variety, but 
+in all cases we looked at, there is some reason to expect that
+they might nevertheless attract different members.
+For example,
 \subreddit{hiphopheads} and \subreddit{Monstercat} are both topically
@@
 has a more geographically local focus (Monstercat is an independent
-electronic music label based in Vancouver).  While both
-\subreddit{MLPLounge} and \subreddit{CasualConversation} are oriented
-towards casual general-topic conversation, the former is geared
-towards members of a niche internet
-sub-culture. % adult fans of the television show My Little Pony. There are other subreddits (mentioned in r/MLPLounge's "about") that are for on-topic discussion of MLP
-Both \subreddit{rocketbeans} and \subreddit{de} are primarily
+electronic music label based in Vancouver). The interactions in both 
+\subreddit{MLPLounge} and \subreddit{CasualConversation} 
+could be described as casual conversation, the former is intended specifically for
+members of a niche internet sub-culture.
+% adult fans of the children's television show My Little Pony.
+There are other subreddits (mentioned in r/MLPLounge's "about") that are for on-topic discussion of MLP
+The \subreddit{exmormon} and \subreddit{Catholicism} communities discuss
+the Mormon and Catholic churches, although their members have different 
+relationships towards those organizations---the former is intended for
+former members of the church, whereas the later is geared towards practicing
+Catholics.
+Finally, both \subreddit{rocketbeans} and \subreddit{de} are primarily
 German-language subreddits, but the former is comprised of fans of a
@@
 The problem of computing \(d(L,S)\) is known as the orthogonal Procrustes problem~\citep{Gower2004}.\footnote{
-This approach has also been used to compare word embeddings across representations \citep[e.g.,][]{Hamilton2016}.}
+This approach has also been used to compare word embeddings across representations \citep[e.g.,][]{Hamilton2016a}.}
 The solution is
@@
 
+There are several recent studies that aim to measure \emph{linguistic distinctiveness}
+at the level of speech community, which is an possible interpretation of 
+of community-stratified information gain the CCLM over its unconditioned counterpart \cref{sec:ig}. Whereas the metrics in previous work are based on lexical frequency
+\citep{OConnor2010,Zhang2018,Lucy2021}, CCLM information gain is capable of capturing 
+distinctiveness at other levels of linguistic analysis. Further work is needed to
+investigate exactly what kinds of variation are captured.
+
+%- DelTredici 2017
+%- Einstein 2010
+%- Baman et al 2014/2014a
+%- Martin 2017 (community2vec), Kumar 2018
+%- Huang 2014
+%- Lucy 2021
+
 \section{Discussion and Conclusion}\label{sec:discussion-conclusion}
@@
 
-Our main result is that the the representation of communities built by
+Our main result is that the representation of communities built by
 CCLM has a positive correlation with user co-occurrence patterns.
-Even though kind of correlation is a core hypothesis of
-socio-linguistics,\todo{what was this effect called?}  we believe that
-this study is the first to test it on topic-organised
-communities. Furthermore, it appears that our method (correlating
+Even though such \emph{homophilic} correlation is a core hypothesis of 
+sociolinguistics (see \citet{Kovacs2020}, for example), 
+we believe that this study is the first to test it at 
+the level of communities of practice using computatial methods.
+Furthermore, it appears that our method (correlating
 linguistic embeddings and social embeddings) is novel. Indeed, even
 though the procrustes method has been used to correlate two sets of
-linguistic emdeddings \emph{for the same model}, we find no evidence
+linguistic embeddings \emph{for the same model}, we find no evidence
 of the method being applied to embeddings for widely different models,
@@
 
+\section{Ethical considerations}
+
+Our work uses pulicly available data from Reddit,
+collected from the API made available by \citet{Baumgartner2020}.
+Additional considerations apply, however 
+(see \citet{Gliniecka2021} for discussion). Reddit users are not,
+in general, aware of the possiiblity that their data will be used 
+for research purposes, and deleted posts can persist in archive formats.
+We are not reseasing any data or trained models as part of this study,
+since the data is publicly available and our models have low enough
+computational requirements as to pose little barrier for erplication.
+
+In the qualitative discussion offered in \cref{sec:storytelling},
+our comparative characterization of the topic, mode of interaction, 
+and language varieties
+used in the pairs of communities were formed by reading comments from the data our 
+language models were trained on. This included Googling words and phrases that 
+were unfamiliar. 
+Where we make claims about the how the community is ``premised'' or what kinds of 
+members it is ``geared towards'' or ``intended for'', these are based on the text
+of the sidebar on the community's Reddit page.
+While we believe this methodology, aggregated over many pairs of communities, 
+is appropriate for making a qualitative comparison
+of the community features encoded by different representations, 
+to make conclusions about \emph{particular} communities based on such a cursory and
+subjective analysis would be dubious and potentially harmful.
+
 \bibliography{paper}
 \bibliographystyle{acl_natbib}
+
+\include{appendix}
 

### 2021-11-29 paper/paper1.tex
first fix

@@
 For this purpose, we introduce (\cref{sec:cclm}) various
-Community-Conditioned Language Models (CCLMs for short),
-which learn embedding of the communites they are conditioned on 
-as a side effect of their language modeling objective.
-We test which architectures best make use of the community information and 
-how and in what ways the resulting embeddings are
-correlated with a social network-based representation of communities(\cref{sec:embedding-analysis}).
+Community-Conditioned Language Models (CCLMs for short). These models
+are conditioned on a vector representation (or embedding), which
+varies for each community. Hence, they learn \emph{community
+  embeddings}.  Even though we test which architectures best make use
+of the community information, our purpose is not to improve models in
+terms of perplexity. Rather we want to extract the best set of
+community embeddings and test in what ways the resulting embeddings
+predict the social structure of subreddits. For this purpose, we test
+the correlation with a social network-based representation of
+communities (\cref{sec:embedding-analysis}).
+
 % , and model the degree to which a

### 2021-11-29 paper/paper1.tex
lstm vs transformer

@@
 unidirectional LSTM \citep{Hochreiter1997} and a Transformer
-\citep{Vaswani2017}.  In either case, the model is organised as a
+\citep{Vaswani2017}. Even though the transformer is dominating LSTM in
+the state of the art, it achieves its dominance partly thanks to the
+availability of very large data sets \jp{Bill: can you add relevant citations?
+  (BERT, GPT3, etc.)},
+which are not available to us.\footnote{Fine-tuning existing models
+  falls outside our methodology, which consists in fundamentally
+  changing the structure of the network by concatenating
+  community embeddings.}.
+
+In either case, the model is organised as a
 standard $3$-layer neural sequence encoder, where the input for the
@@
 (with $\IG_M < 1$).
+
+We additionally observe that the best LSTM model achieves better
+performance than the best Transformer model. This partly contradicts
+the fact that Transformer-based models are the ones used in
+state-of-the art models for mainstream tasks. We propose that the
+cause is that we are training all models from scratch on a relatively
+small dataset. In this situation, the relatively smaller number of
+parameters of the LSTM means that it can generalise better than the
+Transformer.
 

### 2021-12-10 langid_analysis.py
langid analysis

@@ -1,0 +1,29 @@
+import pandas as pd
+from pathlib import Path
+import data
+import numpy as np
+import json
+import os
+from collections import Counter
+
+model_family_dir = Path('model/reddit')
+fields = data.load_fields(model_family_dir)
+comms = fields['community'].vocab.itos
+comms_ = []
+
+langs = {}
+for comm in comms:
+    try:
+        with open(f'data/langid/{comm}.test.txt') as f:
+            langs[comm] = Counter([eval(line)[0] for line in f.readlines()])
+            if sum(langs[comm].values()) == 0:
+                comms_.append(comm)
+    except:
+        comms_.append(comm)
+
+comms = [c for c in comms if not c in comms_]
+
+english_ratio = pd.Series({c: langs[c]['en'] / sum(langs[c].values()) for c in comms})
+
+english_ratio.mean()
+

### 2021-12-10 paper/paper1.tex
clarify relevance of most salient pairs

@@
 of the year 2015.
-Most but not all of these communities (with about 10 exceptions) 
-are mostly English-language.
 Each community corpus consist of \num{42000} randomly selected messages from the year 2015.
@@
 
-Messages were preprocessed as follows. 
+Most of the communities are predominantly English-language.
+Using \texttt{langid.py} \citep{Lui2012}, we observe that a majority of the overall
+comments are classified as English (94.6\% of the test set) and a majority 
+of communities (396 of 510) have more than 95\% of their comments classified as English.\footnote{
+  We performed the classification an the message level and took the language to which \texttt{langid.py}
+  assigned the highest probablity, but it is worth noting that code switching is very common on social media.}
+However, it would be difficult to set a principled cutoff since 12 communities have between
+50 and 95\% of their comments classified as English. 
+Furthermore, relying on a tool like \texttt{langid.py} may bias 
+the dataset away from communities that use a non-hegemonic variety of English.
+For that reason and becasue the amount of non-English data is realtively small,
+we do not attempt to filter the data.
+
+Messages were preprocessed as follows:
 We excluded the content of block quotes, code blocks, and tables.
@@
 concrete sense of what this method yields, we first survey some of the
-most salient community pairs.
+most salient community pairs. We stress that this survey is not meant
+as a rigorous statistical analysis, as we shall see. Rather it is
+meant to give a flavor of discrepancies and similarities existing
+between linguistic and social relations.
 

### 2021-12-12 paper/paper1.tex

--- 2021-11-29 paper/paper1.tex
@@
 their interlocutors, based on joint membership in a community
-\cite{Stalnaker2002, Clark1996}.  Such a \emph{speech community} \citep{Gumperz1972} 
-may be of any size, from the very small, like members of a particular friend group, 
+\cite{Stalnaker2002, Clark1996}.  Such \emph{speech communities} \citep{Gumperz1972} 
+range in size from the very small, like members of a friend group, 
 to the very large, like speakers of English; however as \citet{Eckert1992}
 point out, it is \emph{communities of practice}---defined by mutual social engagement
-in some common activity---that are the primary locus of linguistic variation.
+in a common activity---that are the primary locus of linguistic variation.
 
@@
 % \citep{Hovy2015,Yang2017}. \jp{Move this to related work so we can get to the point quicker.}
+Variation is an important object of study in sociolinguistics,
+and is naturally suited for computational analysis \citep{Nguyen2016}.
 Most previous computational work on linguistic variation 
@@
 \citep{Burger2011,Ciot2013,Bamman2014}, age \citep{Nguyen2013}, and
-geographic location \citep{Eisenstein2010,Bamman2014a}.
-Recent exceptions are \citet{DelTredici2017} who use a modified
+geographic location \citep{Eisenstein2010,Bamman2014a}.%
+\footnote{Recent exceptions are \citet{DelTredici2017} who use a modified
 skip-gram model to show that lexical semantic variation occurs even across 
@@
 \citet{Lucy2021}, who use contextualised word representations to
-investigate community-level variation in word sense distributions.
-
+investigate community-level variation in word sense distributions.}
 In the present work, we investigate linguistic variation
@@
 varies for each community. Hence, they learn \emph{community
-  embeddings}.  Even though we test which architectures best make use
+  embeddings}.  Although we test which architectures best make use
 of the community information, our purpose is not to improve models in
-terms of perplexity. Rather we want to extract the best set of
-community embeddings and test in what ways the resulting embeddings
+terms of perplexity, but rather we want to extract the best set of
+community embeddings and test how the resulting embeddings
 predict the social structure of subreddits. For this purpose, we test
@@
 unidirectional LSTM \citep{Hochreiter1997} and a Transformer
-\citep{Vaswani2017}. Even though the transformer is dominating LSTM in
-the state of the art, it achieves its dominance partly thanks to the
-availability of very large data sets \jp{Bill: can you add relevant citations?
-  (BERT, GPT3, etc.)},
+\citep{Vaswani2017}. Although Transformer-based language models are considered
+state of the art, they achieve dominance partly thanks to the
+availability of very large data sets \citep[e.g.,][]{Devlin2019,Brown2020},
 which are not available to us.\footnote{Fine-tuning existing models
-  falls outside our methodology, which consists in fundamentally
-  changing the structure of the network by concatenating
-  community embeddings.}.
+  is not compatible with our methodology, since we would fundamentally
+  change the structure of the network by concatenating
+  community embeddings.}
 
@@
 layer. What we have described so far does not take community into account
-and as such we call them \emph{uncodnitioned models}, but the same
+and as such we call them \emph{unconditioned models}, but the same
 encoder architecture also forms the core of our conditioned models.
 
-In the conditional models, we add a \emph{community embedding} parameter.
+In the CCLMs, we add a \emph{community embedding} parameter.
 This parameter is concatenated (at each time step) with the hidden layer of the
@@
 
-%TODO: Diagram of model?
-
-\subsection{Data sets}
+\subsection{Data sets}\label{sec:data}
 
@@
   from the archive at \url{https://pushshift.io/}.
-  \cite{Baumgartner2020}.}
+  \cite{Baumgartner2020}. Code for reproducing our dataset, as well as
+ our pre-trained community embeddings are available at URL.} %TODO ADD URL for camera ready
 %
@@
 of the year 2015.
-Most but not all of these communities (with about 10 exceptions) 
-are mostly English-language.
-Each community corpus consist of \num{42000} randomly selected messages from the year 2015.
-For each community, we reserve \num{1000} messages each for development and testing,
+Ignoring empty and deleted comments, we randomly sampled
+\num{42000} messages from 2015 for each community.
+We reserved \num{1000} messages from each community for development and testing,
 leaving a total of \num{20.4}M messages for training.
 
-Messages were preprocessed as follows. 
-We excluded the content of block quotes, code blocks, and tables.
-We removed any markup (formatting) commands from the remaining text, 
+Using \texttt{langid.py} \citep{Lui2012}, we observe that a majority of the overall
+messages are classified as English (94.6\% of the test set) and X of \num{510} communities %TODO
+have more than half of their messages classified as English.\footnote{See \cref{sec:ethics} for futher
+  discussion.}
+
+Messages were preprocessed as follows:
+We excluded the content of block quotes, code blocks, and tables and
+removed markup (formatting) commands, 
 extracting only rendered text.
-We tokenized the messages using the default English model for the SpaCy tokenizer 
+Messages were tokenized using the default English model for the SpaCy tokenizer 
 Version 2.2.3 \citep{Honnibal2017}.
@@
 
-We use a vocabulary size of \num{40000} tokens, including a special
-out-of-vocabulary token.  The vocabulary consists of the most frequent
-tokens across all communities.
-
-For the LSTM architecture, we trained the models on a simple left-to-right language
-modeling task with cross-entropy loss.  Because
+The models use a vocabulary of \num{40000} tokens (including a special
+out-of-vocabulary token), consisting of the most frequent tokens
+across all communities.
+
+We trained the models on a simple 
+auto-regressive language modeling task with cross-entropy loss.  Because
 the transformer operates on all tokens in the sequence at once, the
-language modeling was achieved by masking and incrementally un-masking
-input tokens in a left-to-right fashion (so we have an
-auto-regressive model).  We used the AdamW
+inputs to the model were masked and incrementally un-masked.
+We used the AdamW
 \citep{Loshchilov2019} optimisation algorithm, with an initial
@@
 %
-We used a batch size of \num{256} and a maximum sequence length of
+The batch size was \num{256} and the maximum sequence length set to
 \num{64} tokens, truncating longer messages (16.8\% of messages were
@@
 
-All experiments use models with \num{3} sequence encoder layers,
+All experiments use models with \num{3} encoder layers,
 each with hidden (and token embedding) size of \num{256}. 
 The transformer models had \num{8} attention heads per layer.\footnote{
-  This number of attention heads was chosen in part to give the LSTM and transformer
+  This number of attention heads was chosen to give the LSTM and transformer
   models a comparable number of parameters 
@@
 two epochs in a row, and used the weights from the epoch with the
-smallest validation loss for testing.  In general, transformer models
-were trained for about half as long as the LSTM models (\cref{tab:model-results}, 
-test epoch).
+smallest validation loss for testing.
+
+%In general, transformer models
+%were trained for about half as long as the LSTM models (\cref{tab:model-results}, 
+%test epoch).
 
@@
   together. The messages in these communites adhere to highly regular formats, 
-  which are presumably conventional to the communities.}
+  which are presumably conventional to the community.}
 %
@@
 The boxes indicate upper and lower quartiles, while the whiskers are placed at the upper and lower maximum, with communities more than $1.5\times IQR$ (inter-quartile range) above the upper quartile considered outliers (represented as dots).
-The three most extreme outliers are excluded from this view.}.% \cref{fig:ppl-info-gain}).
+The three most extreme outliers are excluded from this view.}% \cref{fig:ppl-info-gain}).
    %See also table 1 in the supplementary materials % not very useful for this actually since we sort by indisc and it's only one model anyway.
@@
 matter is relevant to a broad range of people.  Conditioning the model
-on community has the most benefit for narrower special-interest
+on community appears to have the most benefit for narrower special-interest
 subreddits, such as those organised around a certain videogame, sports
-team, or subculture.  This makes sense intuitively, since communities
-with more niche topics would tend to have more specialised vocabulary
-and other linguistic patterns.
+team, or subculture.  This supports the idea that communities of practice
+are the primary locus of linguistic variation.
+%makes sense intuitively, since communities
+%with more niche topics would tend to have more specialised vocabulary
+%and other linguistic patterns.
 
@@
 
-In this section we investigate whether (and how much) community
-embeddings obtained from the CCLMs correlate with the community-member
-relationship on Reddit.
+In this section we investigate the degree to which CCLM community
+embeddings correlate with the community-member relationship on Reddit.
 
@@
   observed similar results for other values of $l_c$.}
-with the community embedding created by \citet{Kumar2018},
+with the community embedding created by \citet{Kumar2018},\footnote{
+  Available at \url{https://snap.stanford.edu/data/web-RedditEmbeddings.html}}
 which were generated using using a negative-sampling optimization
@@
 
-Unsuprisingly, the first category (\cref{fig:embedding-differences-examples}, left)yields communities that are qualitatively very similar. 
+Unsuprisingly, the first category (\cref{fig:embedding-differences-examples}, left) yields communities that are qualitatively very similar. 
 The \subreddit{SSBPM} and \subreddit{darksouls} communities are focused
@@
 it was trained, it does not have to be the case that communities with
-a similar representation have a high overlap in their membership,
-but rather that the \emph{kinds of members} of the first community
-are also the \emph{kinds of members} who would belong to the second, 
-based on their membership in other communities.
+a similar representation have a high overlap in membership,
+but rather that members from the first community look similar
+to members from the second community in terms of their \emph{overall}
+community membership profile.
 
@@
 German-language subreddits, but the former is comprised of fans of a
-computer gaming-related YouTube channel, while the later is more
+computer gaming YouTube channel, while the later is more
 general-interest.
@@
 
-\subsection{Comparing embeddings: procrustes method}
+\subsection{Comparing embeddings: Procrustes method}
 \label{sec:statistical-evidence}
@@
 The anecdotal evidence suggests that for some communities, proximity
-in linguistic patterns do match social proximity… and for others it
+in linguistic patterns do match social proximity and for others it
 does not. Unfortunately, it seems hard, if not impossible, to draw a
-general conclusion from this way. In this section, we propose a
-systematic approach thanks to which we can quantify the correlation
+general conclusion from this mode of analysis. In this section, we propose a
+systematic approach with which we can quantify the correlation
 between social proximity and linguistic proximity, and measure its
@@
 \[d(L,S) = n - \mathsf {Tr} (Σ)\] where the matrix \(Σ\) is obtained by the singular value
-decomposition (SVD) \(U^TΣV = LS^T\). Additionally the vectors of \(U\) and \(V\) give
+decomposition (SVD) \(U^TΣV = LS^T\). The vectors of \(U\) and \(V\) give
 the directions of correlation respectively of \(L\) and \(S\). That is, each
@@
 
-As common when doing SVD, we can arrange \(U\), \(V\) and \(Σ\) such
+As is common when doing SVD, we arrange \(U\), \(V\) and \(Σ\) such
 that \(σ_i > σ_j\) iff \(i < j\). Doing so, the largest singular value
@@
 significant correlation between \(L\) and \(S\). A 4-sigma difference
-has less than one percent chance of occurring randomly. In our case:
-we observe a difference between 61 and 68 standard deviations
+has less than one percent chance of occurring randomly. In our case,
+we observe a difference of between 61 and 68 standard deviations
 (\cref{tab:embedding-correlations}). This definitely indicates a
@@
 There are several recent studies that aim to measure \emph{linguistic distinctiveness}
-at the level of speech community, which is an possible interpretation of 
-of community-stratified information gain the CCLM over its unconditioned counterpart \cref{sec:ig}. Whereas the metrics in previous work are based on lexical frequency
-\citep{OConnor2010,Zhang2018,Lucy2021}, CCLM information gain is capable of capturing 
-distinctiveness at other levels of linguistic analysis. Further work is needed to
+at the level of speech community \citep{OConnor2010,Zhang2018,Lucy2021}, 
+which is an possible interpretation of the community-stratified information gain 
+of the CCLM over its unconditioned counterpart \cref{sec:ig}. 
+Whereas the metrics in previous work are based on lexical frequency
+(and in the case of \citet{Lucy2021}, word sense distributions),
+CCLM information gain is capable of capturing 
+distinctiveness at multiple levels of linguistic analysis. 
+However, further work is needed to
 investigate exactly what kinds of variation are captured.
@@
 The pattern of information gain by community is similar across
-architectures.  Communities that benefit most from the conditioned
+architectures;  communities that benefit most from the conditioned
 model behave that way for both the LSTM and transformer.  However,
@@
 difference in information gain between the $l_c=0$ and $l_c=3$ LSTMs 
-seem to be organised around trading collectables or organising virtual meetups
-(\emph{friendsafari},
-\emph{Fireteams},
-\emph{RandomActsOfGaming},
-\emph{randomactsofcsgo},
-\emph{Pokemongiveaway},
-\emph{ACTrade},
-\emph{SVExchange})
-Why was the model where the community information was incorporated before
-the LSTM able to take better advantage of it for these communities in particular? 
-It would be interesting to investigate these differences in future work, since it
+are organised around trading collectables or organising virtual meetups
+(e.g.,
+%\subreddit{friendsafari},
+%\subreddit{Fireteams},
+%\subreddit{RandomActsOfGaming},
+%\subreddit{randomactsofcsgo},
+\subreddit{Pokemongiveaway},
+\subreddit{ACTrade}, and
+\subreddit{SVExchange}).  
+These communities tended to have highly conventionalized ways negotiating trades
+and coordinating meetups.
+%Why was the model where the community information was incorporated before
+%the LSTM able to take better advantage of it for these communities in particular? 
+It would be interesting to investigate these differences further in future work, since it
 could reveal differences in the kind of linguistic variation the
@@
 
-Our main result is that the representation of communities built by
-CCLM has a positive correlation with user co-occurrence patterns.
+Our main result is that community representations learned by
+CCLMs are positively correlation with user co-occurrence patterns.
 Even though such \emph{homophilic} correlation is a core hypothesis of 
@@
 we believe that this study is the first to test it at 
-the level of communities of practice using computatial methods.
+the level of communities of practice using computational methods.
 Furthermore, it appears that our method (correlating
 linguistic embeddings and social embeddings) is novel. Indeed, even
-though the procrustes method has been used to correlate two sets of
+though the Procrustes method has been used to correlate two sets of
 linguistic embeddings \emph{for the same model}, we find no evidence
@@
 
-\section{Ethical considerations}
-
-Our work uses pulicly available data from Reddit,
+\section{Ethical considerations}\label{sec:ethics}
+
+\paragraph{Data privacy}
+Our work uses publicly available data from Reddit,
 collected from the API made available by \citet{Baumgartner2020}.
@@
 (see \citet{Gliniecka2021} for discussion). Reddit users are not,
-in general, aware of the possiiblity that their data will be used 
+in general, aware of the possibility that their data will be used 
 for research purposes, and deleted posts can persist in archive formats.
-We are not reseasing any data or trained models as part of this study,
-since the data is publicly available and our models have low enough
-computational requirements as to pose little barrier for erplication.
-
+We do not release any data, since the it is already publicly available 
+and duplicating the dataset increases the likelihood that deleted 
+posts will persist.
+
+The paper does not include any text that could be linked back to 
+personally identifiable information.
+We do release our trained community embeddings, but the dimensionality 
+is low enough that we believe it would be impossible to reproduce 
+personally identifiable information from them.
+
+\paragraph{Language identification}
+As discussed in section \cref{sec:data}, we decided not to filter our data for non-English 
+comments.
+Although our focus in this paper is intra-language variation, 
+language identification has the potential to introduce bias by reinforcing hegemonic
+language classes and the boundaries between them.
+In our case, filtering out messages classified as non-English would bias the 
+dataset away from non-standard and code-switched language varieties,
+which are of interest in the current work.
+
+Nevertheless, the representations learned by our model are (necessarily)
+relative to the set of communities we chose. 
+Thus the learned representations for non-English communities tend to be more similar to each other 
+than to other communities that use mostly English, even if their predominant languages are not the same.
+This would probably not be the case if the distribution of messages 
+was more varied across hegemonic language classes; our work cannot be used to conclude,
+for example, that there is more variation within English than between Dutch and German.
+
+\paragraph{Subjective analysis}
 In the qualitative discussion offered in \cref{sec:storytelling},

### 2021-12-12 paper/paper1.tex
small edits

@@
 messages are classified as English (94.6\% of the test set) and X of \num{510} communities %TODO
-have more than half of their messages classified as English.\footnote{See \cref{sec:ethics} for futher
-  discussion.}
+have more than half of their messages classified as English.
+Given the small amount of non-English data, 
+we decided that the bias introduced by attempting to filter message
+by language outweighted the potential benefits of doing so.\footnote{
+  See \cref{sec:ethics} for futher discussion.}
 
@@
 personally identifiable information.
-We do release our trained community embeddings, but the dimensionality 
-is low enough that we believe it would be impossible to reproduce 
-personally identifiable information from them.
+We do release our trained community embeddings, but they have 
+low dimensionality and pose a low risk for exposing personally
+identifiable information.
 
 \paragraph{Language identification}
-As discussed in section \cref{sec:data}, we decided not to filter our data for non-English 
+As mentioned in section \cref{sec:data}, we decided not to filter our data for non-English 
 comments.

### 2021-12-14 paper/paper1.tex
update to reply

@@
   and find that they are highly correlated with a social network-based 
-  representation that doesn't use any linguistic input.
+  representation that does not use any linguistic input.
 \end{abstract}
@@
 
-% Not only is linguistic variation an object of inquiry in its own
-% right, but it is also important to consider it in the development of
-% robust, equitable language technology.  Previous work has shown that
-% by taking demographic information into account, models can achieve
-% better performance on various standard NLP tasks
-% \citep{Hovy2015,Yang2017}. \jp{Move this to related work so we can get to the point quicker.}
 Variation is an important object of study in sociolinguistics,
@@
 we decided that the bias introduced by attempting to filter message
-by language outweighted the potential benefits of doing so.\footnote{
+by language outweighted the potential benefits.\footnote{
   See \cref{sec:ethics} for futher discussion.}
@@
 \num{93.58} for the best conditional LSTM model 
-(\cref{fig:comm-stratified-box}; also see supplementary materials for perplexity by community). %TODO keep? 
+(\cref{fig:comm-stratified-box}; also see supplementary materials for perplexity by community). 
 The conditioned models also perform differently across different communities---%
@@
 We consider communities from three different selection criteria: 
-Those with high social \emph{and} linguistic similarity 
+Those with high linguistic \emph{and} social similarity 
 (where the sum of the two is highest),
@@
 That the three football club communities are so closely related points
-to an important point about the social embedding: Because of the way
+to an important point about the social embedding: because of the way
 it was trained, it does not have to be the case that communities with
@@
 experiences; \subreddit{legaladvice}, \subreddit{Cooking}, and
-\subreddit{loseit} all have a knowledge-sharing component to them;
+\subreddit{loseit} all involve sharing knowledge on a particular topic;
 \subreddit{running} and \subreddit{Coffee} are hobby-focused; and
@@
 over-fitting effects by taking the average distance rather than the
-maximum. % JP: TODO is the number from the table. this should probably be rewritten
+maximum. 
 In sum, it is clear that the CCLM embeddings predict some aspect the social-network embeddings--- but far from all of it.
@@
 
-%\begin{figure*}[t]
-%\begin{tikzpicture}
-  %\begin{groupplot}[
-    %group style={group size=2 by 1},
-    %yticklabels={,,}, xticklabels={,,},
-    %legend style={anchor=north, legend columns=1, font=\sffamily\tiny},
-  %]
-    %\nextgroupplot
-    %\commscatter{lstm-3-1-pca1}{lstm-3-1-pca2}
-    %\nextgroupplot
-    %\commscatter{transformer-3-3_pca1}{transformer-3-3_pca2}
-    %\legend{General interest,Videogames,Technology,Sports,Female-focused,Other}
-  %\end{groupplot}
-%\end{tikzpicture}
-%\caption{Plot of embeddings along the 2 principal dimensions of correlation between .... TODO of an LSTM (left, $c=1$) and transformer (right, $c=3$) model.
-%Color correspond the 5-means clustering of .... TODO}
-%\label{fig:comm-pca}
-%\end{figure*}
-
 \section{Related work}\label{sec:related-work}
@@
 
+While the focus of the paper is theoretical,
+computational models of variation can also support 
+robust, equitable language technology.  Previous work has shown that
+speaker demographics can improve performance on standard NLP tasks
+\citep{Hovy2015,Yang2017}. %\jp{Move this to related work so we can get to the point quicker.}
+
 %- DelTredici 2017
@@
 community-specific language, as witnessed by the information gain that
-they exhibit over a baseline, non-conditioned models.
-
-We find that the layer at which the community embedding is taken into
-account ($l_c$) has a weak effect on the information gain and the
+they exhibit over baseline unconditioned models.
+
+We find that the layer depth of the community embedding 
+($l_c$) has a weak effect on the information gain and the
 perplexity of the CCLMs.
@@
 average perplexity of \num{79.13}).
-This could have something to do with the community embedding and linear layer
+This could be due to the community embedding and linear layers
 interrupting the delicate layer normalisation between transformer encoder layers.
@@
 language classes and the boundaries between them.
-In our case, filtering out messages classified as non-English would bias the 
-dataset away from non-standard and code-switched language varieties,
+In our case, filtering out messages classified as non-English would 
+introduce bias by disproportionately removing messages in
+non-standard and code-switched language varieties,
 which are of interest in the current work.
@@
 Nevertheless, the representations learned by our model are (necessarily)
-relative to the set of communities we chose. 
+relative to the other communities in the dataset. 
 Thus the learned representations for non-English communities tend to be more similar to each other 
-than to other communities that use mostly English, even if their predominant languages are not the same.
+than to other communities that use mostly English, even if their predominant language is not the same.
 This would probably not be the case if the distribution of messages 

### 2021-12-14 paper/paper1.tex
fix langid numbers

@@
 Using \texttt{langid.py} \citep{Lui2012}, we observe that a majority of the overall
-messages are classified as English (94.6\% of the test set) and X of \num{510} communities %TODO
+messages are classified as English (95\% of the test set) and 498 of \num{510} communities %TODO
 have more than half of their messages classified as English.

### 2021-12-15 paper/paper1.tex
a
revise up to procrustes

@@
 
-\title{Community-conditioned language models for linguistic variation}
+\title{Vector Representations of Linguistic Communities and Associated Models}
+ % Representations -conditioned language models for linguistic variation
 
@@
 \begin{abstract}
-  Community-level linguistic variation is a core concept in sociolinguistics,
-  but more tools are needed to study variation from a computational perspective.
-  In this paper, we use conditional language models to investigate linguistic variation
+  Community-level linguistic variation is a core concept in sociolinguistics.
+  % but more tools are needed to study variation from a computational perspective.
+  In this paper, we use computational methods to investigate how much linguistic variation corresponds with social connections between communities.
+  Technically, use conditional language models to investigate linguistic variation
   across 510 online communities.
@@
 range in size from the very small, like members of a friend group, 
-to the very large, like speakers of English; however as \citet{Eckert1992}
+to the very large, like speakers of English. However, as \citet{Eckert1992}
 point out, it is \emph{communities of practice}---defined by mutual social engagement
@@
 Variation is an important object of study in sociolinguistics,
-and is naturally suited for computational analysis \citep{Nguyen2016}.
+and is naturally amenable to computational analysis \citep{Nguyen2016}.
 Most previous computational work on linguistic variation 
@@
 in online communities on the social media website Reddit. 
-In contrast to previous work, we investigate variation across 
+This contrasts with previous work: we investigate variation across 
 speech communities, rather than macro-social categories.
@@
 varies for each community. Hence, they learn \emph{community
-  embeddings}.  Although we test which architectures best make use
-of the community information, our purpose is not to improve models in
+  embeddings}.  Even though we report which architectures make best use
+of the community information, our primary purpose is not to improve models in
 terms of perplexity, but rather we want to extract the best set of
 community embeddings and test how the resulting embeddings
-predict the social structure of subreddits. For this purpose, we test
+predict the social structure of subreddits. Technically, we test
 the correlation with a social network-based representation of
@@
 \citep{Vaswani2017}. Although Transformer-based language models are considered
-state of the art, they achieve dominance partly thanks to the
+state-of-the-art, they achieve dominance partly thanks to the
 availability of very large data sets \citep[e.g.,][]{Devlin2019,Brown2020},
 which are not available to us.\footnote{Fine-tuning existing models
-  is not compatible with our methodology, since we would fundamentally
+  is not compatible with our methodology, because we fundamentally
   change the structure of the network by concatenating
-  community embeddings.}
+  community embeddings with hidden states at various levels.} Thus the LSTM is a worthy model to test for us.
 
@@
 Messages were preprocessed as follows:
-We excluded the content of block quotes, code blocks, and tables and
+we excluded the content of block quotes, code blocks, and tables and
 removed markup (formatting) commands, 
@@
 smallest validation loss for testing. 
+Each training epoch took approximately 1.5 hours of GPU time.
 
@@
 
-We observe empirically that across all the models we
+We observe that across all the models we
 tested, communities where conditioning has the least effect tend to be
@@
 subreddits, such as those organised around a certain videogame, sports
-team, or subculture.  This supports the idea that communities of practice
+team, or subculture.  These empirical observations corroborate the idea that communities of practice
 are the primary locus of linguistic variation.
@@
 
-A difficulty is that the social embeddings and linguistic embeddings
-range over completely unrelated spaces, so one cannot use the usual
-cosine similarity metric \emph{across} this boundary. One can,
+When comparing social embeddings and linguistic embeddings, a difficulty is that they
+range over completely unrelated spaces. Thus one cannot use the usual
+cosine similarity metric \emph{between} these spaces. One can,
 however, use cosine similarity between \emph{pairs} of communities,
-and verify the similarities correlated between linguistic to social
-embeddings.  This gives us a way of characterizing the differences
+and verify that the similarities are correlated between linguistic and social
+embeddings.  This gives a way of characterizing the differences
 between the two kinds of community representation. To get a more
@@
 
-We simultaneously have examples of high correlation between social and
-linguistic embeddings, but also of low correlation. This is
-essentially anecdotal evidence: we must resort to statistical tools to
-extract the general trends.
-A straightforward, but ultimately flawed,
-way to measure how similar the two spaces are is to consider each pair
-of communities \((i,j)\), and compute the correlation between the
-cosine similarities of both embeddings.
+In sum, empirical observation simultaneously reveals examples of high
+and low correlation between social and linguistic embeddings.  To
+quantify correlation and extract the general trends, we must resort to
+statistical tools, as we do below.
+
+A straightforward (but ultimately flawed) way to measure how similar
+the two spaces are would be to generalise the above method, by
+consider each pair of communities \((i,j)\), and compute the
+correlation between the cosine similarities of both embeddings.
 

### 2021-12-15 paper/paper1.tex
other revisions

@@
 
-The anecdotal evidence suggests that for some communities, proximity
-in linguistic patterns do match social proximity and for others it
-does not. Unfortunately, it seems hard, if not impossible, to draw a
-general conclusion from this mode of analysis. In this section, we propose a
+% The anecdotal evidence suggests that for some communities, proximity
+% in linguistic patterns do match social proximity and for others it
+% does not. Unfortunately, it seems hard, if not impossible, to draw a
+% general conclusion from this mode of analysis.
+
+In this section, we propose a
 systematic approach with which we can quantify the correlation
@@
 
-The problem with this approach is that even if several dimensions of
-\(L\) and \(S\) are correlated, they will not coincide in the
-\emph{representation}. That is, applying a simple rotation (orthogonal
-transformation) on either matrix widely changes the \(||L-S||_F\)
-correlation metric. Another way to consider the same problem is that,
-applying such a rotation does \emph{not} change the cosine similarities
-considered in the previous section.
+The problem with the above metric is that even if several dimensions
+of \(L\) and \(S\) are correlated, they will not coincide in the
+\emph{representation} of embeddings. That is, applying a simple
+rotation (orthogonal transformation) on either matrix widely changes
+the \(||L-S||_F\) correlation metric.
+
+% Another way to consider the same problem is that,
+% applying such a rotation does \emph{not} change the cosine similarities
+% considered in the previous section.
 
 To make the metric independent of the representation (up to orthogonal
-transformations), we compute the \emph{minimum} distance between
-\(L_i\) and \(S_i\), for any orthogonal matrix \(\Omega\) applied to
-\(L\):
+transformations, which preserve cosine similarities), we compute the
+\emph{minimum} distance between \(L_i\) and \(S_i\), for any
+orthogonal matrix \(\Omega\) applied to \(L\):
 
@@
 
-Along these lines, \citet{OConnor2010} uses a Bayesian generative
+A piece of work with similar goals as ours is that of \citet{OConnor2010}, which uses a Bayesian generative
 model to infer communities from variation in text data.  In contrast
@@
 which is an possible interpretation of the community-stratified information gain 
-of the CCLM over its unconditioned counterpart \cref{sec:ig}. 
+of the CCLM over its unconditioned counterpart (\cref{sec:ig}). 
 Whereas the metrics in previous work are based on lexical frequency
@@
 
-While the focus of the paper is theoretical,
+While the focus of this paper is sociolinguistic aspects,
 computational models of variation can also support 

### 2021-12-16 paper/paper1.tex
fix typo

@@
   % but more tools are needed to study variation from a computational perspective.
-  In this paper, we use computational methods to investigate how much linguistic variation corresponds with social connections between communities.
-  Technically, use conditional language models to investigate linguistic variation
-  across 510 online communities.
-  We probe the community representations learned by the conditional language model
-  and find that they are highly correlated with a social network-based 
+  In this paper, we use computational methods to investigate the degree to which  
+  linguistic variation corresponds with social connections between communities.
+  Technically, we use conditional language models to learn vector representations
+  for 510 online communities.
+  We find that they are highly correlated with a social network-based 
   representation that does not use any linguistic input.

### 2022-09-19 paper/paper1.tex
changes from last round of reviews

@@
 
-\date{}
-
 \begin{document}
+
 \maketitle
@@
 Community-level linguistic variation is a core concept in sociolinguistics.
-  % but more tools are needed to study variation from a computational perspective.
-  In this paper, we use computational methods to investigate the degree to which  
-  linguistic variation corresponds with social connections between communities.
-  Technically, we use conditional language models to learn vector representations
-  for 510 online communities.
-  We find that they are highly correlated with a social network-based 
-  representation that does not use any linguistic input.
+In this paper, we use conditioned neural language models to learn vector representations for 510 online communities.
+We use these representations to measure linguistic variation between communities and investigate the degree to which linguistic variation corresponds with social connections between communities.
+We find that our sociolinguistic embeddings are highly correlated with a social network-based representation that does not use any linguistic input.
 \end{abstract}
@@
 geographic location \citep{Eisenstein2010,Bamman2014a}.%
-\footnote{Recent exceptions are \citet{DelTredici2017} who use a modified
-skip-gram model to show that lexical semantic variation occurs even across 
-different communities organised around the same topic and 
-\citet{Lucy2021}, who use contextualised word representations to
-investigate community-level variation in word sense distributions.}
-In the present work, we investigate linguistic variation
-in online communities on the social media website Reddit. 
-This contrasts with previous work: we investigate variation across 
-speech communities, rather than macro-social categories.
+In the present work, however, we investigate linguistic variation
+across online communities in the social media website Reddit.
 
@@
 are conditioned on a vector representation (or embedding), which
-varies for each community. Hence, they learn \emph{community
-  embeddings}.  Even though we report which architectures make best use
-of the community information, our primary purpose is not to improve models in
-terms of perplexity, but rather we want to extract the best set of
-community embeddings and test how the resulting embeddings
-predict the social structure of subreddits. Technically, we test
-the correlation with a social network-based representation of
+varies by community. Hence, they learn \emph{community
+  embeddings}.  
+We report which architectures make best use
+of the community information (\cref{sec:performance}), 
+however our our primary purpose is not to improve language models in
+terms of perplexity, but rather to extract 
+community embeddings that capture linguistic similarities between
+communities and test how the resulting embeddings
+correspond to the social structure of subreddits. To that end, we test
+the how well the community embeddings correlate with a social network-based representation of
 communities (\cref{sec:embedding-analysis}).
+
+The contributions of this work are twofold. 
+First, we develop a language model-based
+community embedding that we show is correlated with (but still different from) 
+an embedding based on community membership alone.
+Second, the method we describe for testing the correlation between two embeddings
+from different models is, to our knowledge, novel to computational linguistics.
 
@@
 
-We experiment with two kinds of model architecture: a simple
-unidirectional LSTM \citep{Hochreiter1997} and a Transformer
+We experiment with two kinds of model architecture: simple
+unidirectional LSTM \citep{Hochreiter1997} and a masked Transformer
 \citep{Vaswani2017}. Although Transformer-based language models are considered
@@
 
-In the CCLMs, we add a \emph{community embedding} parameter.
+In the CCLMs, we add a \emph{community embedding} parameter, 
+which varies depending on the community of origin of the input sample. 
 This parameter is concatenated (at each time step) with the hidden layer of the
@@
 
-The models use a vocabulary of \num{40000} tokens (including a special
+Models used a vocabulary of \num{40000} tokens (including a special
 out-of-vocabulary token), consisting of the most frequent tokens
@@
 auto-regressive language modeling task with cross-entropy loss.  Because
-the transformer operates on all tokens in the sequence at once, the
+the Transformer operates on all tokens in the sequence at once, the
 inputs to the model were masked and incrementally un-masked.
@@
 \citep{Loshchilov2019} optimisation algorithm, with an initial
-learning rate of \num{0.001}, with no extra control on the decay
+learning rate of \num{0.001} and no extra control on the decay
 of learning rate.
@@
 each with hidden (and token embedding) size of \num{256}. 
-The transformer models had \num{8} attention heads per layer.\footnote{
-  This number of attention heads was chosen to give the LSTM and transformer
+The Transformer models had \num{8} attention heads per layer.\footnote{
+  This number of attention heads was chosen to give the LSTM and Transformer
   models a comparable number of parameters 
@@
 
-\section{CCLM Performance}
+\section{CCLM Performance}\label{sec:performance}
 
@@
 than their respective unconditioned baseline models, 
-meaning information gain above one (\cref{tab:model-results}).
-While the absolute performance ($\Ppl_{M_j}$) of the LSTM models is better,
-the best transformer models have somewhat higher information gain
+(i.e., $\IG_M > 1$, \cref{tab:model-results}).
+While the absolute performance ($\Ppl_{M}$) of the LSTM models is better,
+the best Transformer models have somewhat higher information gain
 than their LSTM counterparts.
@@
 but all of the conditioned models perform similarly well.
-For the transformer, the best model incorporates the community information
+For the Transformer, the best model incorporates the community information
 first, concatenating it directly to the word vectors ($l_c=0$).
 It performs similarly to the model that only integrates the community information
-after all all the transformer layers ($l_c=3$),
+after all all the Transformer layers ($l_c=3$),
 but the two middle-layer models actually perform worse than the unconditioned model
@@
 
-We additionally observe that the best LSTM model achieves better
-performance than the best Transformer model. This partly contradicts
-the fact that Transformer-based models are the ones used in
-state-of-the art models for mainstream tasks. We propose that the
-cause is that we are training all models from scratch on a relatively
-small dataset. In this situation, the relatively smaller number of
-parameters of the LSTM means that it can generalise better than the
-Transformer.
+%We additionally observe that the best LSTM model achieves better
+%performance than the best Transformer model. 
+%This partly contradicts
+%the fact that Transformer-based models are the ones used in
+%state-of-the art models for mainstream tasks. 
+%We propose that the
+%cause is that we are training all models from scratch on a relatively
+%small dataset. In this situation, the relatively smaller number of
+%parameters of the LSTM means that it can generalise better than the
+%Transformer.
 
@@
 \num{93.58} for the best conditional LSTM model 
-(\cref{fig:comm-stratified-box}; also see supplementary materials for perplexity by community). 
+(\cref{fig:comm-stratified-box}; also see \cref{sec:community-level-results} for detailed community-level results). 
 The conditioned models also perform differently across different communities---%
@@
 In this section we investigate the degree to which CCLM community
-embeddings correlate with the community-member relationship on Reddit.
+embeddings correlate with the social network structure of Reddit.
 
@@
 which were generated using using a negative-sampling optimization
-algorithm, with the message author-community co-occurrence matrix as ground
+algorithm, with the author-community co-occurrence matrix as ground
 truth, using data from January 2014 to April 2017.
@@
 information is used to create these embeddings: they only reflect the social
-relationship between communities, via 
-the user-community membership relation. In contrast, CCLM community embeddings depend in no
+relationship between communities via community membership. 
+In contrast, CCLM community embeddings depend in no
 way on which user is the author of any given message: we only use the
@@
 (where linguistic similarity is below these median and linguistic similarity is highest).\footnote{We use the LSTM ($l_c=1$) community vectors for these purposes, 
-  but results attain with the best transformer model.}\footnote{Median similarity
+  but results attain with the best Transformer model.}\footnote{Median similarity
   among pairs of communities was \num{0.177} for the social embedding
-  and \num{0.010} and \num{0.012} for the LSTM and transformer
+  and \num{0.010} and \num{0.012} for the LSTM and Transformer
   linguistic embeddings, respectively.}
@@
 likewise comprised of fans of a particular English football club.
-That the three football club communities are so closely related points
-to an important point about the social embedding: because of the way
-it was trained, it does not have to be the case that communities with
-a similar representation have a high overlap in membership,
-but rather that members from the first community look similar
-to members from the second community in terms of their \emph{overall}
-community membership profile.
+%That the three football club communities are so closely related points
+%to an important point about the social embedding: because of the way
+%it was trained, it does not have to be the case that communities with
+%a similar representation have a high overlap in membership,
+%but rather that members from the first community look similar
+%to members from the second community in terms of their \emph{overall}
+%community membership profile.
 
@@
 The analysis shows positive correlation for both the LSTM ($r=0.438$)
-and transformer ($r=0.452$) linguistic embeddings.\footnote{ By
+and Transformer ($r=0.452$) linguistic embeddings.\footnote{ By
   comparison, the correlation between the two linguistic embeddings is
@@
 
-There are several recent studies that aim to measure \emph{linguistic distinctiveness}
-at the level of speech community \citep{OConnor2010,Zhang2018,Lucy2021}, 
-which is an possible interpretation of the community-stratified information gain 
+Perhaps the most topically methodologically similar work to ours 
+is that of \citet{DelTredici2017}, who use a modified
+skip-gram model to show that lexical semantic variation occurs even across 
+different communities organised around the same topic.
+Their approach does not result in community level representations, however.
+
+There are several other recent studies that aim to measure \emph{linguistic distinctiveness}
+at the level of speech community \citep{OConnor2010,Zhang2018,Lucy2021}.
+Distinctiveness is one possible interpretation of the community-stratified information gain 
 of the CCLM over its unconditioned counterpart (\cref{sec:ig}). 
@@
 (with \num{68.74} for the unconditioned model).
-For transformer models, it varies a bit more, between \num{75.66} and \num{83.53},
+For Transformer models, it varies a bit more, between \num{75.66} and \num{83.53},
 but this seems to be mainly due to the poor performance of the models where
-the community embedding is inserted between transformer layers
-($l_c=2$ and $3$ both test above the unconditioned transformer's 
+the community embedding is inserted between Transformer layers
+($l_c=2$ and $3$ both test above the unconditioned Transformer's 
 average perplexity of \num{79.13}).
-This could be due to the community embedding and linear layers
-interrupting the delicate layer normalisation between transformer encoder layers.
+%This could be due to the community embedding and linear layers
+%interrupting the delicate layer normalisation between transformer encoder layers.
 
@@
 architectures;  communities that benefit most from the conditioned
-model behave that way for both the LSTM and transformer.  However,
+model behave that way for both the LSTM and Transformer.  However,
 there are some differences.  For example, many of the communities with the biggest 
@@
 of the community features encoded by different representations, 
-to make conclusions about \emph{particular} communities based on such a cursory and
-subjective analysis would be dubious and potentially harmful.
+to make conclusions about \emph{particular} communities based on such 
+an analysis would be dubious and potentially harmful.
 
 \bibliography{paper}
-\bibliographystyle{acl_natbib}
 

### 2022-10-24 paper/paper1.tex
some small changes

@@
 % Remove the "review" option to generate the final version.
-\usepackage[review]{acl}
+\usepackage{acl}
 
@@
 
-\title{Vector Representations of Linguistic Communities and Associated Models}
+\title{Conditional Language Models for Community-Level Linguistic Variation}
  % Representations -conditioned language models for linguistic variation
@@
 
-\author{First Author \\
-  Affiliation / Address line 1 \\
-  Affiliation / Address line 2 \\
-  Affiliation / Address line 3 \\
-  \texttt{email@domain} \\\And
-  Second Author \\
-  Affiliation / Address line 1 \\
-  Affiliation / Address line 2 \\
-  Affiliation / Address line 3 \\
-  \texttt{email@domain} \\}
+%\author{Bill Noble\\
+  %Affiliation / Address line 1 \\
+  %Affiliation / Address line 2 \\
+  %Affiliation / Address line 3 \\
+  %\texttt{email@domain} \\\And
+  %Second Author \\
+  %Affiliation / Address line 1 \\
+  %Affiliation / Address line 2 \\
+  %Affiliation / Address line 3 \\
+  %\texttt{email@domain} \\}
+
+\author{Bill Noble \and Jean-Philippe Bernardy\\
+  Centre for Linguistic Theory and Studies in Probability (CLASP) \\
+  Dept. of Philosophy, Linguistics and Theory of Science\\
+  University of Gothenburg \\
+\texttt{\{bill.noble@, jean.philippe.bernardy@\}.gu.se}}
 
@@
 to the very large, like speakers of English. However, as \citet{Eckert1992}
-point out, it is \emph{communities of practice}---defined by mutual social engagement
-in a common activity---that are the primary locus of linguistic variation.
+point out, it is \emph{communities of practice}\,---\,defined by mutual social engagement
+in a common activity\,---\,that are the primary locus of linguistic variation.
 
@@
 The conditional models were given a community embedding with \num{16} dimensions. 
-We experimented with every possible value for $l_c$ in a three-layer model ($l_c\in\{0,1,2,3\}$).
+We experimented with every possible value for $l_c$, the depth of the
+community embedding, in a three-layer model ($l_c\in\{0,1,2,3\}$).
 
@@
 (\cref{fig:comm-stratified-box}; also see \cref{sec:community-level-results} for detailed community-level results). 
-The conditioned models also perform differently across different communities---%
+The conditioned models also perform differently across different communities\,---\,%
 even among the best models, some communities have $\IG_{M_j} < 1$,
@@
 
-\newcommand{\modelboxplot}[3]{
+\newcommand{\modelboxplot}[4]{
   \addplot+[
     boxplot={draw position=#3, box extend=0.3}, 
-    draw=#2, mark=*, mark options={fill=#2, scale=0.5}, solid, fill=#2!10,
+    draw=#2, mark=*, mark options={fill=#2, scale=0.5}, solid, 
+    pattern=#4,
+    pattern color=#2,
     area legend] 
@@
     \foreach \lc in {0, ..., 3}{
-      \modelboxplot{lstm-3-\lc-ig}{violet}{\lc+0.2}
-      \modelboxplot{transformer-3-\lc-ig}{teal}{\lc-0.2}
+      \modelboxplot{lstm-3-\lc-ig}{violet}{\lc+0.2}{crosshatch dots}
+      \modelboxplot{transformer-3-\lc-ig}{teal}{\lc-0.2}{grid}
     \legend{,LSTM, Transformer}
@@
     ]
-    \modelboxplot{lstm-3-ppl}{violet}{0.2}
-    \modelboxplot{transformer-3-ppl}{teal}{-.2}
+    \modelboxplot{lstm-3-ppl}{violet}{0.2}{crosshatch dots}
+    \modelboxplot{transformer-3-ppl}{teal}{-.2}{grid}
   \end{axis}
@@
     ]
-    \addplot[draw=violet,fill=violet!10] table [y=lstm-sim] {floats/#1};
-    \addplot[draw=teal,fill=teal!10] table [y=transformer-sim] {floats/#1};
+    \addplot[draw=violet,pattern color=violet!30, pattern=crosshatch dots] table [y=lstm-sim] {floats/#1};
+    \addplot[draw=teal,pattern color=teal!30, pattern=grid] table [y=transformer-sim] {floats/#1};
     \addplot[draw=red,fill=red!10] table [y=social-sim] {floats/#1};
@@
     ]
-    \addplot[draw=violet,fill=violet!10] table [y=lstm-sim] {floats/high_ling_high_soc.csv};
-    \addplot[draw=teal,fill=teal!10] table [y=transformer-sim] {floats/high_ling_high_soc.csv};
+    \addplot[draw=violet,pattern color=violet!30,pattern=crosshatch dots] table [y=lstm-sim] {floats/high_ling_high_soc.csv};
+    \addplot[draw=teal,pattern color=teal!30,pattern=grid] table [y=transformer-sim] {floats/high_ling_high_soc.csv};
     \addplot[draw=red,fill=red!10] table [y=social-sim] {floats/high_ling_high_soc.csv};
@@
 the Mormon and Catholic churches, although their members have different 
-relationships towards those organizations---the former is intended for
+relationships towards those organizations\,---\,the former is intended for
 former members of the church, whereas the later is geared towards practicing
@@
 to interpret.  It is mostly easy to see why these communities would
-have different linguistic embeddings---%
+have different linguistic embeddings\,---\,%
 in all cases the topics are quite different.  The reason they have
@@
 different patterns in the \emph{social function} that people attribute
-to this particular social media website---%
+to this particular social media website\,---\,%
 people who use Reddit in one way are more likely to belong to
@@
 weakest of correlations. A further flaw is that the data points in
-\(C\) are \emph{not} distributed independently---far from it in fact,
+\(C\) are \emph{not} distributed independently\,---\,far from it in fact,
 since each data point is generated from \num{2} of \num{510}
@@
 of \(L\) and \(S\) are correlated, they will not coincide in the
-\emph{representation} of embeddings. That is, applying a simple
+\emph{representation} of embeddings. That is, re-aligning the embeddings by applying a simple
 rotation (orthogonal transformation) on either matrix widely changes
@@
 the directions of correlation respectively of \(L\) and \(S\). That is, each
-singular value \(\sigma_i\) (the elements of the diagonal matrix \(Σ\)), give a measure of how much
+singular value \(\sigma_i\) (the elements of the diagonal matrix \(Σ\)), gives a measure of how much
 correlation there is between the directions \(U_i\) and \(V_i\).
@@
 it suffices to check if its value is significantly larger than the same
-value for random linguistic embeddings \(L\). Even though this is the
-distribution of \(d(L,S)\) for random embeddings is difficult to
-compute analytically, one can instead evaluate it using a Monte Carlo
+value for random linguistic embeddings \(L'\). The 
+distribution of \(d(L',S)\) for random embeddings is difficult to
+compute analytically, but we can instead evaluate it using a Monte Carlo
 method.
 
-Doing so, we observed that \(d(L,S)\) exhibits a mean of
+Doing so, we observed that \(d(L',S)\) exhibits a mean of
 \(μ_d=431.39\) and a (Bessel's-corrected) standard deviation
@@
 and the cosine similarity with its true social embedding \(S_k\) is
-expected to be \(0.39\) = \((431.39 - 232.18) / 510\) --- accounting for
+expected to be \(0.39\) = \((431.39 - 232.18) / 510\)\,---\,accounting for
 over-fitting effects by taking the average distance rather than the
 maximum. 
-In sum, it is clear that the CCLM embeddings predict some aspect the social-network embeddings--- but far from all of it.
+In sum, it is clear that the CCLM embeddings predict some aspect the social-network embeddings\,---\,but far from all of it.
 
@@
 
-Perhaps the most topically methodologically similar work to ours 
-is that of \citet{DelTredici2017}, who use a modified
-skip-gram model to show that lexical semantic variation occurs even across 
+\citet{DelTredici2017}, use a modified
+skip-gram model to community-level linguistic variation.
+They show that lexical semantic variation occurs even across 
 different communities organised around the same topic.
@@
 
+\section*{Acknowledgements}
+
+This work was supported by grant 2014-39 from the Swedish Research Council (VR) for the establishment of the Centre for Linguistic Theory and Studies in Probability (CLASP) at the University of Gothenburg.
+
 \bibliography{paper}
